### 2020-07-15 arcticgrids.tex
initial commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{Impact of horizontal resolution on the meteorology and climate of the Arctic in CESM2.2}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
\authors{Adam R. Herrington \affil{1}, Marcus Lovferstrom \affil{2}, Peter H. Lauritzen \affil{1} and Andrew Gettelman \affil{1}}

 \affiliation{1}{National Center for Atmospheric Research, 1850 Table Mesa Drive, Boulder, Colorado, USA}
\affiliation{2}{Department of Geosciences, University of Arizona, 1040 E. 4th Street, Tucson, AZ USA}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

%\affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{=name=}{=email address=}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
[ enter your Abstract here ]
\end{abstract}

\section*{Plain Language Summary}
[ enter your Plain Language Summary here or delete this section]

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
\section{Introduction}
%
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

\section{Methods}
\section{Results}

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%

\section{Conclusions}

%Text here ===>>>

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURES
%
% \begin{figure}
% \includegraphics{example.png}
% \caption{caption}
% \end{figure}
%
% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
% \begin{figure}
% \noindent\includegraphics[width=\textwidth]{anothersample.png}
%\caption{caption}
%\label{pngfiguresample}
%\end{figure}
%
%
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
%

%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$}
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% >>>>	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% 	All funding sources related to this work from all authors
%
% 	Any real or perceived financial conflicts of interests for any
%	author
%
% 	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
%
% It is also the appropriate place to thank colleagues and other contributors.
% AGU does not normally allow dedications.

\acknowledgments
This material is based upon work supported by the National Center for Atmospheric Research (NCAR), which is a major facility sponsored by the NSF under Cooperative Agreement 1852977. Computing and data storage resources, including the Cheyenne supercomputer
(doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at NCAR.

The data presented in this manuscript is available at {\url{https://github.com/adamrher/2020-arcticgrids}}.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \bibliography{<name of your .bib file>} don't specify the file extension
%
% don't specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliography{ enter your bibtex bibliography filename here }
\bibliography{bib}

%Reference citation instructions and examples:
%
% Please use ONLY \cite and \citeA for reference citations.
% \cite for parenthetical references
% ...as shown in recent studies (Simpson et al., 2019)
% \citeA for in-text citations
% ...Simpson et al. (2019) have shown...
%
%
%...as shown by \citeA{jskilby}.
%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
%...has been shown \cite{jskilbye}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown by \cite <e.g.,>[and others]{lewin76}.
%
% apacite uses < > for prenotes and [ ] for postnotes
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%

\end{document}

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

### 2020-10-13 arcticgrids.tex
mods
Updates from Overleaf
intro

@@
 
-\title{Impact of horizontal resolution on the meteorology and climate of the Arctic in CESM2.2}
+\title{Impact of grids and dycores in CESM2.2 on the meteorology and climate of the Arctic}
 
@@
 % Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-\authors{Adam R. Herrington \affil{1}, Marcus Lovferstrom \affil{2}, Peter H. Lauritzen \affil{1} and Andrew Gettelman \affil{1}}
+\authors{Adam R. Herrington \affil{1}, Marcus Lofverstrom \affil{2}, Peter H. Lauritzen \affil{1} and Andrew Gettelman \affil{1}}
 
@@
 \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
 
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
+General Circulation Models are a powerful tool for simulating and understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
+Despite their importance, the numerics of GCMs have had to grapple with the \textit{pole-problem}, a numerical instability arising discretizing the equations of motions on a latitude-longitude grid, with the meridians converging at the poles.
+In the 1970's this issue was largely defeated through the wide-spread use of the global spectral-transform method in GCMs.
+But as computing power has increased, local numerical methods have become more desirable for their ability to run efficiently on massively parallel systems.
+As the pole-problem has re-emerged in contemporary climate models, some combination of reduced grids and polar filters are necessary to subdue this instability.
+An alternative approach is the use of unstructured grids.
+Unstructured grids permit quasi-uniform grid spacing globally, thereby eliminating the pole-problem.
+They are more flexible than structured, latitude-longitude grids, allowing for regional grid refinement that may be used to capture higher resolution in the polar regions.
+Regional refinement of polar regions may be desirable as latitude-longitude grids, by virtue of the convergence of meridians, have greater horizontal resolution in polar regions compared to a quasi-uniform grid containing the same degrees of freedom.
 
 \section{Methods}
+\subsection{Grids}
+\subsection{Dynamical cores}
+\subsubsection{Finite-volume model}
+\subsubsection{Spectral-element model}
+\subsection{Physical parameterizations}
+\subsection{Observational datasets}
+\subsubsection{ERA5}
+\subsubsection{LIVVkit 2.1}
+\subsection{TempestExtremes}
+\subsection{StormCompositer}
+
 \section{Results}
+
+\subsection{Tropospheric temperatures}
+\subsection{Inter-annual Variability}
+\subsection{Synoptic-scale storm characteristics}
+\subsection{Orographic gravity waves emanating from Greenland}
+\subsection{Katabatic winds emanating from Greenland}
+\subsection{Greenland surface mass balance}
 

### 2020-10-15 arcticgrids.tex
draft intro, grid figures

@@
 \linenumbers
+
 %%%%%%%
@@
 
-General Circulation Models are a powerful tool for simulating and understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, the numerics of GCMs have had to grapple with the \textit{pole-problem}, a numerical instability arising discretizing the equations of motions on a latitude-longitude grid, with the meridians converging at the poles.
-In the 1970's this issue was largely defeated through the wide-spread use of the global spectral-transform method in GCMs.
+General Circulation Models (GCMs) are a powerful tool for understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
+Despite their importance, the numerics of GCMs have had to grapple with the \textit{pole-problem} \cite{W2007JMSJ}, an economic bottleneck arising from the convergence of meridians into a polar singularity on latitude-longitude grids.
+In the 1970's this issue was largely defeated through wide-spread adaption of the spectral transform method in GCMs, which transforms grid point fields into an isotropic representation in wave space.
 But as computing power has increased, local numerical methods have become more desirable for their ability to run efficiently on massively parallel systems.
-As the pole-problem has re-emerged in contemporary climate models, some combination of reduced grids and polar filters are necessary to subdue this instability.
-An alternative approach is the use of unstructured grids.
+The pole-problem has thus re-emerged in contemporary climate models using latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to subdue this instability \cite{JW2010LNCSE}.
+An alternative approach is to use unstructured grids.
 Unstructured grids permit quasi-uniform grid spacing globally, thereby eliminating the pole-problem.
-They are more flexible than structured, latitude-longitude grids, allowing for regional grid refinement that may be used to capture higher resolution in the polar regions.
-Regional refinement of polar regions may be desirable as latitude-longitude grids, by virtue of the convergence of meridians, have greater horizontal resolution in polar regions compared to a quasi-uniform grid containing the same degrees of freedom.
+They allow for more flexible grid-structures than latitude-longitude grids, and can support regional grid refinement that may be used to capture higher resolution in the polar regions.
+Regional refinement of polar regions may be desirable as latitude-longitude grids, by virtue of the convergence of meridians, have higher horizontal resolution in polar regions compared to quasi-uniform grids containing the same degrees of freedom.
 
-\section{Methods}
+The Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}) supports a diverse number of atmospheric dynamical cores (hereafter referred to as \textit{dycores}).
+These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
+The EUL dycore is the oldest dycore in the Community Atmosphere Model (CAM), the atmospheric component of CESM, and is the least supported of all the dycores.
+FV3 is the newest dycore in CESM, but it was not fully incorporated into the model at the time this work commenced, and so is omitted from this study.
+The authors instead focus on the two most well supported dycores in CESM, SE and FV.
+The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
+
+The Arctic region has unique geography consisting of a polar ocean capped in a layer of sea ice, partially encircled by the northern coasts of the northern hemisphere continents and the large island of Greenland.
+Greenland is covered in a kilometers thick ice sheet, and receives a substantial portion of its snowfall from synoptic-scale storms traversing up the eastern sea-board of North America, and steered in the direction of the Icelandic low.
+Katabatic winds accelerate down the steep slopes of the Greenland Ice Sheet (GrIS) in episodic bursts that can clear out sea ice from its fjords and surrounding coastal region.
+Large volumes of fresh water are drained into the ocean through river outlets in Canada and Russia, ice discharge and melt water runoff from the narrow ablation zones of GrIS.
+Polar Lows form during the cold season and their gale-speed winds can kick up large waves and pose a hazard to mariners' and coastal communities.
+These Arctic processes have proven challenging to represent in GCMs, primarily due to the fine-scales involved \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
+As such, the Arctic serves as a challenging test-bed for comparing the simulated meteorology and climate in the SE and FV dycores.
+
+The FV dycore is assessed using grids with $\Delta x$ of $1^{\circ}$ and $2^{\circ}$ grid spacing ($\Delta x$ refers to the average equatorial grid spacing).
+These are compared to the equivalent  $1^{\circ}$ SE grid, referred to as ne30pg3.
+This grid does not refer to the SE dycore proper, but a variant in which the dry dynamics are solved using the SE method, and tracer advection computed using the conservative semi-Lagrangian advection method \cite <SE-CSLAM;>{LTOUNGK2017MWR}.
+In SE-CSLAM the physical parameterizations are computed on the finite-volume tracer advection grid.
+Optionally, one can compute the physics on a finite-volume grid that is $\frac{3}{2} \times \Delta x$ of the tracer-advection grid \cite{HETAL2019JAMES}.
+All variants of the SE dycore discussed herein are available as part of the CESM2.2 release.
+A comparison of the lower-resolution physics grid version of ne30pg3, referred to as ne30pg2, over the Arctic is included in this study.
+
+The SE dycore (not SE-CSLAM) supports regional grid refinement via it's variable resolution configuration.
+Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
+The ARCTIC grid is a $\Delta x=1^{\circ}$ grid with $\Delta x=\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
+The ARCTICGRIS grid is similar to the ARCTIC grid, but additionally contains a patch covering the big island of Greenland with $\Delta x=\frac{1}{8}^{\circ}$ resolution.
+The Arctic refined grids are depicted in Figure~\ref{fig:vr-grids}, and are available as part of the CESM2.2 release.
+The surface mass balance of the GrIS is compared across all grids and dycores in this study, and the ARCTCGRIS grid is certainly at an advantage due to its higher resolution.
+
+The study is laid out as follows.
+Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study.
+The Arctic refined grids were developed by the author, and this section serves as their official documentation.
+Section~\ref{sec:methods} also contains a description of the experiments along with the observational datasets and post-processing software used for validating the models.
+Section~\ref{sec:results} contains the results of the experiments and Section~\ref{sec:conclusions} provides some discussion and conclusions.
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cccc}
+         \includegraphics[width=60mm]{figs/grid-ARCTIC.pdf}&
+         \includegraphics[width=60mm]{figs/grid-ARCTICGRIS.pdf} \\
+\end{tabular}
+\end{center}
+\caption{.}
+\label{fig:vr-grids}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cc}
+         \includegraphics[width=60mm]{figs/grid-f19.pdf}&
+         \includegraphics[width=60mm]{figs/grid-f09.pdf}\\
+         \includegraphics[width=60mm]{figs/grid-ne30pg2.pdf}&
+         \includegraphics[width=60mm]{figs/grid-ne30pg3.pdf} \\
+\end{tabular}
+\end{center}
+\caption{.}
+\label{fig:uni-grids}
+\end{figure}
+
+\section{Methods}\label{sec:methods}
 \subsection{Grids}
@@
 \subsection{Physical parameterizations}
+\subsection{Experimental desing}
 \subsection{Observational datasets}
@@
 
-\section{Results}
+\section{Results}\label{sec:results}
 
 \subsection{Tropospheric temperatures}
-\subsection{Inter-annual Variability}
+\subsection{Inter-annual variability}
 \subsection{Synoptic-scale storm characteristics}
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusions}
 

### 2020-11-03 arcticgrids.tex
new figs
edits

@@
 General Circulation Models (GCMs) are a powerful tool for understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, the numerics of GCMs have had to grapple with the \textit{pole-problem} \cite{W2007JMSJ}, an economic bottleneck arising from the convergence of meridians into a polar singularity on latitude-longitude grids.
-In the 1970's this issue was largely defeated through wide-spread adaption of the spectral transform method in GCMs, which transforms grid point fields into an isotropic representation in wave space.
+Despite their importance, not all GCMs treat the poles the same way numerically as they do other regions of the globe, owing to the \textit{pole-problem} \cite{W2007JMSJ}.
+The pole-problem refers to instability arising from the convergence of meridians into a polar singularity on latitude-longitude grids.
+In the 1970's this issue was largely defeated through wide-spread adaption of the spectral transform methods in GCMs, which transforms grid point fields into an isotropic representation in wave space.
 But as computing power has increased, local numerical methods have become more desirable for their ability to run efficiently on massively parallel systems.
@@
 
-The Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}) supports a diverse number of atmospheric dynamical cores (hereafter referred to as \textit{dycores}).
+The atmospheric component of the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), the Community Atmosphere Model, version 6.3 (CAM; \url{https://github.com/ESCOMP/CAM/wiki}), supports a diverse number of atmospheric dynamical cores (hereafter referred to as \textit{dycores}).
 These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-The EUL dycore is the oldest dycore in the Community Atmosphere Model (CAM), the atmospheric component of CESM, and is the least supported of all the dycores.
+The EUL dycore is the oldest dycore in CAM, and is the least supported of all the dycores.
 FV3 is the newest dycore in CESM, but it was not fully incorporated into the model at the time this work commenced, and so is omitted from this study.
@@
 
-The Arctic region has unique geography consisting of a polar ocean capped in a layer of sea ice, partially encircled by the northern coasts of the northern hemisphere continents and the large island of Greenland.
+The Arctic region has a unique geography.
+It consists of a polar ocean encircled by the northern coasts of the northern hemisphere continents and the large island of Greenland, and capped by sea ice for half the year.
 Greenland is covered in a kilometers thick ice sheet, and receives a substantial portion of its snowfall from synoptic-scale storms traversing up the eastern sea-board of North America, and steered in the direction of the Icelandic low.
 Katabatic winds accelerate down the steep slopes of the Greenland Ice Sheet (GrIS) in episodic bursts that can clear out sea ice from its fjords and surrounding coastal region.
-Large volumes of fresh water are drained into the ocean through river outlets in Canada and Russia, ice discharge and melt water runoff from the narrow ablation zones of GrIS.
+Large volumes of fresh water are drained into the ocean through river outlets in Canada and Russia, ice discharge and melt water runoff from the periphery of GrIS.
 Polar Lows form during the cold season and their gale-speed winds can kick up large waves and pose a hazard to mariners' and coastal communities.
 These Arctic processes have proven challenging to represent in GCMs, primarily due to the fine-scales involved \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
-As such, the Arctic serves as a challenging test-bed for comparing the simulated meteorology and climate in the SE and FV dycores.
+As such, the Arctic serves as a test-bed for comparing the simulated meteorology and climate in the SE and FV dycores.
 
@@
 Optionally, one can compute the physics on a finite-volume grid that is $\frac{3}{2} \times \Delta x$ of the tracer-advection grid \cite{HETAL2019JAMES}.
+A comparison of the lower-resolution physics grid version of ne30pg3, referred to as ne30pg2, over the Arctic is included in this study.
 All variants of the SE dycore discussed herein are available as part of the CESM2.2 release.
-A comparison of the lower-resolution physics grid version of ne30pg3, referred to as ne30pg2, over the Arctic is included in this study.
 
-The SE dycore (not SE-CSLAM) supports regional grid refinement via it's variable resolution configuration.
+The SE dycore (not SE-CSLAM) supports regional grid refinement via its variable resolution configuration.
 Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
@@
 The ARCTICGRIS grid is similar to the ARCTIC grid, but additionally contains a patch covering the big island of Greenland with $\Delta x=\frac{1}{8}^{\circ}$ resolution.
-The Arctic refined grids are depicted in Figure~\ref{fig:vr-grids}, and are available as part of the CESM2.2 release.
-The surface mass balance of the GrIS is compared across all grids and dycores in this study, and the ARCTCGRIS grid is certainly at an advantage due to its higher resolution.
+These Arctic refined grids are depicted in Figure~\ref{fig:vr-grids}.
+The surface mass balance of the GrIS is compared across all grids and dycores in this study, and the ARCTCGRIS grid is certainly at an advantage due to its higher resolution over the ice sheet.
 
@@
 \section{Conclusions}\label{sec:conclusions}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_violens.pdf}
+\end{center}
+\caption{.}
+\label{fig:pointdiff}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
+\end{center}
+\caption{.}
+\label{fig:tseries}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_PRECT.pdf}
+\end{center}
+\caption{.}
+\label{fig:prect}
+\end{figure}
 

### 2020-11-04 arcticgrids.tex
ok

@@
 General Circulation Models (GCMs) are a powerful tool for understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, not all GCMs treat the poles the same way numerically as they do other regions of the globe, owing to the \textit{pole-problem} \cite{W2007JMSJ}.
+Despite their importance, numerically, not all GCMs treat the polar regions the same way as they do elsewhere in the domain due to the \textit{pole-problem} \cite{W2007JMSJ}.
+The wide variety of grids structures and numerical techniques used to represent polar regions in conventional GCMs are evaluated in this study for their impacts on the meteorology and climate of the Arctic.
+
 The pole-problem refers to instability arising from the convergence of meridians into a polar singularity on latitude-longitude grids.
@@
 The pole-problem has thus re-emerged in contemporary climate models using latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to subdue this instability \cite{JW2010LNCSE}.
+
 An alternative approach is to use unstructured grids.
@@
 Regional refinement of polar regions may be desirable as latitude-longitude grids, by virtue of the convergence of meridians, have higher horizontal resolution in polar regions compared to quasi-uniform grids containing the same degrees of freedom.
+
+The meteorology and climate of the Arctic is charcterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
+The Arctic geography consists of a polar ocean surrounded by the northern coasts of the northern hemisphere continents and the large island of Greenland.
+Greenland is covered in a kilometers thick ice sheet, and receives a substantial portion of its snowfall from synoptic-scale storms traversing up the eastern sea-board of North America that rain out over its steep margins.
+Intense cold season Polar Lows traverse the Arctic's melange of sea-ice and open ocean, inducing gale-speed winds.
+Large volumes of fresh water drain into the ocean through river outlets in Canada and Russia, iceberg discharge and melt water runoff emanating from the margins of GrIS.
+As such, the Arctic serves as a test-bed for comparing the simulated meteorology and climate to different numerical treatments of polar regions.
 
@@
 The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
-
-The Arctic region has a unique geography.
-It consists of a polar ocean encircled by the northern coasts of the northern hemisphere continents and the large island of Greenland, and capped by sea ice for half the year.
-Greenland is covered in a kilometers thick ice sheet, and receives a substantial portion of its snowfall from synoptic-scale storms traversing up the eastern sea-board of North America, and steered in the direction of the Icelandic low.
-Katabatic winds accelerate down the steep slopes of the Greenland Ice Sheet (GrIS) in episodic bursts that can clear out sea ice from its fjords and surrounding coastal region.
-Large volumes of fresh water are drained into the ocean through river outlets in Canada and Russia, ice discharge and melt water runoff from the periphery of GrIS.
-Polar Lows form during the cold season and their gale-speed winds can kick up large waves and pose a hazard to mariners' and coastal communities.
-These Arctic processes have proven challenging to represent in GCMs, primarily due to the fine-scales involved \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
-As such, the Arctic serves as a test-bed for comparing the simulated meteorology and climate in the SE and FV dycores.
 

### 2021-03-10 arcticgrids.tex
Add files via upload

@@
 
-General Circulation Models (GCMs) are a powerful tool for understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, numerically, not all GCMs treat the polar regions the same way as they do elsewhere in the domain due to the \textit{pole-problem} \cite{W2007JMSJ}.
-The wide variety of grids structures and numerical techniques used to represent polar regions in conventional GCMs are evaluated in this study for their impacts on the meteorology and climate of the Arctic.
+General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
+Despite their importance, the numerical treatment of polar regions is handled in vastly-different ways due to the \textit{pole-problem} \cite{W2007JMSJ}.
+The pole-problem refers to instability arising from the convergence of meridians at the polar points on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
+Depending on the numerical design, methods exist to stabilize the pole problem, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain.
+With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past.
+In this study, a spectrum of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), from conventional latitude-longitude grids to unstructured grids with globally uniform grid spacing to unstructured grids with regional refinement over the Arctic, are evaluated to understand the impacts of grids and dycores on the simulated characteristics of the Arctic.
 
-The pole-problem refers to instability arising from the convergence of meridians into a polar singularity on latitude-longitude grids.
-In the 1970's this issue was largely defeated through wide-spread adaption of the spectral transform methods in GCMs, which transforms grid point fields into an isotropic representation in wave space.
+In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs.
+These methods transforms grid point fields into a global, isotropic representation in wave space, where linear operators in the equation set can be solved for exactly.
 But as computing power has increased, local numerical methods have become more desirable for their ability to run efficiently on massively parallel systems.
-The pole-problem has thus re-emerged in contemporary climate models using latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to subdue this instability \cite{JW2010LNCSE}.
+The pole-problem has thus re-emerged in contemporary climate models using latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
+Polar filters are akin to a band-aid; they subdue the growth of unstable modes by introducing additional damping to the solution over polar regions.
+One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this is unlikely to occur in practice.
 
-An alternative approach is to use unstructured grids.
-Unstructured grids permit quasi-uniform grid spacing globally, thereby eliminating the pole-problem.
-They allow for more flexible grid-structures than latitude-longitude grids, and can support regional grid refinement that may be used to capture higher resolution in the polar regions.
-Regional refinement of polar regions may be desirable as latitude-longitude grids, by virtue of the convergence of meridians, have higher horizontal resolution in polar regions compared to quasi-uniform grids containing the same degrees of freedom.
+An alternative approach is unstructured grids.
+Unstructured grids allow for more flexible grid structures than latitude-longitude grids, permitting quasi-uniform grid spacing globally that eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c).
+This grid flexibility also allows for variable-resolution or regional grid refinement.
+Grids can be developed with refinement over polar regions that may in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region.
+But unstructured grids scale more efficiently on parallel systems than latitude-longitude grids, resulting in latitude-longitude grids becoming less common, and unstructured grids more common as computing power continued to increase over time.
 
-The meteorology and climate of the Arctic is charcterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
-The Arctic geography consists of a polar ocean surrounded by the northern coasts of the northern hemisphere continents and the large island of Greenland.
-Greenland is covered in a kilometers thick ice sheet, and receives a substantial portion of its snowfall from synoptic-scale storms traversing up the eastern sea-board of North America that rain out over its steep margins.
-Intense cold season Polar Lows traverse the Arctic's melange of sea-ice and open ocean, inducing gale-speed winds.
-Large volumes of fresh water drain into the ocean through river outlets in Canada and Russia, iceberg discharge and melt water runoff emanating from the margins of GrIS.
-As such, the Arctic serves as a test-bed for comparing the simulated meteorology and climate to different numerical treatments of polar regions.
+The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
+Synoptic scale storms are well represented at typical GCM resolutions, but mesoscale Polar Lows are not.
+These mesoscale systems are prevalent during the cold season, and induce gale-force winds that impact fluxes through the underlying sea-ice/ocean interface.
+The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}).
+While it sits on the largest island in the world (Greenland), GrIS is not well resolved at a typical 100 km grid resolution.
+GrIS ablation zones exert a primary control on the mass balance of the ice sheet, but are too narrow to resolve at conventional resolutions.
+The Arctic is therefore well-suited to understand the impact of different grids and dycores on processes that are marginally resolved at conventional GCM resolutions.
 
-The atmospheric component of the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), the Community Atmosphere Model, version 6.3 (CAM; \url{https://github.com/ESCOMP/CAM/wiki}), supports a diverse number of atmospheric dynamical cores (hereafter referred to as \textit{dycores}).
+The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://github.com/ESCOMP/CAM/wiki}), supports a diverse number of atmospheric dynamical cores.
 These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
@@
 \begin{center}
+\begin{tabular}{cc}
+         \includegraphics[width=60mm]{figs/grid-f19.pdf}&
+         \includegraphics[width=60mm]{figs/grid-f09.pdf}\\
+         \includegraphics[width=60mm]{figs/grid-ne30pg2.pdf}&
+         \includegraphics[width=60mm]{figs/grid-ne30pg3.pdf} \\
+\end{tabular}
+\end{center}
+\caption{.}
+\label{fig:uni-grids}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
 \begin{tabular}{cccc}
@@
 \begin{center}
-\begin{tabular}{cc}
-         \includegraphics[width=60mm]{figs/grid-f19.pdf}&
-         \includegraphics[width=60mm]{figs/grid-f09.pdf}\\
-         \includegraphics[width=60mm]{figs/grid-ne30pg2.pdf}&
-         \includegraphics[width=60mm]{figs/grid-ne30pg3.pdf} \\
-\end{tabular}
+         \includegraphics[width=60mm]{figs/temp_grisres.pdf}
 \end{center}
 \caption{.}
-\label{fig:uni-grids}
+\label{fig:prect}
 \end{figure}

### 2021-03-10 arcticgrids.tex
Updates from Overleaf

--- 2020-11-04 arcticgrids.tex
@@
 
-\title{Impact of grids and dycores in CESM2.2 on the meteorology and climate of the Arctic}
+\title{Impact of grids and dynamical cores in CESM2.2 on the meteorology and climate of the Arctic}
 
@@
 
-General Circulation Models (GCMs) are a powerful tool for understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, numerically, not all GCMs treat the polar regions the same way as they do elsewhere in the domain due to the \textit{pole-problem} \cite{W2007JMSJ}.
-The wide variety of grids structures and numerical techniques used to represent polar regions in conventional GCMs are evaluated in this study for their impacts on the meteorology and climate of the Arctic.
+General Circulation Models (GCMs) are a powerful tool for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change.
+Despite their importance,  not all GCMs treat the polar regions the same way as they do elsewhere in the domain due to the so called \textit{pole-problem} \cite{W2007JMSJ}.
+The wide variety of grids structures and numerical techniques used to represent polar regions in conventional GCMs are evaluated in this study for their impacts on the meteorology and climate of the Arctic, with a special focus on the Greenland Ice Sheet.
 
@@
 
-An alternative approach is to use unstructured grids.
-Unstructured grids permit quasi-uniform grid spacing globally, thereby eliminating the pole-problem.
-They allow for more flexible grid-structures than latitude-longitude grids, and can support regional grid refinement that may be used to capture higher resolution in the polar regions.
+An alternative approach is to use unstructured grids, which eliminate the pole-problem by adopting a quasi-uniform global grid spacing that includes the poles.
+These grids allow for more flexible grid-structures than latitude-longitude grids, and can support regional grid refinement that may be used to capture higher resolution in the polar regions.
 Regional refinement of polar regions may be desirable as latitude-longitude grids, by virtue of the convergence of meridians, have higher horizontal resolution in polar regions compared to quasi-uniform grids containing the same degrees of freedom.
@@
 The meteorology and climate of the Arctic is charcterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
-The Arctic geography consists of a polar ocean surrounded by the northern coasts of the northern hemisphere continents and the large island of Greenland.
+The Arctic geography consists of a polar ocean surrounded by the northern coasts of the Northern Hemisphere continents and the large island of Greenland.
 Greenland is covered in a kilometers thick ice sheet, and receives a substantial portion of its snowfall from synoptic-scale storms traversing up the eastern sea-board of North America that rain out over its steep margins.
 Intense cold season Polar Lows traverse the Arctic's melange of sea-ice and open ocean, inducing gale-speed winds.
-Large volumes of fresh water drain into the ocean through river outlets in Canada and Russia, iceberg discharge and melt water runoff emanating from the margins of GrIS.
+Large volumes of fresh water drain into the ocean through river outlets in Canada and Russia, and iceberg discharge and melt water runoff emanating from the margins of GrIS.
 As such, the Arctic serves as a test-bed for comparing the simulated meteorology and climate to different numerical treatments of polar regions.
+
+The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
 
@@
 These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-The EUL dycore is the oldest dycore in CAM, and is the least supported of all the dycores.
-FV3 is the newest dycore in CESM, but it was not fully incorporated into the model at the time this work commenced, and so is omitted from this study.
+The EUL dycore is the oldest dycore in CAM, and is the least supported of all the dycores in the current model.
+FV3 is the newest dycore in CESM, but it was not fully incorporated into the CAM at the time this work commenced, and so is omitted from this study.
 The authors instead focus on the two most well supported dycores in CESM, SE and FV.
-The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
+%The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
 
@@
 \section{Methods}\label{sec:methods}
+\subsection{Community Earth System Model}
 \subsection{Grids}

### 2021-03-11 arcticgrids.tex
Add files via upload

--- 2021-03-10 arcticgrids.tex
@@
 
-General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the poles, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, the numerical treatment of polar regions is handled in vastly-different ways due to the \textit{pole-problem} \cite{W2007JMSJ}.
+General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change.
+Despite their importance, the numerical treatment of polar regions is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}.
 The pole-problem refers to instability arising from the convergence of meridians at the polar points on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
@@
 With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past.
-In this study, a spectrum of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), from conventional latitude-longitude grids to unstructured grids with globally uniform grid spacing to unstructured grids with regional refinement over the Arctic, are evaluated to understand the impacts of grids and dycores on the simulated characteristics of the Arctic.
+In this study, a spectrum of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), from conventional latitude-longitude grids to unstructured grids with globally uniform grid spacing to unstructured grids with regional refinement over the Arctic, are evaluated to understand the impacts of grids and dycores on the simulated characteristics of the Arctic, with a special focus on the Greenland Ice Sheet.
 
 In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs.
-These methods transforms grid point fields into a global, isotropic representation in wave space, where linear operators in the equation set can be solved for exactly.
-But as computing power has increased, local numerical methods have become more desirable for their ability to run efficiently on massively parallel systems.
+These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators in the equation set can be solved for exactly.
+While spectral transform methods were still commonly used into the 21st century, more recently, local numerical methods have become more desirable for their ability to run efficiently on massively parallel systems.
 The pole-problem has thus re-emerged in contemporary climate models using latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
@@
 
-An alternative approach is unstructured grids.
+An alternative approach is to use unstructured grids.
 Unstructured grids allow for more flexible grid structures than latitude-longitude grids, permitting quasi-uniform grid spacing globally that eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c).
 This grid flexibility also allows for variable-resolution or regional grid refinement.
-Grids can be developed with refinement over polar regions that may in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region.
+Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region.
 But unstructured grids scale more efficiently on parallel systems than latitude-longitude grids, resulting in latitude-longitude grids becoming less common, and unstructured grids more common as computing power continued to increase over time.
@@
 Synoptic scale storms are well represented at typical GCM resolutions, but mesoscale Polar Lows are not.
-These mesoscale systems are prevalent during the cold season, and induce gale-force winds that impact fluxes through the underlying sea-ice/ocean interface.
+These mesoscale systems are prevalent during the cold season, and produce gale-force winds that can induce large fluxes through the underlying sea-ice/ocean interface.
 The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}).
-While it sits on the largest island in the world (Greenland), GrIS is not well resolved at a typical 100 km grid resolution.
-GrIS ablation zones exert a primary control on the mass balance of the ice sheet, but are too narrow to resolve at conventional resolutions.
+While it sits on the largest island in the world (Greenland), GrIS is only marginally resolved at typical GCM resolutions.
+GrIS ablation zones exert a primary control on the mass balance of the ice sheet, but are on the order of 100km wide, and models struggle to resolve these narrow features.
+GrIS topography descends rapidly towards the coast, and these steep margins facilitate orographic precipitation events that contribute mass to the ice sheet, but which are not well resolved in GCMs.
 The Arctic is therefore well-suited to understand the impact of different grids and dycores on processes that are marginally resolved at conventional GCM resolutions.
 
-The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://github.com/ESCOMP/CAM/wiki}), supports a diverse number of atmospheric dynamical cores.
-These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-The EUL dycore is the oldest dycore in CAM, and is the least supported of all the dycores.
-FV3 is the newest dycore in CESM, but it was not fully incorporated into the model at the time this work commenced, and so is omitted from this study.
-The authors instead focus on the two most well supported dycores in CESM, SE and FV.
 The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
-
-The FV dycore is assessed using grids with $\Delta x$ of $1^{\circ}$ and $2^{\circ}$ grid spacing ($\Delta x$ refers to the average equatorial grid spacing).
-These are compared to the equivalent  $1^{\circ}$ SE grid, referred to as ne30pg3.
-This grid does not refer to the SE dycore proper, but a variant in which the dry dynamics are solved using the SE method, and tracer advection computed using the conservative semi-Lagrangian advection method \cite <SE-CSLAM;>{LTOUNGK2017MWR}.
-In SE-CSLAM the physical parameterizations are computed on the finite-volume tracer advection grid.
-Optionally, one can compute the physics on a finite-volume grid that is $\frac{3}{2} \times \Delta x$ of the tracer-advection grid \cite{HETAL2019JAMES}.
-A comparison of the lower-resolution physics grid version of ne30pg3, referred to as ne30pg2, over the Arctic is included in this study.
-All variants of the SE dycore discussed herein are available as part of the CESM2.2 release.
-
-The SE dycore (not SE-CSLAM) supports regional grid refinement via its variable resolution configuration.
-Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
-The ARCTIC grid is a $\Delta x=1^{\circ}$ grid with $\Delta x=\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
-The ARCTICGRIS grid is similar to the ARCTIC grid, but additionally contains a patch covering the big island of Greenland with $\Delta x=\frac{1}{8}^{\circ}$ resolution.
-These Arctic refined grids are depicted in Figure~\ref{fig:vr-grids}.
-The surface mass balance of the GrIS is compared across all grids and dycores in this study, and the ARCTCGRIS grid is certainly at an advantage due to its higher resolution over the ice sheet.
-
 The study is laid out as follows.
@@
 \subsection{Dynamical cores}
+The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://github.com/ESCOMP/CAM/wiki}), supports a diverse number of atmospheric dynamical cores.
+These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
+The EUL dycore is the oldest dycore in CAM, and is the least supported of all the dycores in the current model.
+FV3 is the newest dycore in CESM, but it was not fully incorporated into CAM at the time this work commenced, and so is omitted from this study.
+The authors instead focus on the two most well supported dycores in CESM, SE and FV.
+The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
 \subsubsection{Finite-volume model}
+The FV dycore is assessed using grids with $\Delta x$ of $1^{\circ}$ and $2^{\circ}$ grid spacing ($\Delta x$ refers to the average equatorial grid spacing).
+These are compared to the equivalent  $1^{\circ}$ SE grid, referred to as ne30pg3.
+This grid does not refer to the SE dycore proper, but a variant in which the dry dynamics are solved using the SE method, and tracer advection computed using the conservative semi-Lagrangian advection method \cite <SE-CSLAM;>{LTOUNGK2017MWR}.
+In SE-CSLAM the physical parameterizations are computed on the finite-volume tracer advection grid.
+Optionally, one can compute the physics on a finite-volume grid that is $\frac{3}{2} \times \Delta x$ of the tracer-advection grid \cite{HETAL2019JAMES}.
+A comparison of the lower-resolution physics grid version of ne30pg3, referred to as ne30pg2, over the Arctic is included in this study.
+All variants of the SE dycore discussed herein are available as part of the CESM2.2 release.
 \subsubsection{Spectral-element model}
+The SE dycore (not SE-CSLAM) supports regional grid refinement via its variable resolution configuration.
+Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
+The ARCTIC grid is a $\Delta x=1^{\circ}$ grid with $\Delta x=\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
+The ARCTICGRIS grid is similar to the ARCTIC grid, but additionally contains a patch covering the big island of Greenland with $\Delta x=\frac{1}{8}^{\circ}$ resolution.
+These Arctic refined grids are depicted in Figure~\ref{fig:vr-grids}.
+The surface mass balance of the GrIS is compared across all grids and dycores in this study, and the ARCTCGRIS grid is certainly at an advantage due to its higher resolution over the ice sheet.
 \subsection{Physical parameterizations}

### 2021-03-11 arcticgrids.tex
new figures

@@
 \begin{center}
-         \includegraphics[width=60mm]{figs/temp_grisres.pdf}
+\begin{tabular}{cccc}
+         \includegraphics[width=60mm]{figs/temp_grisres_dxoverglc.pdf}&
+         \includegraphics[width=60mm]{figs/temp_grisres_error.pdf} \\
+\end{tabular}
 \end{center}
@@
 \begin{center}
+         \includegraphics[width=130mm]{figs/temp_PRECT.pdf}
+\end{center}
+\caption{.}
+\label{fig:prect}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_tseries_GRIS_map_TO_f19.pdf}
+\end{center}
+\caption{.}
+\label{fig:tseries}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
          \includegraphics[width=130mm]{figs/temp_violens.pdf}
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
+\begin{tabular}{cccc}
+         \includegraphics[width=100mm]{figs/temp_ktransect.pdf} \\
+         \includegraphics[width=100mm]{figs/temp_btransect.pdf} \\
+\end{tabular}
 \end{center}
 \caption{.}
-\label{fig:tseries}
+\label{fig:prect}
 \end{figure}
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_PRECT.pdf}
+\begin{tabular}{cccc}
+         \includegraphics[width=60mm]{figs/temp_zsmb_btrans_obsperiod.pdf}&
+         \includegraphics[width=60mm]{figs/temp_zsmb_ktrans_obsperiod.pdf} \\
+\end{tabular}
 \end{center}

### 2021-03-11 arcticgrids.tex
some text mods

@@
 The pole-problem has thus re-emerged in contemporary climate models using latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
-Polar filters are akin to a band-aid; they subdue the growth of unstable modes by introducing additional damping to the solution over polar regions.
-One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this is unlikely to occur in practice.
+Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
+One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems an unlikely outcome in practice.
 
 An alternative approach is to use unstructured grids.
-Unstructured grids allow for more flexible grid structures than latitude-longitude grids, permitting quasi-uniform grid spacing globally that eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c).
+Unstructured grids allow for more flexible grid structures, permitting quasi-uniform grid spacing globally that eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c).
 This grid flexibility also allows for variable-resolution or regional grid refinement.
@@
 The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}).
-While it sits on the largest island in the world (Greenland), GrIS is only marginally resolved at typical GCM resolutions.
+While it blankets the largest island in the world (Greenland), GrIS is only marginally resolved at typical GCM resolutions.
 GrIS ablation zones exert a primary control on the mass balance of the ice sheet, but are on the order of 100km wide, and models struggle to resolve these narrow features.
-GrIS topography descends rapidly towards the coast, and these steep margins facilitate orographic precipitation events that contribute mass to the ice sheet, but which are not well resolved in GCMs.
+GrIS elevations descends rapidly toward the coasts, resulting in steep margins that facilitate orographic precipitation events, but which are not well resolved in GCMs.
 The Arctic is therefore well-suited to understand the impact of different grids and dycores on processes that are marginally resolved at conventional GCM resolutions.
 
-The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
-The study is laid out as follows.
+The goal of this study is to characterize the representation of high-latitiude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
+The manuscript is laid out as follows.
 Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study.

### 2021-03-11 arcticgrids.tex
new words

@@
 With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past.
-In this study, a spectrum of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), from conventional latitude-longitude grids to unstructured grids with globally uniform grid spacing to unstructured grids with regional refinement over the Arctic, are evaluated to understand the impacts of grids and dycores on the simulated characteristics of the Arctic, with a special focus on the Greenland Ice Sheet.
+In this study, a spectrum of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), from conventional latitude-longitude grids to unstructured grids with globally uniform grid spacing to unstructured grids with regional refinement over the Arctic, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the Greenland Ice Sheet.
 
@@
 These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators in the equation set can be solved for exactly.
-While spectral transform methods were still commonly used into the 21st century, more recently, local numerical methods have become more desirable for their ability to run efficiently on massively parallel systems.
+While spectral transform methods were still commonly used into the 21st century, more recently, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
 The pole-problem has thus re-emerged in contemporary climate models using latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
@@
 \begin{tabular}{cccc}
-         \includegraphics[width=60mm]{figs/temp_zsmb_btrans_obsperiod.pdf}&
-         \includegraphics[width=60mm]{figs/temp_zsmb_ktrans_obsperiod.pdf} \\
+         \includegraphics[width=60mm]{figs/temp_zsmb_ktrans_obsperiod.pdf}&
+         \includegraphics[width=60mm]{figs/temp_zsmb_btrans_obsperiod.pdf} \\
 \end{tabular}

### 2021-03-12 arcticgrids.tex
more text

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
@@
 The pole-problem refers to instability arising from the convergence of meridians at the polar points on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
-Depending on the numerical design, methods exist to stabilize the pole problem, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain.
+Depending on the numerics, methods exist to stabilize the pole problem, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain.
 With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past.
@@
 These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators in the equation set can be solved for exactly.
-While spectral transform methods were still commonly used into the 21st century, more recently, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
-The pole-problem has thus re-emerged in contemporary climate models using latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
+While spectral transform methods are still commonly used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
+The pole-problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
 Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
-One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems an unlikely outcome in practice.
+One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems an unlikely outcome and is explored further in this study.
 
@@
 
-The goal of this study is to characterize the representation of high-latitiude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
+The goal of this study is to characterize the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
 The manuscript is laid out as follows.
 Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study.
-The Arctic refined grids were developed by the author, and this section serves as their official documentation.
-Section~\ref{sec:methods} also contains a description of the experiments along with the observational datasets and post-processing software used for validating the models.
+The Arctic refined grids were developed by the authors, and this section serves as their official documentation in CESM2.2.
+Section~\ref{sec:methods} also contains a description of the experiments along with the observational datasets and post-processing software used for evaluating the models.
 Section~\ref{sec:results} contains the results of the experiments and Section~\ref{sec:conclusions} provides some discussion and conclusions.
@@
 \section{Methods}\label{sec:methods}
-\subsection{Grids}
-\subsection{Dynamical cores}
-The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://github.com/ESCOMP/CAM/wiki}), supports a diverse number of atmospheric dynamical cores.
+\subsection{Grids and dynamical cores}
+
+The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://ncar.github.io/CAM/doc/build/html/index.html}), supports a diverse number of atmospheric dynamical cores.
 These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-The EUL dycore is the oldest dycore in CAM, and is the least supported of all the dycores in the current model.
-FV3 is the newest dycore in CESM, but it was not fully incorporated into CAM at the time this work commenced, and so is omitted from this study.
-The authors instead focus on the two most well supported dycores in CESM, SE and FV.
-The goal of this study is to characterize the representation of polar regions using the SE and FV dycores, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
+The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores in the current model.
+FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced, and so is omitted from this study.
+The authors instead focus on the two most well supported dycores in CAM, SE and FV.
+
 \subsubsection{Finite-volume model}
-The FV dycore is assessed using grids with $\Delta x$ of $1^{\circ}$ and $2^{\circ}$ grid spacing ($\Delta x$ refers to the average equatorial grid spacing).
-These are compared to the equivalent  $1^{\circ}$ SE grid, referred to as ne30pg3.
-This grid does not refer to the SE dycore proper, but a variant in which the dry dynamics are solved using the SE method, and tracer advection computed using the conservative semi-Lagrangian advection method \cite <SE-CSLAM;>{LTOUNGK2017MWR}.
-In SE-CSLAM the physical parameterizations are computed on the finite-volume tracer advection grid.
-Optionally, one can compute the physics on a finite-volume grid that is $\frac{3}{2} \times \Delta x$ of the tracer-advection grid \cite{HETAL2019JAMES}.
-A comparison of the lower-resolution physics grid version of ne30pg3, referred to as ne30pg2, over the Arctic is included in this study.
+
+The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a global latitude-longitude grid.
+The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to Eulerian reference grid in the vertical.
+Hyperviscous damping is applied to the divergent modes while Laplacian damping is applied to momentum in the top few layers, referred to as a ``sponge layer." A polar filter is used to avoid computational instability due to the convergence of the meridians, permitting the use of a more practical time-step.
+It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically poleward.
+The FV dycore is run with grids $\Delta x$ of $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
+
+\subsubsection{Spectral-element model}
+
+The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method.
+The computational domain is defined by a cubed-sphere grid that is tiled with quasi-equal area quadralateral elements (e.g., Figure~\ref{fig:vr-grids}).
+Each element contains a fourth order basis set in the two horizontal directions, and the solution defined at the roots of the basis functions, the Gauss-Lobatto-Lagendre (GLL) quadrature points.
+This results in 16 GLL point values within each element, with 12 of the points lying on the element boundary.
+This compact support allows for efficient parallelization up to one element for task.
+Communication between elements happens via the direct stiffness summation, which applies a numerical flux to the element boundaries to reconcile overlapping nodal point values, and reulting in a continuous global basis set.
+
+As with the FV dycore, the dyanmics evolve in floating Lagrangian layers that are subsequently mapped to a Eulerian reference grid.
+The 2D dynamics have no implicit dissipation and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors.
+Laplacian damping is applied in the sponge layer, similar to the FV dycore.
+
+The SE dycore supports regional grid refinement via its variable-resolution configuration.
+Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
+The software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}) was used to generate these meshes.
+The $ARCTIC$ grid is a $\Delta x=1^{\circ}$ grid with $\Delta x=\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
+The $ARCTICGRIS$ grid is similar to the $ARCTIC$ grid, but additionally contains a patch covering the big island of Greenland with $\Delta x=\frac{1}{8}^{\circ}$ resolution.
+Both grids are depicted in Figure~\ref{fig:vr-grids}.
+
+For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Mulit-tracer tranport scheme is used instead \cite[CSLAM]{LTOUNGK2017MWR}.
+CSLAM has improved tracer property preservation and accelerated multi-tracer transport.
+It uses a seperate grid from the spectral-element dynamics, through dividing each element into $3\times3$ quasi-equal area control volumes.
+The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated at the GLL nodal points \cite{HL2018MWR}.
+
+The $1^{\circ}$ CAM-SE-CSLAM grid is referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c).
+The $ne$ indicates the number of elements along an edge of a cubed-sphere face, and the $pg$ denotes the number of control volumes along the edge of an element used for computing the physical parameterizations.
+Optionally, one can compute the physics on a finite-volume grid that is $\frac{3}{2} \times \Delta x$ of the tracer-advection grid, found through dividing the elmenet into $2\times2$ control volumes \cite[(Figure~\ref{fig:uni-grids}d)]{HETAL2019JAMES}.
+A comparison of this lower-resolution physics grid $ne30pg2$ over the Arctic is included in this study.
 All variants of the SE dycore discussed herein are available as part of the CESM2.2 release.
-\subsubsection{Spectral-element model}
-The SE dycore (not SE-CSLAM) supports regional grid refinement via its variable resolution configuration.
-Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
-The ARCTIC grid is a $\Delta x=1^{\circ}$ grid with $\Delta x=\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
-The ARCTICGRIS grid is similar to the ARCTIC grid, but additionally contains a patch covering the big island of Greenland with $\Delta x=\frac{1}{8}^{\circ}$ resolution.
-These Arctic refined grids are depicted in Figure~\ref{fig:vr-grids}.
-The surface mass balance of the GrIS is compared across all grids and dycores in this study, and the ARCTCGRIS grid is certainly at an advantage due to its higher resolution over the ice sheet.
+
 \subsection{Physical parameterizations}
-\subsection{Experimental desing}
+
+The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used for all simulations in this study.
+CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite[CLUBB]{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme.
+CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions.
+Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} inlcuding convective momentum transport \cite{RSG2010JAS}.
+PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method driven by the orientation of topographic ridges at the sub-grid scale.
+
+\subsection{Experimental design}
+
+All grids and dycores are run using an identical transient 1979-1998 AMIP-style configuration, with prescribed monthly SST/sea-ice after \cite{CESMSST}.
+This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2.
+The Community Land Model (CLM5; \url{https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/index.html}) calculates the surface energy balance at each land tile within a grid cell, which is used to compute snow and bare ice melting to inform the surface mass balance (SMB) of a glacier unit \cite{VKETAL2020JGR}.
+Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation.
+The SMB of the GrIS was spun-up for the variable-resolution grids by forcing CLM5 offline, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years.
+The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.
+
 \subsection{Observational datasets}

### 2021-03-12 arcticgrids.tex
new words

@@
 Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
-One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems an unlikely outcome and is explored further in this study.
+One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems an unlikely and is explored further in this study.
 
@@
 \section{Methods}\label{sec:methods}
-\subsection{Grids and dynamical cores}
+\subsection{Dynamical cores}
 
@@
 
-\subsubsection{Finite-volume model}
+\subsubsection{Finite-volume dynamical core}
 
-The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a global latitude-longitude grid.
-The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to Eulerian reference grid in the vertical.
-Hyperviscous damping is applied to the divergent modes while Laplacian damping is applied to momentum in the top few layers, referred to as a ``sponge layer." A polar filter is used to avoid computational instability due to the convergence of the meridians, permitting the use of a more practical time-step.
-It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically poleward.
-The FV dycore is run with grids $\Delta x$ of $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
+The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical latitude-longitude grid \cite{LR1997QJR}.
+The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate.
+Hyperviscous damping is applied to the divergent modes while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC}.
+A polar filter is used to avoid computational instability due to the convergence of the meridians, allowing for a more practical time-step.
+It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically poleward \cite{ST1995GEOS}.
 
-\subsubsection{Spectral-element model}
+\subsubsection{Spectral-element dynamical core}
 
-The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method.
-The computational domain is defined by a cubed-sphere grid that is tiled with quasi-equal area quadralateral elements (e.g., Figure~\ref{fig:vr-grids}).
-Each element contains a fourth order basis set in the two horizontal directions, and the solution defined at the roots of the basis functions, the Gauss-Lobatto-Lagendre (GLL) quadrature points.
-This results in 16 GLL point values within each element, with 12 of the points lying on the element boundary.
-This compact support allows for efficient parallelization up to one element for task.
-Communication between elements happens via the direct stiffness summation, which applies a numerical flux to the element boundaries to reconcile overlapping nodal point values, and reulting in a continuous global basis set.
+The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,DetAl2012IJHPCA,LetAl2018JAMES}.
+The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}).
+Each element contains a fourth order basis set in the two horizontal directions, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Lagendre (GLL) quadrature points.
+This results in 16 GLL point values within each element, with 12 of the points lying on the (shared) element boundary.
+Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal point values to produces a continuous global basis set.
 
-As with the FV dycore, the dyanmics evolve in floating Lagrangian layers that are subsequently mapped to a Eulerian reference grid.
-The 2D dynamics have no implicit dissipation and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors.
-Laplacian damping is applied in the sponge layer, similar to the FV dycore.
+As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to a Eulerian reference grid.
+A dry mass vertical coordinate was more recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}.
+The 2D dynamics have no implicit dissipation and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors \cite{DetAl2012IJHPCA}.
+Laplacian damping is applied in the sponge layer.
 
-The SE dycore supports regional grid refinement via its variable-resolution configuration.
-Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
-The software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}) was used to generate these meshes.
-The $ARCTIC$ grid is a $\Delta x=1^{\circ}$ grid with $\Delta x=\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
-The $ARCTICGRIS$ grid is similar to the $ARCTIC$ grid, but additionally contains a patch covering the big island of Greenland with $\Delta x=\frac{1}{8}^{\circ}$ resolution.
-Both grids are depicted in Figure~\ref{fig:vr-grids}.
+The SE dycore supports regional grid refinement via its variable-resolution configuration, requiring two enhancements over uniform resolution grids.
+(1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing their strength  by about an order of magnitude per doubling of resolution.
+A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements.
+(2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes.
+The NCAR topography software \cite{gmdd-8-4623-2015} was modified to scale the smoothing radius by the local element size.
 
-For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Mulit-tracer tranport scheme is used instead \cite[CSLAM]{LTOUNGK2017MWR}.
+For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Mulit-tracer tranport scheme is used instead \cite<CSLAM;>{LTOUNGK2017MWR}.
 CSLAM has improved tracer property preservation and accelerated multi-tracer transport.
@@
 
-The $1^{\circ}$ CAM-SE-CSLAM grid is referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c).
-The $ne$ indicates the number of elements along an edge of a cubed-sphere face, and the $pg$ denotes the number of control volumes along the edge of an element used for computing the physical parameterizations.
-Optionally, one can compute the physics on a finite-volume grid that is $\frac{3}{2} \times \Delta x$ of the tracer-advection grid, found through dividing the elmenet into $2\times2$ control volumes \cite[(Figure~\ref{fig:uni-grids}d)]{HETAL2019JAMES}.
-A comparison of this lower-resolution physics grid $ne30pg2$ over the Arctic is included in this study.
-All variants of the SE dycore discussed herein are available as part of the CESM2.2 release.
+\subsection{Grids}
+
+Six grid are evaluated in this study (Table X).
+The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
+The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
+An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, in which the physics are computed on a grid that is $\frac{3}{2} \times$ larger than $pg3$ grid spacing, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.
+
+Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
+The software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}) was used to generate these meshes.
+The $ARCTIC$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
+The $ARCTICGRIS$ grid is identical to the $ARCTIC$ grid, but contains an additional patch covering the big island of Greenland with $\frac{1}{8}^{\circ}$ resolution.
+Both of these grids are depicted in Figure~\ref{fig:vr-grids}.
 
@@
 The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used for all simulations in this study.
-CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite[CLUBB]{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme.
+CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme.
 CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions.
 Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} inlcuding convective momentum transport \cite{RSG2010JAS}.
-PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method driven by the orientation of topographic ridges at the sub-grid scale.
+PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.
 
@@
 This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2.
-The Community Land Model (CLM5; \url{https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/index.html}) calculates the surface energy balance at each land tile within a grid cell, which is used to compute snow and bare ice melting to inform the surface mass balance (SMB) of a glacier unit \cite{VKETAL2020JGR}.
+The Community Land Model (CLM5; \url{https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/index.html}), which uses the same grid as the atmosphere grid, calculates the surface energy balance at each land tile within a grid cell, which is used to compute snow and bare ice melting to inform the surface mass balance (SMB) of a glacier unit \cite{VKETAL2020JGR}.
 Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation.
-The SMB of the GrIS was spun-up for the variable-resolution grids by forcing CLM5 offline, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years.
+Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up  by forcing CLM5 offline, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years.
 The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.

### 2021-03-13 arcticgrids.tex
updates

@@
 
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{cccc}
-         \includegraphics[width=60mm]{figs/temp_grisres_dxoverglc.pdf}&
-         \includegraphics[width=60mm]{figs/temp_grisres_error.pdf} \\
-\end{tabular}
-\end{center}
-\caption{.}
-\label{fig:prect}
-\end{figure}
-
 \section{Methods}\label{sec:methods}
@@
 
-The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,DetAl2012IJHPCA,LetAl2018JAMES}.
+The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,DetAl2012IJHPCA}.
 The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}).
 Each element contains a fourth order basis set in the two horizontal directions, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Lagendre (GLL) quadrature points.
-This results in 16 GLL point values within each element, with 12 of the points lying on the (shared) element boundary.
-Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal point values to produces a continuous global basis set.
+This results in 16 GLL nodal points within each element, with 12 of the points lying on the (shared) element boundary.
+Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal point values and produces a continuous global basis set.
 
@@
 The SE dycore supports regional grid refinement via its variable-resolution configuration, requiring two enhancements over uniform resolution grids.
-(1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing their strength  by about an order of magnitude per doubling of resolution.
+(1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by about an order of magnitude per halving of the grid spacing.
 A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements.
-(2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes.
-The NCAR topography software \cite{gmdd-8-4623-2015} was modified to scale the smoothing radius by the local element size.
+(2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so  the NCAR topography software \cite{gmdd-8-4623-2015} was modified to scale the smoothing radius by the local element size.
 
@@
 The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
-An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, in which the physics are computed on a grid that is $\frac{3}{2} \times$ larger than $pg3$ grid spacing, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.
+An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical paramerizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.
 
-Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic.
-The software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}) was used to generate these meshes.
+Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic (Figure~\ref{fig:vr-grids}).
+The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
 The $ARCTIC$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
 The $ARCTICGRIS$ grid is identical to the $ARCTIC$ grid, but contains an additional patch covering the big island of Greenland with $\frac{1}{8}^{\circ}$ resolution.
-Both of these grids are depicted in Figure~\ref{fig:vr-grids}.
 
@@
 
-\subsection{Observational datasets}
-\subsubsection{ERA5}
-\subsubsection{LIVVkit 2.1}
-\subsection{TempestExtremes}
-\subsection{StormCompositer}
-
 \section{Results}\label{sec:results}
@@
 \subsection{Tropospheric temperatures}
+
+Before delving into the simulated characteristics of the Arctic, an understanding of the global mean differences between the grids and dycores are assessed, Figure~\ref{fig:dT-lores} shows the climatological, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
+The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.This is common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
+The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the mid-latitudes in the $f09$ grid.
+
+As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}).
+The stratosphere has a different signal, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
+The differences in temperature between $ne30pg3$ and $ne30pg2$ are small, with a slight warming near the tropopause at high latitudes.
+This is consistent with the similar climates using these grids shown in \cite{HETAL2019JAMES}.
+
+In comparing the variable-resolution grids to the uniform resolution grids is convoluted because not only is the resolution higher in refinement region, but the physics time-step is shorter in the variable-resolution runs.
+An additional $ne30pg3$ simulations was run with physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$.
+Figure~\ref{fig:dT-hires} shows the changes in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$.
+A similar warming response to increase resolution occurs when the time-step is reduced, and the mechanism is similar in that this faciliates greater condensational heating by CLUBB.
+Figure~\ref{fig:dT-hires} shows difference in climatolgical temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
+The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.
+
 \subsection{Inter-annual variability}
@@
 \section{Conclusions}\label{sec:conclusions}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-lores.pdf}
+\end{center}
+\caption{.}
+\label{fig:dT-lores}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB_phytstep_JJA.pdf}
+\end{center}
+\caption{.}
+\label{fig:dT-hires}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cccc}
+         \includegraphics[width=60mm]{figs/temp_grisres_dxoverglc.pdf}&
+         \includegraphics[width=60mm]{figs/temp_grisres_error.pdf} \\
+\end{tabular}
+\end{center}
+\caption{.}
+\label{fig:prect}
+\end{figure}
 

### 2021-03-14 arcticgrids.tex
update

@@
 
+The CAM convention is that the physics package determines the vertical resolution.
+Since all runs use CAM6 physics, all grids and dycores use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$.
+The physics time-step, in contrast, is dependent on grid resolution.
+Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step should be reduced to avoid large time truncation errors \cite{HR2018JAMES}.
+The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids.
+
 \subsection{Experimental design}
@@
 Before delving into the simulated characteristics of the Arctic, an understanding of the global mean differences between the grids and dycores are assessed, Figure~\ref{fig:dT-lores} shows the climatological, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
-The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.This is common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
+The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.
+This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
 The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the mid-latitudes in the $f09$ grid.
@@
 As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}).
-The stratosphere has a different signal, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
+The stratosphere has a different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
 The differences in temperature between $ne30pg3$ and $ne30pg2$ are small, with a slight warming near the tropopause at high latitudes.
-This is consistent with the similar climates using these grids shown in \cite{HETAL2019JAMES}.
+This is consistent with the similar climates found between these grids in \cite{HETAL2019JAMES}.
 
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB_phytstep_JJA.pdf}
+         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
 \end{center}

### 2021-05-05 arcticgrids.tex
new integrate smb fig

@@
 General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, the numerical treatment of polar regions is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}.
-The pole-problem refers to instability arising from the convergence of meridians at the polar points on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
-Depending on the numerics, methods exist to stabilize the pole problem, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain.
+Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}, which refers to numerical instability arising from the convergence of meridians at the polar points on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
+Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain.
 With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past.
-In this study, a spectrum of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), from conventional latitude-longitude grids to unstructured grids with globally uniform grid spacing to unstructured grids with regional refinement over the Arctic, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the Greenland Ice Sheet.
+In this study, a spectrum of model grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}) are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the meteorology over the Greenland Ice Sheet.
 
 In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs.
-These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators in the equation set can be solved for exactly.
-While spectral transform methods are still commonly used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
+These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g.
+horizontal derivatives) in the equation set can be solved for exactly.
+While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
 The pole-problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
 Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
-One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems an unlikely and is explored further in this study.
+One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems unlikely (describe why) and is explored further in this study.
 
-An alternative approach is to use unstructured grids.
-Unstructured grids allow for more flexible grid structures, permitting quasi-uniform grid spacing globally that eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c).
-This grid flexibility also allows for variable-resolution or regional grid refinement.
-Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region.
-But unstructured grids scale more efficiently on parallel systems than latitude-longitude grids, resulting in latitude-longitude grids becoming less common, and unstructured grids more common as computing power continued to increase over time.
+An alternative approach is to use unstructured grids, which allow for more flexible grid structures that permit quasi-uniform grid spacing globally and eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c).
+This grid flexibility also allows for variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}).
+Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region (the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space).
+However, unstructured grids scale more efficiently on parallel systems than latitude-longitude grids, likely resulting in a greater prevalence of unstructured grids as computing power continued to increase over time.
 
 The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
-Synoptic scale storms are well represented at typical GCM resolutions, but mesoscale Polar Lows are not.
-These mesoscale systems are prevalent during the cold season, and produce gale-force winds that can induce large fluxes through the underlying sea-ice/ocean interface.
+For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions.
+These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface.
 The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}).
-While it blankets the largest island in the world (Greenland), GrIS is only marginally resolved at typical GCM resolutions.
-GrIS ablation zones exert a primary control on the mass balance of the ice sheet, but are on the order of 100km wide, and models struggle to resolve these narrow features.
-GrIS elevations descends rapidly toward the coasts, resulting in steep margins that facilitate orographic precipitation events, but which are not well resolved in GCMs.
-The Arctic is therefore well-suited to understand the impact of different grids and dycores on processes that are marginally resolved at conventional GCM resolutions.
+While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and melting) are only partially resolved at typical GCM resolutions.
+For example, GrIS precipitation is typically confined to the ice-sheet margins (predominately the southeastern margin) where orographic precipitation is generated by steep topographic slopes.
+Moreover, GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation.
+GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.
 
-The goal of this study is to characterize the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in very different ways.
-The manuscript is laid out as follows.
-Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study.
+The goal of this study is to characterize the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in different ways.
+The manuscript is laid out as follows: Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study.
 The Arctic refined grids were developed by the authors, and this section serves as their official documentation in CESM2.2.
-Section~\ref{sec:methods} also contains a description of the experiments along with the observational datasets and post-processing software used for evaluating the models.
-Section~\ref{sec:results} contains the results of the experiments and Section~\ref{sec:conclusions} provides some discussion and conclusions.
+Section~\ref{sec:methods} also contains a description of the experiments along with reanalysis datasets and post-processing software used for evaluating the model simulations.
+Section~\ref{sec:results} contains the results of the experiments, followed by Section~\ref{sec:conclusions} that provides a general discussion and conclusions.
 
@@
 The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}).
-Each element contains a fourth order basis set in the two horizontal directions, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Lagendre (GLL) quadrature points.
+Each element contains a fourth order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points.
 This results in 16 GLL nodal points within each element, with 12 of the points lying on the (shared) element boundary.
-Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal point values and produces a continuous global basis set.
+Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal values and produces a continuous global basis set.
 
@@
 
-In comparing the variable-resolution grids to the uniform resolution grids is convoluted because not only is the resolution higher in refinement region, but the physics time-step is shorter in the variable-resolution runs.
-An additional $ne30pg3$ simulations was run with physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$.
-Figure~\ref{fig:dT-hires} shows the changes in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$.
-A similar warming response to increase resolution occurs when the time-step is reduced, and the mechanism is similar in that this faciliates greater condensational heating by CLUBB.
-Figure~\ref{fig:dT-hires} shows difference in climatolgical temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
+Comparing the variable-resolution grids to the uniform resolution grids is complicated because we are simultaneously increase the grid resolution and reducing the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}.
+An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$.
+Figure~\ref{fig:dT-hires} shows the change in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$.
+A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB.
+Figure~\ref{fig:dT-hires} shows the difference in climatological temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
 The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_tseries_GRIS_map_TO_f19.pdf}
+         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
 \end{center}

### 2021-05-05 arcticgrids.tex
Updates from Overleaf

@@
 The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
-For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions.
+For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees (REF: 2013 IPCC report, 9.1.3.3, Jablonowski and Williamson), mesoscale Polar Lows are not well resolved at these resolutions.
 These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface.
@@
 Moreover, GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation.
-GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.
+GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite[e.g.,][]{Lofverstrom_EA_2020_JAMES}.
 
@@
 
-The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://ncar.github.io/CAM/doc/build/html/index.html}), supports a diverse number of atmospheric dynamical cores.
-These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores in the current model.
-FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced, and so is omitted from this study.
-The authors instead focus on the two most well supported dycores in CAM, SE and FV.
+The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://ncar.github.io/CAM/doc/build/html/index.html}), supports a number of different atmospheric dynamical cores.
+These include dycores using latitude-longitude grids, such as finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
+The EUL dycore is the oldest dycore in CAM and the least supported of all the dycores in the current model.
+FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced; both the EUL and FV3 dycores are omitted from this study.
+As such, the results presented in this study are comparing the performance of the SE and FV dycores.
 
-\subsubsection{Finite-volume dynamical core}
+\subsubsection{Finite-volume (FV) dynamical core}
 
@@
 Hyperviscous damping is applied to the divergent modes while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC}.
-A polar filter is used to avoid computational instability due to the convergence of the meridians, allowing for a more practical time-step.
-It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically poleward \cite{ST1995GEOS}.
+A zonal Fourier filter --- with damping coefficients increasing monotonically in the poleward direction \cite{ST1995GEOS} --- is applied to suppress numerical instabilities due to grid convergence at the North and South Pole.
+This so called \textit{polar filer} allows for a longer time-step, but it has the disadvantage that atmospheric fields are artificially smoothed at high latitudes.
 
-\subsubsection{Spectral-element dynamical core}
+\subsubsection{Spectral-element (SE) dynamical core}
 
@@
 The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}).
-Each element contains a fourth order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points.
+Each element contains a fourth order basis set in the two horizontal directions, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Lagendre (GLL) quadrature points.
 This results in 16 GLL nodal points within each element, with 12 of the points lying on the (shared) element boundary.
-Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal values and produces a continuous global basis set.
+Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal point values and produces a continuous global basis set.
 
-As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to a Eulerian reference grid.
-A dry mass vertical coordinate was more recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}.
+As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to an Eulerian reference grid.
+A dry mass vertical coordinate was recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}.
 The 2D dynamics have no implicit dissipation and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors \cite{DetAl2012IJHPCA}.
@@
 
-For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Mulit-tracer tranport scheme is used instead \cite<CSLAM;>{LTOUNGK2017MWR}.
+For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) is used instead \cite<CSLAM;>{LTOUNGK2017MWR}.
 CSLAM has improved tracer property preservation and accelerated multi-tracer transport.
-It uses a seperate grid from the spectral-element dynamics, through dividing each element into $3\times3$ quasi-equal area control volumes.
+It uses a separate grid from the spectral-element dynamics, through dividing each element into $3\times3$ control volumes with quasi-equal area.
 The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated at the GLL nodal points \cite{HL2018MWR}.
@@
 
-The CAM convention is that the physics package determines the vertical resolution.
-Since all runs use CAM6 physics, all grids and dycores use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$.
-The physics time-step, in contrast, is dependent on grid resolution.
-Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step should be reduced to avoid large time truncation errors \cite{HR2018JAMES}.
-The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids.
-
 \subsection{Experimental design}
@@
 Before delving into the simulated characteristics of the Arctic, an understanding of the global mean differences between the grids and dycores are assessed, Figure~\ref{fig:dT-lores} shows the climatological, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
-The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.
-This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
+The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.This is common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
 The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the mid-latitudes in the $f09$ grid.
@@
 As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}).
-The stratosphere has a different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
+The stratosphere has a different signal, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
 The differences in temperature between $ne30pg3$ and $ne30pg2$ are small, with a slight warming near the tropopause at high latitudes.
-This is consistent with the similar climates found between these grids in \cite{HETAL2019JAMES}.
+This is consistent with the similar climates using these grids shown in \cite{HETAL2019JAMES}.
 
-Comparing the variable-resolution grids to the uniform resolution grids is complicated because we are simultaneously increase the grid resolution and reducing the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}.
-An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$.
-Figure~\ref{fig:dT-hires} shows the change in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$.
-A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB.
-Figure~\ref{fig:dT-hires} shows the difference in climatological temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
+In comparing the variable-resolution grids to the uniform resolution grids is convoluted because not only is the resolution higher in refinement region, but the physics time-step is shorter in the variable-resolution runs.
+An additional $ne30pg3$ simulations was run with physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$.
+Figure~\ref{fig:dT-hires} shows the changes in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$.
+A similar warming response to increase resolution occurs when the time-step is reduced, and the mechanism is similar in that this faciliates greater condensational heating by CLUBB.
+Figure~\ref{fig:dT-hires} shows difference in climatolgical temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
 The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
+         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB_phytstep_JJA.pdf}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
+         \includegraphics[width=100mm]{figs/temp_tseries_GRIS_map_TO_f19.pdf}
 \end{center}

### 2021-05-06 arcticgrids.tex
figure caption

--- 2021-05-05 arcticgrids.tex
@@
 \end{center}
-\caption{.}
+\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to RACMO3.2.
+The raw fields are mapped to two target low resolution grids, f19 \& ne30pg2, and using two different remapping methods, conservative ESMF and high order TempestRemap.
+The remapped values are then integrated over the ice mask of the target grid.
+This gives four time-series for each simulation (three for f19 \& ne30pg2), with the mean value given by the solid line and shading spanning the extent of the remapped solutions.}
 \label{fig:tseries}

### 2021-05-07 arcticgrids.tex
marcus edits

@@
 
-The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://ncar.github.io/CAM/doc/build/html/index.html}), supports a diverse number of atmospheric dynamical cores.
-These range from dycores using latitude-longitude grids, the finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, the spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
+The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://ncar.github.io/CAM/doc/build/html/index.html}), supports a number of different atmospheric dynamical cores.
+These include dycores using latitude-longitude grids, such as finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
 The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores in the current model.
-FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced, and so is omitted from this study.
-The authors instead focus on the two most well supported dycores in CAM, SE and FV.
+FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced; both the EUL and FV3 dycores are omitted from this study.
+As such, the results presented in this study are comparing the performance of the SE and FV dycores.
 
-\subsubsection{Finite-volume dynamical core}
+\subsubsection{Finite-volume (FV) dynamical core}
 
@@
 A polar filter is used to avoid computational instability due to the convergence of the meridians, allowing for a more practical time-step.
-It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically poleward \cite{ST1995GEOS}.
+It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the poleward direction \cite{ST1995GEOS}.
 
-\subsubsection{Spectral-element dynamical core}
+\subsubsection{Spectral-element (SE) dynamical core}
 
@@
 
-As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to a Eulerian reference grid.
+As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to an Eulerian reference grid.
 A dry mass vertical coordinate was more recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}.
@@
 
-For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Mulit-tracer tranport scheme is used instead \cite<CSLAM;>{LTOUNGK2017MWR}.
+For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) is used instead \cite<>{LTOUNGK2017MWR}.
 CSLAM has improved tracer property preservation and accelerated multi-tracer transport.
-It uses a seperate grid from the spectral-element dynamics, through dividing each element into $3\times3$ quasi-equal area control volumes.
+It uses a separate grid from the spectral-element dynamics, through dividing each element into $3\times3$ control volumes with quasi-equal area.
 The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated at the GLL nodal points \cite{HL2018MWR}.

### 2021-05-13 arcticgrids.tex
zwally basin figure
new racmo version in fig 6

@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_PRECT.pdf}
+         \includegraphics[width=130mm]{figs/temp_diffRACMO23p2_5p5km.pdf}
 \end{center}
-\caption{.}
+\caption{Climatological (1979-1998) annual precipitation rate bias relative to the RACMO3.2p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
 \label{fig:prect}
@@
 \label{fig:tseries}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_tseries_BASIN_uqmap.pdf}
+\end{center}
+\caption{.}
+\label{fig:zwally}
 \end{figure}

### 2021-10-14 arcticgrids.tex
new racmo version
new climo smb components figure

@@
 \end{center}
-\caption{Climatological (1979-1998) annual precipitation rate bias relative to the RACMO3.2p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
+\caption{Climatological (1979-1998) annual precipitation rate bias relative to the RACMO2.3p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
 \label{fig:prect}
@@
 \label{fig:tseries}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_climo_GRIS.pdf}
+\end{center}
+\caption{.}
+\label{fig:climo}
 \end{figure}

### 2021-12-21 arcticgrids.tex
writing

@@
 The Community Land Model (CLM5; \url{https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/index.html}), which uses the same grid as the atmosphere grid, calculates the surface energy balance at each land tile within a grid cell, which is used to compute snow and bare ice melting to inform the surface mass balance (SMB) of a glacier unit \cite{VKETAL2020JGR}.
+For land ice tiles coinciding with the Antarctic and Greenland Ice Sheets, the surface energy and mass balance are evaluated at ten elevation classes, and integrated over sub-grid hypsometric area-elevation bins for more accurate estimation of the surface mass balance of the big ice sheets.
 Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation.
@@
 
-Before delving into the simulated characteristics of the Arctic, an understanding of the global mean differences between the grids and dycores are assessed, Figure~\ref{fig:dT-lores} shows the climatological, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
+Before delving into the simulated characteristics of the Arctic, the global mean differences between the grids and dycores are assessed, Figure~\ref{fig:dT-lores} shows the climatological, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
 The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.
@@
 
-\subsection{Inter-annual variability}
 \subsection{Synoptic-scale storm characteristics}
+Synoptic storms are identified and analyzed using TempestExtremes \cite{UETAL2021}.
+The new $ARCTIC$ grid contains $\frac{1}{4}^{\circ}$ refinement north of about $45^{\circ}$ latitude, and so the storm tracker is applied to this region for the $ARCTIC$ and $ne30pg3$ run to identify differences in storm characteristics due to horizontal resolution.
+Figure~\ref{fig:comp-mean} shows the mean precipitation rates averaged over all January storms identified by TempestExtremes.
+The iconic comma structure of the mid-latitude cyclones is simulated in $ne30pg3$ and $ARCTIC$ grids, with the magnitudes about the same in these two grids, with perhaps a marginal increase in precipitation rates in the storm center of the $ARCTIC$ grid.
+For good measure, the $ne30pg3^{*}$ run is also plotted, and looks more-or-less identical to the $ne30pg3$ run.
+
+As has been previously reported, horizontal resolution can have large impacts on extreme precipitation events.
+Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with synoptic storms, by month.
+The $ARCTIC$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
+This is primarily due to an increase in resolution; the $ne30pg3^{*}$ only marginally increases the precipitation rates compared with $ne30pg3$.
+It should also be noted that the PDFs are evaluated on an identical composite grid, and so storm statistics are not impacted by differences in dataset resolution.
+
+The extreme precipitation rates in the $ARCTIC$ run are closer to the ERA5 reanalysis than in $ne30pg3$ (Figure~\ref{fig:comp-pdf}).
+It's difficult to know the extent that the extreme precipitation rates in ERA5 are constrained by data assimilation, or whether these precipitation rates are due to using a similar $\frac{1}{4}^{\circ}$ model as the $ARCTIC$ grid.
+However, precipitation rates in $\frac{1}{4}^{\circ}$ models tend to produce more skillful extreme events \cite{OETAL2016JAMES}.
+
 \subsection{Orographic gravity waves emanating from Greenland}
@@
 \label{fig:dT-hires}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
+\end{center}
+\caption{.}
+\label{fig:comp-mean}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_composite_ge45N_pdf.pdf}
+\end{center}
+\caption{.}
+\label{fig:comp-pdf}
 \end{figure}

### 2021-12-23 arcticgrids.tex
writing

@@
 Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation.
-Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up  by forcing CLM5 offline, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years.
+Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up  by forcing CLM5 standalone, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years.
 The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.
+
+\subsection{SMB Analysis}
+
+A common high resolution dataset is used to generate the GrIS boundary conditions using the CLM dataset creation tools.
+Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change over the GrIS.
+Figure~\ref{fig:grisdx} shows the GrIS ice mask area across the different grids, as a function of the number of grid points.
+Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset.
+RACMO, however, uses a smaller ice mask, about $3\%$ smaller than the raw CLM dataset.
+While Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, we seek to compare these integrated quantities with RACMO.
+Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks.
+Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ low resolution grids.
+In addition, two remapping packages are used; ESMF conservative and a TempestRemap high-order, monotone algorithm.
+In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.
 
@@
 
-Comparing the variable-resolution grids to the uniform resolution grids is complicated because we are simultaneously increase the grid resolution and reducing the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}.
+Comparing the variable-resolution grids to the uniform resolution grids is complicated because we simultaneously increase the grid resolution and reduce the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}.
 An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$.
@@
 This is primarily due to an increase in resolution; the $ne30pg3^{*}$ only marginally increases the precipitation rates compared with $ne30pg3$.
-It should also be noted that the PDFs are evaluated on an identical composite grid, and so storm statistics are not impacted by differences in dataset resolution.
+The $f09$ and $f19$ grids looks similar to $ne30pg3$, but with $f09$ having more frequent extreme precipitation events than $f19$ (not shown).
+It should also be noted that the PDFs are evaluated on an identical composite grid for all runs, and so storm statistics are not impacted by differences in output resolution.
 
@@
 \subsection{Greenland surface mass balance}
+
+The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study.
+As the accuracy of the computed SMB is sensitive to grid resolution, Figure~\ref{fig:grisdx} shows the average grid spacing for all six grids in this study.
+The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $ARCTICGRIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO model.
+The $ne30pg3$ grid has an average $\Delta x=111.2~km$, which is substantially more coarse than the $f09$ grid, having an average $\Delta x=60~km$.
+This is interesting because $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, but due to the convergence of meridians the finite-volume model has enhanced resolution over GrIS.
+The $ARCTIC$ grid has an average grid spacing over GrIS of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
+The $ARCTICGRIS$ grid is about twice as expensive as the $ARCTIC$ grid.
 
@@
 \caption{.}
-\label{fig:prect}
+\label{fig:grisdx}
 \end{figure}

### 2022-03-12 arcticgrids.tex
added results text

@@
 
-\title{Impact of grids and dynamical cores in CESM2.2 on the meteorology and climate of the Arctic}
+\title{Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet}
 
@@
 
+The SMB is the sum of the mass source term, accumulation, or equivalently, precipitation, and the mass sink term, ablation.
+Ablation can be expressed as total runoff, which is a combination of liquid precipitation and snow and ice melt.
+Not all liquid precipitation becomes runoff; rain may penetrate pore spaces in the firn layer and freeze.
+These processes are represented by various components in CESM, and the surface mass balance is computed by CLM.
+CLM uses a downscaling technique to account for sub-grid variability in SMB.
+In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation model.
+The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, allowing for the precipitation rates to be repartitioned into solid or liquid based on the temperature of each EC.
+This technique allows for a unique surface energy balance and surface mass balance at each EC.
+Integrating over all ECs using the area weights then provides a more accurate SMB.
+For a more detailed description of how the surface mass balance is computed in the CESM, see \cite{KETAL2020JAMES}.
+
+Simulated SMB are compared against several observational data products.
+RACMO2.3 11km and RACMO2.3p2 5.5km regional simulations forced by reanalysis at it's domain boundaries, performs very skillfully relative to observations and is therefore considered an ideal modeling target \cite{NETAL2019SCIENCE}.
+The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit} maintains a repository of snow pit and and ice core SMB measurements, as well as the IceBridge radar accumulation dataset.
+
 A common high resolution dataset is used to generate the GrIS boundary conditions using the CLM dataset creation tools.
@@
 
-Before delving into the simulated characteristics of the Arctic, the global mean differences between the grids and dycores are assessed, Figure~\ref{fig:dT-lores} shows the climatological, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
+Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed.
+Figure~\ref{fig:dT-lores} shows the climatological, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
 The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.
@@
 
-\subsection{Synoptic-scale storm characteristics}
+\subsection{Greenland surface mass balance}
+
+The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study.
+As the accuracy of the computed SMB is sensitive to grid resolution, Figure~\ref{fig:grisdx} shows the average grid spacing for all six grids in this study.
+The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $ARCTICGRIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO model.
+The $ne30pg3$ grid has an average $\Delta x=111.2~km$, which is substantially more coarse than the $f09$ grid, having an average $\Delta x=60~km$.
+This is interesting because $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, but due to the convergence of meridians the finite-volume model has enhanced resolution over GrIS.
+The $ARCTIC$ grid has an average grid spacing over GrIS of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
+The $ARCTICGRIS$ grid is about twice as expensive as the $ARCTIC$ grid.
+
+The climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:grisdx}, defined as differences from the RACMO2.3p2 5.5km solution \cite{NETAL2019SCIENCE}.
+What sticks out is that the uniform $1^{\circ}-2^{\circ}$ grids have positive bias centered over the southern dome.
+The $ARCTIC$ runs improves this bias substantially, and is all but eliminated in $ARCTCIGRIS$ run.
+The southern dome bias is therefore due to inadequate horizontal resolution, which is consistent with the results of \cite{VETAL2018TC}.
+Southeast Greenland has the largest accumulation rates in GrIS due to synoptic systems moving in from the southeast, which become orographically lifted by the steep southeast ice sheet margin, dumping large amounts of precipitation along the southeast coast.
+At lower horizontal resolutions, the topography is too smooth and large amounts of moisture penetrates further inland, incorrectly dumping large amounts of precipitation onto the interior of the ice sheet.
+
+Figure~\ref{fig:tseries} shows time-series of integrated precipitation and snow/ice melt over the GrIS in the various different grids and dycores, with both RACMO2.3 and 2.3p2 show in black.
+The 1979-1998 climatological mean values are provided as circles on the right side of the panels.
+The uniform $1^{\circ}-2^{\circ}$ grids all show a distinctive positive bias in precipitation due to this over-prediction of interior precipitation rates.
+The the variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO.The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +170 Gt/yr and $ne30pg2$, +230 Gt/yr.
+The results suggest that uniform resolution spectral-element grids have larger biases than the finite-volume grids, consistent with spectral-element grids having a coarser representation of GrIS (Figure~\ref{fig:grisdx}).
+While it seems likely that $ne30pg2$ has the largest bias of all the grids, it's difficult to say for certain that the $ne30pg3$ precipitation biases are worse than the finite-volume grids, owing to the use of a single realization for each grid.
+
+The combined snow/ice melt integrated over the GrIS is given by the bottom panel of Figure~\ref{fig:tseries}.
+The $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO.
+The $ARCTICGRIS$ grid over predicts melting by about 150 Gt/yr.
+This is likely due to an anomalously warm lower troposphere during the summer, relative to the $ARCTIC$ run (not shown).
+The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, with a positive melt bias between 70-90 Gt/yr.
+The spectral-element grids have the largest positive melt bias, between 210-230 Gt/yr.
+
+Figure~\ref{fig:pointdiff} shows the distribution of point-wise differences from LIVVkit observational database, as a violin plot.
+The IceBridge dataset is exclusively from the interior of the ice sheet and represents accumulation rates.
+The uniform $1^{\circ}-2^{\circ}$ grids have similar median errors of about +35-50 mm.w.e, while the variable-resolution errors are noticeably smaller.
+The in-situ observations in the accumulation zone, shown in the middle plot looks very similar to the IceBridge errors, and so provides confidence that the variable-resolution grids are outperforming the uniform grids in the interior accumulation zone.
+The errors evaluated at in-situ ablation zone measurements are in tension with the RACMO results in Figure~\ref{fig:tseries}, indicating that the uniform $1^{\circ}-2^{\circ}$ grids perform similarly, the $ARCTIC$ grid showing a reduction in bias and the $ARCTICGRIS$ having the smallest bias of all the grids.
+However, the in-situ ablation measurements are sparse in time and space, and so there is a large uncertainty in extending these results to the entirety of the ablation zone.
+
+The infamous k-transect mass balance profile is contained in the LIVVkit database, and shown in Figure~\ref{fig:trans}, along with the all the simulations.
+The in-situ observation points seem to be nicely replicated by the $ARCTCIGRIS$ run, whereas the $ARCTIC$ grid seems to biased positive in the higher-elevations of the ablation zone.
+The $f09$ grid is surprisingly competitive with the variable-resolution grids, capturing a realistic slope of the elevation-SMB curve.
+The elevation-SMB slopes of the uniform spectral-element and $f19$ grid simulate too shallow a slope in the higher elevation regions of the k-transect.
+Figure~\ref{ztrans} shows the representation of the surface of the ice sheet along these transects in the simulations, compared to a high resolution elevation dataset.
+The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
+The $f09$ grid is a bit of an exception, since the grid cells become very narrow in the meridional direction at high latitudes, and so a larger number of grid cells can populate the transect, consistent with its skillful representation of the ablation zone.
+
+What the authors refer to as the ``b-transect" in northwest Greenland (Figure~\ref{fig:trans}) is characterized by orographic precipitation, resulting in the accumulation zone extending down to the ice margin.
+The variable resolution grids perform relatively well, with larger SMBs at lower elevations where the precipitation rates are highest, whereas the $1^{\circ}-2^{\circ}$ grids underestimate the SMB in these lower elevation regions.
+Only the variable resolution grids can capture the local reduction in SMB in the 1500-2000 m region.
+The skill of the variable-resolution grids is clearly related to the accurate representation of this steep transect, while also capturing the protrusion around 1500-2000 m that coincides with the local minimum in SMB (Figure~\ref{ztrans}).
+
+\subsection{Impact of horizontal resolution on extremes}
 Synoptic storms are identified and analyzed using TempestExtremes \cite{UETAL2021}.
@@
 \subsection{Katabatic winds emanating from Greenland}
-\subsection{Greenland surface mass balance}
-
-The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study.
-As the accuracy of the computed SMB is sensitive to grid resolution, Figure~\ref{fig:grisdx} shows the average grid spacing for all six grids in this study.
-The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $ARCTICGRIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO model.
-The $ne30pg3$ grid has an average $\Delta x=111.2~km$, which is substantially more coarse than the $f09$ grid, having an average $\Delta x=60~km$.
-This is interesting because $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, but due to the convergence of meridians the finite-volume model has enhanced resolution over GrIS.
-The $ARCTIC$ grid has an average grid spacing over GrIS of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
-The $ARCTICGRIS$ grid is about twice as expensive as the $ARCTIC$ grid.
 
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
-\end{center}
-\caption{.}
-\label{fig:comp-mean}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_composite_ge45N_pdf.pdf}
-\end{center}
-\caption{.}
-\label{fig:comp-pdf}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
 \begin{tabular}{cccc}
@@
 \end{center}
-\caption{Climatological (1979-1998) annual precipitation rate bias relative to the RACMO2.3p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
+\caption{Climatological (1979-1998) annual precipitation rate bias (in mm/yr) relative to the RACMO2.3p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
 \label{fig:prect}
@@
 
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=100mm]{figs/temp_climo_GRIS.pdf}
-\end{center}
-\caption{.}
-\label{fig:climo}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=100mm]{figs/temp_tseries_BASIN_uqmap.pdf}
-\end{center}
-\caption{.}
-\label{fig:zwally}
-\end{figure}
+%\begin{figure}[t]
+%\begin{center}
+%         \includegraphics[width=100mm]{figs/temp_tseries_BASIN_uqmap.pdf}
+%\end{center}
+%\caption{.}
+%\label{fig:zwally}
+%\end{figure}
 
@@
 \begin{tabular}{cccc}
+         \includegraphics[width=25mm]{figs/temp_icemask.pdf} \\
+         \includegraphics[width=50mm]{figs/temp_zsmb_ktrans_obsperiod.pdf}&
+         \includegraphics[width=50mm]{figs/temp_zsmb_btrans_obsperiod.pdf} \\
+\end{tabular}
+\end{center}
+\caption{.}
+\label{fig:trans}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cccc}
          \includegraphics[width=100mm]{figs/temp_ktransect.pdf} \\
@@
 \caption{.}
-\label{fig:prect}
+\label{fig:ztrans}
 \end{figure}
@@
 \begin{center}
-\begin{tabular}{cccc}
-         \includegraphics[width=60mm]{figs/temp_zsmb_ktrans_obsperiod.pdf}&
-         \includegraphics[width=60mm]{figs/temp_zsmb_btrans_obsperiod.pdf} \\
-\end{tabular}
+         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
 \end{center}
 \caption{.}
-\label{fig:prect}
+\label{fig:comp-mean}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_composite_ge45N_pdf.pdf}
+\end{center}
+\caption{.}
+\label{fig:comp-pdf}
 \end{figure}

### 2022-03-12 arcticgrids.tex
updated figures

@@
 
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_grisres.pdf}
+\end{center}
+\caption{The spatial properties of the GrIS for different grids in this study.
+(Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks.}
+\label{fig:grisdx}
+\end{figure}
+
 Simulated SMB are compared against several observational data products.
@@
 \label{fig:dT-hires}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{cccc}
-         \includegraphics[width=60mm]{figs/temp_grisres_dxoverglc.pdf}&
-         \includegraphics[width=60mm]{figs/temp_grisres_error.pdf} \\
-\end{tabular}
-\end{center}
-\caption{.}
-\label{fig:grisdx}
 \end{figure}

### 2022-03-13 arcticgrids.tex
mods

@@
 \end{center}
-\caption{.}
+\caption{Computational grids for the uniform $1^{\circ}-2^{\circ}$ grids in this study.}
 \label{fig:uni-grids}
@@
 \end{center}
-\caption{.}
+\caption{Spectral-element grid for the variable-resolution ARCTIC grid in this study. Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
 \label{fig:vr-grids}
@@
 
-The SMB is the sum of the mass source term, accumulation, or equivalently, precipitation, and the mass sink term, ablation.
-Ablation can be expressed as total runoff, which is a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation becomes runoff; rain may penetrate pore spaces in the firn layer and freeze.
-These processes are represented by various components in CESM, and the surface mass balance is computed by CLM.
+The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation.
+Ablation can be expressed as sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
+Not all liquid precipitation becomes runoff; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface.
+All these processes are represented by different components in CESM, but the surface mass balance proper is computed by CLM.
 CLM uses a downscaling technique to account for sub-grid variability in SMB.
@@
 \end{center}
-\caption{The spatial properties of the GrIS for different grids in this study.
+\caption{The spatial properties of the GrIS as represented by different grids in this study.
 (Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks.}

### 2022-03-16 arcticgrids.tex
updates

@@
 General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}, which refers to numerical instability arising from the convergence of meridians at the polar points on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
+Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}, which refers to numerical instability arising from the convergence of meridians to polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
 Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain.
@@
 
+Initial simulations with the $ne30pg3$ spectral-element grid produced a weaker shortwave cloud forcing relative to the tuned up finite-volume dycore.
+All runs with the spectral-element dycore have two CLUBB parameter changes in order to provide a more realistic shortwave cloud forcing.
+These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and c14, reduced from 2.2 to 1.6.
+Briefly, the $gamma$ parameter scales the width of the sub-grid distribution of vertical velocity, and c14 controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy.
+For a thorough explanation of how CLUBB parameters impact the simulated climate, the reader is referred to \cite{GETAL2015JAMES}.
+
 \subsection{Experimental design}
@@
 Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation.
-Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up  by forcing CLM5 standalone, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years.
+Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up by forcing CLM in standalone mode, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years.
 The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.
@@
 The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation.
-Ablation can be expressed as sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
+Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
 Not all liquid precipitation becomes runoff; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface.
-All these processes are represented by different components in CESM, but the surface mass balance proper is computed by CLM.
+All these processes are represented by different components in CESM, and the surface mass balance is computed by CLM.
 CLM uses a downscaling technique to account for sub-grid variability in SMB.
@@
 Integrating over all ECs using the area weights then provides a more accurate SMB.
-For a more detailed description of how the surface mass balance is computed in the CESM, see \cite{KETAL2020JAMES}.
+For a detailed description of how the SMB is computed in the CESM, the reader is referred to \cite{SETAL2019TC,KETAL2020JAMES}.
 
@@
 Simulated SMB are compared against several observational data products.
-RACMO2.3 11km and RACMO2.3p2 5.5km regional simulations forced by reanalysis at it's domain boundaries, performs very skillfully relative to observations and is therefore considered an ideal modeling target \cite{NETAL2019SCIENCE}.
+RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations that have been forced by reanalysis at its domain boundaries.
+The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.
 The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit} maintains a repository of snow pit and and ice core SMB measurements, as well as the IceBridge radar accumulation dataset.
+The LIVVkit dataset is compared against model simulations by finding the nearest grid cell center to the location of each observation.
 
 A common high resolution dataset is used to generate the GrIS boundary conditions using the CLM dataset creation tools.
-Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change over the GrIS.
+Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS.
 Figure~\ref{fig:grisdx} shows the GrIS ice mask area across the different grids, as a function of the number of grid points.
 Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset.
-RACMO, however, uses a smaller ice mask, about $3\%$ smaller than the raw CLM dataset.
+RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset.
 While Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, we seek to compare these integrated quantities with RACMO.
 Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks.
-Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ low resolution grids.
-In addition, two remapping packages are used; ESMF conservative and a TempestRemap high-order, monotone algorithm.
+Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$, the lowest resolution grid for both dycores considered in this study.
+In addition, two remapping algorithms are used; ESMF conservative and TempestRemap high-order, monotone algorithm.
 In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.
@@
 Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed.
-Figure~\ref{fig:dT-lores} shows the climatological, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
+Figure~\ref{fig:dT-lores} shows plots of the 1979-1998 annual mean, zonal mean height expressed as differences between the uniform resolution grids and dycores.
 The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.
@@
 The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the mid-latitudes in the $f09$ grid.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-lores.pdf}
+\end{center}
+\caption{.}
+\label{fig:dT-lores}
+\end{figure}
 
@@
 The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
+\end{center}
+\caption{.}
+\label{fig:dT-hires}
+\end{figure}
+
+It's useful to understand summer temperature biases, instead of annual means, due to its control on ice/snow melt \cite{O2001JAM,HT2008Paleo}.
+Figure~\ref{fig:dThyps} shows the 1979-1998 lower troposphere summer temperature bias relative to ERA5.
+It is computed from the 500 hPa-1000 hPa geopotential thickness, solving for the layer mean virtual temperature using the hypsometric equation.
+The results generally track with the analysis of the zonal mean height plots; increasing resolution from $f19$ to $f09$ leads to a warmer climate, and the $1^{\circ}$ spectral-elements grids are warmer than the finite-volume grids.
+The summer temperatures in the finite-volume grids are persistently colder than ERA5 at high latitudes, whereas the $1^{\circ}$ spectral-element grids are warmer than ERA5 at only very high-latitudes, north of $80^{\circ}$.
+The $ARCTIC$ grid has similar summer temperatures to the $1^{\circ}$ spectral-element grids, except for a region extending from eastern N.Atlantic to northern Eurasia.
+The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic region, but in particular west of Greenland in the Baffin Island/Hudson Bay region, extending westward to the northern coast of Alaska.
+
+Some of these temperature anomalies may be related to summer shortwave cloud forcing differences across the different grids and dycores.
+Figure~\ref{fig:SWCF} shows the summer short-wave cloud forcing climatology in the runs.
+The uniform $1^{\circ}-2^{\circ}$ grids have similar cloud forcing, in the range of -85 to -90 W/m2 averaged north of $40^{\circ}$ latitude.
+The $ARCTIC$ grid has a weaker cloud forcing of about -80 W/m2, with less negative cloud forcing over the north pole, and the poleward side of the northern hemisphere land masses.
+The $ARCTCGRIS$ grid has much weaker cloud forcing at -70 W/m2, being systematically lower over most of the region plotted.
+While the cloud forcing to the west of Greenland is reduced in $ARCTCIGRIS$, there are also anomalous southerly winds in this region during summer (not shown), and so the large warm bias to the west of Greenland in Figure~\ref{fig:dThyps} is likely a combination of both cloud radiative and circulation changes.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_contours_diffERA5_Thyps.pdf}
+\end{center}
+\caption{.}
+\label{fig:dThyps}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_contours_diffCERES_SWCF.pdf}
+\end{center}
+\caption{.}
+\label{fig:SWCF}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_FSDS.pdf}
+\end{center}
+\caption{.}
+\label{fig:FSDS}
+\end{figure}
 
@@
 The in-situ observations in the accumulation zone, shown in the middle plot looks very similar to the IceBridge errors, and so provides confidence that the variable-resolution grids are outperforming the uniform grids in the interior accumulation zone.
-The errors evaluated at in-situ ablation zone measurements are in tension with the RACMO results in Figure~\ref{fig:tseries}, indicating that the uniform $1^{\circ}-2^{\circ}$ grids perform similarly, the $ARCTIC$ grid showing a reduction in bias and the $ARCTICGRIS$ having the smallest bias of all the grids.
-However, the in-situ ablation measurements are sparse in time and space, and so there is a large uncertainty in extending these results to the entirety of the ablation zone.
+The errors evaluated at in-situ ablation zone measurements are in tension with the RACMO results in Figure~\ref{fig:tseries}.
+They indicate that the uniform $1^{\circ}-2^{\circ}$ grids perform similarly, the $ARCTIC$ grid showing a reduction in bias and the $ARCTICGRIS$ having the smallest bias of all grids.
+However, the in-situ ablation measurements are sparse in time and space, and so there is a large amount of uncertainty in extending these results to the entirety of the ablation zone.
 
-The infamous k-transect mass balance profile is contained in the LIVVkit database, and shown in Figure~\ref{fig:trans}, along with the all the simulations.
+The infamous k-transect mass balance profile is contained in the LIVVkit database, and shown in Figure~\ref{fig:trans}, along with all the simulations.
 The in-situ observation points seem to be nicely replicated by the $ARCTCIGRIS$ run, whereas the $ARCTIC$ grid seems to biased positive in the higher-elevations of the ablation zone.
 The $f09$ grid is surprisingly competitive with the variable-resolution grids, capturing a realistic slope of the elevation-SMB curve.
-The elevation-SMB slopes of the uniform spectral-element and $f19$ grid simulate too shallow a slope in the higher elevation regions of the k-transect.
+The elevation-SMB slopes of the uniform spectral-element and $f19$ grid are too shallow, in particular at the higher elevation regions of the k-transect.
 Figure~\ref{ztrans} shows the representation of the surface of the ice sheet along these transects in the simulations, compared to a high resolution elevation dataset.
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $f09$ grid is a bit of an exception, since the grid cells become very narrow in the meridional direction at high latitudes, and so a larger number of grid cells can populate the transect, consistent with its skillful representation of the ablation zone.
+The $f09$ grid is a bit of an exception.
+The $f09$ grid cells become very narrow in the meridional direction at high latitudes, and so a larger number of grid cells can populate the east-west transect, consistent with its skillful representation of the ablation zone compared with RACMO.
 
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-lores.pdf}
-\end{center}
-\caption{.}
-\label{fig:dT-lores}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
-\end{center}
-\caption{.}
-\label{fig:dT-hires}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_diffRACMO23p2_5p5km.pdf}
+         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_CLOUD_PRECIP.pdf}
 \end{center}

### 2022-03-20 arcticgrids.tex
updates

@@
 % Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-\authors{Adam R. Herrington \affil{1}, Marcus Lofverstrom \affil{2}, Peter H. Lauritzen \affil{1} and Andrew Gettelman \affil{1}}
+\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} $\ $and Andrew Gettelman\affil{1}}
 
@@
 General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}, which refers to numerical instability arising from the convergence of meridians to polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
+Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}.
+The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
 Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain.
 With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past.
-In this study, a spectrum of model grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}) are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the meteorology over the Greenland Ice Sheet.
+In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.
 
@@
 While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
-The pole-problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
+The pole problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
 Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
-One might expect that this additional damping reduces the effective resolution such that the resolved scales are similar across the entire domain, but this seems unlikely (describe why) and is explored further in this study.
+This additional damping reduces the effective resolution in polar regions, and the resolved scales are approximately the same everywhere on the grid.
 
 An alternative approach is to use unstructured grids, which allow for more flexible grid structures that permit quasi-uniform grid spacing globally and eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c).
-This grid flexibility also allows for variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}).
+This grid flexibility also permits variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}).
 Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region (the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space).
-However, unstructured grids scale more efficiently on parallel systems than latitude-longitude grids, likely resulting in a greater prevalence of unstructured grids as computing power continued to increase over time.
+Unstructured grids also scale more efficiently on parallel systems than latitude-longitude grids, likely resulting in a greater prevalence of unstructured grids as computing power continued to increase over time.
 
@@
 The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}).
-While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and melting) are only partially resolved at typical GCM resolutions.
+While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions.
 For example, GrIS precipitation is typically confined to the ice-sheet margins (predominately the southeastern margin) where orographic precipitation is generated by steep topographic slopes.
-Moreover, GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation.
+GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation.
 GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.
@@
 
-The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 (CAM; \url{https://ncar.github.io/CAM/doc/build/html/index.html}), supports a number of different atmospheric dynamical cores.
+The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 \cite<CAM;>{CAM63}, supports a number of different atmospheric dynamical cores.
 These include dycores using latitude-longitude grids, such as finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores in the current model.
+The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores.
 FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced; both the EUL and FV3 dycores are omitted from this study.
@@
 
-The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,DetAl2012IJHPCA}.
+The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}.
 The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}).
@@
 The SE dycore supports regional grid refinement via its variable-resolution configuration, requiring two enhancements over uniform resolution grids.
-(1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by about an order of magnitude per halving of the grid spacing.
+(1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
 A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements.
-(2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so  the NCAR topography software \cite{gmdd-8-4623-2015} was modified to scale the smoothing radius by the local element size.
+(2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size.
 
@@
 
-Six grid are evaluated in this study (Table X).
+Six grid are evaluated in this study (Table~\ref{tbl:table1}).
 The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
@@
 
+ \begin{table*}
+ \centering
+ \scriptsize
+ \begin{tabular}{lccccc}
+   \hline
+   grid name & dycore & $\Delta x_{eq}$ (km) & $\Delta x_{refine}$ (km) & $\Delta t_{phys}$ (s) \\ 
+   \hline
+   $f19$ & FV & 278 & - &1800 \\
+   $f09$ & FV & 139 & - &1800 \\
+   $ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
+   $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
+   $ARCTIC$ & SE & 111 & 28 & 450 \\
+   $ARCTCIGRIS$ & SE & 111 & 14 & 225 \\
+ \hline
+ \end{tabular}
+  \caption{Grids and dycores used in this study.
+$\Delta x_{eq}$ refers to average equatorial grid spacing, $\Delta x_{refine}$ refers to grid spacing in the refined region (if applicable) and $\Delta t_{phys}$ refers to the physics time-step.
+The dycore abbreviation FV refers to the finite-volume dycore, SE the spectral-element dycore and SE-CSLAM the spectral-element dycore w/ CSLAM tracer advection.}
+ \label{tbl:table1}
+ \end{table*}
+
 \subsection{Physical parameterizations}
 
-The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used for all simulations in this study.
+The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used in all simulations in this study.
 CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme.
@@
 
-The CAM convention is that the physics package determines the vertical resolution.
-Since all runs use CAM6 physics, all grids and dycores use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$.
-The physics time-step, in contrast, is dependent on grid resolution.
-Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step should be reduced to avoid large time truncation errors \cite{HR2018JAMES}.
-The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids.
-
-Initial simulations with the $ne30pg3$ spectral-element grid produced a weaker shortwave cloud forcing relative to the tuned up finite-volume dycore.
-All runs with the spectral-element dycore have two CLUBB parameter changes in order to provide a more realistic shortwave cloud forcing.
+All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$.
+The physics time-step is dependent on grid resolution.
+Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}.
+The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids (Table~\ref{tbl:table1}.
+
+Initial simulations with the $ne30pg3$ spectral-element grid produced weaker shortwave cloud forcing relative to the tuned up finite-volume dycore.
+All runs with the spectral-element dycore have two CLUBB parameter changes in order to provide a more realistic cloud forcing and top-of-atmosphere radiation balance.
 These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and c14, reduced from 2.2 to 1.6.
@@
 This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2.
-The Community Land Model (CLM5; \url{https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/index.html}), which uses the same grid as the atmosphere grid, calculates the surface energy balance at each land tile within a grid cell, which is used to compute snow and bare ice melting to inform the surface mass balance (SMB) of a glacier unit \cite{VKETAL2020JGR}.
-For land ice tiles coinciding with the Antarctic and Greenland Ice Sheets, the surface energy and mass balance are evaluated at ten elevation classes, and integrated over sub-grid hypsometric area-elevation bins for more accurate estimation of the surface mass balance of the big ice sheets.
-Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation.
-Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up by forcing CLM in standalone mode, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years.
-The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.
-
-\subsection{SMB Analysis}
-
+
+The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study.
 The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation.
 Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation becomes runoff; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface.
-All these processes are represented by different components in CESM, and the surface mass balance is computed by CLM.
-CLM uses a downscaling technique to account for sub-grid variability in SMB.
-In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation model.
-The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, allowing for the precipitation rates to be repartitioned into solid or liquid based on the temperature of each EC.
-This technique allows for a unique surface energy balance and surface mass balance at each EC.
-Integrating over all ECs using the area weights then provides a more accurate SMB.
-For a detailed description of how the SMB is computed in the CESM, the reader is referred to \cite{SETAL2019TC,KETAL2020JAMES}.
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_grisres.pdf}
-\end{center}
-\caption{The spatial properties of the GrIS as represented by different grids in this study.
-(Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks.}
-\label{fig:grisdx}
-\end{figure}
-
-Simulated SMB are compared against several observational data products.
-RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations that have been forced by reanalysis at its domain boundaries.
+Not all liquid precipitation becomes runs off the ice sheet; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface.
+These processes are represented by different components in CESM, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
+
+CLM runs on the same grid as the atmosphere, but also uses a downscaling technique to account for sub-grid variability in SMB.
+In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation dataset.
+The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, and the precipitation rates from CAM are repartitioned into solid or liquid based on the temperature of the EC.
+Ice accumulation is modeled as a capping flux, or snow in excess of a 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of ice.
+A unique surface energy balance and SMB is computed for each EC.
+Integrating over all ECs using the area weights provides a more accurate SMB.
+For a more detailed description of how the SMB is computed in CESM, the reader is referred to \cite{CISM1,SETAL2019TC,KETAL2020JAMES}.
+
+Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up by forcing CLM in standalone mode, cycling over a 20 year $ARCTIC$ $FHIST$ run for about 500 years.
+The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.
+
+\subsection{Observational Datasets}
+
+Several observational datasets are used in this study to understand the performance of the simulations.
+A list of the datasets used in this study are shown in Table~\ref{tbl:table2}.
+Several of these products (ERA5, CALIPSO and CERES) are near-global gridded datasets commonly used to evaluate GCMs.
+Surface mass balance datasets are gathered from several sources.
+RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalysis at its domain boundaries.
 The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.
@@
 
-A common high resolution dataset is used to generate the GrIS boundary conditions using the CLM dataset creation tools.
+ \begin{table*}
+ \centering
+ \scriptsize
+ \begin{tabular}{lccc}
+   \hline
+   data product & years used in this study & citation \\ 
+   \hline
+   ERA5 & 1979-1998 & \cite{ERA5} \\
+   CERES-EBAF ED4.1 & 2003-2020 & \cite{CERES-EBAF} \\
+   CALIPSO-GOCCP & 2006-2017 & \cite{CALIPSO-GOCCP} \\
+   RACMO2.3 & 1979-1998 & \cite{NETAL2015TC} \\
+   RACMO2.3p2 & 1979-1998 & \cite{NETAL2019SCIENCE} \\
+   LIVVkit,v2 & 1979-1998 & \cite{LIVVkit} \\
+ \hline
+ \end{tabular}
+  \caption{Description of observational datasets used in this study.}
+ \label{tbl:table2}
+ \end{table*}
+
+\subsection{SMB Analysis}
+
+A common high resolution dataset is used to generate the GrIS boundary conditions in all grids, using the CLM dataset creation tools.
 Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS.
@@
 RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset.
-While Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, we seek to compare these integrated quantities with RACMO.
+
+Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, but we seek to compare these integrated quantities to RACMO2.3.
 Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks.
@@
 
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_grisres.pdf}
+\end{center}
+\caption{The spatial properties of the GrIS as represented by different grids in this study.
+(Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks.}
+\label{fig:grisdx}
+\end{figure}
+
 \section{Results}\label{sec:results}
@@
 Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed.
-Figure~\ref{fig:dT-lores} shows plots of the 1979-1998 annual mean, zonal mean height expressed as differences between the uniform resolution grids and dycores.
+Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
 The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.
 This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
-The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the mid-latitudes in the $f09$ grid.
+The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the low and mid-latitudes in the $f09$ grid.
 
@@
 The stratosphere has a different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
-The differences in temperature between $ne30pg3$ and $ne30pg2$ are small, with a slight warming near the tropopause at high latitudes.
+Figure~\ref{fig:dT-lores} also shows differences in temperature between $ne30pg3$ and $ne30pg2$, which are small, although there is a slight warming near the tropopause at high latitudes.
 This is consistent with the similar climates found between these grids in \cite{HETAL2019JAMES}.
@@
 An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$.
-Figure~\ref{fig:dT-hires} shows the change in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$.
+Figure~\ref{fig:dT-hires} shows the change in the climatological summer temperatures in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$.
 A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB.
-Figure~\ref{fig:dT-hires} shows the difference in climatological temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
+Figure~\ref{fig:dT-hires} shows the difference in climatological summer temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
 The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.
@@
 The summer temperatures in the finite-volume grids are persistently colder than ERA5 at high latitudes, whereas the $1^{\circ}$ spectral-element grids are warmer than ERA5 at only very high-latitudes, north of $80^{\circ}$.
-The $ARCTIC$ grid has similar summer temperatures to the $1^{\circ}$ spectral-element grids, except for a region extending from eastern N.Atlantic to northern Eurasia.
-The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic region, but in particular west of Greenland in the Baffin Island/Hudson Bay region, extending westward to the northern coast of Alaska.
+All grid illustrate a north-south gradient in bias over Greenland, in which the summer temperature bias becomes more positive in the northward direction.
+This pattern is also evident in the 2m summer temperature bias over Greenland (not shown).
+
+The $ARCTIC$ grid has similar summer temperatures to the $1^{\circ}$ spectral-element grids, but it is a bit warmer over northern Eurasia and the North Pole.
+An anomalous cooling patch forms to the west of Greenland, centered over Baffin Island.
+The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic region, but approximately maintains the same pattern of summer temperature bias as in the $ARCTIC$ grid.
 
 Some of these temperature anomalies may be related to summer shortwave cloud forcing differences across the different grids and dycores.
-Figure~\ref{fig:SWCF} shows the summer short-wave cloud forcing climatology in the runs.
-The uniform $1^{\circ}-2^{\circ}$ grids have similar cloud forcing, in the range of -85 to -90 W/m2 averaged north of $40^{\circ}$ latitude.
-The $ARCTIC$ grid has a weaker cloud forcing of about -80 W/m2, with less negative cloud forcing over the north pole, and the poleward side of the northern hemisphere land masses.
-The $ARCTCGRIS$ grid has much weaker cloud forcing at -70 W/m2, being systematically lower over most of the region plotted.
-While the cloud forcing to the west of Greenland is reduced in $ARCTCIGRIS$, there are also anomalous southerly winds in this region during summer (not shown), and so the large warm bias to the west of Greenland in Figure~\ref{fig:dThyps} is likely a combination of both cloud radiative and circulation changes.
+Figure~\ref{fig:SWCF} shows the summer short-wave cloud forcing bias in the runs, using the CERES-EBAF product.
+All the uniform $1^{\circ}-2^{\circ}$ grids have similar biases, with the clouds reflecting 20-40 W/m2 too much shortwave radiation over a wide swath of the Arctic, primarily over the land masses.
+There's also a halo of low cloud forcing bias around the oceanic perimeter of Greenland.
+The $ARCTIC$ grid has much smaller cloud forcing biases over the Arctic land masses, although still too reflective, whereas the $ARCTCGRIS$ grid vastly improves the cloud forcing bias over Eurasia, and improves the bias over N.America compared to the $ARCTIC$ grid.
+In both variable-resolution grids, the halo of too weak cloud forcing bias around the perimeter of Greenland is absent.
+
+While the summer cloud forcing biases are consistent with the summer temperature biases in Figure~\ref{fig:dThyps} --regions where clouds are too reflective coincide with regions that are too cold-- it is not clear whether the cold biases are caused by the cloud biases, or whether the cold biases amplify the cloud forcing bias.
 
@@
 
+\subsection{Shortwave radiation over Greenland}
+
+In addition to summer temperatures, shortwave radiation is also an important determinant of snow/ice melt.
+Figure~\ref{fig:FSDS} shows the summer incident shortwave radiation bias at the surface, zoomed in over Greenland.
+The top panel computes the bias using the CERES-EBAF dataset, and the bottom panel using RACMO2.3p2 dataset.
+This halo of excessive incident shortwave radiation around the coasts of Greenland is apparent in both datasets, consistent with the shortwave cloud forcing biases in Figure~\ref{fig:SWCF}.
+
+The interior of the ice sheet receives too little shortwave radiation in the coarser grids.
+In the variable-resolution grids, both the interior deficit in shortwave and the excessive shortwave around the oceanic perimeter of Greenland are improved.
+This suggests that the coarse grids clouds are too thick in the interior of Greenland, and too thin around the perimeter of Greenland, and that increasing horizontal resolution balances out these biases.
+This is consistent with total summer cloud fraction bias, computed from the CALIPSO cloud dataset (Figure~\ref{fig:prect}).
+Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates any changes arising from any single layer due to the maximum overlap assumption used to compute this quantity.
+Despite the attenuated signal, the total cloud fraction does indicate a reduction in cloud coverage in the interior, and an increase in cloudiness about the oceanic perimeter, in the variable-resolution grids.
+
+The agreement of the cloud biases over Greenland in multiple independent datasets indicates this is a robust feature of the coarser grids.
+The reduction of these cloud biases in the variable-resolution grids suggests they are a result of insufficient horizontal resolution in coarse grids.
+
 \begin{figure}[t]
@@
 
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_CLOUD_PRECIP.pdf}
+\end{center}
+\caption{Climatological (1979-1998) annual precipitation rate bias (in mm/yr) relative to the RACMO2.3p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
+\label{fig:prect}
+\end{figure}
+
 \subsection{Greenland surface mass balance}
 
-The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study.
-As the accuracy of the computed SMB is sensitive to grid resolution, Figure~\ref{fig:grisdx} shows the average grid spacing for all six grids in this study.
-The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $ARCTICGRIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO model.
-The $ne30pg3$ grid has an average $\Delta x=111.2~km$, which is substantially more coarse than the $f09$ grid, having an average $\Delta x=60~km$.
+The accuracy of the simulated SMB is expected to be sensitive to grid resolution.
+Figure~\ref{fig:prect} shows the average grid spacing over the Greenland Ice Sheet (GrIS) in all six grids in this study.
+The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $ARCTICGRIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO2.3 grid.
+The $ne30pg3$ grid has an average $\Delta x=111.2~km$, which is substantially more coarse than the $f09$ grid, with an average $\Delta x=60~km$.
 This is interesting because $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, but due to the convergence of meridians the finite-volume model has enhanced resolution over GrIS.
-The $ARCTIC$ grid has an average grid spacing over GrIS of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
-The $ARCTICGRIS$ grid is about twice as expensive as the $ARCTIC$ grid.
-
-The climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:grisdx}, defined as differences from the RACMO2.3p2 5.5km solution \cite{NETAL2019SCIENCE}.
-What sticks out is that the uniform $1^{\circ}-2^{\circ}$ grids have positive bias centered over the southern dome.
-The $ARCTIC$ runs improves this bias substantially, and is all but eliminated in $ARCTCIGRIS$ run.
-The southern dome bias is therefore due to inadequate horizontal resolution, which is consistent with the results of \cite{VETAL2018TC}.
-Southeast Greenland has the largest accumulation rates in GrIS due to synoptic systems moving in from the southeast, which become orographically lifted by the steep southeast ice sheet margin, dumping large amounts of precipitation along the southeast coast.
-At lower horizontal resolutions, the topography is too smooth and large amounts of moisture penetrates further inland, incorrectly dumping large amounts of precipitation onto the interior of the ice sheet.
-
-Figure~\ref{fig:tseries} shows time-series of integrated precipitation and snow/ice melt over the GrIS in the various different grids and dycores, with both RACMO2.3 and 2.3p2 show in black.
+The $ARCTIC$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models (whereas the $ARCTICGRIS$ grid is about twice as expensive as the $ARCTIC$ grid).
+
+The summer climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:grisdx}, expressed as the fractional difference from the RACMO2.3p2 solution.
+What sticks out is that the coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
+The $ARCTIC$ run improves this bias substantially, and the $ARCTICGIS$ run improves this bias further.
+This suggests the southern dome bias is due to inadequate horizontal resolution, which is consistent with the original GrIS variable-resolution experiments in \cite{VETAL2018TC}.
+
+Southeast Greenland has the largest accumulation rates in GrIS due to synoptic systems moving in from the southeast.
+These systems are orographically lifted by the steep southeast ice sheet margin, dumping large amounts of precipitation along the southeast coast.
+At lower horizontal resolutions, the topography is too smooth and large amounts of moisture penetrates further inland, incorrectly dumping precipitation onto the interior of the ice sheet.
+A similar bias occurs in northwest Greenland, in particular during the summer, when it's common for synoptic systems to arrive from the southwest.
+The ability of the variable-resolution grids to more accurately simulate the orographic precipitation process in Greenland is consistent with all the cloud results up to this point.
+Since the precipitation centers move from the interior towards the coasts, and even out over the ocean with increasing resolution, the cloud decks should, and are, moved accordingly, eliminating this halo of low cloud bias around the oceanic perimeter of Greenland.
+
+Figure~\ref{fig:tseries} shows time-series of annually integrated precipitation and snow/ice melt over the GrIS in the various different grids and dycores, with both versions of RACMO shown in black.
 The 1979-1998 climatological mean values are provided as circles on the right side of the panels.
 The uniform $1^{\circ}-2^{\circ}$ grids all show a distinctive positive bias in precipitation due to this over-prediction of interior precipitation rates.
-The the variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO.The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +170 Gt/yr and $ne30pg2$, +230 Gt/yr.
+The  variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO.The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$, +230 Gt/yr.
 The results suggest that uniform resolution spectral-element grids have larger biases than the finite-volume grids, consistent with spectral-element grids having a coarser representation of GrIS (Figure~\ref{fig:grisdx}).
-While it seems likely that $ne30pg2$ has the largest bias of all the grids, it's difficult to say for certain that the $ne30pg3$ precipitation biases are worse than the finite-volume grids, owing to the use of a single realization for each grid.
-
-The combined snow/ice melt integrated over the GrIS is given by the bottom panel of Figure~\ref{fig:tseries}.
+
+The combined annual snow/ice melt integrated over the GrIS is given by the bottom panel of Figure~\ref{fig:tseries}.
 The $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO.
-The $ARCTICGRIS$ grid over predicts melting by about 150 Gt/yr.
-This is likely due to an anomalously warm lower troposphere during the summer, relative to the $ARCTIC$ run (not shown).
-The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, with a positive melt bias between 70-90 Gt/yr.
-The spectral-element grids have the largest positive melt bias, between 210-230 Gt/yr.
-
-Figure~\ref{fig:pointdiff} shows the distribution of point-wise differences from LIVVkit observational database, as a violin plot.
-The IceBridge dataset is exclusively from the interior of the ice sheet and represents accumulation rates.
-The uniform $1^{\circ}-2^{\circ}$ grids have similar median errors of about +35-50 mm.w.e, while the variable-resolution errors are noticeably smaller.
-The in-situ observations in the accumulation zone, shown in the middle plot looks very similar to the IceBridge errors, and so provides confidence that the variable-resolution grids are outperforming the uniform grids in the interior accumulation zone.
-The errors evaluated at in-situ ablation zone measurements are in tension with the RACMO results in Figure~\ref{fig:tseries}.
-They indicate that the uniform $1^{\circ}-2^{\circ}$ grids perform similarly, the $ARCTIC$ grid showing a reduction in bias and the $ARCTICGRIS$ having the smallest bias of all grids.
-However, the in-situ ablation measurements are sparse in time and space, and so there is a large amount of uncertainty in extending these results to the entirety of the ablation zone.
-
-The infamous k-transect mass balance profile is contained in the LIVVkit database, and shown in Figure~\ref{fig:trans}, along with all the simulations.
-The in-situ observation points seem to be nicely replicated by the $ARCTCIGRIS$ run, whereas the $ARCTIC$ grid seems to biased positive in the higher-elevations of the ablation zone.
-The $f09$ grid is surprisingly competitive with the variable-resolution grids, capturing a realistic slope of the elevation-SMB curve.
-The elevation-SMB slopes of the uniform spectral-element and $f19$ grid are too shallow, in particular at the higher elevation regions of the k-transect.
-Figure~\ref{ztrans} shows the representation of the surface of the ice sheet along these transects in the simulations, compared to a high resolution elevation dataset.
-The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $f09$ grid is a bit of an exception.
-The $f09$ grid cells become very narrow in the meridional direction at high latitudes, and so a larger number of grid cells can populate the east-west transect, consistent with its skillful representation of the ablation zone compared with RACMO.
-
-What the authors refer to as the ``b-transect" in northwest Greenland (Figure~\ref{fig:trans}) is characterized by orographic precipitation, resulting in the accumulation zone extending down to the ice margin.
-The variable resolution grids perform relatively well, with larger SMBs at lower elevations where the precipitation rates are highest, whereas the $1^{\circ}-2^{\circ}$ grids underestimate the SMB in these lower elevation regions.
-Only the variable resolution grids can capture the local reduction in SMB in the 1500-2000 m region.
-The skill of the variable-resolution grids is clearly related to the accurate representation of this steep transect, while also capturing the protrusion around 1500-2000 m that coincides with the local minimum in SMB (Figure~\ref{ztrans}).
-
-\subsection{Impact of horizontal resolution on extremes}
-Synoptic storms are identified and analyzed using TempestExtremes \cite{UETAL2021}.
-The new $ARCTIC$ grid contains $\frac{1}{4}^{\circ}$ refinement north of about $45^{\circ}$ latitude, and so the storm tracker is applied to this region for the $ARCTIC$ and $ne30pg3$ run to identify differences in storm characteristics due to horizontal resolution.
-Figure~\ref{fig:comp-mean} shows the mean precipitation rates averaged over all January storms identified by TempestExtremes.
-The iconic comma structure of the mid-latitude cyclones is simulated in $ne30pg3$ and $ARCTIC$ grids, with the magnitudes about the same in these two grids, with perhaps a marginal increase in precipitation rates in the storm center of the $ARCTIC$ grid.
-For good measure, the $ne30pg3^{*}$ run is also plotted, and looks more-or-less identical to the $ne30pg3$ run.
-
-As has been previously reported, horizontal resolution can have large impacts on extreme precipitation events.
-Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with synoptic storms, by month.
-The $ARCTIC$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
-This is primarily due to an increase in resolution; the $ne30pg3^{*}$ only marginally increases the precipitation rates compared with $ne30pg3$.
-The $f09$ and $f19$ grids looks similar to $ne30pg3$, but with $f09$ having more frequent extreme precipitation events than $f19$ (not shown).
-It should also be noted that the PDFs are evaluated on an identical composite grid for all runs, and so storm statistics are not impacted by differences in output resolution.
-
-The extreme precipitation rates in the $ARCTIC$ run are closer to the ERA5 reanalysis than in $ne30pg3$ (Figure~\ref{fig:comp-pdf}).
-It's difficult to know the extent that the extreme precipitation rates in ERA5 are constrained by data assimilation, or whether these precipitation rates are due to using a similar $\frac{1}{4}^{\circ}$ model as the $ARCTIC$ grid.
-However, precipitation rates in $\frac{1}{4}^{\circ}$ models tend to produce more skillful extreme events \cite{OETAL2016JAMES}.
-
-\subsection{Orographic gravity waves emanating from Greenland}
-\subsection{Katabatic winds emanating from Greenland}
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-
-\section{Conclusions}\label{sec:conclusions}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_CLOUD_PRECIP.pdf}
-\end{center}
-\caption{Climatological (1979-1998) annual precipitation rate bias (in mm/yr) relative to the RACMO2.3p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
-\label{fig:prect}
-\end{figure}
+The $ARCTICGRIS$ grid over predicts melting by about 125 Gt/yr.
+This is likely due to an anomalously warm lower troposphere during the summer, relative to the $ARCTIC$ run (Figure~\ref{fig:dThyps}).
+The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, simulating too much melt by about 70-90 Gt/yr.
+The spectral-element grids have the largest positive melt bias, between 200-220 Gt/yr.
+It is more difficult to attribute these differences to resolution alone, since the finite-volumes grids have colder summer temperatures than the uniform resolution spectral-element grids.
+However, that the $ARCTCIGRIS$ grid has the warmest summer temperatures, yet has a lower melting bias than the uniform spectral-element grids, suggests that increasing resolution improves the ablation process.
+
+ \begin{table*}
+ \centering
+ \scriptsize
+ \begin{tabular}{lccc}
+   \hline
+   grid name & accumulation (Gt/yr) & ice/snow melt (Gt/yr) \\ 
+   \hline
+   $RACMO$ & 768.5 & -347.2 \\
+   \hline
+   $f19$ & 882.5 & -440.3 \\
+   $f09$ & 874.8 & -418.4 \\
+   $ne30pg2$ & 1000. & -549.4 \\
+   $ne30pg3$ & 934.9 & -568.8 \\
+   $ARCTIC$ & 795.9 & -367.3 \\
+   $ARCTICGRIS$ & 708.7 & -471.6 \\
+ \hline
+ \end{tabular}
+  \caption{1979-1998 Surface Mass Balance of the Greenland Ice Sheet.}
+ \label{tbl:table3}
+ \end{table*}
 
@@
 \end{figure}
+
+Figure~\ref{fig:pointdiff} shows the distribution of point-wise differences from LIVVkit observational database as violin plots.
+The IceBridge dataset is exclusively from the interior of the ice sheet and represents accumulation rates.
+The uniform $1^{\circ}-2^{\circ}$ grids have similar median errors of about +35-50 mm.w.e, while the variable-resolution errors are noticeably smaller.
+The in-situ observations in the accumulation zone, shown in the middle plot looks very similar to the IceBridge errors, providing confidence that the variable-resolution grids are outperforming the uniform grids in the interior accumulation zone.
+The errors evaluated at in-situ ablation zone measurements are in tension with the RACMO results in Figure~\ref{fig:tseries}.
+They indicate that the uniform $1^{\circ}-2^{\circ}$ grids perform similarly, and that the $ARCTICGRIS$ performs best, and an improvement over the $ARCTIC$ grid.
+However, the in-situ ablation measurements are sparse in time and space, and so there is a large amount of uncertainty in extending these results to the entirety of the ablation zone.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_violens.pdf}
+\end{center}
+\caption{.}
+\label{fig:pointdiff}
+\end{figure}
+
+The ``k-transect" is perhaps the most well studied transect in Greenland, as it has been regularly monitored since the 1950s.
+The LIVVkit has compiled the k-transect observations, shown in Figure~\ref{fig:trans} as elevation vs.
+SMB along the transect, with the model's simulated transects shown as well.
+The in-situ observation points are nicely replicated by the $ARCTCIGRIS$ run, whereas the $ARCTIC$ grid is biased positive in the higher-elevations of the ablation zone.
+The $f09$ grid is surprisingly competitive with the variable-resolution grids, capturing a realistic slope of the elevation-SMB curve, although it exhibits a similar positive bias as the $ARCTIC$ grid.
+The elevation-SMB slopes of the uniform spectral-element and $f19$ grid are too shallow, in particular at the higher elevation regions of the transect.
+
+Figure~\ref{fig:ztrans} shows the representation of the surface of the ice sheet along these transects in the different grids, compared to the high resolution dataset used to generate CAM topography boundary conditions.
+The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
+The $f09$ grid is a bit of an exception --since the grid cells become very narrow in the meridional direction at high latitudes, a larger number of grid cells can populate the east-west transect, consistent with its skillful representation of the ablation zone.
+
+What the authors refer to as the ``b-transect" in northwest Greenland (Figure~\ref{fig:trans}) is characterized by orographic precipitation, resulting in the accumulation zone extending down to the ice margin.
+The variable resolution grids perform relatively well, with larger SMBs at lower elevations where the precipitation rates are highest, whereas the $1^{\circ}-2^{\circ}$ grids underestimate the SMB in these lower elevation regions.
+Only the variable resolution grids can capture the local reduction in SMB in the 1500-2000 m region.
+The skill of the variable-resolution grids is clearly related to the more accurate representation of the steepness of the transect, while also capturing the protrusion around 1500-2000 m that coincides with the local minimum in SMB (Figure~\ref{fig:ztrans}).
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_transect_obsperiod.pdf}
+\end{center}
+\caption{.}
+\label{fig:trans}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cccc}
+         \includegraphics[width=100mm]{figs/temp_ktransect.pdf} \\
+         \includegraphics[width=100mm]{figs/temp_btransect.pdf} \\
+\end{tabular}
+\end{center}
+\caption{.}
+\label{fig:ztrans}
+\end{figure}
+
+To give an idea of what the ablation zones look like in the highest resolution $ARCTCIGRIS$ grid, Figure~\ref{fig:viz} is a still from a visualization produced by the authors.
+It shows the daily mean surface mass balance during the height of the melt season in year 1981 of the simulation.
+The ablation zone appears to be well resolved in the $ARCTICGRIS$ grid.
+The full visualization is publicly available on youtube.com\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/Vis1923.jpg}
+\end{center}
+\caption{.}
+\label{fig:viz}
+\end{figure}
+
+\subsection{Impact of horizontal resolution on extremes}
+
+Synoptic storms are tracked and analyzed using TempestExtremes \cite{UETAL2021}.
+As the $ARCTIC$ grid contains $\frac{1}{4}^{\circ}$ refinement north of about $45^{\circ}$ latitude, the storm tracker is applied to this region for the $ARCTIC$ and $ne30pg3$ run to identify differences in storm characteristics due to horizontal resolution.
+Figure~\ref{fig:comp-mean} shows the mean precipitation rates averaged over all January storms identified by TempestExtremes.
+The iconic comma structure of the synoptic cyclones is simulated in $ne30pg3$ and $ARCTIC$ grids, with the magnitudes about the same in these two grids, with perhaps a marginal increase in precipitation rates in the storm center of the $ARCTIC$ grid.
+For good measure, the $ne30pg3^{*}$ run is also plotted, and looks more-or-less identical to the $ne30pg3$ run.
+
+As has been previously reported, horizontal resolution can have large impacts on extreme precipitation events.
+Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with synoptic storms, by month.
+The PDFs are constructed by sampling all the precipitation rates within $30^{\circ}$ of the storm center, for each point on the storm track and for all storms.
+The PDFs are evaluated on an identical composite grid for all runs, and so storm statistics are not impacted by differences in output resolution.
+The $ARCTIC$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
+This is primarily due to an increase in resolution, and not the reduced physics times-step; the $ne30pg3^{*}$ only marginally increases the precipitation rates compared with $ne30pg3$.
+
+The extreme precipitation rates in the $ARCTIC$ run are closer to the ERA5 reanalysis than in $ne30pg3$ (Figure~\ref{fig:comp-pdf}).
+It's difficult to know the extent that the extreme precipitation rates in ERA5 are constrained by data assimilation, or whether these precipitation rates are due to using a similar $\frac{1}{4}^{\circ}$ model as the $ARCTIC$ grid.
+However, it is well documented that $\frac{1}{4}^{\circ}$ models are more skillful at simulating extreme events \cite{BetAl2013JC,OETAL2016JAMES}, and so this is an additional benefit of the variable-resolution grids.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
+\end{center}
+\caption{.}
+\label{fig:comp-mean}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_composite_ge45N_pdf.pdf}
+\end{center}
+\caption{.}
+\label{fig:comp-pdf}
+\end{figure}
+
+\section{Conclusions}\label{sec:conclusions}
+
+Six grids from different dynamical cores in CESM2.2 are evaluated in an AMIP-style configuration for their performance over the Arctic and in simulating the Greenland Ice Sheet (GrIS) surface mass balance (SMB).
+The $1-2^{\circ}$ finite-volume grids have enhanced resolution over Polar regions due to the convergence of meridian lines, although a polar filter is employed to prevent features from forming at this higher resolution.
+Spectral-element grids comparable to the resolution of the finite-volume grids have an isotropic grid structure, meaning the grid resolution is similar over the entire domain.
+Two variable-resolution grids were developed and introduced into CESM2.2 as part of this work.
+They use the spectral-element dycore; the $ARCTIC$ grid has $\frac{1}{4}^{\circ}$ refinement over the broader Arctic, whereas the $ARCTICGRIS$ grid is identical except it has an $\frac{1}{8}^{\circ}$ patch of refinement over Greenland.
+
+In general, the finite-volume grids have colder summer temperatures over the Arctic, and the spectral-element grids (incl.
+the variable-resolution grids) are warmer.
+The cloud biases in all the uniform resolution grids, whether finite-volume or spectral-elements, are similar, in general being too cloudy over Arctic land masses.
+The variable-resolution grids largely improve the cloud biases.
+It should be emphasized that our analysis is specific to the Arctic summer due to its relevance to melt rates over GrIS; improved clouds in the Arctic should not be taken to mean that lower latitude regions have an improved cloud field as well.
+
+At the regional level, a halo of negative cloud bias is found around the oceanic perimeter of Greenland in all $1-2^{\circ}$ grids, and is absent in the variable-resolution grids.
+This halo bias coincides with a positive cloud bias over the interior of the ice sheet.
+This has been traced back to the inadequacy of the lower resolution grids to resolve the orographic precipitation process in Greenland.
+Synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins due to overly smooth topography in the $1-2^{\circ}$ grids.
+As as result, moisture penetrates accross the ice margin and dumps excess precipitation into the interior of the GrIS, instead of being concentrated closer to the coastal margins as indicated by observations.
+This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter of Greenland.
+The agreement of different observational data products on this bias lends confidence  in attributing the cause of the precipitation and cloud biases.
+The variable-resolution grids compare more favorably to the observations and indicate the orographic precipitation process in Greenland is largely resolved.
+
+The primary source and sink terms of the SMB equation are integrated over the GrIS, and evaluated in the six grids.
+The uniform $1-2^{\circ}$ grids all have large positive accumulation biases owing to their inability to resolve the orographic precipitation process.
+The uniform spectral-element grids have larger accumulation biases, suggesting that the finite-volume grids are more skillful at resolving the precipitation processes due to their finer grid spacing over Greenland, despite the polar filter.
+The variable resolution grids have the most accurate accumulation rates of all the grids.
+
+The primary mass sink term of the GrIS, ice/snow melt, expresses similar biases as the accumulation rates.
+The uniform resolution spectral-element grids melt too much, while the finite-volume grids have reduced biases.
+It's more difficult to attribute these biases to grid resolution alone; the finite-volume grids have colder summers that is consistent with their lower melt bias.
+However, the $ARCTICGRIS$ grid has the warmest summer temperatures of all grids, yet it has reduced melting compared to the uniform resolution spectral-element grids.
+This suggests that grid resolution is responsible for a large fraction of the melt biases.
+That the melt process would be sensitive to grid resolution is intuitive, since local elevation impacts on temperature and melting is largely handled by the land model, which does not have a pole problem because there is no horizontal dynamics.
+The larger number of grid cells in the ablation zone in the finite-volume grids seems to more accurately resolve the melt process relative to the uniform resolution spectral-element grids, in which the ablation zone is made up of substantially fewer grid cells.
+
+% \section{Materials and Methods}
+% Here is text on Materials and Methods.
+%
+% \subsection{A descriptive heading about methods}
+% More about Methods.
+%
+% \section{Data} (Or section title might be a descriptive heading about data)
+%
+% \section{Results} (Or section title might be a descriptive heading about the
+% results)
+%
 
@@
 %\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_violens.pdf}
-\end{center}
-\caption{.}
-\label{fig:pointdiff}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{cccc}
-         \includegraphics[width=25mm]{figs/temp_icemask.pdf} \\
-         \includegraphics[width=50mm]{figs/temp_zsmb_ktrans_obsperiod.pdf}&
-         \includegraphics[width=50mm]{figs/temp_zsmb_btrans_obsperiod.pdf} \\
-\end{tabular}
-\end{center}
-\caption{.}
-\label{fig:trans}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{cccc}
-         \includegraphics[width=100mm]{figs/temp_ktransect.pdf} \\
-         \includegraphics[width=100mm]{figs/temp_btransect.pdf} \\
-\end{tabular}
-\end{center}
-\caption{.}
-\label{fig:ztrans}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
-\end{center}
-\caption{.}
-\label{fig:comp-mean}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_composite_ge45N_pdf.pdf}
-\end{center}
-\caption{.}
-\label{fig:comp-pdf}
-\end{figure}
 

### 2022-03-22 arcticgrids.tex
new basin plot

@@
 
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_tseries_BASIN.pdf}
+\end{center}
+\caption{.}
+\label{fig:basin}
+\end{figure}
+
 Figure~\ref{fig:pointdiff} shows the distribution of point-wise differences from LIVVkit observational database as violin plots.
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_transect_obsperiod.pdf}
+         \includegraphics[width=130mm]{figs/temp_transect_obsperiod.pdf}
 \end{center}
@@
 \begin{tabular}{cccc}
-         \includegraphics[width=100mm]{figs/temp_ktransect.pdf} \\
-         \includegraphics[width=100mm]{figs/temp_btransect.pdf} \\
+         \includegraphics[width=130mm]{figs/temp_ktransect.pdf} \\
+         \includegraphics[width=130mm]{figs/temp_btransect.pdf} \\
 \end{tabular}

### 2022-03-24 arcticgrids.tex
new melt figure

@@
 
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO_melt.pdf}
+\end{center}
+\caption{.}
+\label{fig:melt}
+\end{figure}
+
 Figure~\ref{fig:pointdiff} shows the distribution of point-wise differences from LIVVkit observational database as violin plots.
@@
 
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_violens.pdf}
-\end{center}
-\caption{.}
-\label{fig:pointdiff}
-\end{figure}
+%\begin{figure}[t]
+%\begin{center}
+%         \includegraphics[width=130mm]{figs/temp_violens.pdf}
+%\end{center}
+%\caption{.}
+%\label{fig:pointdiff}
+%\end{figure}
 
@@
 
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_transect_obsperiod.pdf}
-\end{center}
-\caption{.}
-\label{fig:trans}
-\end{figure}
+%\begin{figure}[t]
+%\begin{center}
+%         \includegraphics[width=130mm]{figs/temp_transect_obsperiod.pdf}
+%\end{center}
+%\caption{.}
+%\label{fig:trans}
+%\end{figure}
 
@@
 
-\subsection{Impact of horizontal resolution on extremes}
+\subsection{Precipitation extremes}
 
@@
 
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
-\end{center}
-\caption{.}
-\label{fig:comp-mean}
-\end{figure}
+%\begin{figure}[t]
+%\begin{center}
+%         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
+%\end{center}
+%\caption{.}
+%\label{fig:comp-mean}
+%\end{figure}
 

### 2022-03-26 arcticgrids.tex
new elev figure
melt bias plot

@@
 With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past.
-In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.
+In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.
 
@@
 Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
-This additional damping reduces the effective resolution in polar regions, and the resolved scales are approximately the same everywhere on the grid.
+This additional damping reduces the effective resolution in polar regions such that the resolved scales are approximately the same everywhere on the grid.
 
@@
 While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions.
-For example, GrIS precipitation is typically confined to the ice-sheet margins (predominately the southeastern margin) where orographic precipitation is generated by steep topographic slopes.
+For example, GrIS precipitation is concentrated at the ice-sheet margins, where orographic precipitation is generated by steep topographic slopes.
 GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation.
-GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.
+GCMs struggle to resolve the magnitude and extent of these features \cite{P2010CC,VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.
 
@@
 The Arctic refined grids were developed by the authors, and this section serves as their official documentation in CESM2.2.
-Section~\ref{sec:methods} also contains a description of the experiments along with reanalysis datasets and post-processing software used for evaluating the model simulations.
+Section~\ref{sec:methods} also contains a description of the experiments along with the datasets and methods for evaluating the model simulations.
 Section~\ref{sec:results} contains the results of the experiments, followed by Section~\ref{sec:conclusions} that provides a general discussion and conclusions.
@@
 The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
-An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical paramerizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.
+An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical parameterizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.
 
 Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic (Figure~\ref{fig:vr-grids}).
+This manuscript serves as the official documentation of these grids.
 The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
@@
    $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
+   $ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 \\
    $ARCTIC$ & SE & 111 & 28 & 450 \\
@@
 
+The physics time-step should depend on grid resolution.
+Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}.
+The $ARCTIC$ and $ARCTICGRIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in the standard $1^{\circ}-2^{\circ}$ grids (Table~\ref{tbl:table1}).
+
+All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$.
+However, any grid or dycore can in principle be run with a higher-top and/or with finer vertical resolution.
+
 \subsection{Physical parameterizations}
 
-The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used in all simulations in this study.
-CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme.
+The CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{CAM63} is used in all simulations in this study.
+CAM version 6 physics and later is most notably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme.
 CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions.
-Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} inlcuding convective momentum transport \cite{RSG2010JAS}.
+Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
 PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.
-
-All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$.
-The physics time-step is dependent on grid resolution.
-Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}.
-The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids (Table~\ref{tbl:table1}.
 
@@
 Several of these products (ERA5, CALIPSO and CERES) are near-global gridded datasets commonly used to evaluate GCMs.
-Surface mass balance datasets are gathered from several sources.
-RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalysis at its domain boundaries.
+Surface mass balance datasets were initially gathered from several sources.
+RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalyses at its domain boundaries.
 The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.
-The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit} maintains a repository of snow pit and and ice core SMB measurements, as well as the IceBridge radar accumulation dataset.
-The LIVVkit dataset is compared against model simulations by finding the nearest grid cell center to the location of each observation.
+
+In-situ SMB (snow pit and ice cores) and radar accumulation datasets (e.g., IceBridge) are conveniently maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}.
+However, these point-wise measurements are difficult to compare to model output spanning several different grids, and would be more useful if the surface mass balance from the individual EC's were available from the model output, which they are not.
+A nearest neighbor technique was used to do an initial analysis, and indicated that the model biases are similar to those computed using the RACMO datasets.
+Owing to the uncertainty of comparing gridded fields to point-wise measurements, and the lack of additional information it adds with regard to model biases, these datasets were omitted from the analysis in this study.
 
@@
  \scriptsize
- \begin{tabular}{lccc}
+ \begin{tabular}{lcccc}
    \hline
-   data product & years used in this study & citation \\ 
+   data product & years used in this study & resolution & citation \\ 
    \hline
-   ERA5 & 1979-1998 & \cite{ERA5} \\
-   CERES-EBAF ED4.1 & 2003-2020 & \cite{CERES-EBAF} \\
-   CALIPSO-GOCCP & 2006-2017 & \cite{CALIPSO-GOCCP} \\
-   RACMO2.3 & 1979-1998 & \cite{NETAL2015TC} \\
-   RACMO2.3p2 & 1979-1998 & \cite{NETAL2019SCIENCE} \\
-   LIVVkit,v2 & 1979-1998 & \cite{LIVVkit} \\
+   ERA5 & 1979-1998 & $\frac{1}{4}^{\circ}$ & \citeA{ERA5} \\
+   CERES-EBAF ED4.1 & 2003-2020 & $1^{\circ}$ & \citeA{CERES-EBAF} \\
+   CALIPSO-GOCCP & 2006-2017 & $1^{\circ}$ & \citeA{CALIPSO-GOCCP} \\
+   RACMO2.3 & 1979-1998 & 11 km & \citeA{NETAL2015TC} \\
+   RACMO2.3p2 & 1979-1998 & 5.5 km & \citeA{NETAL2019SCIENCE} \\
  \hline
@@
 Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS.
-Figure~\ref{fig:grisdx} shows the GrIS ice mask area across the different grids, as a function of the number of grid points.
+Figure~\ref{fig:grisdx} shows the GrIS ice mask area errors (computed as percent differences from the high resolution dataset) across the different grids, and as a function of the number of grid points.
 Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset.
@@
 
-Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, but we seek to compare these integrated quantities to RACMO2.3.
+The area errors in Figure~\ref{fig:grisdx} don't seem large, but even small changes in ice mask area can lead to large differences in integrated SMB \cite{HETAL2022TC}.
 Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks.
-Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$, the lowest resolution grid for both dycores considered in this study.
-In addition, two remapping algorithms are used; ESMF conservative and TempestRemap high-order, monotone algorithm.
+Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ grids, the lowest resolution grid for each dycore considered in this study.
+In addition, two remapping algorithms are used; Earth System Modeling Framework \cite<ESMF;>{ESMF} first-order conservative and TempestRemap \cite{TempestRemap} high-order, monotone algorithm.
 In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.
@@
 \caption{The spatial properties of the GrIS as represented by different grids in this study.
-(Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks.}
+(Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
 \label{fig:grisdx}
@@
 The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.
-This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
+This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \citeA{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
 The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the low and mid-latitudes in the $f09$ grid.
@@
 \end{center}
-\caption{.}
+\caption{1979-1998 annual mean (left column) temperature and (right column) CLUBB temperature tendencies in zonal mean height space, expressed as differences between the various $1^{\circ}-1^{\circ}$ grids.
+The thick grey and magenta lines are the tropopause for control run and the test run, respectively.}
 \label{fig:dT-lores}
@@
 As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}).
-The stratosphere has a different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
+The stratosphere has a uniquely different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
 Figure~\ref{fig:dT-lores} also shows differences in temperature between $ne30pg3$ and $ne30pg2$, which are small, although there is a slight warming near the tropopause at high latitudes.
-This is consistent with the similar climates found between these grids in \cite{HETAL2019JAMES}.
+This is consistent with the similar climates found between these grids in \citeA{HETAL2019JAMES}.
 
@@
 Figure~\ref{fig:dT-hires} shows the difference in climatological summer temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
-The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.
+The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the comparison.
+
+Figure~\ref{fig:dT-hires} indicates that the $ARCTICGRIS$ grid is a lot warmer than the $ARCTIC$ grid in the Arctic.
+Part of this may be due to the shorter physics time-step; a similar physics-time-step sensitivity in the $ne30pg3^{*}$ comparison is evident in the CLUBB tendencies in the $ARCTCGRIS$ grid.
+However, the temperature response is disproportionately large relative to the change in CLUBB tendencies.
+This summer warming appears to be a result of variations in the stationary wave pattern, with anomalous southerly winds occurring to the west of Greenland (not shown).
+This dynamic response is extremely interesting because other than the physics time-step, the only difference between $ARCTICGRIS$ and the $ARCTIC$ grid is the doubling of resolution of Greenland.
+This behavior will be explored more fully in a future study.
 
@@
 \end{center}
-\caption{.}
+\caption{As in Figure~\ref{fig:dT-lores} but for the short-time step experiment and the variable-resolution grids.
+The fields plotted are for the climatological northern hemisphere summer.
+The focus on summer is because that is when the resolution response is largest, and the refined regions are located in the northern hemisphere.}
 \label{fig:dT-hires}
@@
 
-It's useful to understand summer temperature biases, instead of annual means, due to its control on ice/snow melt \cite{O2001JAM,HT2008Paleo}.
+It's useful to understand summer temperature biases due to its control on ice/snow melt over the Greenland Ice Sheet \cite<GrIS;>{O2001JAM}.
 Figure~\ref{fig:dThyps} shows the 1979-1998 lower troposphere summer temperature bias relative to ERA5.
@@
 The results generally track with the analysis of the zonal mean height plots; increasing resolution from $f19$ to $f09$ leads to a warmer climate, and the $1^{\circ}$ spectral-elements grids are warmer than the finite-volume grids.
-The summer temperatures in the finite-volume grids are persistently colder than ERA5 at high latitudes, whereas the $1^{\circ}$ spectral-element grids are warmer than ERA5 at only very high-latitudes, north of $80^{\circ}$.
-All grid illustrate a north-south gradient in bias over Greenland, in which the summer temperature bias becomes more positive in the northward direction.
+The summer temperatures in the finite-volume grids are persistently colder than ERA5, whereas the $1^{\circ}$ spectral-element grids are not as cold, and are actually warmer than ERA5 at very high-latitudes, north of $75^{\circ}$.
+All grids illustrate a north-south gradient in bias over Greenland, in which the summer temperature bias becomes more positive on the northern side of the ice sheet.
 This pattern is also evident in the 2m summer temperature bias over Greenland (not shown).
@@
 An anomalous cooling patch forms to the west of Greenland, centered over Baffin Island.
-The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic region, but approximately maintains the same pattern of summer temperature bias as in the $ARCTIC$ grid.
+The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic, but maintains a similar spatail pattern of summer temperature bias as in the $ARCTIC$ grid.
 
@@
 \end{center}
-\caption{.}
+\caption{1979-1998 lower troposphere, northern hemisphere summer virtual temperature biases, computed as the difference from ERA5.
+Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation.
+ERA5 is very high-resolution at high-latitudes owing to the use of a $\frac{1}{4}^{\circ}$ latitude-longitude grid, and so all model differences are computed after mapping ERA5 to their respective model grids.}
 \label{fig:dThyps}
@@
 \end{center}
-\caption{.}
+\caption{1979-1998 northern hemisphere summer shortwave cloud forcing bias, relative to the CERES-EBAF gridded dataset.
+Differences are computed after mapping all model output to the $1^{\circ}$ CERES-EBAF grid.}
 \label{fig:SWCF}
@@
 This suggests that the coarse grids clouds are too thick in the interior of Greenland, and too thin around the perimeter of Greenland, and that increasing horizontal resolution balances out these biases.
-This is consistent with total summer cloud fraction bias, computed from the CALIPSO cloud dataset (Figure~\ref{fig:prect}).
+This is consistent with total summer cloud fraction bias, computed from the CALIPSO-GOCCP cloud dataset (Figure~\ref{fig:prect}).
 Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates any changes arising from any single layer due to the maximum overlap assumption used to compute this quantity.
@@
 
-The agreement of the cloud biases over Greenland in multiple independent datasets indicates this is a robust feature of the coarser grids.
-The reduction of these cloud biases in the variable-resolution grids suggests they are a result of insufficient horizontal resolution in coarse grids.
+The agreement of the cloud biases in and around Greenland from multiple independent datasets indicates this is a robust feature of the coarser grids.
+The reduction of these cloud biases in the variable-resolution grids suggests they are a result of insufficient horizontal resolution in the coarse grids.
 
@@
 \end{center}
-\caption{.}
+\caption{1979-1998 northern hemisphere summer, incident shortwave radiation bias, computed as the difference (top) from CERES-EBAF, and (bottom) RACMO2.3p2 dataset.
+The differences in the top panel are found by mapping the model output to the $1^{\circ}$ CERES-EBAF grid, and differences on the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
+Note that the averaging period for the CERES-EBAF panels, 2003-2020, is different from the averaging period for the model results.}
 \label{fig:FSDS}
@@
 \end{center}
-\caption{Climatological (1979-1998) annual precipitation rate bias (in mm/yr) relative to the RACMO2.3p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
+\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias, relative to the CALIPSO-GOCCP dataset, and (bottom) precipitation rate bias, relative to the RACMO2.3p2 dataset.
+The CALIPSO-GOCCP differences are computed after mapping all model output to the $1^{\circ}$ grid, whereas the RACMO differences are computed after mapping the RACMO dataset to the individual model grids.
+Note that the averaging period for the CALIPSO-GOCCP panels, 2006-2017, is different than the model averaging period.}
 \label{fig:prect}
@@
 The uniform $1^{\circ}-2^{\circ}$ grids all show a distinctive positive bias in precipitation due to this over-prediction of interior precipitation rates.
-The  variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO.The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$, +230 Gt/yr.
+The  variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO.
+The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$, +230 Gt/yr.
 The results suggest that uniform resolution spectral-element grids have larger biases than the finite-volume grids, consistent with spectral-element grids having a coarser representation of GrIS (Figure~\ref{fig:grisdx}).
 
-The combined annual snow/ice melt integrated over the GrIS is given by the bottom panel of Figure~\ref{fig:tseries}.
+The combined annual snow/ice melt integrated over the GrIS is shown in the bottom panel of Figure~\ref{fig:tseries}.
 The $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO.
@@
 This is likely due to an anomalously warm lower troposphere during the summer, relative to the $ARCTIC$ run (Figure~\ref{fig:dThyps}).
-The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, simulating too much melt by about 70-90 Gt/yr.
+The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, overestimating melt by only 70-90 Gt/yr.
 The spectral-element grids have the largest positive melt bias, between 200-220 Gt/yr.
@@
 However, that the $ARCTCIGRIS$ grid has the warmest summer temperatures, yet has a lower melting bias than the uniform spectral-element grids, suggests that increasing resolution improves the ablation process.
+
+To get an idea of the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \cite{RM2012GRL}.
+The uncertainty due to differences in basin area is larger than for GrIS wide integrals, owing to the differences in basin boundaries as represented on the $f09$ and $ne30pg2$ grids (Figure~\ref{fig:melt}), which pertains to the grids in which all integrals are evaluated.
+Nonetheless, the southeast and southwest basins have the largest accumulation totals.
+All the basins exhibit a monotonic reduction of accumulation rates with increasing grid resolution, with some exceptions.
+The $ARCTCIGRIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins.
+Conversely, the $ARCTCIGIS$ precipitation is closest to the RACMO value in the large southwest basin.
+
+The basin integrated melt rates in Figure~\ref{fig:basin} show a different grid sensitivity than the precipitation rates.
+Of note is that the uniform-resolution spectral-element grids have the largest positive bias in melt rates, in all basins, with the $ARCTICGRIS$ grid a close second.
+This is somewhat unexpected as the $ARCTCIGRIS$ grid has the warmest lower troposphere summer temperatures (Figure~\ref{fig:dThyps}) and incident shortwave radiation (Figure~\ref{fig:dThyps}), yet it has less melting than the uniform-resolution spectral element grids.
+Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture local elevation effects as a result of a better representation of the low-elevation ablation zones.
+A more accurate proxy for the impact of temperature on melting is positive degree days \cite<PDD;>{B1984Z}.
+PDD is a metric that accumulates the near surface temperature in $C^{\circ}$ for any day in which the temperature is above freezing.
+PDD is non-linear in mean monthly near surface temperature \cite{R1991P}, and we compute it using monthly means 2-meter temperature using the method of \citeA{CG2005JG}.
+
+The basin integrated PDD melt estimate is shown in Figure~\ref{fig:basin}.
+One interesting feature is that in the big southeast and southwest basin (and all the other western basins), the $ne30pg3$ grid has larger PDD based melt than the $ARCTCGRIS$ grid.
+The southwest basin also predicts a rather large PDD based melt in the finite-volume grids, relative to the $ARCTCIGRIS$ grid.
+The PDD plots indicate that the near surface temperatures expected to contribute to melt is not well approximated by the summer lower troposphere temperatures in Figure~\ref{fig:dThyps}.
+
+Figure~\ref{fig:melt} shows the biases in the combined ice/snow melt as map plots.
+This plot is quite revealing as it indicates that the largest melt biases occur on the southeast and northwest coast, where large coarse grid cells overlap with the ocean.
+One possibility is that these problematic grid cells are situated at lower elevations then the ice sheet surface exists in nature, leading to too warm of a state and overestimating the melt.
+Figure~\ref{fig:ztrans} shows the representation of the surface of the ice sheet along two transects in the different grids, compared to the high resolution dataset used to generate CAM topography boundary conditions (GMTED).
+The two transects are shown in Figure~\ref{fig:melt}; the infamous "k-transect" in southwest Greenland and a transect extending from the central dome down to the highly productive Kangerlussuaq glacier on the southeast coast.
+The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
+The $f09$ grid is a bit of an exception; the grid cells become very narrow in the meridional direction at high latitudes, a larger number of grid cells can populate the east-west transect.
+The variable-resolution grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of ice sheets.
+
+While the two transects in Figure~\ref{fig:ztrans} only two anecdotes, they indicate that the ice sheet surface is not systematically lower than the true surface in the coarse grids, near the ice margins.
+Rather the smoothing of the raw topography, necessary to prevent coarse models from instigating grid-scale models, makes the ice sheet margin more gradual, causing the lower-elevation ablation zones to extend further away and outside the true ice sheet margin, resulting in the coarse cells in the ablation zone situated higher than the actual ice surface.
+This is another clue; the overextending of the ice sheet out over the ice margin means that these grid cells have multiple surface types within them.
+That is, It is these mixed ice/land/ocean grid cells tend to have the largest positive melt bias.
+It seems likely based on the transect that some of these largely biased coarse cells only have small fractions of ice, and so the state of the grid cell may be predominantly controlled by the land/ocean surface types.
 
@@
 
-%\begin{figure}[t]
-%\begin{center}
-%         \includegraphics[width=130mm]{figs/temp_transect_obsperiod.pdf}
-%\end{center}
-%\caption{.}
-%\label{fig:trans}
-%\end{figure}
-
 \begin{figure}[t]
@@
 \begin{tabular}{cccc}
-         \includegraphics[width=130mm]{figs/temp_ktransect.pdf} \\
-         \includegraphics[width=130mm]{figs/temp_btransect.pdf} \\
+         \includegraphics[width=130mm]{figs/temp_zsects.pdf}
 \end{tabular}
@@
 \label{fig:ztrans}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cccc}
+         \includegraphics[width=100mm]{figs/temp_xy_diffRACMO_melt.pdf}
+\end{tabular}
+\end{center}
+\caption{.}
+\label{fig:bias}
 \end{figure}
@@
 
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/Vis1923.jpg}
-\end{center}
-\caption{.}
-\label{fig:viz}
-\end{figure}
+%\begin{figure}[t]
+%\begin{center}
+%         \includegraphics[width=130mm]{figs/Vis1923.jpg}
+%\end{center}
+%\caption{.}
+%\label{fig:viz}
+%\end{figure}
 

### 2022-03-30 arcticgrids.tex
new melt bias figure
paper completed; ready for coauthor comments

@@
 % Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} $\ $and Andrew Gettelman\affil{1}}
+\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, Marcus Lofverstrom\affil{2}, Andrew Gettelman\affil{1}}
 
@@
 
-\correspondingauthor{=name=}{=email address=}
+\correspondingauthor{Adam Herrington}{aherring@ucar.edu}
 
@@
 \begin{keypoints}
-\item enter point 1 here
-\item enter point 2 here
-\item enter point 3 here
+\item The transition from latitude-longitude to unstructured grids in CESM negatively impacts the surface mass balance of the Greenland Ice Sheet.
+\item Two new Arctic refined mesh configurations have been developed and made available in CESM2.2
+\item The Arctic meshes provide improvements to the surface mass balance over all $1^{\circ}$ grids, including latitude-longitude grids.
 \end{keypoints}
@@
 \begin{abstract}
-[ enter your Abstract here ]
+
+Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS).
+The more conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both the accumulation and ablation processes over the GrIS.
+Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their greater degrees of freedom in representing the GrIS.
+Two Arctic refined-meshes, containing $\frac{1}{4}^{\circ}$ and $\frac{1}{8}^{\circ}$ refinement over Greenland, are documented in this study as newly supported configurations in CESM2.2.
+The Arctic meshes provide substantial improvements to the simulated cloud field in the Arctic and over Greenland, resulting in a skillful representation of the accumulation and ablation processes, and a more realistic GrIS surface mass balance compared with the conventional, uniform resolution grids.
+As CESM is in the process of transitioning away from latitude-longitude grids, these new Arctic refined meshes serve to recover, and improve upon the representation of polar processes in CESM.
+
 \end{abstract}
@@
 The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study.
-The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation.
+The SMB is the sum of the mass accumulation term,  i.e., precipitation, and the mass sink term, ablation.
 Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation becomes runs off the ice sheet; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface.
+Not all liquid precipitation runs-off the ice sheet; rain can penetrate pore spaces in the snowpack/firn layer and freeze, representing an accumulation of mass.
 These processes are represented by different components in CESM, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
@@
 Integrating over all ECs using the area weights provides a more accurate SMB.
-For a more detailed description of how the SMB is computed in CESM, the reader is referred to \cite{CISM1,SETAL2019TC,KETAL2020JAMES}.
+For a more detailed description of how the SMB is computed in CESM, the reader is referred to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES}.
 
@@
 
-Several observational datasets are used in this study to understand the performance of the simulations.
-A list of the datasets used in this study are shown in Table~\ref{tbl:table2}.
-Several of these products (ERA5, CALIPSO and CERES) are near-global gridded datasets commonly used to evaluate GCMs.
+Several observational datasets are used in this study to gauge the performance of the simulations.
+A list of the datasets used in this study are provided in Table~\ref{tbl:table2}.
+Several of these products are near-global gridded datasets commonly used to evaluate GCMs.
 Surface mass balance datasets were initially gathered from several sources.
-RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalyses at its domain boundaries.
-The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.
+RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA40, ERA interim and ERA5 renalysis products at its domain boundaries.
+The RACMO simulations have been shown to perform skillfully against observations and is often used as an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.
 
@@
 
-\subsection{SMB Analysis}
-
-A common high resolution dataset is used to generate the GrIS boundary conditions in all grids, using the CLM dataset creation tools.
-Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS.
-Figure~\ref{fig:grisdx} shows the GrIS ice mask area errors (computed as percent differences from the high resolution dataset) across the different grids, and as a function of the number of grid points.
-Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset.
-RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset.
-
-The area errors in Figure~\ref{fig:grisdx} don't seem large, but even small changes in ice mask area can lead to large differences in integrated SMB \cite{HETAL2022TC}.
-Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks.
-Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ grids, the lowest resolution grid for each dycore considered in this study.
-In addition, two remapping algorithms are used; Earth System Modeling Framework \cite<ESMF;>{ESMF} first-order conservative and TempestRemap \cite{TempestRemap} high-order, monotone algorithm.
-In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_grisres.pdf}
+\subsection{SMB Analysis}\label{sec:SMB}
+
+We seek to integrate components of the SMB over a GrIS ice mask and diagnose their contributions to the mass budget of the GrIS.
+However, the ice masks vary across all the grids, especially in comparison to the RACMO3.2 ice mask which is about $3\%$ smaller than the reference dataset (Figure~\ref{fig:grisdx}).
+The area errors in Figure~\ref{fig:grisdx} may not seem large, but even order $1-2\%$ percent ice mask area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
+CLM's dataset creation tools generates the model ice mask by mapping a raw high resolution dataset containing the GrIS boundary conditions, to the target grid using the Earth System Modeling Framework \cite<ESMF;>{ESMF} first-order conservative remapping algorithm.
+The figure suggests the mapping errors are less than $1.5\%$ across just the model grids.
+
+In this study we have taken a common ice mask approach by mapping all model fields to the lowest resolution grids and integrating over their respective low resolution ice masks.
+A low resolution common ice mask is a conservative decision and is largely because we seek to use first-order remapping algorithms for mapping fields to the common ice mask, which is not reliable in the downscaling direction, i.e., mapping to a higher resolution grid than the source grid.
+Two remapping algorithms are used; ESMF first-order conservative and TempestRemap \cite{TempestRemap} high-order monotone algorithm.
+Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ grids, the lowest resolution grid for each grid coordinate type.
+In all, each integrated quantity is evaluated (at most) four times per grid to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.
+
+We consider two different approaches for integrating mapped fields over the common ice mask.
+\begin{enumerate}
+\item Map the patch level quantities, or the state over the ice fractional component of the grid cell, to the common grid, and integrate the mapped field over the common ice mask. \label{lbl:label1}
+\item Map the grid cell mean quantities to the common grid, and integrate the mapped field over the common ice sheet. \label{lbl:label2}
+\end{enumerate}
+
+Method~\ref{lbl:label1} introduces errors that is related to differences in the ice mask between the source and target grid.
+Since the lower resolution ice masks tend to have larger GrIS areas (Figure~\ref{fig:grisdx}), we expect these errors to inflate the resulting integrals.
+Method~\ref{lbl:label2} assumes the patch level values are identical to the grid mean values.
+This has the effect of homogenizing ice mask specific states over the entire grid cell.
+This is especially detrimental to the ice melt quantity, since ice melt can only occur over the ice mask.
+The accumulation term (combined solid plus liquid precipitation), however, should not be adversely impacted by method~\ref{lbl:label2} since it is not directly tied to the ice mask, but instead controlled by larger resolved scales.
+One could also argue that method~\ref{lbl:label1} is less accurate for the accumulation term, since it does not map to the new ice mask as cleanly as in method~\ref{lbl:label2}.
+
+Method~\ref{lbl:label2} may be necessary in situations where only the grid mean values are available, and the patch level values cannot be reconstructed.
+This is unfortunately the case for the RACMO datasets; ice melt cannot be reconstructed at the ice patch level.
+Method~\ref{lbl:label2} is therefore used in this study, and we should expect a dissipative effect when diagnosing the melt rates.
+As we will show later, this method still preserves the relative differences between the different grids and in comparison with the RACMO datasets.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
 \end{center}
@@
 Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation.
-ERA5 is very high-resolution at high-latitudes owing to the use of a $\frac{1}{4}^{\circ}$ latitude-longitude grid, and so all model differences are computed after mapping ERA5 to their respective model grids.}
+Differences are computed after mapping the ERA5 data to the finite-volume grids since the geopotential field is only available on the output tapes in the spectral-element runs that have been interpolated to the $f09$ grid, inline.}
 \label{fig:dThyps}
@@
 The CALIPSO-GOCCP differences are computed after mapping all model output to the $1^{\circ}$ grid, whereas the RACMO differences are computed after mapping the RACMO dataset to the individual model grids.
-Note that the averaging period for the CALIPSO-GOCCP panels, 2006-2017, is different than the model averaging period.}
+Note that the averaging period for the CALIPSO-GOCCP panels, 2006-2017, is different than the model averaging period.
+\color{red}{ARH - still trying to fix the layout of this figure so the label bars aren't on top of the panels}.}
 \label{fig:prect}
@@
 
-The summer climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:grisdx}, expressed as the fractional difference from the RACMO2.3p2 solution.
-What sticks out is that the coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
-The $ARCTIC$ run improves this bias substantially, and the $ARCTICGIS$ run improves this bias further.
+The summer climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:prect}, expressed as the fractional difference from the RACMO2.3p2 solution.
+It is evident that the coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
+The $ARCTIC$ run improves this bias substantially, and the $ARCTICGRIS$ run improves this bias further.
 This suggests the southern dome bias is due to inadequate horizontal resolution, which is consistent with the original GrIS variable-resolution experiments in \cite{VETAL2018TC}.
 
-Southeast Greenland has the largest accumulation rates in GrIS due to synoptic systems moving in from the southeast.
-These systems are orographically lifted by the steep southeast ice sheet margin, dumping large amounts of precipitation along the southeast coast.
-At lower horizontal resolutions, the topography is too smooth and large amounts of moisture penetrates further inland, incorrectly dumping precipitation onto the interior of the ice sheet.
-A similar bias occurs in northwest Greenland, in particular during the summer, when it's common for synoptic systems to arrive from the southwest.
+Large accumulation rates in GrIS result from synoptic systems arriving from the south.
+These systems are orographiclly lifted at the ice sheet margin, especially the steep slopes in southeast Greenland, concentrating large amounts of precipitation about the ice margin.
+At lower horizontal resolutions the topography is too smooth and anomalous moisture penetrates inland, erroneously dumping precipitation onto the southern dome.
 The ability of the variable-resolution grids to more accurately simulate the orographic precipitation process in Greenland is consistent with all the cloud results up to this point.
-Since the precipitation centers move from the interior towards the coasts, and even out over the ocean with increasing resolution, the cloud decks should, and are, moved accordingly, eliminating this halo of low cloud bias around the oceanic perimeter of Greenland.
+As the precipitation centers move from the interior towards the margins, and even out over the ocean with increasing resolution, the cloud decks should, and are, moved accordingly.
+Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, \ref{fig:prect} clearly illustrate a reduction in cloud amount in the interior and an increased cloudiness around the oceanic perimeter of Greenland, in comparing the uniform $1^{\circ}-2^{\circ}$ runs to the variable-resolution runs.
+
+ \begin{table*}
+ \centering
+ \scriptsize
+ \begin{tabular}{lccccc}
+   \hline
+   grid name & accumulation & total melt & runoff & sublimation & SMB \\ 
+   \hline
+   $RACMO$ & 768.5 (733.5) & -347.2 (-436.4) & -221.7 (-258.5) & -36.5 (-38.8) & 510.3 (436.2) \\
+   \hline
+   $f19$ & 882.5 (913.5) & -440.3 (-546.5) & -283.7 (-284.3) & -36.6 (-37.5) & 562.2 (591.7) \\
+   $f09$ & 874.8 (882.1) & -418.4 (-482.3) & -255.0 (-212.3) & -38.1 (-37.4) & 581.7 (632.4) \\
+   $ne30pg2$ & 1000. (973.4) & -549.4 (-647.3) & -383.9 (-347.0) & -33.4 (-32.1) & 582.7 (594.3) \\
+   $ne30pg3$ & 934.9 (909.3) & -568.8 (-686.7) & -356.2 (-330.1) & -34.4 (-32.6) & 544.3 (546.6) \\
+   $ARCTIC$ & 795.9 (818.6) & -367.3 (-436.8) & -208.9 (-194.2) & -44.1 (-43.9) & 542.9 (580.5) \\
+   $ARCTICGRIS$ & 708.7 (747.3) & -471.6 (-610.4) & -261.1 (-307.8) & -50.7 (-51.8) & 396.9 (387.7) \\
+ \hline
+ \end{tabular}
+  \caption{1979-1998 surface mass balance of the Greenland Ice Sheet in Gt/yr.
+Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.}
+ \label{tbl:table3}
+ \end{table*}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
+\end{center}
+\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to the RACMO datasets.
+The time-series were generated using the common ice mask approach, which results in up to 4 ensembles, with the mean value given by the solid line and shading spanning the extent of the ensemble members.}
+\label{fig:tseries}
+\end{figure}
+
+Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components in all the grids, and compared with RACMO.
+Note the values in the table averages over the ensemble of common ice masks and regridding methods described in section~\ref{sec:SMB}.
+The RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}) using the same common ice mask approach.
+The SMB is computed from the individual components as:
+\begin{equation}
+SMB = accumulation + runoff + sublimation \label{eq:SMB}
+\end{equation}
+and all terms have consistent sign conventions (positive values contribute mass, negative values reduce mass).
+Note that this glaciological SMB differs from the internally computed SMB in CLM; see \citeA{KETAL2020JAMES} for an explanation of the differences.
+
+Table~\ref{tbl:table3} also contain's (in parentheses) the SMB components derived from evaluating the integrals on each model's native grid and ice mask.
+Of note is the large reduction in ice/melt rates using the common ice mask approach compared with that computed on the native grid.
+This illustrates the dissipative impact of the common ice mask approach discussed in section~\ref{sec:SMB}.
+The errors are large because it is mixed ice/ice-free grid cells straddling the ice sheet margins that are more prone to errors, which are commonly in the ablation zone and have large melt rates.
+Contrast this with the differences in integrated precipitation, where the differences between the native approach and common ice mask approach is much less.
+As discussed in the methods, this is because the combined solid/liquid precipitation rates are not directly tied to the ice mask, whereas ice melt is exclusively tied to the ice mask.
 
 Figure~\ref{fig:tseries} shows time-series of annually integrated precipitation and snow/ice melt over the GrIS in the various different grids and dycores, with both versions of RACMO shown in black.
-The 1979-1998 climatological mean values are provided as circles on the right side of the panels.
+The 1979-1998 climatological mean values, tabulated in Table~\ref{tbl:table3}, are also provided as circles on the right side of the panels.
 The uniform $1^{\circ}-2^{\circ}$ grids all show a distinctive positive bias in precipitation due to this over-prediction of interior precipitation rates.
@@
 
-The combined annual snow/ice melt integrated over the GrIS is shown in the bottom panel of Figure~\ref{fig:tseries}.
-The $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO.
+The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO.
 The $ARCTICGRIS$ grid over predicts melting by about 125 Gt/yr.
@@
 The spectral-element grids have the largest positive melt bias, between 200-220 Gt/yr.
-It is more difficult to attribute these differences to resolution alone, since the finite-volumes grids have colder summer temperatures than the uniform resolution spectral-element grids.
-However, that the $ARCTCIGRIS$ grid has the warmest summer temperatures, yet has a lower melting bias than the uniform spectral-element grids, suggests that increasing resolution improves the ablation process.
-
-To get an idea of the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \cite{RM2012GRL}.
-The uncertainty due to differences in basin area is larger than for GrIS wide integrals, owing to the differences in basin boundaries as represented on the $f09$ and $ne30pg2$ grids (Figure~\ref{fig:melt}), which pertains to the grids in which all integrals are evaluated.
-Nonetheless, the southeast and southwest basins have the largest accumulation totals.
-All the basins exhibit a monotonic reduction of accumulation rates with increasing grid resolution, with some exceptions.
-The $ARCTCIGRIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins.
-Conversely, the $ARCTCIGIS$ precipitation is closest to the RACMO value in the large southwest basin.
-
-The basin integrated melt rates in Figure~\ref{fig:basin} show a different grid sensitivity than the precipitation rates.
-Of note is that the uniform-resolution spectral-element grids have the largest positive bias in melt rates, in all basins, with the $ARCTICGRIS$ grid a close second.
-This is somewhat unexpected as the $ARCTCIGRIS$ grid has the warmest lower troposphere summer temperatures (Figure~\ref{fig:dThyps}) and incident shortwave radiation (Figure~\ref{fig:dThyps}), yet it has less melting than the uniform-resolution spectral element grids.
-Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture local elevation effects as a result of a better representation of the low-elevation ablation zones.
-A more accurate proxy for the impact of temperature on melting is positive degree days \cite<PDD;>{B1984Z}.
-PDD is a metric that accumulates the near surface temperature in $C^{\circ}$ for any day in which the temperature is above freezing.
-PDD is non-linear in mean monthly near surface temperature \cite{R1991P}, and we compute it using monthly means 2-meter temperature using the method of \citeA{CG2005JG}.
+It is more difficult to attribute these differences to resolution alone, since the finite-volume grids have colder summer temperatures than the uniform resolution spectral-element grids.
+
+To understand the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \citeA{RM2012GRL}.
+The uncertainty due to differences in basin area is larger than for GrIS wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which can be seen in the $f19$ and $ne30pg2$ panels of Figure~\ref{fig:melt}.
+Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the largest accumulation totals.
+All basins exhibit a monotonic reduction of accumulation rates with increasing grid resolution, with some exceptions.
+The $ARCTCIGRIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to the RACMO precipitation amounts in the large southwest basin.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_tseries_BASIN.pdf}
+\end{center}
+\caption{1979-1998 basin integrated components of the SMB; (top) precipitation, (middle) ice/snow melt and (bottom) ice/snow melt estimated from the PDD method.
+Whiskers span the max/min of the 4 ensemble members generated from the common ice mask approach.
+Basin definitions are after \citeA{RM2012GRL}, and are found on the common ice masks using a nearest neighbor approach, and shown in Figure~\ref{fig:melt}.}
+\label{fig:basin}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO_melt.pdf}
+\end{center}
+\caption{1979-1998 ice/snow melt biases (in mm/yr) relative to RACMO2.3p2, evaluated on the native model grids.
+The \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid.
+Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $f19$ and $ne30pg2$ panels, in computing the basin-scale integrals.
+Blue lines in the $f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
+\label{fig:melt}
+\end{figure}
+
+The basin integrated melt rates in Figure~\ref{fig:basin} show a dycore dependent-type pattern.
+The uniform-resolution spectral-element grids have the largest positive bias in melt rates, in all basins.
+The $ARCTICGRIS$ grid is a close second, while the finite-volume grids have systematically smaller melt-rates.
+The ``second-place" standing of $ARCTICGRIS$ is somewhat unexpected as this grid has the warmest lower troposphere summer temperatures (Figure~\ref{fig:dThyps}) and incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the uniform-resolution spectral element grids.
+
+Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones.
+A more accurate proxy for the impact of temperature on melting is positive degree-days \cite<PDD;>{B1984Z}.
+PDD is a metric that accumulates the near surface temperature in $^{\circ}C$ for any day in which the temperature is above freezing.
+PDD is non-linear in mean monthly temperature \cite{R1991P}, and we compute it using monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}C$, and a degree-day factor of 5 $mm \ d^{-1}$ $^{\circ}C^{-1}$.
 
 The basin integrated PDD melt estimate is shown in Figure~\ref{fig:basin}.
-One interesting feature is that in the big southeast and southwest basin (and all the other western basins), the $ne30pg3$ grid has larger PDD based melt than the $ARCTCGRIS$ grid.
+One interesting feature is that in the big southeast and southwest basins (and all the other western basins), the $ne30pg3$ grid has larger PDD based melt than the $ARCTCGRIS$ grid.
 The southwest basin also predicts a rather large PDD based melt in the finite-volume grids, relative to the $ARCTCIGRIS$ grid.
@@
 One possibility is that these problematic grid cells are situated at lower elevations then the ice sheet surface exists in nature, leading to too warm of a state and overestimating the melt.
-Figure~\ref{fig:ztrans} shows the representation of the surface of the ice sheet along two transects in the different grids, compared to the high resolution dataset used to generate CAM topography boundary conditions (GMTED).
+Figure~\ref{fig:ztrans} shows the representation of the surface of the ice sheet along two transects in the different grids, compared to the high resolution dataset used to generate CAM topography boundary conditions \cite{GMTED2010,gmdd-8-4623-2015}.
 The two transects are shown in Figure~\ref{fig:melt}; the infamous "k-transect" in southwest Greenland and a transect extending from the central dome down to the highly productive Kangerlussuaq glacier on the southeast coast.
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $f09$ grid is a bit of an exception; the grid cells become very narrow in the meridional direction at high latitudes, a larger number of grid cells can populate the east-west transect.
-The variable-resolution grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of ice sheets.
-
-While the two transects in Figure~\ref{fig:ztrans} only two anecdotes, they indicate that the ice sheet surface is not systematically lower than the true surface in the coarse grids, near the ice margins.
-Rather the smoothing of the raw topography, necessary to prevent coarse models from instigating grid-scale models, makes the ice sheet margin more gradual, causing the lower-elevation ablation zones to extend further away and outside the true ice sheet margin, resulting in the coarse cells in the ablation zone situated higher than the actual ice surface.
-This is another clue; the overextending of the ice sheet out over the ice margin means that these grid cells have multiple surface types within them.
-That is, It is these mixed ice/land/ocean grid cells tend to have the largest positive melt bias.
-It seems likely based on the transect that some of these largely biased coarse cells only have small fractions of ice, and so the state of the grid cell may be predominantly controlled by the land/ocean surface types.
-
- \begin{table*}
- \centering
- \scriptsize
- \begin{tabular}{lccc}
-   \hline
-   grid name & accumulation (Gt/yr) & ice/snow melt (Gt/yr) \\ 
-   \hline
-   $RACMO$ & 768.5 & -347.2 \\
-   \hline
-   $f19$ & 882.5 & -440.3 \\
-   $f09$ & 874.8 & -418.4 \\
-   $ne30pg2$ & 1000. & -549.4 \\
-   $ne30pg3$ & 934.9 & -568.8 \\
-   $ARCTIC$ & 795.9 & -367.3 \\
-   $ARCTICGRIS$ & 708.7 & -471.6 \\
- \hline
- \end{tabular}
-  \caption{1979-1998 Surface Mass Balance of the Greenland Ice Sheet.}
- \label{tbl:table3}
- \end{table*}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
-\end{center}
-\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to RACMO3.2.
-The raw fields are mapped to two target low resolution grids, f19 \& ne30pg2, and using two different remapping methods, conservative ESMF and high order TempestRemap.
-The remapped values are then integrated over the ice mask of the target grid.
-This gives four time-series for each simulation (three for f19 \& ne30pg2), with the mean value given by the solid line and shading spanning the extent of the remapped solutions.}
-\label{fig:tseries}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=100mm]{figs/temp_tseries_BASIN.pdf}
-\end{center}
-\caption{.}
-\label{fig:basin}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO_melt.pdf}
-\end{center}
-\caption{.}
-\label{fig:melt}
-\end{figure}
-
-Figure~\ref{fig:pointdiff} shows the distribution of point-wise differences from LIVVkit observational database as violin plots.
-The IceBridge dataset is exclusively from the interior of the ice sheet and represents accumulation rates.
-The uniform $1^{\circ}-2^{\circ}$ grids have similar median errors of about +35-50 mm.w.e, while the variable-resolution errors are noticeably smaller.
-The in-situ observations in the accumulation zone, shown in the middle plot looks very similar to the IceBridge errors, providing confidence that the variable-resolution grids are outperforming the uniform grids in the interior accumulation zone.
-The errors evaluated at in-situ ablation zone measurements are in tension with the RACMO results in Figure~\ref{fig:tseries}.
-They indicate that the uniform $1^{\circ}-2^{\circ}$ grids perform similarly, and that the $ARCTICGRIS$ performs best, and an improvement over the $ARCTIC$ grid.
-However, the in-situ ablation measurements are sparse in time and space, and so there is a large amount of uncertainty in extending these results to the entirety of the ablation zone.
-
-%\begin{figure}[t]
-%\begin{center}
-%         \includegraphics[width=130mm]{figs/temp_violens.pdf}
-%\end{center}
-%\caption{.}
-%\label{fig:pointdiff}
-%\end{figure}
-
-The ``k-transect" is perhaps the most well studied transect in Greenland, as it has been regularly monitored since the 1950s.
-The LIVVkit has compiled the k-transect observations, shown in Figure~\ref{fig:trans} as elevation vs.
-SMB along the transect, with the model's simulated transects shown as well.
-The in-situ observation points are nicely replicated by the $ARCTCIGRIS$ run, whereas the $ARCTIC$ grid is biased positive in the higher-elevations of the ablation zone.
-The $f09$ grid is surprisingly competitive with the variable-resolution grids, capturing a realistic slope of the elevation-SMB curve, although it exhibits a similar positive bias as the $ARCTIC$ grid.
-The elevation-SMB slopes of the uniform spectral-element and $f19$ grid are too shallow, in particular at the higher elevation regions of the transect.
-
-Figure~\ref{fig:ztrans} shows the representation of the surface of the ice sheet along these transects in the different grids, compared to the high resolution dataset used to generate CAM topography boundary conditions.
-The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $f09$ grid is a bit of an exception --since the grid cells become very narrow in the meridional direction at high latitudes, a larger number of grid cells can populate the east-west transect, consistent with its skillful representation of the ablation zone.
-
-What the authors refer to as the ``b-transect" in northwest Greenland (Figure~\ref{fig:trans}) is characterized by orographic precipitation, resulting in the accumulation zone extending down to the ice margin.
-The variable resolution grids perform relatively well, with larger SMBs at lower elevations where the precipitation rates are highest, whereas the $1^{\circ}-2^{\circ}$ grids underestimate the SMB in these lower elevation regions.
-Only the variable resolution grids can capture the local reduction in SMB in the 1500-2000 m region.
-The skill of the variable-resolution grids is clearly related to the more accurate representation of the steepness of the transect, while also capturing the protrusion around 1500-2000 m that coincides with the local minimum in SMB (Figure~\ref{fig:ztrans}).
+The $f09$ grid is a bit of an exception; the grid cells become very narrow in the meridional direction at high latitudes and a larger number of grid cells populate the east-west transect.
+The variable-resolution grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of the GrIS margins.
 
@@
 \end{center}
-\caption{.}
+\caption{Model surface elevation along the (top) k-transect, and (bottom) a transect spanning the central dome, down to the Kangerlussuaq glacier in southeast Greenland, for all model grids.
+The reference surface (GMTED) is a 1 km surface elevation dataset used for generating the CAM topography boundary conditions.
+}
 \label{fig:ztrans}
@@
 
+While the transects in Figure~\ref{fig:ztrans} only provide two anecdotes, they indicate that the ice sheet surface is not systematically lower than the true surface in the coarse grids, in the ablation zones.
+Rather the smoothing of the raw topography, necessary to prevent the model from instigating grid-scale modes, flattens the ice sheet, causing the lower-elevation ablation zones to extend further outside the true ice sheet margin, and are situated higher than the actual ice surface.
+The $f19$ grid has the smoothest topography since it's dynamics are coarsest (whereas $f09$, $ne30pg2$ and $ne30pg3$ use identical smoothing), and has the flattest ice sheet.
+This suggests that if anything, coarser models will tend to elevate the ablation zones and depress melt rates.
+
+Figure~\ref{fig:ztrans} also labels the ice margin boundary, which nicely illustrates that the ablation zone occurs over very narrow horizontal distances where the ice sheet rapidly plunges down to sea-level.
+Due to this abrupt transition, coarse grids will commonly represent the ablation zone with mixed grid cells containing mixtures of the ice and ice-free regions.
+We hypothesize that the reason that the coarser models tend to have large melt biases, despite the $ARCTCIGRIS$ having the warmest troposphere, is because summer melting is confined to these mixed ice/land/ocean grid cells.
+CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only see a homogenized state due to volume averaging over the sub-grid mixture.
+The implication is that the ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice patch of a grid cell.
+
+Figure~\ref{fig:bias} shows mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region.
+Note the common ice mask approach is used to compute the errors, meaning all fields are mapped to the common ice masks and the grid cell ice fraction is defined by the common ice masks.
+The figure shows that coarser grids generally have two peaks in ice fraction space; a bump in positive melting errors on the 0-20$\%$ range, and another in the fully ice-covered ice cells.
+Also shown are the $\pm$ one standard deviation of the biases for each bin.
+They indicate that the biases in 0-20$\%$ bins are mostly contained in the positive bias region (a fractional bias greater than 0), whereas the fully-covered ice cells have a wider distribution, with many grid cells also containing negative melting biases.
+The excessive melting that occurs in 0-20$\%$ ice fraction bins supports our hypothesis that the prevalence of mixed-grid cells in representing the ablation zone in coarse grids is responsible for their large melt bias.
+
 \begin{figure}[t]
@@
 \begin{tabular}{cccc}
-         \includegraphics[width=100mm]{figs/temp_xy_diffRACMO_melt.pdf}
+         \includegraphics[width=80mm]{figs/temp_xy_diffRACMO_melt.pdf}
 \end{tabular}
 \end{center}
-\caption{.}
+\caption{Fractional melt bias over the GrIS, computed relative to the RACMO datasets using the common ice mask approach, and conditionally sampled by grid cell ice fraction provided by the common ice masks.
+Solid lines are the mean of the distribution with $\pm$ one standard deviation expressed by shading.}
 \label{fig:bias}
@@
 
-To give an idea of what the ablation zones look like in the highest resolution $ARCTCIGRIS$ grid, Figure~\ref{fig:viz} is a still from a visualization produced by the authors.
-It shows the daily mean surface mass balance during the height of the melt season in year 1981 of the simulation.
-The ablation zone appears to be well resolved in the $ARCTICGRIS$ grid.
-The full visualization is publicly available on youtube.com\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}.
-
-%\begin{figure}[t]
-%\begin{center}
-%         \includegraphics[width=130mm]{figs/Vis1923.jpg}
-%\end{center}
-%\caption{.}
-%\label{fig:viz}
-%\end{figure}
-
 \subsection{Precipitation extremes}
 
-Synoptic storms are tracked and analyzed using TempestExtremes \cite{UETAL2021}.
+Synoptic storms are tracked using TempestExtremes \cite{UETAL2021}.
 As the $ARCTIC$ grid contains $\frac{1}{4}^{\circ}$ refinement north of about $45^{\circ}$ latitude, the storm tracker is applied to this region for the $ARCTIC$ and $ne30pg3$ run to identify differences in storm characteristics due to horizontal resolution.
-Figure~\ref{fig:comp-mean} shows the mean precipitation rates averaged over all January storms identified by TempestExtremes.
-The iconic comma structure of the synoptic cyclones is simulated in $ne30pg3$ and $ARCTIC$ grids, with the magnitudes about the same in these two grids, with perhaps a marginal increase in precipitation rates in the storm center of the $ARCTIC$ grid.
-For good measure, the $ne30pg3^{*}$ run is also plotted, and looks more-or-less identical to the $ne30pg3$ run.
-
-As has been previously reported, horizontal resolution can have large impacts on extreme precipitation events.
-Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with synoptic storms, by month.
+The composite mean precipitation maps are similar between the two grids, and exhibit the iconic comma structure of synoptic cyclones (not shown).
+
+Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with the storms, by month.
 The PDFs are constructed by sampling all the precipitation rates within $30^{\circ}$ of the storm center, for each point on the storm track and for all storms.
@@
 The $ARCTIC$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
-This is primarily due to an increase in resolution, and not the reduced physics times-step; the $ne30pg3^{*}$ only marginally increases the precipitation rates compared with $ne30pg3$.
+This is primarily due to an increase in resolution, and not the reduced physics times-step; the $ne30pg3^{*}$ run only marginally increases the extreme precipitation rates compared with $ne30pg3$ (Figure~\ref{fig:comp-pdf}).
 
@@
 
-%\begin{figure}[t]
-%\begin{center}
-%         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
-%\end{center}
-%\caption{.}
-%\label{fig:comp-mean}
-%\end{figure}
-
 \begin{figure}[t]
@@
 \end{center}
-\caption{.}
+\caption{PDFs of the total precipitation rate associated with tracked storms, by month, in the $ne30pg3$, $ne30pg3^{*}$ and $ARCTIC$ runs, and compared with the ERA5 dataset.}
 \label{fig:comp-pdf}
@@
 Six grids from different dynamical cores in CESM2.2 are evaluated in an AMIP-style configuration for their performance over the Arctic and in simulating the Greenland Ice Sheet (GrIS) surface mass balance (SMB).
-The $1-2^{\circ}$ finite-volume grids have enhanced resolution over Polar regions due to the convergence of meridian lines, although a polar filter is employed to prevent features from forming at this higher resolution.
+The $1-2^{\circ}$ finite-volume grids have enhanced resolution over polar regions due to the convergence of meridian lines, although a polar filter is employed to prevent atmospheric features from forming at this higher resolution.
 Spectral-element grids comparable to the resolution of the finite-volume grids have an isotropic grid structure, meaning the grid resolution is similar over the entire domain.
@@
 
-In general, the finite-volume grids have colder summer temperatures over the Arctic, and the spectral-element grids (incl.
-the variable-resolution grids) are warmer.
+In general, the finite-volume grids have colder summer temperatures over the Arctic compared with the spectral-element grids (incl.
+the variable-resolution grids).
 The cloud biases in all the uniform resolution grids, whether finite-volume or spectral-elements, are similar, in general being too cloudy over Arctic land masses.
@@
 This halo bias coincides with a positive cloud bias over the interior of the ice sheet.
-This has been traced back to the inadequacy of the lower resolution grids to resolve the orographic precipitation process in Greenland.
+This pattern has been traced back to the inadequacy of the lower resolution grids to resolve the orographic precipitation process in Greenland.
 Synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins due to overly smooth topography in the $1-2^{\circ}$ grids.
-As as result, moisture penetrates accross the ice margin and dumps excess precipitation into the interior of the GrIS, instead of being concentrated closer to the coastal margins as indicated by observations.
+As as result, moisture penetrates and dumps excess precipitation into the interior of the GrIS, instead of being concentrated closer to the coastal margins as indicated by observations.
 This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter of Greenland.
@@
 The uniform $1-2^{\circ}$ grids all have large positive accumulation biases owing to their inability to resolve the orographic precipitation process.
-The uniform spectral-element grids have larger accumulation biases, suggesting that the finite-volume grids are more skillful at resolving the precipitation processes due to their finer grid spacing over Greenland, despite the polar filter.
+The uniform spectral-element grids have larger accumulation biases, suggesting that the finite-volume grids are more skillful at resolving the precipitation processes due to their finer grid spacing over Greenland, despite a polar filter.
 The variable resolution grids have the most accurate accumulation rates of all the grids.
@@
 This suggests that grid resolution is responsible for a large fraction of the melt biases.
-That the melt process would be sensitive to grid resolution is intuitive, since local elevation impacts on temperature and melting is largely handled by the land model, which does not have a pole problem because there is no horizontal dynamics.
-The larger number of grid cells in the ablation zone in the finite-volume grids seems to more accurately resolve the melt process relative to the uniform resolution spectral-element grids, in which the ablation zone is made up of substantially fewer grid cells.
+A mechanism is proposed; that coarse grids represent the ablation zones using mixed surface-type grid cells, and the ice-free patches of the cells may largely determining the volume mean state and lead to a warm bias over the ice covered patches of the grid cell.
+This mechanism is supported by analysis of melt biases binned by grid cell ice fraction.
+
+The $ARCTIC$ grid provides substantial improvements to the simulated climate of the Arctic, such as precipitation extremes and the surface mass balance over Greenland, compared with the uniform $1^{\circ}-2^{\circ}$ grids.
+The $ARCTICGRIS$ grid has the most realistic cloud and precipitation fields, but suffers from excessive summer temperatures.
+The $1^{\circ}$ finite-volume model provides a surprisingly realistic surface mass balance, likely due to the larger degrees of freedom used to represent the GrIS in latitude-longitude grids.
+In particular, the large number of grid cells representing the ablation zone leads to a reduction in mixed ice-covered/ice-free grid cells that seem to be so detrimental to the ablation process in the other uniform resolution grids.
+
+As modeling systems move away from latitude-longitude grids, and towards quasi-uniform unstructured grids, it is worth taking stock as to whether this will adversely impact the skill of the simulated climate in polar regions.
+In our case, we have found that the $1^{\circ}$ finite-volume model has clear advantages over the $1^{\circ}$ spectral element model in simulating the surface mass balance of the GrIS.
+While this finding is not sufficient to slow-down the already ongoing transition towards unstructured grids in CESM, largely driven by gains in computational efficiency, it has inspired us to develop alternative configurations that recover the fidelity of polar climate in latitude-longitude grids.
+We have shown in this study (and a prior companion study \cite{VETAL2018TC}), that for CESM, the inclusion of Arctic refined meshes can substantially improve the simulated mass balances of the GrIS, even compared with the finite-volume $1^{\circ}$ grid.
+This should provide the CESM modeling community with some comfort; the ongoing transition away from latitude-longitude grids will not adversely impact CESM's reputation as a state-of-the-art tool for simulating and understanding polar processes.
+
+We are actively working to develop a configuration of the $ARCTIC$ mesh that is fully-coupled with the CESM ocean model, and the Community Ice Sheet Model (CISM) to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise.
+A fully-coupled pre-industrial control configuration has already been developed and vetted, and should be incorporated into CESM in the near-future.
+A visualization of the $ARCTCGRIS$ has been developed and made available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}} in order to advertise some of the stunning capabilities of the refined mesh approach.
+A snapshot of this visualization is provided in Figure~\ref{fig:viz}, illustrating mesoscale katabatic winds descending down the southeastern slopes of GrIS.
+We are optimistic about the polar science opportunities provided by these new Arctic mesh configurations, and our open to collaborating with the CESM community to assist in anyway we can.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/Vis1472.jpg}
+\end{center}
+\caption{Snapshot of the lowest model level streamlines from the $ARCTICGRIS$ visualization, with color shading denoting the wind magnitude.}
+\label{fig:viz}
+\end{figure}
 
@@
 This material is based upon work supported by the National Center for Atmospheric Research (NCAR), which is a major facility sponsored by the NSF under Cooperative Agreement 1852977.
-Computing and data storage resources, including the Cheyenne supercomputer
-(doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at NCAR.
+Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at NCAR.
+A.
+Herrington is especially grateful to W.
+Lipscomb (NCAR) for providing crucial feedback in his ASP postdoc application, that ultimately funded this study.
+A.
+Herrington also thanks Matt Rehme (NCAR/CISL) for his role in generating the $ARCTICGRIS$ visualization available on youtube ({\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}).
 

### 2022-04-05 arcticgrids.tex
bring big overleaf edits into github

--- 2022-03-24 arcticgrids.tex
@@
 % Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} $\ $and Andrew Gettelman\affil{1}}
+\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} \ and Andrew Gettelman\affil{1} }
 
@@
 \begin{keypoints}
-\item enter point 1 here
-\item enter point 2 here
-\item enter point 3 here
+\item The transition from latitude-longitude to unstructured grids in CESM2.2 adversely impacts the surface mass balance of the Greenland Ice Sheet.
+\item Two Arctic refined mesh configurations have been developed and made available in CESM2.2.
+\item The Arctic meshes substantially improves the surface mass balance over all $1^{\circ}$ grids, including the latitude-longitude grids.
 \end{keypoints}
@@
 \begin{abstract}
-[ enter your Abstract here ]
+
+Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS).
+The more conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both accumulation and ablation over the GrIS.
+Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their greater degrees of freedom in representing the GrIS.
+Two Arctic-refined meshes, with $\frac{1}{4}^{\circ}$ and $\frac{1}{8}^{\circ}$ refinement over Greenland, are documented as newly supported configurations in CESM2.2.
+The Arctic meshes substantially improve the simulated clouds and precipitation rates in the Arctic.
+Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance.
+As CESM is in the process of transitioning away from latitude-longitude grids, these new Arctic refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids.
+
 \end{abstract}
@@
 
-General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high-latitudes, which are among the most sensitive regions on Earth to global and environmental change.
-Despite their importance, the numerical treatment of polar regions in GCMs is handled in vastly-different ways due to the so-called \textit{pole-problem} \cite{W2007JMSJ}.
-The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a).
-Depending on the numerics, methods exist to suppress this instability, and latitude-longitude grids may be advantageous for polar processes as structures can be represented with more degrees of freedom than elsewhere in the computational domain.
-With the recent trend towards globally uniform unstructured grids, any potential benefits of latitude-longitude grids on polar regions may become a relic of the past.
-In this study a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 (CESM; \url{http://www.cesm.ucar.edu/models/cesm2/}), including brand new variable-resolution grids, are evaluated to understand their impacts on the simulated characteristics of the Arctic, with a special focus on the climate and surface mass balance of the Greenland Ice Sheet.
-
-In the 1970's the pole problem was largely defeated through wide-spread adoption of efficient spectral transform methods in GCMs.
-These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g.
-horizontal derivatives) in the equation set can be solved for exactly.
-While spectral transform methods are still used in the 21st century, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
-The pole problem has thus re-emerged in contemporary climate models that use latitude-longitude grids, and some combination of reduced grids and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
-Polar filters are akin to a band-aid; they subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
-This additional damping reduces the effective resolution in polar regions, and the resolved scales are approximately the same everywhere on the grid.
-
-An alternative approach is to use unstructured grids, which allow for more flexible grid structures that permit quasi-uniform grid spacing globally and eliminates the pole-problem entirely (e.g., Figure~\ref{fig:uni-grids}c).
-This grid flexibility also permits variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}).
-Grids can be developed with refinement over polar regions that could in principle make up for any loss in polar resolution in transitioning away from latitude-longitude grids (e.g., Figure~\ref{fig:vr-grids}), although this comes at the cost of a smaller CFL-limiting time-step in the refined region (the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space).
-Unstructured grids also scale more efficiently on parallel systems than latitude-longitude grids, likely resulting in a greater prevalence of unstructured grids as computing power continued to increase over time.
-
-The meteorology and climate of the Arctic is characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
-For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale Polar Lows are not well resolved at these resolutions.
-These mesoscale systems are prevalent during the cold season and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface.
-The Arctic also contains the Greenland Ice Sheet (hereafter denoted as \textit{GrIS}).
-While it blankets the largest island in the world (Greenland), many of the processes that control the GrIS annual surface mass balance (the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions.
-For example, GrIS precipitation is typically confined to the ice-sheet margins (predominately the southeastern margin) where orographic precipitation is generated by steep topographic slopes.
-GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically 10s to 100\,km wide and confined to low-level areas or regions with limited precipitation.
-GCMs struggle to resolve the magnitude and extent of these features \cite{VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.
-
-The goal of this study is to characterize the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these models treat the high-latitudes, e.g., the pole-problem, in different ways.
-The manuscript is laid out as follows: Section~\ref{sec:methods} consists of documentation of the grids, dycores and physical parameterizations used in this study.
-The Arctic refined grids were developed by the authors, and this section serves as their official documentation in CESM2.2.
-Section~\ref{sec:methods} also contains a description of the experiments along with reanalysis datasets and post-processing software used for evaluating the model simulations.
-Section~\ref{sec:results} contains the results of the experiments, followed by Section~\ref{sec:conclusions} that provides a general discussion and conclusions.
+General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change.
+GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole-problem} \cite{W2007JMSJ}.
+The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids).
+Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain.
+With the recent trend towards globally uniform unstructured grids, the benefits of lat-lon grids in polar regions may be lost.
+{\color{blue} (WHL: Would it be worth mentioning that quasi-uniform grids also have benefits in polar regions, e.g.
+by removing elongated grid cells?)}{\color{red} (ARH: I don't think it's worth getting into that here.
+You'll see later in the manuscript that b/c the cells are narrower in the E-W direction, more cells can fit into the big ablation zone on the western margin.)}{\color{purple}(PHL: another advantage of lat-lon grids is that the coordinate lines are orthogonal!)}
+In this study we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate.
+We focus on the climate and surface mass balance of the Greenland Ice Sheet.
+
+In the 1970s the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs. 
+{\color{blue}(WHL: Add refs)}
+These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the equation set can be solved exactly.
+While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
+The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids
+{\color{blue}(define 'reduced grids')}
+and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
+Polar filters subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
+This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid.
+We emphasize \textit{approximately} since it's at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering.
+
+{\color{blue} (WHL: This raises the question: How do lat-lon grids better resolve polar regions at a given nominal global resolution, if the filtering makes them effectively coarser?
+Does the damping degrade some processes more than others?)}{\color{red} (ARH - The phrase "approximately the same" doing a lot of heavy lifting here b/c I frankly don't know.
+The results from this paper suggest that the effective resolution is finer over polar regions, but this is a bit hand-wavy, despite my attempts to quantify it more exactly, and based on the precipitation rates having smaller biases.
+Also, to be clear, effective resolution is specific to atmospheric features; the land model is free to take advantage of these additional grids cells representing the ablation zone, and they certainly do.
+Perhaps I can add something to the effect of ``approximately the same everywhere on the grid, although marginal increases in effective resolution over the polar regions cannot be ruled out...")\\
+
+{\color{purple}PHL: These are good points!
+I think it is useful to separate dynamics from physics in this discussion.
+Physics is called at more columns on the lat-lon grid as one approaches the poles.
+There is no "polar filtering" in physics so effectively the physics resolution is much higher than on a the cubed-sphere grid as one approaches the mid-latitudes and poles.
+Dynamics, on the other hand, filters out smaller scales so effectively it should only retain the 1 degree scale.
+That said, you will get a more accurate solution since you have more grid-points to represent waves (even though smaller scales are filtered) compared to the cubed-sphere grid.
+I think that is pretty clear in flow over orography.
+One can compare the lat-lon grid setup to a spectral model where one keeps viscosity coefficients fixed as one increased horizontal resolution - in such a setup one will get convergence of the dynamics!}}
+
+An alternative approach is to use unstructured grids, which permit quasi-uniform grid spacing globally and eliminates the pole-problem (e.g., Figure~\ref{fig:uni-grids}c).
+{\color{blue} (WHL: Add 1 or 2 refs)}
+This grid flexibility also allows variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}, hereafter abbreviated as \textit{VR}).
+In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).
+However, this refinement comes at the cost of a smaller CFL-limited time step in the refined region.
+(The CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.) Unstructured grids also scale more efficiently on parallel systems than lat-lon grids, another advantage as computing power increases over time.
+
+The meteorology and climate of the Arctic are characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
+For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale polar lows are not well resolved at these resolutions.
+These mesoscale systems are prevalent during winter and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface.
+The Arctic also contains the Greenland Ice Sheet (GrIS), which blankets the worlds's largest island.
+Many of the processes that control the GrIS annual surface mass balance (SMB; the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions.
+For example, GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation.
+GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
+GCMs struggle to resolve the magnitude and extent of these features \cite{P2010CC,VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.
+{\color{blue} (WHL: I suggest adding some language explaining why advance is more prevalent than retreat.
+E.g.:  ``… of these features.
+For example, CESM2 typically underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component.''  Could cite Lofverstrom et al.
+(2020).)}{\color{red} (ARH: I may delete the reference to this anecdote of advancing in northern GrIS altogether b/c I've been unable to show whether it's related to resolution.
+I don't have confidence defining a direction of the SMB bias due to resolution because resolution inflates both source and sink terms; the net effect can go either way.)}
+
+This study compares the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these two dycores treat high latitudes (i.e., the pole problem) in different ways.
+It also serves to document the Arctic-refined grids developed by the authors for CESM2.2.
+Section~\ref{sec:methods} documents the grids, dycores, and physical parameterizations used in this study, and also describes the experiments, datasets, and evaluation methods.
+Section~\ref{sec:results} analyzes the results of the experiments, and Section~\ref{sec:conclusions} provides a general discussion and conclusions.
 
@@
 \end{center}
-\caption{Spectral-element grid for the variable-resolution ARCTIC grid in this study. Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
+\caption{Spectral-element grid for the VR Arctic grids in this study. Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
 \label{fig:vr-grids}
@@
 
-The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 \cite<CAM;>{CAM63}, supports a number of different atmospheric dynamical cores.
-These include dycores using latitude-longitude grids, such as finite-volume \cite<FV;>{L2004MWR} and eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-The EUL dycore is the oldest dycore in CAM, and the least supported of all the dycores.
-FV3 is the newest dycore in CAM, but it was not fully incorporated at the time this work commenced; both the EUL and FV3 dycores are omitted from this study.
-As such, the results presented in this study are comparing the performance of the SE and FV dycores.
+The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 \cite<CAM;>{CAM63}, supports several different atmospheric dynamical cores.
+These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
+The EUL dycore is the oldest and least supported dycore in CAM.
+FV3 is the newest dycore, but was not fully incorporated when this work begin.
+This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores.
+
+[phl: maybe mention that FV was used for IPCC simulations; SE in iHESP]
 
@@
 
-The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical latitude-longitude grid \cite{LR1997QJR}.
-The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate.
-Hyperviscous damping is applied to the divergent modes while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC}.
-A polar filter is used to avoid computational instability due to the convergence of the meridians, allowing for a more practical time-step.
+The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}.
+The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate.
+Hyperviscous damping is applied to the divergent modes, while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC} [phl: as far as I know the Laplacian-like damping is not invoked in lower resolution setups - was implemented for 1/4 applications where the polar night jet became excessively strong].
+A polar filter damps computational instability due to the convergence of meridians, allowing a longer time step.
 It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the poleward direction \cite{ST1995GEOS}.
@@
 The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}.
-The computational domain is a cubed-sphere grid tiled with quadrilateral elements (e.g., Figure~\ref{fig:vr-grids}).
-Each element contains a fourth order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points.
-This results in 16 GLL nodal points within each element, with 12 of the points lying on the (shared) element boundary.
-Communication between elements happens via the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries that reconciles overlapping nodal values and produces a continuous global basis set.
+The computational domain is a cubed-sphere grid tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
+Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points.
+This results in 16 GLL nodal points per element, with 12 of the points lying on the (shared) element boundary.
+Communication between elements uses the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal values and produce a continuous global basis set.
 
 As with the FV dycore, the dynamics evolve in floating Lagrangian layers that are subsequently mapped to an Eulerian reference grid.
-A dry mass vertical coordinate was more recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}.
-The 2D dynamics have no implicit dissipation and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors \cite{DetAl2012IJHPCA}.
+A dry mass vertical coordinate was recently implemented for thermodynamic consistency with condensates \cite{LetAl2018JAMES}.
+The 2D dynamics have no implicit dissipation, and so hyperviscosity operators are applied to all prognostic variables to remove spurious numerical errors \cite{DetAl2012IJHPCA}.
 Laplacian damping is applied in the sponge layer.
 
-The SE dycore supports regional grid refinement via its variable-resolution configuration, requiring two enhancements over uniform resolution grids.
-(1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
+The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements to uniform-resolution grids: (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
 A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements.
@@
 
-For spectral-element grids with quasi-uniform grid spacing, a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) is used instead \cite<>{LTOUNGK2017MWR}.
+For spectral-element grids with quasi-uniform grid spacing, we use a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}.
+{\color{blue} (WHL: I'm not sure what CSLAM is a variant of, since I don't see a tracer transport scheme described above.
+How is tracer transport handled on VR grids?
+If it's not too much detail, why does CSLAM not work on VR grids?)} {\color{red} (ARH: fair point.
+the spectral-element section is lacking in general, so keeping this here as a reminder to explain the spectral-element method is used for tracer transport in ``vanilla" cam-se, and we only support vanilla var-res.
+PHL is planning on adding a description of enhancements to cam-se introduced in CESM2.2, since this turns out to be the only paper documenting CESM2.2.)}
 CSLAM has improved tracer property preservation and accelerated multi-tracer transport.
-It uses a separate grid from the spectral-element dynamics, through dividing each element into $3\times3$ control volumes with quasi-equal area.
+It uses a separate grid from the spectral-element dynamics, dividing each element into $3\times3$ control volumes with quasi-equal area.
 The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated at the GLL nodal points \cite{HL2018MWR}.
@@
 
-Six grid are evaluated in this study (Table~\ref{tbl:table1}).
+We evaluate six grids in this study (Table~\ref{tbl:table1}).
 The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
-The $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid is also run, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
-An additional $1^{\circ}$ CAM-SE-CSLAM grid is run, but with the physical paramerizations computed on a grid that contains $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.
-
-Two variable resolution meshes were developed as part of the CESM2.2 release that contains grid refinement over the Arctic (Figure~\ref{fig:vr-grids}).
+We also run the $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
+We run an additional $1^{\circ}$ CAM-SE-CSLAM grid with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.
+
+Two VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic (Figure~\ref{fig:vr-grids}).
+This paper serves as the official documentation of these grids.
 The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
-The $ARCTIC$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
-The $ARCTICGRIS$ grid is identical to the $ARCTIC$ grid, but contains an additional patch covering the big island of Greenland with $\frac{1}{8}^{\circ}$ resolution.
+The $Arctic$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
+The $Arctic-GrIS$ grid is identical to the $Arctic$ grid, but with an additional patch covering the island of Greenland with $\frac{1}{8}^{\circ}$ resolution.
 
@@
    \hline
-   grid name & dycore & $\Delta x_{eq}$ (km) & $\Delta x_{refine}$ (km) & $\Delta t_{phys}$ (s) \\ 
+   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{refine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) \\ 
    \hline
@@
    $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
-   $ARCTIC$ & SE & 111 & 28 & 450 \\
-   $ARCTCIGRIS$ & SE & 111 & 14 & 225 \\
+   $ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 \\
+   $Arctic$ & SE & 111 & 28 & 450 \\
+   $Arctic-GrIS$ & SE & 111 & 14 & 225 \\
  \hline
@@
   \caption{Grids and dycores used in this study.
-$\Delta x_{eq}$ refers to average equatorial grid spacing, $\Delta x_{refine}$ refers to grid spacing in the refined region (if applicable) and $\Delta t_{phys}$ refers to the physics time-step.
-The dycore abbreviation FV refers to the finite-volume dycore, SE the spectral-element dycore and SE-CSLAM the spectral-element dycore w/ CSLAM tracer advection.}
+$\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{refine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step.
+FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.
+We use the $ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.}
  \label{tbl:table1}
@@
 
+The physics time step should depend on grid resolution.
+Increased horizontal resolution permits faster vertical velocities that reduce characteristic time scales, so the physics time step is reduced to avoid large time truncation errors \cite{HR2018JAMES}.
+The $ARCTIC$ and $Arctic-GrIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time step relative to the default 1800 s time step for the standard $1^{\circ}$ and $2^{\circ}$ grids (Table~\ref{tbl:table1}).
+
+All grids and dycores in this study use 32 levels in the vertical, with a model top of about 1~hPa or 40~km.
+However, any grid or dycore can in principle be run with a higher top or finer vertical resolution.
+
 \subsection{Physical parameterizations}
 
-The CAM6 physical parameterization package (hereafter referred to as the \textit{physics}; \url{https://ncar.github.io/CAM/doc/build/html/index.html}) is used in all simulations in this study.
-CAM6 physics is most noteably different from it's predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a PBL, shallow convection and cloud macrophysics scheme.
-CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory in representing cloud ice for improved cloud-aerosol interactions.
-Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} inlcuding convective momentum transport \cite{RSG2010JAS}.
-PBL form drag is modeled after \cite{BBW2004QJRMS} and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.
-
-All grids and dycores in this study use 32 levels in the vertical, with a model top of about $1 \ hPa$ or about $40 \ km$.
-The physics time-step is dependent on grid resolution.
-Increases in horizontal resolution permit faster vertical velocities that reduce characteristic time-scales, and so the physics time-step is reduced to avoid large time truncation errors \cite{HR2018JAMES}.
-The $ARCTIC$ and $ARCTICGRIS$ grids are therefore run with a $4\times$ and $8\times$ reduction in physics time-step relative to the default 1800 s time-step used in coarser, uniform resolution grids (Table~\ref{tbl:table1}.
-
-Initial simulations with the $ne30pg3$ spectral-element grid produced weaker shortwave cloud forcing relative to the tuned up finite-volume dycore.
-All runs with the spectral-element dycore have two CLUBB parameter changes in order to provide a more realistic cloud forcing and top-of-atmosphere radiation balance.
-These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and c14, reduced from 2.2 to 1.6.
-Briefly, the $gamma$ parameter scales the width of the sub-grid distribution of vertical velocity, and c14 controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy.
-For a thorough explanation of how CLUBB parameters impact the simulated climate, the reader is referred to \cite{GETAL2015JAMES}.
+All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{CAM63}.
+The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer (PBL), shallow convection, and cloud macrophysics scheme.
+CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
+Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
+PBL form drag is modeled after \citeA{BBW2004QJRMS}.
+Orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.
+
+Initial simulations with the $ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore.
+All runs with the SE dycore therefore have two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance.
+These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and $c14$, reduced from 2.2 to 1.6.
+Briefly, $gamma$ scales the width of the sub-grid distribution of vertical velocity, and $c14$ controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy.
+For a thorough explanation of how CLUBB parameters impact the simulated climate, see \cite{GETAL2015JAMES}.
 
@@
 
-All grids and dycores are run using an identical transient 1979-1998 AMIP-style configuration, with prescribed monthly SST/sea-ice after \cite{CESMSST}.
-This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2.
-
-The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study.
-The SMB is the sum of the mass source term, accumulation (i.e., precipitation), and the mass sink term, ablation.
-Ablation can be expressed as evaporation/sublimation plus total runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation becomes runs off the ice sheet; rain may penetrate pore spaces in the firn layer and freeze, forming ice lenses in the subsurface.
-These processes are represented by different components in CESM, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
-
-CLM runs on the same grid as the atmosphere, but also uses a downscaling technique to account for sub-grid variability in SMB.
-In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (EC), weighted by their respective area fractions at each EC, which is derived from a high resolution GrIS elevation dataset.
-The near surface air temperature, humidity and air density are calculated at each EC using an assumed lapse rate and the elevation difference from the grid mean, and the precipitation rates from CAM are repartitioned into solid or liquid based on the temperature of the EC.
-Ice accumulation is modeled as a capping flux, or snow in excess of a 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of ice.
-A unique surface energy balance and SMB is computed for each EC.
-Integrating over all ECs using the area weights provides a more accurate SMB.
-For a more detailed description of how the SMB is computed in CESM, the reader is referred to \cite{CISM1,SETAL2019TC,KETAL2020JAMES}.
-
-Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up by forcing CLM in standalone mode, cycling over a 20 year $ARCTIC$ $FHIST$ run for about 500 years.
-The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.
+We run all grids and dycores using an identical transient 1979-1998 Atmospheric Model Inter-comparison Project configuration (AMIP), with prescribed monthly SST and sea ice following \citeA{CESMSST}.
+In CESM terminology, AMIP simuations use the $FHIST$ computational set and run out of the box in CESM2.2.
+
+All grids and dycores simulate the GrIS surface mass balance (SMB).
+The SMB is the sum of the mass accumulation term,  i.e., precipitation, and the mass sink term, ablation.
+Ablation is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
+Not all liquid precipitation runs off the ice sheet; rain can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the snow mass.
+These processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
+
+CLM runs on the same grid as the atmosphere, using a downscaling technique to account for sub-grid variability in SMB.
+In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (ECs), each with a distinct surface energy balance and SMB.
+The area fraction of each EC is derived from a high-resolution GrIS elevation dataset.
+The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean.
+Precipitation from CAM is repartitioned into solid or liquid based on the surface temperature of the EC; precipitation falls as snow for $T < -2^\circ$~C, as rain for $T > 0^\circ$~C,  and as a rain/snow mix between $-2^\circ$~C and $0^\circ$~C.
+Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent.
+Any snow above the 10-m cap is routed to the ocean as solid runoff.
+Refreezing of liquid within the snowpack is an additional source of ice.
+Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
+For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES}.
+
+Changes in ice depth, but not snow depth, count toward the SMB.
+That is, snow accumulation above the 10-m cap contributes a positive SMB, and surface ice melting (after melting of the overlying snow) yields a negative SMB.
+Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over a 20-year $Arctic$ $FHIST$ run for about 500 years.
+The uniform-resolution grids are initialized with the SMB from an existing $f09$ spun-up initial condition.
 
@@
 
-Several observational datasets are used in this study to understand the performance of the simulations.
-A list of the datasets used in this study are shown in Table~\ref{tbl:table2}.
-Several of these products (ERA5, CALIPSO and CERES) are near-global gridded datasets commonly used to evaluate GCMs.
-Surface mass balance datasets are gathered from several sources.
-RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA interim and ERA5 renalysis at its domain boundaries.
-The RACMO simulations have been shown to performs very skillfully against observations and is therefore considered an ideal modeling target \cite{NETAL2015TC,NETAL2019SCIENCE}.
-The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit} maintains a repository of snow pit and and ice core SMB measurements, as well as the IceBridge radar accumulation dataset.
-The LIVVkit dataset is compared against model simulations by finding the nearest grid cell center to the location of each observation.
+We use several observational datasets (Table~\ref{tbl:table2}) to gauge the performance of the simulations.
+Several products are near-global gridded datasets commonly used to evaluate GCMs.
+We gathered SMB datasets from several sources.
+RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA40, ERA interim and ERA5 renalysis products at the domain boundaries.
+The RACMO simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite{NETAL2015TC,NETAL2019SCIENCE}.
+
+In-situ SMB (snow pit and ice cores) and radar accumulation datasets (e.g., IceBridge) are maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}.
+However, these point-wise measurements are difficult to compare to model output spanning several different grids, especially since the SMB from each EC is not available from the model output.
+We used a nearest-neighbor technique for an initial analysis, which showed that the model biases are similar to those computed using the RACMO datasets.
+Because of the uncertainty of comparing gridded fields to point-wise measurements, and the lack of information added with regard to model biases, we omitted these datasets from our analysis.
 
@@
 
-\subsection{SMB Analysis}
-
-A common high resolution dataset is used to generate the GrIS boundary conditions in all grids, using the CLM dataset creation tools.
-Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change of the GrIS.
-Figure~\ref{fig:grisdx} shows the GrIS ice mask area across the different grids, as a function of the number of grid points.
-Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset.
-RACMO2.3, however, uses a smaller ice mask, about $3\%$ smaller than the raw ice mask dataset.
-
-Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, but we seek to compare these integrated quantities to RACMO2.3.
-Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks.
-Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$, the lowest resolution grid for both dycores considered in this study.
-In addition, two remapping algorithms are used; ESMF conservative and TempestRemap high-order, monotone algorithm.
-In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_grisres.pdf}
+\subsection{SMB Analysis \color{red}{ARH - this section is under construction}}\label{sec:SMB}
+
+We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
+However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}).
+The area errors in Figure~\ref{fig:grisdx} may not seem large, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
+CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework first-order conservative remapping algorithm \cite<ESMF;>{ESMF}.
+The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
 \end{center}
 \caption{The spatial properties of the GrIS as represented by different grids in this study.
-(Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks.}
+(Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
 \label{fig:grisdx}
@@
 
+We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $f19$ and $ne30pg2$ grids, and integrating over these low-resolution ice masks.
+The use of low-resolution common ice masks is a conservative decision and is largely because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not reliable in the downscaling direction, i.e., mapping to a higher-resolution grid than the source grid.
+We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
+Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $f19$ and $ne30pg2$ masks.
+Thus, we evaluate an integrated quantity on a given grid up to four times to estimate the uncertainty due to differences in grid type and remapping algorithms.
+
+The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets.
+The SMB for total water can be expressed as:
+\begin{equation}
+SMB = precipitation + runoff + evaporation/sublimation, \label{eq:SMB}
+\end{equation}
+where all terms have consistent sign conventions (positive values contribute mass, negative values reduce mass).
+The precipitation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
+Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined snow plus ice melt.
+Note that this SMB expression is different from the internally computed SMB in CLM \cite{KETAL2020JAMES}.
+
+We consider two approaches for mapping and integrating the SMB components over the common ice masks:
+\begin{enumerate}
+\item Map the grid-cell mean quantities to the common grid, and integrate the mapped fields over the common ice masks. \label{lbl:label1}
+\item Map the patch-level quantities (i.e., the state over the ice fractional component of the grid cell) to the common grid, and integrate the mapped fields over the common ice masks.
+\label{lbl:label2}
+\end{enumerate}
+
+Note that we are mapping to low-resolution grids that have larger GrIS areas (Figure~\ref{fig:grisdx}).
+Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
+While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
+Method~\ref{lbl:label2}, in contrast, does not map values outside the source ice mask to the common ice mask.
+This tends to be dissipative, since mapping to the low-resolution grid will average a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
+However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.
+%The dilution of the these patch level fields caused by mapping to the common ice mask are then inflated due to integrating over the larger ice mask.
+
+The different error characteristics of the two methods are used to diversify the ensemble.
+Each of the four regridding combinations (with conservative and high-order remapping to the $f09$ and $ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
+Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
+This will degrade the SMB estimates using method~\ref{lbl:label2}, but note that we are more interested in the behavior of the processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
+
 \section{Results}\label{sec:results}
@@
 
-Before delving into the simulated characteristics of the Arctic, the global mean differences between the various grids and dycores are assessed.
-Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between the uniform resolution grids and dycores.
-The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes and throughout the depth of the troposphere.
-This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}, and \cite{HK2020QJRMS} has shown that this occurs in CAM due to greater resolved vertical velocities that in turn, facilitate greater condensational heating in the macrophyiscs routine in CLUBB.
-The right columns in Figure~\ref{fig:dT-lores} supports this interpretation, which shows an increase in the climatological CLUBB heating in the low and mid-latitudes in the $f09$ grid.
+Before delving into the simulated Arctic, we assess the global mean differences between the various grids and dycores.
+Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between uniform-resolution grids and dycores.
+The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes throughout the depth of the troposphere.
+This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}.
+\citeA{HK2020QJRMS} have shown that this occurs in CAM due to greater resolved vertical velocities which, in turn, drive more condensational heating in the CLUBB macrophyiscs.
+The right columns in Figure~\ref{fig:dT-lores} support this interpretation, showing an increase in the climatological CLUBB heating in low and mid-latitudes on the $f09$ grid.
 
@@
 \end{center}
-\caption{.}
+\caption{1979-1998 annual mean temperature (left column) and CLUBB temperature tendencies (right column) in zonal mean height space, expressed as differences between the various $1^{\circ}-2^{\circ}$ grids.
+The thick grey and magenta lines are the tropopause for the control run and the test run, respectively.}
 \label{fig:dT-lores}
@@
 
-As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so a modest, resolution-like sensitivity occurs in which $ne30pg3$ is warmer than $f09$ (Figure~\ref{fig:dT-lores}).
-The stratosphere has a different response, in which $ne30pg3$ is much cooler than $f09$ in the mid-to-high latitudes.
-Figure~\ref{fig:dT-lores} also shows differences in temperature between $ne30pg3$ and $ne30pg2$, which are small, although there is a slight warming near the tropopause at high latitudes.
-This is consistent with the similar climates found between these grids in \cite{HETAL2019JAMES}.
-
-Comparing the variable-resolution grids to the uniform resolution grids is complicated because we simultaneously increase the grid resolution and reduce the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}.
-An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$.
-Figure~\ref{fig:dT-hires} shows the change in the climatological summer temperatures in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$.
-A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB.
+As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so the $ne30pg3$ troposphere is modestly warmer than $f09$ (Figure~\ref{fig:dT-lores}).
+The stratosphere responds differently, with $ne30pg3$ much cooler than $f09$ in the mid-to-high latitudes.
+Figure~\ref{fig:dT-lores} also shows small temperature differences between $ne30pg3$ and $ne30pg2$, with  $ne30pg3$ slightly warmer near the tropopause at high latitudes.
+This is consistent with the similar climates found for these two grids by \citeA{HETAL2019JAMES}.
+
+Comparing the VR grids to the uniform-resolution grids is complicated because we simultaneously increase the resolution and reduce the physics time-step, both of which impact the solution \cite{W2008TELLUS}.
+We therefore run an additional $ne30pg3$ simulation with the shorter physics time step used in the $Arctic$ grid (450~s), referred to as $ne30pg3^{*}$.
+Figure~\ref{fig:dT-hires} shows the difference between $ne30pg3^{*}$ and $ne30pg3$ for climatological summer temperatures in zonal-mean height space.
+Much of the troposphere is warmer with the reduced time step, and the mechanism is similar in that the shorter time step increases condensational heating by CLUBB.
 Figure~\ref{fig:dT-hires} shows the difference in climatological summer temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
-The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.
+With the same physics time step on each grid, the greater condensational heating and warmer temperatures are confined to the refined Arctic region.
+
+Figure~\ref{fig:dT-hires} shows that the $Arctic-GrIS$ grid is much warmer than the $Arctic$ grid in the Arctic summer. This may be due, in part, to the shorter physics time step,
+%; a similar physics-time-step sensitivity in the $ne30pg3^{*}$ comparison is evident in the CLUBB tendencies in the $ARCTCGRIS$ grid.
+but the temperature response is too large to be explained by the CLUBB changes alone.
+This summer warming appears to be a result of variations in the stationary wave pattern, with anomalous southerly winds to the west of Greenland (not shown).
+This dynamic response is interesting, because other than the physics time step, the only difference between the $Arctic-GrIS$ and $Arctic$ runs is the doubling of resolution over Greenland.
+This behavior will be explored further in a future study.
 
@@
 \end{center}
-\caption{.}
+\caption{As in Figure~\ref{fig:dT-lores} but for the short-time-step experiment and the VR grids.
+The fields plotted are for the climatological northern hemisphere summer.
+We focus on summer because that is when the resolution response is largest, and the refined regions are located in the northern hemisphere.}
 \label{fig:dT-hires}
@@
 
-It's useful to understand summer temperature biases, instead of annual means, due to its control on ice/snow melt \cite{O2001JAM,HT2008Paleo}.
-Figure~\ref{fig:dThyps} shows the 1979-1998 lower troposphere summer temperature bias relative to ERA5.
-It is computed from the 500 hPa-1000 hPa geopotential thickness, solving for the layer mean virtual temperature using the hypsometric equation.
-The results generally track with the analysis of the zonal mean height plots; increasing resolution from $f19$ to $f09$ leads to a warmer climate, and the $1^{\circ}$ spectral-elements grids are warmer than the finite-volume grids.
-The summer temperatures in the finite-volume grids are persistently colder than ERA5 at high latitudes, whereas the $1^{\circ}$ spectral-element grids are warmer than ERA5 at only very high-latitudes, north of $80^{\circ}$.
-All grid illustrate a north-south gradient in bias over Greenland, in which the summer temperature bias becomes more positive in the northward direction.
+It is useful to understand summer temperature biases due to their control on ice and snow melt over the Greenland Ice Sheet \cite<GrIS;>{O2001JAM}.
+Figure~\ref{fig:dThyps} shows the 1979--1998 lower troposphere summer temperature bias relative to ERA5.
+It is computed from the 500--1000 hPa geopotential thickness, solving for the layer-mean virtual temperature using the hypsometric equation.
+The results are consistent with the zonal mean height plots; increasing resolution from $f19$ to $f09$ warms the climate, and the $1^{\circ}$ SE grids are warmer than the FV grids.
+The FV summer temperatures are persistently colder than ERA5, whereas the $1^{\circ}$ SE grids are not as cold, and are actually warmer than ERA5 at high-latitudes, north of $75^{\circ}$.
+All grids show a north-south gradient in bias over Greenland, with the summer temperature bias more positive for the northern part of the ice sheet.
 This pattern is also evident in the 2m summer temperature bias over Greenland (not shown).
 
-The $ARCTIC$ grid has similar summer temperatures to the $1^{\circ}$ spectral-element grids, but it is a bit warmer over northern Eurasia and the North Pole.
+The $Arctic$ grid has summer temperatures similar to the $1^{\circ}$ SE grids, but is slightly warmer over northern Eurasia and the North Pole.
 An anomalous cooling patch forms to the west of Greenland, centered over Baffin Island.
-The $ARCTICGRIS$ grid is warmer than the $ARCTIC$ grid over most of the Arctic region, but approximately maintains the same pattern of summer temperature bias as in the $ARCTIC$ grid.
-
-Some of these temperature anomalies may be related to summer shortwave cloud forcing differences across the different grids and dycores.
-Figure~\ref{fig:SWCF} shows the summer short-wave cloud forcing bias in the runs, using the CERES-EBAF product.
-All the uniform $1^{\circ}-2^{\circ}$ grids have similar biases, with the clouds reflecting 20-40 W/m2 too much shortwave radiation over a wide swath of the Arctic, primarily over the land masses.
-There's also a halo of low cloud forcing bias around the oceanic perimeter of Greenland.
-The $ARCTIC$ grid has much smaller cloud forcing biases over the Arctic land masses, although still too reflective, whereas the $ARCTCGRIS$ grid vastly improves the cloud forcing bias over Eurasia, and improves the bias over N.America compared to the $ARCTIC$ grid.
-In both variable-resolution grids, the halo of too weak cloud forcing bias around the perimeter of Greenland is absent.
-
-While the summer cloud forcing biases are consistent with the summer temperature biases in Figure~\ref{fig:dThyps} --regions where clouds are too reflective coincide with regions that are too cold-- it is not clear whether the cold biases are caused by the cloud biases, or whether the cold biases amplify the cloud forcing bias.
+The $Arctic-GrIS$ grid is warmer than the $Arctic$ grid over most of the Arctic, but with a similar spatial pattern of summer temperature bias.
+
+Some of these temperature differences may be related to different summer shortwave cloud forcing for the various grids and dycores.
+Figure~\ref{fig:SWCF} shows the summer shortwave cloud forcing bias in six runs, using the CERES-EBAF product.
+A negative bias corresponds to excessive reflection and cooling.
+The uniform grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses.
+There is also a halo of positive bias (clouds not reflective enough) around the ocean perimeter of Greenland.
+The $Arctic$ grid has much smaller cloud forcing biases over the Arctic land masses, but is still too reflective over Alaska, the Canadian Archipelago, and parts of Eurasia.
+Compared to the $Arctic$ grid, the $ARCTCGRIS$ grid vastly reduces the cloud forcing bias over Eurasia, and also improves the bias over North America.
+In both VR grids, the halo of positive forcing bias around the perimeter of Greenland is absent.
+
+While the summer cloud forcing biases are consistent with the summer temperature biases in Figure~\ref{fig:dThyps} -- regions where clouds are too reflective coincide with regions that are too cold -- it is not clear whether the cold biases are caused by the cloud biases, or whether the cold biases amplify the cloud forcing bias.
 
@@
 \end{center}
-\caption{.}
+\caption{1979-1998 lower troposphere, northern hemisphere summer virtual temperature biases, computed as the difference from ERA5.
+Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation.
+Differences are computed after mapping the ERA5 data to the finite-volume grids since the geopotential field is only available on the output tapes in the spectral-element runs that have been interpolated to the $f09$ grid, inline.}
 \label{fig:dThyps}
@@
 
-In addition to summer temperatures, shortwave radiation is also an important determinant of snow/ice melt.
+In addition to summer temperatures, shortwave radiation is an important determinant of snow and ice melt.
 Figure~\ref{fig:FSDS} shows the summer incident shortwave radiation bias at the surface, zoomed in over Greenland.
-The top panel computes the bias using the CERES-EBAF dataset, and the bottom panel using RACMO2.3p2 dataset.
-This halo of excessive incident shortwave radiation around the coasts of Greenland is apparent in both datasets, consistent with the shortwave cloud forcing biases in Figure~\ref{fig:SWCF}.
-
-The interior of the ice sheet receives too little shortwave radiation in the coarser grids.
-In the variable-resolution grids, both the interior deficit in shortwave and the excessive shortwave around the oceanic perimeter of Greenland are improved.
-This suggests that the coarse grids clouds are too thick in the interior of Greenland, and too thin around the perimeter of Greenland, and that increasing horizontal resolution balances out these biases.
-This is consistent with total summer cloud fraction bias, computed from the CALIPSO cloud dataset (Figure~\ref{fig:prect}).
-Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates any changes arising from any single layer due to the maximum overlap assumption used to compute this quantity.
-Despite the attenuated signal, the total cloud fraction does indicate a reduction in cloud coverage in the interior, and an increase in cloudiness about the oceanic perimeter, in the variable-resolution grids.
-
-The agreement of the cloud biases over Greenland in multiple independent datasets indicates this is a robust feature of the coarser grids.
-The reduction of these cloud biases in the variable-resolution grids suggests they are a result of insufficient horizontal resolution in coarse grids.
+The top panel shows the bias relative to the CERES-EBAF dataset, and the bottom panel relative to the RACMO2.3p2 dataset.
+The halo of excessive incident shortwave radiation around the coasts of Greenland is apparent for both datasets in relation to the coarser grids, consistent with the shortwave cloud forcing biases in Figure~\ref{fig:SWCF}.
+
+The ice sheet interior receives too little shortwave radiation on the coarser grids.
+On the VR grids, both the interior shortwave deficit and the excessive shortwave around the ocean perimeter are improved.
+This suggests that the coarse-grid clouds are too thick in the Greenland interior and too thin around the perimeter, and that increasing horizontal resolution reduces these biases.
+This is consistent with the total summer cloud fraction bias, computed from the CALIPSO-GOCCP cloud dataset  and shown in Figure~\ref{fig:prect}.
+Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates the changes arising from any single layer due to the maximum overlap assumption used to compute this quantity.
+Despite the attenuated signal, the total cloud fraction for the VR grids does indicate reduced cloud coverage in the interior and increased cloudiness around the ocean perimeter.
+
+The agreement of the cloud biases in and around Greenland from multiple independent datasets shows that the biases are a robust feature of the coarser grids.
+The reduced biases on the VR grids suggest that the coarse-grid biases are a result of insufficient horizontal resolution.
 
@@
 \end{center}
-\caption{Climatological (1979-1998) annual precipitation rate bias (in mm/yr) relative to the RACMO2.3p2 5.5km resolution data product \cite{NETAL2019SCIENCE}.}
+\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias, relative to the CALIPSO-GOCCP dataset, and (bottom) precipitation rate bias, relative to the RACMO2.3p2 dataset.
+The CALIPSO-GOCCP differences are computed after mapping all model output to the $1^{\circ}$ grid, whereas the RACMO differences are computed after mapping the RACMO dataset to the individual model grids.
+Note that the averaging period for the CALIPSO-GOCCP panels, 2006-2017, is different than the model averaging period.
+\color{red}{ARH - still trying to fix the layout of this figure so the label bars aren't on top of the panels}.}
 \label{fig:prect}
@@
 Figure~\ref{fig:prect} shows the average grid spacing over the Greenland Ice Sheet (GrIS) in all six grids in this study.
-The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $ARCTICGRIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO2.3 grid.
-The $ne30pg3$ grid has an average $\Delta x=111.2~km$, which is substantially more coarse than the $f09$ grid, with an average $\Delta x=60~km$.
-This is interesting because $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, but due to the convergence of meridians the finite-volume model has enhanced resolution over GrIS.
-The $ARCTIC$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models (whereas the $ARCTICGRIS$ grid is about twice as expensive as the $ARCTIC$ grid).
-
-The summer climatological mean precipitation bias over GrIS is shown in Figure~\ref{fig:grisdx}, expressed as the fractional difference from the RACMO2.3p2 solution.
-What sticks out is that the coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
-The $ARCTIC$ run improves this bias substantially, and the $ARCTICGIS$ run improves this bias further.
-This suggests the southern dome bias is due to inadequate horizontal resolution, which is consistent with the original GrIS variable-resolution experiments in \cite{VETAL2018TC}.
-
-Southeast Greenland has the largest accumulation rates in GrIS due to synoptic systems moving in from the southeast.
-These systems are orographically lifted by the steep southeast ice sheet margin, dumping large amounts of precipitation along the southeast coast.
-At lower horizontal resolutions, the topography is too smooth and large amounts of moisture penetrates further inland, incorrectly dumping precipitation onto the interior of the ice sheet.
-A similar bias occurs in northwest Greenland, in particular during the summer, when it's common for synoptic systems to arrive from the southwest.
-The ability of the variable-resolution grids to more accurately simulate the orographic precipitation process in Greenland is consistent with all the cloud results up to this point.
-Since the precipitation centers move from the interior towards the coasts, and even out over the ocean with increasing resolution, the cloud decks should, and are, moved accordingly, eliminating this halo of low cloud bias around the oceanic perimeter of Greenland.
-
-Figure~\ref{fig:tseries} shows time-series of annually integrated precipitation and snow/ice melt over the GrIS in the various different grids and dycores, with both versions of RACMO shown in black.
-The 1979-1998 climatological mean values are provided as circles on the right side of the panels.
-The uniform $1^{\circ}-2^{\circ}$ grids all show a distinctive positive bias in precipitation due to this over-prediction of interior precipitation rates.
-The  variable-resolution grids have the smallest precipitation biases, providing a comparable solution to RACMO.The $f19$ and $f09$ perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$, +230 Gt/yr.
-The results suggest that uniform resolution spectral-element grids have larger biases than the finite-volume grids, consistent with spectral-element grids having a coarser representation of GrIS (Figure~\ref{fig:grisdx}).
-
-The combined annual snow/ice melt integrated over the GrIS is given by the bottom panel of Figure~\ref{fig:tseries}.
-The $ARCTIC$ grid simulates the most realistic melt rates, with all other grids tending to have larger melt rates than RACMO.
-The $ARCTICGRIS$ grid over predicts melting by about 125 Gt/yr.
-This is likely due to an anomalously warm lower troposphere during the summer, relative to the $ARCTIC$ run (Figure~\ref{fig:dThyps}).
-The $f19$ and $f09$ melting rates are improved over $ARCTICGRIS$, simulating too much melt by about 70-90 Gt/yr.
-The spectral-element grids have the largest positive melt bias, between 200-220 Gt/yr.
-It is more difficult to attribute these differences to resolution alone, since the finite-volumes grids have colder summer temperatures than the uniform resolution spectral-element grids.
-However, that the $ARCTCIGRIS$ grid has the warmest summer temperatures, yet has a lower melting bias than the uniform spectral-element grids, suggests that increasing resolution improves the ablation process.
+The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $Arctic-GrIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO2.3 grid.
+The $ne30pg3$ grid has an average $\Delta x=111.2~km$, substantially coarser than the $f09$ grid, with an average $\Delta x=60~km$.
+Although $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
+The $Arctic$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
+The $Arctic-GrIS$ grid is about twice as expensive as the $Arctic$ grid.
+
+The lower panels of Figure~\ref{fig:prect} show the summer climatological mean precipitation bias over the GrIS, expressed as the fractional difference from the RACMO2.3p2 solution.
+The coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
+The $Arctic$ grid reduces this bias substantially, and the $Arctic-GrIS$ grid reduces it further.
+This suggests that the southern dome bias arises from inadequate horizontal resolution, consistent with the original GrIS VR experiments in \citeA{VETAL2018TC}.
+
+Large GrIS accumulation rates result from synoptic systems arriving from the south.
+These systems are orographically lifted at the ice sheet margin, especially over the steep slopes in southeast Greenland, concentrating heavy precipitation near the ice margin.
+At lower resolutions, the topography is too smooth and moisture penetrates inland, erroneously dumping precipitation onto the southern dome.
+The ability of the VR grids to more accurately simulate orographic precipitation is consistent with the cloud results above.
+As the precipitation centers move from the interior toward the margins, and even out over the ocean with increasing resolution, the cloud decks move accordingly.
+Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, and \ref{fig:prect} clearly illustrate a shift in clouds from the interior to the ocean perimeter with increasing resolution.
 
@@
  \scriptsize
- \begin{tabular}{lccc}
+ \begin{tabular}{lccccc}
    \hline
-   grid name & accumulation (Gt/yr) & ice/snow melt (Gt/yr) \\ 
+   grid name & accumulation & total melt & runoff & sublimation & SMB \\ 
    \hline
-   $RACMO$ & 768.5 & -347.2 \\
+   $RACMO$ & 768.5 (733.5) & -347.2 (-436.4) & -221.7 (-258.5) & -36.5 (-38.8) & 510.3 (436.2) \\
    \hline
-   $f19$ & 882.5 & -440.3 \\
-   $f09$ & 874.8 & -418.4 \\
-   $ne30pg2$ & 1000. & -549.4 \\
-   $ne30pg3$ & 934.9 & -568.8 \\
-   $ARCTIC$ & 795.9 & -367.3 \\
-   $ARCTICGRIS$ & 708.7 & -471.6 \\
+   $f19$ & 882.5 (913.5) & -440.3 (-546.5) & -283.7 (-284.3) & -36.6 (-37.5) & 562.2 (591.7) \\
+   $f09$ & 874.8 (882.1) & -418.4 (-482.3) & -255.0 (-212.3) & -38.1 (-37.4) & 581.7 (632.4) \\
+   $ne30pg2$ & 1000. (973.4) & -549.4 (-647.3) & -383.9 (-347.0) & -33.4 (-32.1) & 582.7 (594.3) \\
+   $ne30pg3$ & 934.9 (909.3) & -568.8 (-686.7) & -356.2 (-330.1) & -34.4 (-32.6) & 544.3 (546.6) \\
+   $Arctic$ & 795.9 (818.6) & -367.3 (-436.8) & -208.9 (-194.2) & -44.1 (-43.9) & 542.9 (580.5) \\
+   $Arctic-GrIS$ & 708.7 (747.3) & -471.6 (-610.4) & -261.1 (-307.8) & -50.7 (-51.8) & 396.9 (387.7) \\
  \hline
  \end{tabular}
-  \caption{1979-1998 Surface Mass Balance of the Greenland Ice Sheet.}
+  \caption{1979-1998 surface mass balance of the Greenland Ice Sheet in Gt/yr.
+Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.}
  \label{tbl:table3}
@@
 \end{center}
-\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to RACMO3.2.
-The raw fields are mapped to two target low resolution grids, f19 \& ne30pg2, and using two different remapping methods, conservative ESMF and high order TempestRemap.
-The remapped values are then integrated over the ice mask of the target grid.
-This gives four time-series for each simulation (three for f19 \& ne30pg2), with the mean value given by the solid line and shading spanning the extent of the remapped solutions.}
+\caption{Time-series of annual (solid+liquid) precipitation (top) and annual runoff (bottom) integrated over the Greenland Ice Sheet for all six simulations and compared to the RACMO datasets.
+The time-series were generated using the common ice mask approach, which results in up to 4 ensembles, with the mean value given by the solid line and shading spanning the extent of the ensemble members.}
 \label{fig:tseries}
@@
 
+Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO.
+The CESM values are averages over the ensemble of common ice masks and regridding methods described in section~\ref{sec:SMB}, and the RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}) using the same common-ice-mask approach.
+Table~\ref{tbl:table3} also contains (in parentheses) the SMB components derived from evaluating the integrals on each model's native grid and ice mask.
+Of note is the large reduction in melt rates using the common-ice-mask approach compared to the native grid, illustrating the dissipation discussed in section~\ref{sec:SMB}.
+The errors are greatest in partially ice-covered grid cells straddling the ice sheet margins, in the ablation zone where melt rate are large.
+For integrated precipitation, the differences between the native and common-ice-mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
+
+Figure~\ref{fig:tseries} shows time series of annually integrated precipitation and snow/ice melt over the GrIS for the various different grids and dycores, with both versions of RACMO shown in black.
+The 1979-1998 climatological mean values, listed in Table~\ref{tbl:table3}, are shown as circles on the right side of the panels.
+The uniform $1^{\circ}-2^{\circ}$ grids have positive precipitation biases in the interior, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
+The $f19$ and $f09$ grids perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$ by +230 Gt/yr.
+The larger biases on the uniform-resolution SE grids relative to the FV grids are consistent with the coarser GrIS resolution on the SE grids (Figure~\ref{fig:grisdx}).
+
+The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $Arctic$ grid simulates the most realistic melt rates, with the other grids having more melt than RACMO.
+The $Arctic-GrIS$ grid overpredicts melting by about 125 Gt/yr.
+This is likely due to an anomalously warm lower troposphere during the summer, relative to the $Arctic$ run (Figure~\ref{fig:dThyps}).
+The $f19$ and $f09$ melting rates are improved over $Arctic-GrIS$, overestimating melt by only 70--90 Gt/yr.
+The SE grids have the largest positive melt bias, between 200--220 Gt/yr.
+It is more difficult to attribute these differences to resolution alone, since the FV grids have colder summer temperatures than the uniform-resolution SE grids.
+
+To illustrate the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \citeA{RM2012GRL}.
+The uncertainty due to differences in basin area is larger than for GrIS-wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which are shown in the $f19$ and $ne30pg2$ panels of Figure~\ref{fig:melt}.
+Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the most accumulation.
+In all basins, accumulations drops monotonically with increasing grid resolution, with some exceptions.
+The $Arctic-GrIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to the RACMO precipitation in the large southwest basin.
+
 \begin{figure}[t]
@@
 \end{center}
-\caption{.}
+\caption{1979--1998 basin integrated components of the SMB; (top) precipitation, (middle) ice/snow melt and (bottom) ice/snow melt estimated from the PDD method.
+Whiskers span the max/min of the four ensemble members generated from the common-ice-mask approach.
+Basin definitions are after \citeA{RM2012GRL}, and are found on the common ice masks using a nearest neighbor approach, and shown in Figure~\ref{fig:melt}.}
 \label{fig:basin}
@@
 \end{center}
-\caption{.}
+\caption{1979--1998 ice/snow melt biases (in mm/yr) relative to RACMO2.3p2, evaluated on the native model grids.
+The \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid.
+Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $f19$ and $ne30pg2$ panels, in computing the basin-scale integrals.
+Blue lines in the $f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
 \label{fig:melt}
@@
 
-Figure~\ref{fig:pointdiff} shows the distribution of point-wise differences from LIVVkit observational database as violin plots.
-The IceBridge dataset is exclusively from the interior of the ice sheet and represents accumulation rates.
-The uniform $1^{\circ}-2^{\circ}$ grids have similar median errors of about +35-50 mm.w.e, while the variable-resolution errors are noticeably smaller.
-The in-situ observations in the accumulation zone, shown in the middle plot looks very similar to the IceBridge errors, providing confidence that the variable-resolution grids are outperforming the uniform grids in the interior accumulation zone.
-The errors evaluated at in-situ ablation zone measurements are in tension with the RACMO results in Figure~\ref{fig:tseries}.
-They indicate that the uniform $1^{\circ}-2^{\circ}$ grids perform similarly, and that the $ARCTICGRIS$ performs best, and an improvement over the $ARCTIC$ grid.
-However, the in-situ ablation measurements are sparse in time and space, and so there is a large amount of uncertainty in extending these results to the entirety of the ablation zone.
-
-%\begin{figure}[t]
-%\begin{center}
-%         \includegraphics[width=130mm]{figs/temp_violens.pdf}
-%\end{center}
-%\caption{.}
-%\label{fig:pointdiff}
-%\end{figure}
-
-The ``k-transect" is perhaps the most well studied transect in Greenland, as it has been regularly monitored since the 1950s.
-The LIVVkit has compiled the k-transect observations, shown in Figure~\ref{fig:trans} as elevation vs.
-SMB along the transect, with the model's simulated transects shown as well.
-The in-situ observation points are nicely replicated by the $ARCTCIGRIS$ run, whereas the $ARCTIC$ grid is biased positive in the higher-elevations of the ablation zone.
-The $f09$ grid is surprisingly competitive with the variable-resolution grids, capturing a realistic slope of the elevation-SMB curve, although it exhibits a similar positive bias as the $ARCTIC$ grid.
-The elevation-SMB slopes of the uniform spectral-element and $f19$ grid are too shallow, in particular at the higher elevation regions of the transect.
-
-Figure~\ref{fig:ztrans} shows the representation of the surface of the ice sheet along these transects in the different grids, compared to the high resolution dataset used to generate CAM topography boundary conditions.
-The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $f09$ grid is a bit of an exception --since the grid cells become very narrow in the meridional direction at high latitudes, a larger number of grid cells can populate the east-west transect, consistent with its skillful representation of the ablation zone.
-
-What the authors refer to as the ``b-transect" in northwest Greenland (Figure~\ref{fig:trans}) is characterized by orographic precipitation, resulting in the accumulation zone extending down to the ice margin.
-The variable resolution grids perform relatively well, with larger SMBs at lower elevations where the precipitation rates are highest, whereas the $1^{\circ}-2^{\circ}$ grids underestimate the SMB in these lower elevation regions.
-Only the variable resolution grids can capture the local reduction in SMB in the 1500-2000 m region.
-The skill of the variable-resolution grids is clearly related to the more accurate representation of the steepness of the transect, while also capturing the protrusion around 1500-2000 m that coincides with the local minimum in SMB (Figure~\ref{fig:ztrans}).
-
-%\begin{figure}[t]
-%\begin{center}
-%         \includegraphics[width=130mm]{figs/temp_transect_obsperiod.pdf}
-%\end{center}
-%\caption{.}
-%\label{fig:trans}
-%\end{figure}
+The basin-integrated melt rates in Figure~\ref{fig:basin} depend on the dycore.
+The uniform-resolution SE grids have the largest positive biases in all basins.
+The $Arctic-GrIS$ grid is a close second, while the FV grids have systematically smaller melt-rates.
+The ``second-place" standing of $Arctic-GrIS$ is somewhat unexpected, as this grid has the warmest lower-troposphere summer temperatures (Figure~\ref{fig:dThyps}) and greatest incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the uniform-resolution SE grids.
+
+Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones.
+Positive degree-days \cite<PDD;>{B1984Z}, which accumulate the near-surface temperature in $^{\circ}C$ for days with temperature above freezing, are a more accurate proxy.
+PDD is nonlinear in mean monthly temperature \cite{R1991P}.
+We compute it from monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}$C and a degree-day factor of 5~mm~d$^{-1}$ $^{\circ}$C$^{-1}$.
+
+Figure~\ref{fig:basin} shows the basin-integrated PDD melt estimate.
+In the large southeast and southwest basins (and all the other western basins), the $ne30pg3$ grid has larger PDD-based melt than the $ARCTCGRIS$ grid.
+The FV grids also have large PDD-based melt in the southwest basin, relative to $Arctic-GrIS$.
+The PDD plots indicate that the near-surface temperatures which contribute to melt are not well approximated by the summer lower-troposphere temperatures in Figure~\ref{fig:dThyps}.
+
+Figure~\ref{fig:melt} presents the biases in the combined ice/snow melt as map plots.
+These plots show that the largest melt biases are on the southeast and northwest coasts, where large coarse-grid cells overlap with the ocean.
+One possibility is that these problematic grid cells are situated at lower elevations than the true ice sheet surface, leading to a warm bias and too much melt.
+Figure~\ref{fig:ztrans} shows the representation of the ice sheet surface along two transects on the different grids, compared to the high-resolution dataset used to generate CAM topographic boundary conditions \cite{GMTED2010,gmdd-8-4623-2015}.
+The two transects are shown in Figure~\ref{fig:melt}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast.
+The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a few grid cells populating the transect.
+The $f09$ grid is a bit of an exception for the K-transect, with grid cells becoming narrow in the meridional direction at high latitudes.
+The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of the GrIS margins.
 
@@
 \begin{tabular}{cccc}
-         \includegraphics[width=130mm]{figs/temp_ktransect.pdf} \\
-         \includegraphics[width=130mm]{figs/temp_btransect.pdf} \\
+         \includegraphics[width=130mm]{figs/temp_zsects.pdf}
 \end{tabular}
 \end{center}
-\caption{.}
+\caption{Model surface elevation along the (top) K-transect, and (bottom) a transect spanning the central dome down to the Kangerlussuaq glacier in southeast Greenland, for all model grids.
+The reference surface (GMTED) is a 1 km surface elevation dataset used for generating the CAM topographic boundary conditions.
+}
 \label{fig:ztrans}
@@
 
-To give an idea of what the ablation zones look like in the highest resolution $ARCTCIGRIS$ grid, Figure~\ref{fig:viz} is a still from a visualization produced by the authors.
-It shows the daily mean surface mass balance during the height of the melt season in year 1981 of the simulation.
-The ablation zone appears to be well resolved in the $ARCTICGRIS$ grid.
-The full visualization is publicly available on youtube.com\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}.
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/Vis1923.jpg}
-\end{center}
-\caption{.}
-\label{fig:viz}
+The transects in Figure~\ref{fig:ztrans} show that the ice sheet surface on the coarse grids is not systematically lower than the true surface in ablation zones.
+Rather, the smoothing of the raw topography, necessary to prevent the model from instigating grid-scale modes, flattens the ice sheet, causing the lower-elevation ablation zones to extend beyond the true ice sheet margin, where they lie above the actual ice surface.
+The $f19$ grid has the smoothest topography since its dynamics are coarsest (whereas $f09$, $ne30pg2$ and $ne30pg3$ use identical smoothing), and has the flattest ice sheet.
+This suggests that if anything, coarser models will tend to elevate the ablation zones and depress melt rates.
+
+Figure~\ref{fig:ztrans} also shows the ice margin boundary, illustrating that the ablation zone lies in a narrow horizontal band where the ice sheet rapidly plunges to sea-level.
+Due to this abrupt transition, coarse grids will commonly represent the ablation zone with grid cells containing mixtures of ice-covered and ice-free regions.
+We hypothesize that coarser models have larger melt biases
+%despite the $ARCTCIGRIS$ having the warmest troposphere,
+because summer melting is confined to these mixed ice/land/ocean grid cells.
+CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees a homogenized state due to volume averaging over the sub-grid mixture.
+Thus, warm ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice-covered patch.
+
+Figure~\ref{fig:bias} shows mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region.
+Errors are computed using the common-ice-mask approach, meaning that all fields are mapped to the common masks, which define the grid cell ice fraction.
+The figure shows that coarser grids generally have two peaks in ice fraction space; a bump in positive melting errors in the 0-20$\%$ range, and another in the fully ice-covered cells.
+Also shown are the $\pm 1$~standard deviation of the biases for each bin.
+They indicate that the biases in 0-20$\%$ bins are mostly contained in the positive bias region (a fractional bias greater than 0), whereas the fully-covered ice cells have a wider distribution, with many grid cells also containing negative melting biases.
+The excessive melting in the 0-20$\%$ ice fraction bins supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone on coarse grids is responsible for their large melt bias.
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cccc}
+         \includegraphics[width=80mm]{figs/temp_xy_diffRACMO_melt.pdf}
+\end{tabular}
+\end{center}
+\caption{Fractional melt bias over the GrIS, computed relative to the RACMO datasets using the common ice mask approach, and conditionally sampled by grid cell ice fraction provided by the common ice masks.
+Solid lines are the mean of the distribution with $\pm$ one standard deviation expressed by shading.}
+\label{fig:bias}
 \end{figure}
@@
 
-Synoptic storms are tracked and analyzed using TempestExtremes \cite{UETAL2021}.
-As the $ARCTIC$ grid contains $\frac{1}{4}^{\circ}$ refinement north of about $45^{\circ}$ latitude, the storm tracker is applied to this region for the $ARCTIC$ and $ne30pg3$ run to identify differences in storm characteristics due to horizontal resolution.
-Figure~\ref{fig:comp-mean} shows the mean precipitation rates averaged over all January storms identified by TempestExtremes.
-The iconic comma structure of the synoptic cyclones is simulated in $ne30pg3$ and $ARCTIC$ grids, with the magnitudes about the same in these two grids, with perhaps a marginal increase in precipitation rates in the storm center of the $ARCTIC$ grid.
-For good measure, the $ne30pg3^{*}$ run is also plotted, and looks more-or-less identical to the $ne30pg3$ run.
-
-As has been previously reported, horizontal resolution can have large impacts on extreme precipitation events.
-Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with synoptic storms, by month.
+Synoptic storms are tracked using TempestExtremes atmospheric feature detection software \cite{UETAL2021}.
+As the $Arctic$ grid contains $\frac{1}{4}^{\circ}$ refinement north of about $45^{\circ}$ latitude, the storm tracker is applied to this region for the $Arctic$ and $ne30pg3$ runs to identify differences in storm characteristics due to horizontal resolution.
+The composite mean precipitation maps are similar between the two grids, and exhibit the iconic comma structure of synoptic cyclones (not shown).
+
+Figure~\ref{fig:comp-pdf} shows monthly PDFs of the precipitation rates associated with storms.
 The PDFs are constructed by sampling all the precipitation rates within $30^{\circ}$ of the storm center, for each point on the storm track and for all storms.
 The PDFs are evaluated on an identical composite grid for all runs, and so storm statistics are not impacted by differences in output resolution.
-The $ARCTIC$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
-This is primarily due to an increase in resolution, and not the reduced physics times-step; the $ne30pg3^{*}$ only marginally increases the precipitation rates compared with $ne30pg3$.
-
-The extreme precipitation rates in the $ARCTIC$ run are closer to the ERA5 reanalysis than in $ne30pg3$ (Figure~\ref{fig:comp-pdf}).
-It's difficult to know the extent that the extreme precipitation rates in ERA5 are constrained by data assimilation, or whether these precipitation rates are due to using a similar $\frac{1}{4}^{\circ}$ model as the $ARCTIC$ grid.
-However, it is well documented that $\frac{1}{4}^{\circ}$ models are more skillful at simulating extreme events \cite{BetAl2013JC,OETAL2016JAMES}, and so this is an additional benefit of the variable-resolution grids.
-
-%\begin{figure}[t]
-%\begin{center}
-%         \includegraphics[width=130mm]{figs/temp_composite_ge45N_JAN-PRECT.pdf}
-%\end{center}
-%\caption{.}
-%\label{fig:comp-mean}
-%\end{figure}
+The $Arctic$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
+This is primarily due to increased resolution and not the reduced physics times-step; the $ne30pg3^{*}$ run only marginally increases the extreme precipitation rates compared with $ne30pg3$ (Figure~\ref{fig:comp-pdf}).
+
+The extreme precipitation rates in the $Arctic$ run are closer than $ne30pg3$ to the ERA5 reanalysis (Figure~\ref{fig:comp-pdf}).
+It is difficult to know how much the extreme precipitation rates in ERA5 are constrained by data assimilation, or whether these precipitation rates are due to using a similar $\frac{1}{4}^{\circ}$ model as the $Arctic$ grid.
+However, it is well documented that $\frac{1}{4}^{\circ}$ models are more skillful at simulating extreme events \cite{BetAl2013JC,OETAL2016JAMES}.
+This is an additional benefit of the VR grids.
 
@@
 \end{center}
-\caption{.}
+\caption{PDFs of the total precipitation rate associated with tracked storms, by month, in the $ne30pg3$, $ne30pg3^{*}$ and $Arctic$ runs, and compared with the ERA5 dataset.}
 \label{fig:comp-pdf}
@@
 
-Six grids from different dynamical cores in CESM2.2 are evaluated in an AMIP-style configuration for their performance over the Arctic and in simulating the Greenland Ice Sheet (GrIS) surface mass balance (SMB).
-The $1-2^{\circ}$ finite-volume grids have enhanced resolution over Polar regions due to the convergence of meridian lines, although a polar filter is employed to prevent features from forming at this higher resolution.
+Running CESM2.2 in an AMIP-style configuration, we have evaluated six grids from two dynamical cores for their performance over the Arctic and in simulating the Greenland Ice Sheet (GrIS) surface mass balance (SMB).
+The $1-2^{\circ}$ finite-volume grids have enhanced resolution over polar regions due to the convergence of meridian lines, although a polar filter is used to prevent spurious atmospheric features from forming at this higher resolution.
 Spectral-element grids comparable to the resolution of the finite-volume grids have an isotropic grid structure, meaning the grid resolution is similar over the entire domain.
-Two variable-resolution grids were developed and introduced into CESM2.2 as part of this work.
-They use the spectral-element dycore; the $ARCTIC$ grid has $\frac{1}{4}^{\circ}$ refinement over the broader Arctic, whereas the $ARCTICGRIS$ grid is identical except it has an $\frac{1}{8}^{\circ}$ patch of refinement over Greenland.
-
-In general, the finite-volume grids have colder summer temperatures over the Arctic, and the spectral-element grids (incl.
-the variable-resolution grids) are warmer.
-The cloud biases in all the uniform resolution grids, whether finite-volume or spectral-elements, are similar, in general being too cloudy over Arctic land masses.
-The variable-resolution grids largely improve the cloud biases.
-It should be emphasized that our analysis is specific to the Arctic summer due to its relevance to melt rates over GrIS; improved clouds in the Arctic should not be taken to mean that lower latitude regions have an improved cloud field as well.
-
-At the regional level, a halo of negative cloud bias is found around the oceanic perimeter of Greenland in all $1-2^{\circ}$ grids, and is absent in the variable-resolution grids.
-This halo bias coincides with a positive cloud bias over the interior of the ice sheet.
-This has been traced back to the inadequacy of the lower resolution grids to resolve the orographic precipitation process in Greenland.
-Synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins due to overly smooth topography in the $1-2^{\circ}$ grids.
-As as result, moisture penetrates accross the ice margin and dumps excess precipitation into the interior of the GrIS, instead of being concentrated closer to the coastal margins as indicated by observations.
-This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter of Greenland.
-The agreement of different observational data products on this bias lends confidence  in attributing the cause of the precipitation and cloud biases.
-The variable-resolution grids compare more favorably to the observations and indicate the orographic precipitation process in Greenland is largely resolved.
-
-The primary source and sink terms of the SMB equation are integrated over the GrIS, and evaluated in the six grids.
-The uniform $1-2^{\circ}$ grids all have large positive accumulation biases owing to their inability to resolve the orographic precipitation process.
-The uniform spectral-element grids have larger accumulation biases, suggesting that the finite-volume grids are more skillful at resolving the precipitation processes due to their finer grid spacing over Greenland, despite the polar filter.
-The variable resolution grids have the most accurate accumulation rates of all the grids.
-
-The primary mass sink term of the GrIS, ice/snow melt, expresses similar biases as the accumulation rates.
-The uniform resolution spectral-element grids melt too much, while the finite-volume grids have reduced biases.
-It's more difficult to attribute these biases to grid resolution alone; the finite-volume grids have colder summers that is consistent with their lower melt bias.
-However, the $ARCTICGRIS$ grid has the warmest summer temperatures of all grids, yet it has reduced melting compared to the uniform resolution spectral-element grids.
+We developed two VR grids and introduced them into CESM2.2 as part of this work.
+Both use the spectral-element dycore; the $Arctic$ grid has $\frac{1}{4}^{\circ}$ refinement over the broader Arctic, whereas the $Arctic-GrIS$ grid is identical except for a $\frac{1}{8}^{\circ}$ patch of refinement over Greenland.
+
+In general, the FV grids have colder summer temperatures over the Arctic compared with the SE grids (including the VR grids).
+The cloud biases in all the uniform-resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.
+The VR grids reduce the cloud biases.
+It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates; improved clouds in the Arctic do not imply improved clouds at lower latitudes.
+
+At the regional level, there is a halo of negative cloud bias around the ocean perimeter of Greenland on all $1-2^{\circ}$ grids, but not the VR grids.
+This halo bias coincides with a positive cloud bias over the ice sheet interior.
+This pattern has been traced to the inadequacy of the coarser grids for resolving orographic precipitation.
+With overly smooth topography on the $1-2^{\circ}$ grids, synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins.
+As a result, moisture penetrates and dumps excess precipitation into the GrIS interior, instead of being concentrated near the coastal margins as shown by observations.
+This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter.
+The agreement of different observational data products on this bias lends confidence in the attribution of causes.
+The VR grids compare better to the observations and show that orographic precipitation in Greenland is largely resolved.
+
+We integrated the primary source and sink terms of the SMB equation over the GrIS for each of the six grids.
+The uniform $1-2^{\circ}$ grids have large positive accumulation biases because they fail to resolve orographic precipitation.
+The uniform SE grids have larger accumulation biases, suggesting that the FV grids are more skillful for precipitation due to finer resolution over Greenland, despite a polar filter.
+The VR grids have the most accurate accumulation rates of all the grids.
+
+The primary mass sink term of the GrIS, ice/snow melt, has similar biases.
+The uniform resolution SE grids have too much melt, while the FV grids have smaller biases.
+It is difficult to attribute these biases to grid resolution alone.
+The FV grids have colder summers, consistent with their lower melt bias.
+However, the $Arctic-GrIS$ grid has the warmest summer temperatures of all grids, yet it has less melting than the uniform-resolution SE grids.
 This suggests that grid resolution is responsible for a large fraction of the melt biases.
-That the melt process would be sensitive to grid resolution is intuitive, since local elevation impacts on temperature and melting is largely handled by the land model, which does not have a pole problem because there is no horizontal dynamics.
-The larger number of grid cells in the ablation zone in the finite-volume grids seems to more accurately resolve the melt process relative to the uniform resolution spectral-element grids, in which the ablation zone is made up of substantially fewer grid cells.
+We propose a mechanism: Coarse grids represent ablation zones using grid cells with mixed surface types, ice-covered and ice-free.
+The warmer ice-free patches may largely determine the mean state, leading to a warm bias over the ice-covered patches of the grid cell.
+This mechanism is supported by analysis of melt biases binned by grid-cell ice fraction.
+
+The $Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the Greenland SMB, compared to the uniform $1^{\circ}-2^{\circ}$ grids.
+The $Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but its summer temperatures are too warm.
+The $1^{\circ}$ FV model gives a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids.
+In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other uniform-resolution grids.
+
+As modeling systems move away from lat-lon grids towards quasi-uniform unstructured grids, it is worth taking stock of whether this will degrade the simulated polar climate.
+We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model in simulating the surface mass balance of the GrIS.
+This finding will not interrupt the ongoing transition towards unstructured grids in CESM, largely driven by gains in computational efficiency, but it has inspired us to develop alternative configurations that recover or improve on the fidelity of polar climate.
+We have shown here (and in a prior companion study \cite{VETAL2018TC}) that for CESM, Arctic-refined meshes can substantially improve the simulated mass balances of the GrIS, even compared to the $1^{\circ}$ grid.
+This should reassure the CESM modeling community that the ongoing transition away from lat-lon grids will not adversely impact CESM's usefuless as a state-of-the-art tool for simulating and understanding polar processes.
+{\color{blue} (WHL: This last sentence may be too sanguine.
+Yes, we can recover the fidelity of Arctic simulations using VR grids, but (so far) only at a considerable cost in cpu-hours.
+This points to the need for an intermediate resolution that is more affordable, and/or model development or tuning that reduces the biases on coarse grids.)}
+
+We are  working to develop a configuration of the $Arctic$ grid that is fully-coupled with the CESM ocean and sea ice components and the Community Ice Sheet Model (CISM), to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise.
+A fully-coupled pre-industrial control configuration has already been developed and vetted.
+%, and should be incorporated into CESM in the near future.
+We have also developed a visualization of the $ARCTCGRIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}.
+%in order to advertise some of the stunning capabilities of the refined mesh approach.
+Figure~\ref{fig:viz} shows a snapshot of this visualization, illustrating mesoscale katabatic winds descending the southeastern slopes of GrIS.
+%We are optimistic about the polar science opportunities provided by these new Arctic mesh configurations, and our open to collaborating with the CESM community to assist in anyway we can.
+These new grids and configurations will provide new opportunities for CESM polar science.
+{\color{blue} (WHL: I replaced the previous last sentence because it seemed too much like an advertisement.
+However, this new ending seems weak.
+I wonder if we should say something about future work motivated by this study, for instance investigating grids and parameterizations that provide some of the same benefits as these VR grids but at lower cost.)}
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/Vis1472.jpg}
+\end{center}
+\caption{Snapshot of the lowest model level streamlines from the $Arctic-GrIS$ visualization, with color shading denoting the wind magnitude.}
+\label{fig:viz}
+\end{figure}
 
@@
 %
-%\appendix
-%\section{Here is a sample appendix}
+\appendix
+\section{Details on spectra-element dynamical core improvements since the CESM2.0 release}
+{\color{red}{PHL working on this!}}\\
+
+Since the CESM2.0 release of the spectral-element dynamical core documented in \citeA{LetAl2018JAMES} some important algorithmic improvements have been implemented and released with CESM2.2.
+These pertain mainly to the flow over orography that, for the spectral-element dynamical core, can lead to noise aligned with the element boundaries \cite{HL2018MWR}.
+\subsection{Reference profiles}
+Significant improvement in removing noise for flow over orography can be achieved by using reference profiles for temperature and pressure
+\begin{eqnarray}
+  T^{(ref)}&=&T_0+T_1 \Pi^{(ref)},\\
+  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}T_0}\right)}
+\end{eqnarray}
+\cite{SJ1991QJRMS} where $g$ gravity, $T_0\equiv 288K$, $T_1=\Gamma_0 T_0 c_p^{(d)}/g$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential.
+The reference Exner function is
+\begin{equation}
+   \Pi^{(ref)}=\left( \frac{p^{(ref)}}{p_0}\right)^{\kappa}
+\end{equation}
+where $\kappa=\frac{R_d}{c_p^{(d)}}$ (the reference surface pressure $p_0=1000hPa$ and at each model level the reference pressure $p^{(ref)}$ is computed from $p_s^{(ref)}$ and the standard hybrid coefficients
+\begin{equation}
+    p^{(ref)}(\eta) = A(\eta)+B(\eta)p_s^{(ref)},
+\end{equation}
+where $A$ and $B$ are the standard hydrid coefficients (using a dry-mass generalized vertical mass coordinate $\eta$).
+These reference profiles are subtracted from the prognostic temperature and pressure-level-thickness states before applying hyperviscosity:
+\begin{eqnarray*}
+   \hbox{CESM2.0}&\rightarrow&\hbox{CESM2.2}\\
+   && \\
+  \nabla_\eta T &\rightarrow& \nabla_\eta \left(T-T^{(ref)}\right),\\ 
+    \nabla_\eta \delta p^{(d)} &\rightarrow& \nabla_\eta \left(\delta p^{(d)}-\delta p^{(ref)}\right).
+\end{eqnarray*}
+This reduces spurious transport of temperature and mass up/down-slope due to hyperviscosity.
+
+\subsection{Rewritting the pressure gradient force (PGF)}
+In the CESM2.0 we used to used the following (standard) form of the pressure gradient term:
+\begin{equation}
+\label{eq:pgf}
+    \nabla_{\eta }\Phi+\frac{1}{\rho}\nabla_{\eta }p,
+\end{equation}
+where $\Phi$ is geopotential and $\rho=\frac{R_dT_v}{p}$ is density and $\eta $ is the dry-mass vertical coordinate \cite<for details see >{LetAl2018JAMES}.
+To alleviate noise for flow over orography we switched to an Exner pressure formulation, that is, we use that (\ref{eq:pgf}) can be written in terms of the Exner pressure
+\begin{equation}\label{eq:pgf2}
+    \nabla_{\eta }\Phi+c_p^{(d)}\theta_v\nabla_{\eta }\Pi,
+\end{equation}
+where the Exner pressure is
+\begin{equation}
+    \Pi\equiv \left( \frac{p}{p_0}\right)^{\kappa}.
+\end{equation}
+The derivation showing that (\ref{eq:pgf}) and (\ref{eq:pgf2}) are equivalent is shown here:
+\begin{eqnarray*}
+c_p^{(d)}\theta_v\nabla_{\eta }\Pi &=& c_p^{(d)}\theta_v\nabla_{\eta }\left( \frac{p}{p_0}\right)^{\kappa},\\
+&=& c_p^{(d)}\theta_v\kappa\left( \frac{p}{p_0}\right)^{\kappa-1} \nabla_{\eta }\left( \frac{p}{p_0}\right),\\
+&=& c_p^{(d)}\theta_v\kappa\Pi\left( \frac{p_0}{p}\right) \nabla_{\eta }\left(\frac{p}{p_0}\right),\\
+&=& \frac{c_p^{(d)}\theta_v\kappa\Pi}{p} \nabla_{\eta }p,\\
+&=& \frac{R_d\theta_v\Pi}{p} \nabla_{\eta }p,\\
+&=& \frac{R_dT_v}{p} \nabla_{\eta }p,\\
+&=& \frac{1}{\rho} \nabla_{\eta }p,
+\end{eqnarray*}
+\subsection{Results}
+One year averages of $OMEGA500$ (vertical pressure velocity at 500hPa) has been found to be a useful quantity to detect spurious up or down-drafts induced by steep orography.
 
@@
 This material is based upon work supported by the National Center for Atmospheric Research (NCAR), which is a major facility sponsored by the NSF under Cooperative Agreement 1852977.
-Computing and data storage resources, including the Cheyenne supercomputer
-(doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at NCAR.
+Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at NCAR.
+A.
+Herrington thanks Matt Rehme (NCAR/CISL) for his role in generating the $Arctic-GrIS$ visualization available on youtube ({\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}).
 

### 2022-04-06 arcticgrids.tex
Updates from Overleaf

@@
 \citeA{HK2020QJRMS} have shown that this occurs in CAM due to greater resolved vertical velocities which, in turn, drive more condensational heating in the CLUBB macrophyiscs.
-The right columns in Figure~\ref{fig:dT-lores} support this interpretation, showing an increase in the climatological CLUBB heating in low and mid-latitudes on the $f09$ grid.
+The right columns in Figure~\ref{fig:dT-lores} support this interpretation, showing an increase in the climatological CLUBB heating in low ({\color{blue}{looks more like subtropics}}) and mid-latitudes on the $f09$ grid.
 
@@
 Figure~\ref{fig:dThyps} shows the 1979--1998 lower troposphere summer temperature bias relative to ERA5.
-It is computed from the 500--1000 hPa geopotential thickness, solving for the layer-mean virtual temperature using the hypsometric equation.
+It is computed from the 500--1000 hPa geopotential thickness, solving for the layer-mean virtual temperature using the hypsometric equation ({\color{blue}{Might be good to write this down, for those of us who studied atmospheric science a really long time ago}}).
 The results are consistent with the zonal mean height plots; increasing resolution from $f19$ to $f09$ warms the climate, and the $1^{\circ}$ SE grids are warmer than the FV grids.
@@
 A negative bias corresponds to excessive reflection and cooling.
-The uniform grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses.
+The uniform grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses ({\color{blue}{please explain what SWCF in caption.}}).
 There is also a halo of positive bias (clouds not reflective enough) around the ocean perimeter of Greenland.
@@
 While the summer cloud forcing biases are consistent with the summer temperature biases in Figure~\ref{fig:dThyps} -- regions where clouds are too reflective coincide with regions that are too cold -- it is not clear whether the cold biases are caused by the cloud biases, or whether the cold biases amplify the cloud forcing bias.
+({\color{blue}{This ends abruptly.
+Anything else to add?}})
 
@@
 The accuracy of the simulated SMB is expected to be sensitive to grid resolution.
-Figure~\ref{fig:prect} shows the average grid spacing over the Greenland Ice Sheet (GrIS) in all six grids in this study.
+Figure~\ref{fig:prect}b shows the average grid spacing over the Greenland Ice Sheet (GrIS) in all six grids in this study.
 The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $Arctic-GrIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO2.3 grid.
@@
 The $Arctic-GrIS$ grid is about twice as expensive as the $Arctic$ grid.
+({\color{blue}{Wondering if this paragraph would go better in an earlier section.}})
 
@@
 It is more difficult to attribute these differences to resolution alone, since the FV grids have colder summer temperatures than the uniform-resolution SE grids.
+({\color{blue}{Not sure I understand this sentence.
+FV is both cooler and higher-resolution than SE, so one might suspect that cooler T is, in fact, connected to higher resolution.}})
 
@@
 
-Figure~\ref{fig:basin} shows the basin-integrated PDD melt estimate.
+Figure~\ref{fig:basin}c shows the basin-integrated PDD melt estimate.
 In the large southeast and southwest basins (and all the other western basins), the $ne30pg3$ grid has larger PDD-based melt than the $ARCTCGRIS$ grid.
@@
 Thus, warm ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice-covered patch.
+({\color{blue}{This is an interesting conclusion pointing to the need for better treatment of surface inhomogeneity in CAM.
+This might be a way to compensate for coarse resolution in future CESM versions?}})
 
@@
 Errors are computed using the common-ice-mask approach, meaning that all fields are mapped to the common masks, which define the grid cell ice fraction.
-The figure shows that coarser grids generally have two peaks in ice fraction space; a bump in positive melting errors in the 0-20$\%$ range, and another in the fully ice-covered cells.
+The figure shows ({\color{blue}{Any idea why the errors are smaller in the cells with intermediate ice fraction?
+I wouldn't have expected this.}}) that coarser grids generally have two peaks in ice fraction space; a bump in positive melting errors in the 0-20$\%$ range, and another in the fully ice-covered cells.
 Also shown are the $\pm 1$~standard deviation of the biases for each bin.
@@
 
-At the regional level, there is a halo of negative cloud bias around the ocean perimeter of Greenland on all $1-2^{\circ}$ grids, but not the VR grids.
+At the regional level, there is a halo of negative cloud bias (/color{blue}{I got confused about signs here, because section 3 talks about a halo of positive cloud SW forcing bias, which corresponds to a negative bias in cloud amounts.
+Maybe replace 'cloud bias' with 'cloudiness bias' or something similar?}) around the ocean perimeter of Greenland on all $1-2^{\circ}$ grids, but not the VR grids.
 This halo bias coincides with a positive cloud bias over the ice sheet interior.

### 2022-04-18 arcticgrids.tex
coauthor edits

@@
 % Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, William H. Lipscomb\affil{1}, Marcus Lofverstrom\affil{2} \ and Andrew Gettelman\affil{1} }
+\authors{Adam R. Herrington\affil{1}, Peter H. Lauritzen\affil{1}, Marcus Lofverstrom\affil{2}, William H. Lipscomb\affil{1}, Andrew Gettelman\affil{1} \ and Mark A. Taylor\affil{3}}
 
  \affiliation{1}{National Center for Atmospheric Research, 1850 Table Mesa Drive, Boulder, Colorado, USA}
-\affiliation{2}{Department of Geosciences, University of Arizona, 1040 E. 4th Street, Tucson, AZ USA}
-
+\affiliation{2}{Department of Geosciences, University of Arizona, 1040 E. 4th Street, Tucson, Arizona USA}
+\affiliation{3}{Sandia National Laboratories, Albuquerque, New Mexico, USA}
 % \affiliation{1}{First Affiliation}
@@
 \begin{keypoints}
-\item The transition from latitude-longitude to unstructured grids in CESM2.2 adversely impacts the surface mass balance of the Greenland Ice Sheet.
-\item Two Arctic refined mesh configurations have been developed and made available in CESM2.2.
-\item The Arctic meshes substantially improves the surface mass balance over all $1^{\circ}$ grids, including the latitude-longitude grids.
+\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two new Arctic refined mesh configurations.
+\item Uniform resolution grids degrade the surface mass balance of the Greenland Ice Sheet compared with equivalent low resolution latitude-longitude grids.
+\item The refined Arctic meshes substantially improve the surface mass balance over all low resolution grids.
 \end{keypoints}
@@
 Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their greater degrees of freedom in representing the GrIS.
-Two Arctic-refined meshes, with $\frac{1}{4}^{\circ}$ and $\frac{1}{8}^{\circ}$ refinement over Greenland, are documented as newly supported configurations in CESM2.2.
+Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, are documented as newly supported configurations in CESM2.2.
 The Arctic meshes substantially improve the simulated clouds and precipitation rates in the Arctic.
 Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance.
-As CESM is in the process of transitioning away from latitude-longitude grids, these new Arctic refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids.
+As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids.
 
@@
 
-\section*{Plain Language Summary}
-[ enter your Plain Language Summary here or delete this section]
+%\section*{Plain Language Summary}
+%[ enter your Plain Language Summary here or delete this section]
 
@@
 Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain.
-With the recent trend towards globally uniform unstructured grids, the benefits of lat-lon grids in polar regions may be lost.
-{\color{blue} (WHL: Would it be worth mentioning that quasi-uniform grids also have benefits in polar regions, e.g.
-by removing elongated grid cells?)}{\color{red} (ARH: I don't think it's worth getting into that here.
-You'll see later in the manuscript that b/c the cells are narrower in the E-W direction, more cells can fit into the big ablation zone on the western margin.)}{\color{purple}(PHL: another advantage of lat-lon grids is that the coordinate lines are orthogonal!)}
+With the recent trend towards globally uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost.
 In this study we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate.
-We focus on the climate and surface mass balance of the Greenland Ice Sheet.
-
-In the 1970s the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs. 
+We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.
+
+In the 1970s, the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs. 
 {\color{blue}(WHL: Add refs)}
-These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the equation set can be solved exactly.
+These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the (truncated) equation set can be solved exactly.
 While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
@@
 This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid.
-We emphasize \textit{approximately} since it's at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering.
-
-{\color{blue} (WHL: This raises the question: How do lat-lon grids better resolve polar regions at a given nominal global resolution, if the filtering makes them effectively coarser?
+We emphasize \textit{approximately} since it's at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.
+
+Dycores built on lat-lon grids have some advantages over unstructured grids.
+Lat-lon coordinate lines are orthogonal, and aligned with zonally symmetric circulations that characterize many large-scale features of Earth's atmosphere.
+\citeA{LJTN2010JAMES} has experimented with rotating lat-lon models such that their coordinate lines no longer align with an idealized, zonally balanced circulation.
+For the finite-volume lat-lon dycore considered in this paper (hereafter \textit{FV}), numerical errors were shown to be largest when the polar singularity is rotated into the baroclinic zone ($45^{\circ}$N latitude), generating spurious wave growth much earlier on in the simulation than for other rotation angles.
+This illustrates the advantages of coordinate surfaces aligned with latitude bands, albeit an extreme example where the polar singularity and the polar filter are also contributing to the spurious wave growth.
+The unstructured grids all generate spurious baroclinic waves earlier on in the simulations than the (unrotated) lat-lon models, although the unstructured model considered in this paper, the spectral-element dycore (hereafter \textit{SE}), holds a balanced zonal flow without wave growth appreciably longer than the rotated FV experiments, regardless of how the SE grid is rotated \cite{LJTN2010JAMES}.
+
+The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which have large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}.
+Unstructured grids support quasi-uniform grid spacing globally, and there is no pole-problem (e.g., Figure~\ref{fig:uni-grids}c).
+Unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure.
+{\color{blue} (WHL: Add 1 or 2 refs)} This grid flexibility allows for variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
+In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).
+However, refinement comes at the cost of a smaller CFL-limited time step in the refined region.
+(The CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.)
+
+It is important to emphasize that the pole-problem is a feature of the dycore.
+Polar filters do not directly interfere with the physical parameterizations, nor do they have any bearing on the surface models; e.g., the land model can take full advantage of the greater number of grid cells in polar regions on lat-lon grids.
+This is particularly relevant for the surface mass balance of the Greenland Ice Sheet (\textit{SMB}; the integrated sum of precipitation and runoff), which relies on hydrological processes handled by the land model.
+
+The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales that are difficult to represent in GCMs \cite{P2010CC}.
+GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation.
+The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias (Pollard 2000) \cite{VETAL2018TC}.
+GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
+These narrow ablation zones are not fully resolved in low resolution GCMs, and may further degrade the simulated SMB.
+For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.
+
+Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates (citations for MM5, RACMO and MAR).
+The limited area domain used by RCMs permits the use of high resolution grids able to resolve the SMB processes, and can skillfully simulate the GrIS SMB.
+However, unlike GCMs, RCMs are not a freely evolving system and boundary conditions must be prescribed at the lateral boundaries.
+The inability of the RCM solution to influence larger-scale dynamics outside the RCM domain (due to the prescribed boundary conditions) severely limits this approach from properly representing the role of the GrIS in the climate system.
+
+In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} utilized the VR capabilities of the SE dycore in CESM, generating a grid where Greenland is represented with $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution.
+The simulated SMB compared favorably to the SMB from RCMs and observations.
+The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios, in the freely evolving GCM framework.
+
+The SE dycore has been an available option since CESM, version 1, but has been under active development ever since.
+This includes the switch to a dry-mass vertical coordinate \cite{LetAl2018JAMES} and incorporation of an accelerated multi-tracer transport scheme \cite{LTOUNGK2017MWR}, made available in CESM2.
+This paper documents several additional enhancements to the SE dycore as part of the release of CESM2.2.
+These include three new VR configurations (Figure~\ref{fig:vr-grids}), two Arctic meshes and a Contiguous United-States mesh ($\tt CONUS$; featured in \citeA{GetAl2017JAMES,MUSCIA}).
+While there are many published studies using VR in CESM \cite<e.g., >{ZetAl2014JC,RHUZ2016JAMC,bambach}, these studies either used development code or collaborated closely with model developers.
+CESM2.2 is the first code release that contains out of the box VR functionality in CESM.
+
+This study compares the representation of high-latitude regions using the SE and FV dycores in CESM2.2 (see description below), as these two dycores treat high latitudes (i.e., the pole problem) in different ways.
+Section~\ref{sec:methods} documents the grids, dycores, and physical parameterizations used in this study, and also describes the experiments, datasets, and evaluation methods.
+Section~\ref{sec:results} analyzes the results of the experiments, and Section~\ref{sec:conclusions} provides a general discussion and conclusions.
+
+%{\color{blue} (WHL: Would it be worth mentioning that quasi-uniform grids also have benefits in polar regions, e.g.
+by removing elongated grid cells?)}{\color{red} (ARH: I don't think it's worth getting into that here.
+You'll see later in the manuscript that b/c the cells are narrower in the E-W direction, more cells can fit into the big ablation zone on the western margin.)}{\color{purple}(PHL: another advantage of lat-lon grids is that the coordinate lines are orthogonal!)}
+
+%{\color{blue} (WHL: This raises the question: How do lat-lon grids better resolve polar regions at a given nominal global resolution, if the filtering makes them effectively coarser?
 Does the damping degrade some processes more than others?)}{\color{red} (ARH - The phrase "approximately the same" doing a lot of heavy lifting here b/c I frankly don't know.
@@
 
-{\color{purple}PHL: These are good points!
+%{\color{purple}PHL: These are good points!
 I think it is useful to separate dynamics from physics in this discussion.
@@
 
-An alternative approach is to use unstructured grids, which permit quasi-uniform grid spacing globally and eliminates the pole-problem (e.g., Figure~\ref{fig:uni-grids}c).
-{\color{blue} (WHL: Add 1 or 2 refs)}
-This grid flexibility also allows variable-resolution or regional grid refinement (e.g., Figure~\ref{fig:vr-grids}, hereafter abbreviated as \textit{VR}).
-In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).
-However, this refinement comes at the cost of a smaller CFL-limited time step in the refined region.
-(The CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.) Unstructured grids also scale more efficiently on parallel systems than lat-lon grids, another advantage as computing power increases over time.
-
-The meteorology and climate of the Arctic are characterized by a range of processes and scales that are difficult to represent in GCMs \cite{BETAL2001MWR,SG2017MWR,VETAL2018TC}.
-For example, while synoptic scale storms are generally well represented at typical GCM resolutions of 1 to 2 degrees \cite{JW2006QJR,IPCCAR5}, mesoscale polar lows are not well resolved at these resolutions.
-These mesoscale systems are prevalent during winter and produce gale-force winds that can induce large heat and moisture fluxes through the underlying sea-ice/ocean interface.
-The Arctic also contains the Greenland Ice Sheet (GrIS), which blankets the worlds's largest island.
-Many of the processes that control the GrIS annual surface mass balance (SMB; the integrated sum of precipitation and runoff) are only partially resolved at typical GCM resolutions.
-For example, GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation.
-GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
-GCMs struggle to resolve the magnitude and extent of these features \cite{P2010CC,VETAL2018TC}, which can lead to unrealistic ice sheet growth in models with an interactive ice sheet component \cite<e.g.,>{LETAL2020JAMES}.
-{\color{blue} (WHL: I suggest adding some language explaining why advance is more prevalent than retreat.
-E.g.:  ``… of these features.
-For example, CESM2 typically underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component.''  Could cite Lofverstrom et al.
-(2020).)}{\color{red} (ARH: I may delete the reference to this anecdote of advancing in northern GrIS altogether b/c I've been unable to show whether it's related to resolution.
-I don't have confidence defining a direction of the SMB bias due to resolution because resolution inflates both source and sink terms; the net effect can go either way.)}
-
-This study compares the representation of high-latitude regions using the spectral-element and finite-volume dycores in CESM2.2, as these two dycores treat high latitudes (i.e., the pole problem) in different ways.
-It also serves to document the Arctic-refined grids developed by the authors for CESM2.2.
-Section~\ref{sec:methods} documents the grids, dycores, and physical parameterizations used in this study, and also describes the experiments, datasets, and evaluation methods.
-Section~\ref{sec:results} analyzes the results of the experiments, and Section~\ref{sec:conclusions} provides a general discussion and conclusions.
-
 \begin{figure}[t]
@@
 \end{center}
-\caption{Computational grids for the uniform $1^{\circ}-2^{\circ}$ grids in this study.}
+\caption{Computational grids for the uniform $1^{\circ}-2^{\circ}$ grids in this study.
+Grids names after Table~\ref{tbl:table1}, (a) $\tt f19$, (b) $\tt f09$, (c) $\tt ne30pg2$ and (d) $\tt ne30pg3$.}
 \label{fig:uni-grids}
@@
 \end{center}
-\caption{Spectral-element grid for the VR Arctic grids in this study. Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
+\caption{Spectral-element grid for the VR grids in this study, (a)$\tt Arctic$, (b) $\tt Arctic-GrIS$ and (a) $\tt CONUS$.
+Note that this is not the computational grid; each element has $3\times3$ independent grid points.
+{\color{red}{add CONUS grid}}}
 \label{fig:vr-grids}
@@
 
-The atmospheric component of CESM2.2, the Community Atmosphere Model, version 6.3 \cite<CAM;>{CAM63}, supports several different atmospheric dynamical cores.
+The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM;>{Gettelman2019GRL}, supports several different atmospheric dynamical cores.
 These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-The EUL dycore is the oldest and least supported dycore in CAM.
-FV3 is the newest dycore, but was not fully incorporated when this work begin.
-This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores.
+This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores {\color{green}{Andrew - maybe not mention EUL or FV3?}}.
 
@@
 A polar filter damps computational instability due to the convergence of meridians, allowing a longer time step.
-It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the poleward direction \cite{ST1995GEOS}.
+It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the meridional direction \cite{ST1995GEOS}.
+The form of the filter is designed to slow down the propagation of large zonal wave-numbers to satisfy the CFL condition of the shortest resolved wave at some reference latitude.
 
@@
 The computational domain is a cubed-sphere grid tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
-Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre (GLL) quadrature points.
-This results in 16 GLL nodal points per element, with 12 of the points lying on the (shared) element boundary.
+Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre quadrature points.
+This results in 16 nodal points per element, with 12 of the points lying on the (shared) element boundary.
 Communication between elements uses the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal values and produce a continuous global basis set.
@@
 
-For spectral-element grids with quasi-uniform grid spacing, we use a variant in which tracer advection is computed using the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}.
-{\color{blue} (WHL: I'm not sure what CSLAM is a variant of, since I don't see a tracer transport scheme described above.
-How is tracer transport handled on VR grids?
-If it's not too much detail, why does CSLAM not work on VR grids?)} {\color{red} (ARH: fair point.
-the spectral-element section is lacking in general, so keeping this here as a reminder to explain the spectral-element method is used for tracer transport in ``vanilla" cam-se, and we only support vanilla var-res.
-PHL is planning on adding a description of enhancements to cam-se introduced in CESM2.2, since this turns out to be the only paper documenting CESM2.2.)}
+For SE grids with quasi-uniform grid spacing, The SE tracer transport scheme is replaced with the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}.
 CSLAM has improved tracer property preservation and accelerated multi-tracer transport.
 It uses a separate grid from the spectral-element dynamics, dividing each element into $3\times3$ control volumes with quasi-equal area.
-The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated at the GLL nodal points \cite{HL2018MWR}.
+The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated Gauss-Lobatto-Legendre points \cite{HL2018MWR}.
+{\color{purplse}{Rene/Andrew - maybe elaborate further on the advantages of CSLAM?}}
 
@@
 
-We evaluate six grids in this study (Table~\ref{tbl:table1}).
-The FV dycore is run with $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $f09$ and $f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
-We also run the $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid, referred to as $ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
-We run an additional $1^{\circ}$ CAM-SE-CSLAM grid with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>{HETAL2019JAMES}.
+We evaluate model simulations on six different grids in this study (Table~\ref{tbl:table1}).
+The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
+We also run the $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
+We run an additional $1^{\circ}$ CAM-SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d; note CSLAM is still run on the $3\times3$ control volume grid;>{HETAL2019JAMES}.
 
@@
 The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
-The $Arctic$ grid is a $1^{\circ}$ grid with $\frac{1}{4}^{\circ}$ regional refinement over the broader Arctic region.
-The $Arctic-GrIS$ grid is identical to the $Arctic$ grid, but with an additional patch covering the island of Greenland with $\frac{1}{8}^{\circ}$ resolution.
+The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region.
+The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution.
 
@@
    \hline
-   $f19$ & FV & 278 & - &1800 \\
-   $f09$ & FV & 139 & - &1800 \\
-   $ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
-   $ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
-   $ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 \\
-   $Arctic$ & SE & 111 & 28 & 450 \\
-   $Arctic-GrIS$ & SE & 111 & 14 & 225 \\
+   $\tt f19$ & FV & 278 & - &1800 \\
+   $\tt f09$ & FV & 139 & - &1800 \\
+   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
+   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
+   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 \\
+   $\tt Arctic$ & SE & 111 & 28 & 450 \\
+   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 \\
  \hline
@@
 FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.
-We use the $ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.}
+We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.{\color{purple}{Rene - maybe include model cost?}}}
  \label{tbl:table1}
@@
 
-The physics time step should depend on grid resolution.
+The physics time step depends on the grid resolution.
 Increased horizontal resolution permits faster vertical velocities that reduce characteristic time scales, so the physics time step is reduced to avoid large time truncation errors \cite{HR2018JAMES}.
-The $ARCTIC$ and $Arctic-GrIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time step relative to the default 1800 s time step for the standard $1^{\circ}$ and $2^{\circ}$ grids (Table~\ref{tbl:table1}).
+The $\tt Arctic$ and $\tt Arctic-GrIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time step relative to the default 1800 s time step for the standard $1^{\circ}$ and $2^{\circ}$ grids (Table~\ref{tbl:table1}).
 
 All grids and dycores in this study use 32 levels in the vertical, with a model top of about 1~hPa or 40~km.
-However, any grid or dycore can in principle be run with a higher top or finer vertical resolution.
+However, note that any grid or dycore can in principle be run with a higher model top or finer vertical resolution.
 
@@
 
-All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{CAM63}.
-The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer (PBL), shallow convection, and cloud macrophysics scheme.
-CLUBB is coupled with the MG2 microphysics scheme \cite{MG2}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
+All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{Gettelman2019GRL}.
+The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer, shallow convection, and cloud macrophysics scheme.
+CLUBB is coupled with the MG2 microphysics scheme \cite{MG2,Morrison2015}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
 Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
-PBL form drag is modeled after \citeA{BBW2004QJRMS}.
-Orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale.
-
-Initial simulations with the $ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore.
-All runs with the SE dycore therefore have two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance.
-These are CLUBB's $gamma$ parameter, reduced from 0.308 to 0.270, and $c14$, reduced from 2.2 to 1.6.
-Briefly, $gamma$ scales the width of the sub-grid distribution of vertical velocity, and $c14$ controls the strength of the damping term in the equation for the horizontal component of turbulent kinetic energy.
+Boundary layer form drag is modeled after \citeA{BBW2004QJRMS}, and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale (the ridge orientation is derived from a high-resolution, global topography dataset \cite{Danielson_Gesch_2011_GMTED2010}).
+
+Initial simulations with the $\tt ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore.
+The SE dycore in CESM2.2 therefore has two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance.
+We reduced the width of the sub-grid distribution of vertical velocity ($\tt clubb_gamma$ is reduced from 0.308 to 0.270) and also reduced the strength of the damping for horizonatal component of turbulent energy ($\tt clubb_c14$ is reduced from 2.2 to 1.6) to increase cloudiness.
 For a thorough explanation of how CLUBB parameters impact the simulated climate, see \cite{GETAL2015JAMES}.
@@
 
-We run all grids and dycores using an identical transient 1979-1998 Atmospheric Model Inter-comparison Project configuration (AMIP), with prescribed monthly SST and sea ice following \citeA{CESMSST}.
+All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
 In CESM terminology, AMIP simuations use the $FHIST$ computational set and run out of the box in CESM2.2.
-
-All grids and dycores simulate the GrIS surface mass balance (SMB).
-The SMB is the sum of the mass accumulation term,  i.e., precipitation, and the mass sink term, ablation.
+{\color{grey}{Marcus - this is the only part that doesn't talk about CLM/CISM coupling.
+Move somewhere else and change section name?}}
+
+All grids and dycores simulate the GrIS Surface Mass Balance (SMB).
+The SMB is the sum of mass accumulation of precipitation and mass loss from ablation.
 Ablation is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation runs off the ice sheet; rain can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the snow mass.
+Not all liquid precipitation runs off the ice sheet; liquid runoff (which is composed of both rain or ice/snow melt) can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass.
 These processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
@@
 The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean.
-Precipitation from CAM is repartitioned into solid or liquid based on the surface temperature of the EC; precipitation falls as snow for $T < -2^\circ$~C, as rain for $T > 0^\circ$~C,  and as a rain/snow mix between $-2^\circ$~C and $0^\circ$~C.
+Precipitation from CAM is repartitioned into solid or liquid based on the surface temperature of the EC; precipitation falls as snow for temperatures between $T < -2^\circ$~C, as rain for $T > 0^\circ$~C,  and as a linear combination of rain and snow for temperatures between $-2^\circ$~C and $0^\circ$~C.
 Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent.
 Any snow above the 10-m cap is routed to the ocean as solid runoff.
-Refreezing of liquid within the snowpack is an additional source of ice.
+Refreezing of liquid water within the snowpack is an additional source of ice.
 Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
-For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES}.
+For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{Muntjewerf_EA_2021_coupling,CISM1,SETAL2019TC,KETAL2020JAMES}.
+{\color{grey}{Marcus - mention you're not running CISM.}}
 
@@
 That is, snow accumulation above the 10-m cap contributes a positive SMB, and surface ice melting (after melting of the overlying snow) yields a negative SMB.
-Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over a 20-year $Arctic$ $FHIST$ run for about 500 years.
-The uniform-resolution grids are initialized with the SMB from an existing $f09$ spun-up initial condition.
+Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over a 20-year $\tt Arctic$ $FHIST$ run for about 500 years.
+The uniform-resolution grids are initialized with the SMB from an existing $\tt f09$ spun-up initial condition.
 
@@
 
-We use several observational datasets (Table~\ref{tbl:table2}) to gauge the performance of the simulations.
-Several products are near-global gridded datasets commonly used to evaluate GCMs.
+We use several observational datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations.
 We gathered SMB datasets from several sources.
-RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA40, ERA interim and ERA5 renalysis products at the domain boundaries.
-The RACMO simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite{NETAL2015TC,NETAL2019SCIENCE}.
+RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA40, ERA interim, and ERA5 renalysis products at the domain boundaries.
+The Regional Atmospheric Climate Model (\textit{RACMO}) simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite{NETAL2015TC,NETAL2019SCIENCE}.
 
 In-situ SMB (snow pit and ice cores) and radar accumulation datasets (e.g., IceBridge) are maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}.
-However, these point-wise measurements are difficult to compare to model output spanning several different grids, especially since the SMB from each EC is not available from the model output.
+However, these point-wise measurements are difficult to compare to model output spanning several different grids, especially since the SMB from each elevation class is not available from the model output.
 We used a nearest-neighbor technique for an initial analysis, which showed that the model biases are similar to those computed using the RACMO datasets.
 Because of the uncertainty of comparing gridded fields to point-wise measurements, and the lack of information added with regard to model biases, we omitted these datasets from our analysis.
+{\color{purple}{Andrew - maybe do show them, briefly?}}
 
@@
    \hline
-   ERA5 & 1979-1998 & $\frac{1}{4}^{\circ}$ & \citeA{ERA5} \\
+   ERA5 & 1979-1998 & $1/4^{\circ}$ & \citeA{ERA5} \\
    CERES-EBAF ED4.1 & 2003-2020 & $1^{\circ}$ & \citeA{CERES-EBAF} \\
@@
 
-\subsection{SMB Analysis \color{red}{ARH - this section is under construction}}\label{sec:SMB}
+\subsection{SMB Analysis}\label{sec:SMB}
 
@@
 The area errors in Figure~\ref{fig:grisdx} may not seem large, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
-CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework first-order conservative remapping algorithm \cite<ESMF;>{ESMF}.
+CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite<ESMF;>{ESMF}.
 The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
@@
 
-We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $f19$ and $ne30pg2$ grids, and integrating over these low-resolution ice masks.
+We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $\tt f19$ and $\tt ne30pg2$ grids, and integrating over these low-resolution ice masks.
 The use of low-resolution common ice masks is a conservative decision and is largely because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not reliable in the downscaling direction, i.e., mapping to a higher-resolution grid than the source grid.
 We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
-Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $f19$ and $ne30pg2$ masks.
+Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $\tt f19$ and $\tt ne30pg2$ masks.
 Thus, we evaluate an integrated quantity on a given grid up to four times to estimate the uncertainty due to differences in grid type and remapping algorithms.
@@
 While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
-Method~\ref{lbl:label2}, in contrast, does not map values outside the source ice mask to the common ice mask.
-This tends to be dissipative, since mapping to the low-resolution grid will average a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
-However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.
+This underestimation is systematic in method~\ref{lbl:label2}; all variables are exclusive to the ice mask.
+Method~\ref{lbl:label2} is more dissipative, since mapping to the low-resolution grid will average a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
+However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.
 %The dilution of the these patch level fields caused by mapping to the common ice mask are then inflated due to integrating over the larger ice mask.
@@
 The different error characteristics of the two methods are used to diversify the ensemble.
-Each of the four regridding combinations (with conservative and high-order remapping to the $f09$ and $ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
+Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
 Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
@@
 Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between uniform-resolution grids and dycores.
-The $f09$ grid is warmer than the $f19$ grid, primarily in the mid-to-high latitudes throughout the depth of the troposphere.
+The $\tt f09$ grid is warmer than the $\tt f19$ grid, primarily in the mid-to-high latitudes throughout the depth of the troposphere.
 This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}.
 \citeA{HK2020QJRMS} have shown that this occurs in CAM due to greater resolved vertical velocities which, in turn, drive more condensational heating in the CLUBB macrophyiscs.
-The right columns in Figure~\ref{fig:dT-lores} support this interpretation, showing an increase in the climatological CLUBB heating in low ({\color{blue}{looks more like subtropics}}) and mid-latitudes on the $f09$ grid.
+The right columns in Figure~\ref{fig:dT-lores} support this interpretation, showing an increase in the climatological CLUBB heating in low ({\color{blue}{looks more like subtropics}}) and mid-latitudes on the $\tt f09$ grid.
 
@@
 
-As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so the $ne30pg3$ troposphere is modestly warmer than $f09$ (Figure~\ref{fig:dT-lores}).
-The stratosphere responds differently, with $ne30pg3$ much cooler than $f09$ in the mid-to-high latitudes.
-Figure~\ref{fig:dT-lores} also shows small temperature differences between $ne30pg3$ and $ne30pg2$, with  $ne30pg3$ slightly warmer near the tropopause at high latitudes.
+As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so the $\tt ne30pg3$ troposphere is modestly warmer than $\tt f09$ (Figure~\ref{fig:dT-lores}).
+The stratosphere responds differently, with $\tt ne30pg3$ much cooler than $\tt f09$ in the mid-to-high latitudes.
+Figure~\ref{fig:dT-lores} also shows small temperature differences between $\tt ne30pg3$ and $\tt ne30pg2$, with  $\tt ne30pg3$ slightly warmer near the tropopause at high latitudes.
 This is consistent with the similar climates found for these two grids by \citeA{HETAL2019JAMES}.
@@
 Comparing the VR grids to the uniform-resolution grids is complicated because we simultaneously increase the resolution and reduce the physics time-step, both of which impact the solution \cite{W2008TELLUS}.
-We therefore run an additional $ne30pg3$ simulation with the shorter physics time step used in the $Arctic$ grid (450~s), referred to as $ne30pg3^{*}$.
-Figure~\ref{fig:dT-hires} shows the difference between $ne30pg3^{*}$ and $ne30pg3$ for climatological summer temperatures in zonal-mean height space.
+We therefore run an additional $\tt ne30pg3$ simulation with the shorter physics time step used in the $\tt Arctic$ grid (450~s), referred to as $\tt ne30pg3^{*}$.
+Figure~\ref{fig:dT-hires}a shows the difference between $\tt ne30pg3^{*}$ and $\tt ne30pg3$ for climatological summer temperatures in zonal-mean height space.
 Much of the troposphere is warmer with the reduced time step, and the mechanism is similar in that the shorter time step increases condensational heating by CLUBB.
-Figure~\ref{fig:dT-hires} shows the difference in climatological summer temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid.
+Figure~\ref{fig:dT-hires}b shows the difference in climatological summer temperature between the $\tt Arctic$ grid and the $\tt ne30pg3^{*}$ grid.
 With the same physics time step on each grid, the greater condensational heating and warmer temperatures are confined to the refined Arctic region.
 
-Figure~\ref{fig:dT-hires} shows that the $Arctic-GrIS$ grid is much warmer than the $Arctic$ grid in the Arctic summer. This may be due, in part, to the shorter physics time step,
-%; a similar physics-time-step sensitivity in the $ne30pg3^{*}$ comparison is evident in the CLUBB tendencies in the $ARCTCGRIS$ grid.
+Figure~\ref{fig:dT-hires}c shows that the $\tt Arctic-GrIS$ grid is much warmer than the $\tt Arctic$ grid in the Arctic summer. This may be due, in part, to the shorter physics time step,
+%; a similar physics-time-step sensitivity in the $\tt ne30pg3^{*}$ comparison is evident in the CLUBB tendencies in the $ARCTCGRIS$ grid.
 but the temperature response is too large to be explained by the CLUBB changes alone.
 This summer warming appears to be a result of variations in the stationary wave pattern, with anomalous southerly winds to the west of Greenland (not shown).
-This dynamic response is interesting, because other than the physics time step, the only difference between the $Arctic-GrIS$ and $Arctic$ runs is the doubling of resolution over Greenland.
+This dynamic response is interesting, because other than the physics time step, the only difference between the $\tt Arctic-GrIS$ and $\tt Arctic$ runs is the doubling of resolution over Greenland.
 This behavior will be explored further in a future study.
@@
 It is computed from the 500--1000 hPa geopotential thickness, solving for the layer-mean virtual temperature using the hypsometric equation ({\color{blue}{Might be good to write this down, for those of us who studied atmospheric science a really long time ago}}).
-The results are consistent with the zonal mean height plots; increasing resolution from $f19$ to $f09$ warms the climate, and the $1^{\circ}$ SE grids are warmer than the FV grids.
+The results are consistent with the zonal mean height plots; increasing resolution from $\tt f19$ to $\tt f09$ warms the climate, and the $1^{\circ}$ SE grids are warmer than the FV grids.
 The FV summer temperatures are persistently colder than ERA5, whereas the $1^{\circ}$ SE grids are not as cold, and are actually warmer than ERA5 at high-latitudes, north of $75^{\circ}$.
@@
 
-The $Arctic$ grid has summer temperatures similar to the $1^{\circ}$ SE grids, but is slightly warmer over northern Eurasia and the North Pole.
+The $\tt Arctic$ grid has summer temperatures similar to the $1^{\circ}$ SE grids, but is slightly warmer over northern Eurasia and the North Pole.
 An anomalous cooling patch forms to the west of Greenland, centered over Baffin Island.
-The $Arctic-GrIS$ grid is warmer than the $Arctic$ grid over most of the Arctic, but with a similar spatial pattern of summer temperature bias.
+The $\tt Arctic-GrIS$ grid is warmer than the $\tt Arctic$ grid over most of the Arctic, but with a similar spatial pattern of summer temperature bias.
 
@@
 There is also a halo of positive bias (clouds not reflective enough) around the ocean perimeter of Greenland.
-The $Arctic$ grid has much smaller cloud forcing biases over the Arctic land masses, but is still too reflective over Alaska, the Canadian Archipelago, and parts of Eurasia.
-Compared to the $Arctic$ grid, the $ARCTCGRIS$ grid vastly reduces the cloud forcing bias over Eurasia, and also improves the bias over North America.
+The $\tt Arctic$ grid has much smaller cloud forcing biases over the Arctic land masses, but is still too reflective over Alaska, the Canadian Archipelago, and parts of Eurasia.
+Compared to the $\tt Arctic$ grid, the $\tt Arctic-GrIS$ grid vastly reduces the cloud forcing bias over Eurasia, and also improves the bias over North America.
 In both VR grids, the halo of positive forcing bias around the perimeter of Greenland is absent.
@@
 ({\color{blue}{This ends abruptly.
-Anything else to add?}})
+Anything else to add?}} {\color{purple}{Causation I agree is hard to prove due to energy transport, but in summer cloud fraction is probably playing a role.
+The likely other thing that could be playing a role is surface albedo (absorbed solar).
+Could note that.}})
 
@@
 Lower troposphere layer mean virtual temperature is derived from the 1000 hPa - 500h Pa geopotential thickness, using the hypsometric equation.
-Differences are computed after mapping the ERA5 data to the finite-volume grids since the geopotential field is only available on the output tapes in the spectral-element runs that have been interpolated to the $f09$ grid, inline.}
+Differences are computed after mapping the ERA5 data to the finite-volume grids since the geopotential field is only available on the output tapes in the spectral-element runs that have been interpolated to the $\tt f09$ grid, inline.}
 \label{fig:dThyps}
@@
 \end{center}
-\caption{1979-1998 northern hemisphere summer shortwave cloud forcing bias, relative to the CERES-EBAF gridded dataset.
+\caption{1979-1998 Northern Hemisphere summer shortwave cloud forcing bias, relative to the CERES-EBAF gridded dataset.
 Differences are computed after mapping all model output to the $1^{\circ}$ CERES-EBAF grid.}
@@
 The CALIPSO-GOCCP differences are computed after mapping all model output to the $1^{\circ}$ grid, whereas the RACMO differences are computed after mapping the RACMO dataset to the individual model grids.
-Note that the averaging period for the CALIPSO-GOCCP panels, 2006-2017, is different than the model averaging period.
+Note that the averaging period for the CALIPSO-GOCCP panels (2006-2017) is different than the model averaging period.
 \color{red}{ARH - still trying to fix the layout of this figure so the label bars aren't on top of the panels}.}
@@
 The accuracy of the simulated SMB is expected to be sensitive to grid resolution.
-Figure~\ref{fig:prect}b shows the average grid spacing over the Greenland Ice Sheet (GrIS) in all six grids in this study.
-The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $Arctic-GrIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO2.3 grid.
-The $ne30pg3$ grid has an average $\Delta x=111.2~km$, substantially coarser than the $f09$ grid, with an average $\Delta x=60~km$.
-Although $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
-The $Arctic$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
-The $Arctic-GrIS$ grid is about twice as expensive as the $Arctic$ grid.
-({\color{blue}{Wondering if this paragraph would go better in an earlier section.}})
+Figure~\ref{fig:gridsdx}b shows the average grid spacing over the Greenland Ice Sheet (GrIS) in all six grids in this study.
+The $\tt ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO2.3 grid.
+The $\tt ne30pg3$ grid has an average $\Delta x=111.2~km$, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60~km$.
+Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
+The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
+The $\tt Arctic-GrIS$ grid is about twice as expensive as the $\tt Arctic$ grid.
+({\color{blue}{Wondering if this paragraph would go better in an earlier section.
+Rene agrees}})
 
@@
 The coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
-The $Arctic$ grid reduces this bias substantially, and the $Arctic-GrIS$ grid reduces it further.
+The $\tt Arctic$ grid reduces this bias substantially, and the $\tt Arctic-GrIS$ grid reduces it further.
 This suggests that the southern dome bias arises from inadequate horizontal resolution, consistent with the original GrIS VR experiments in \citeA{VETAL2018TC}.
@@
    \hline
-   $f19$ & 882.5 (913.5) & -440.3 (-546.5) & -283.7 (-284.3) & -36.6 (-37.5) & 562.2 (591.7) \\
-   $f09$ & 874.8 (882.1) & -418.4 (-482.3) & -255.0 (-212.3) & -38.1 (-37.4) & 581.7 (632.4) \\
-   $ne30pg2$ & 1000. (973.4) & -549.4 (-647.3) & -383.9 (-347.0) & -33.4 (-32.1) & 582.7 (594.3) \\
-   $ne30pg3$ & 934.9 (909.3) & -568.8 (-686.7) & -356.2 (-330.1) & -34.4 (-32.6) & 544.3 (546.6) \\
-   $Arctic$ & 795.9 (818.6) & -367.3 (-436.8) & -208.9 (-194.2) & -44.1 (-43.9) & 542.9 (580.5) \\
-   $Arctic-GrIS$ & 708.7 (747.3) & -471.6 (-610.4) & -261.1 (-307.8) & -50.7 (-51.8) & 396.9 (387.7) \\
+   $\tt f19$ & 882.5 (913.5) & -440.3 (-546.5) & -283.7 (-284.3) & -36.6 (-37.5) & 562.2 (591.7) \\
+   $\tt f09$ & 874.8 (882.1) & -418.4 (-482.3) & -255.0 (-212.3) & -38.1 (-37.4) & 581.7 (632.4) \\
+   $\tt ne30pg2$ & 1000. (973.4) & -549.4 (-647.3) & -383.9 (-347.0) & -33.4 (-32.1) & 582.7 (594.3) \\
+   $\tt ne30pg3$ & 934.9 (909.3) & -568.8 (-686.7) & -356.2 (-330.1) & -34.4 (-32.6) & 544.3 (546.6) \\
+   $\tt Arctic$ & 795.9 (818.6) & -367.3 (-436.8) & -208.9 (-194.2) & -44.1 (-43.9) & 542.9 (580.5) \\
+   $\tt Arctic-GrIS$ & 708.7 (747.3) & -471.6 (-610.4) & -261.1 (-307.8) & -50.7 (-51.8) & 396.9 (387.7) \\
  \hline
@@
   \caption{1979-1998 surface mass balance of the Greenland Ice Sheet in Gt/yr.
-Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.}
+Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.
+{\color{red}{ARH - these numbers need updating ...
+they are only for method 1.}}}
  \label{tbl:table3}
@@
 
-Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO.
+Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO {\color{purple}{Andrew - do you want to talk about the values at all?}}.
 The CESM values are averages over the ensemble of common ice masks and regridding methods described in section~\ref{sec:SMB}, and the RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}) using the same common-ice-mask approach.
@@
 The uniform $1^{\circ}-2^{\circ}$ grids have positive precipitation biases in the interior, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
-The $f19$ and $f09$ grids perform similarly, with +110 Gt/yr bias, whereas $ne30pg3$ is biased by about +165 Gt/yr and $ne30pg2$ by +230 Gt/yr.
+The $\tt f19$ and $\tt f09$ grids perform similarly, with +110 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +165 Gt/yr and $\tt ne30pg2$ by +230 Gt/yr.
 The larger biases on the uniform-resolution SE grids relative to the FV grids are consistent with the coarser GrIS resolution on the SE grids (Figure~\ref{fig:grisdx}).
 
-The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $Arctic$ grid simulates the most realistic melt rates, with the other grids having more melt than RACMO.
-The $Arctic-GrIS$ grid overpredicts melting by about 125 Gt/yr.
-This is likely due to an anomalously warm lower troposphere during the summer, relative to the $Arctic$ run (Figure~\ref{fig:dThyps}).
-The $f19$ and $f09$ melting rates are improved over $Arctic-GrIS$, overestimating melt by only 70--90 Gt/yr.
+The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $\tt Arctic$ grid simulates the most realistic melt rates, with the other grids having more melt than RACMO.
+The $\tt Arctic-GrIS$ grid overpredicts melting by about 125 Gt/yr.
+This is likely due to an anomalously warm lower troposphere during the summer, relative to the $\tt Arctic$ run (Figure~\ref{fig:dThyps}).
+The $\tt f19$ and $\tt f09$ melting rates are improved over $\tt Arctic-GrIS$, overestimating melt by only 70--90 Gt/yr.
 The SE grids have the largest positive melt bias, between 200--220 Gt/yr.
@@
 To illustrate the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \citeA{RM2012GRL}.
-The uncertainty due to differences in basin area is larger than for GrIS-wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which are shown in the $f19$ and $ne30pg2$ panels of Figure~\ref{fig:melt}.
+The uncertainty due to differences in basin area is larger than for GrIS-wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which are shown in the $\tt f19$ and $\tt ne30pg2$ panels of Figure~\ref{fig:melt}.
 Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the most accumulation.
 In all basins, accumulations drops monotonically with increasing grid resolution, with some exceptions.
-The $Arctic-GrIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to the RACMO precipitation in the large southwest basin.
+The $\tt Arctic-GrIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to the RACMO precipitation in the large southwest basin.
 
@@
 The \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid.
-Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $f19$ and $ne30pg2$ panels, in computing the basin-scale integrals.
-Blue lines in the $f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
+Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $\tt f19$ and $\tt ne30pg2$ panels, in computing the basin-scale integrals.
+Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
 \label{fig:melt}
@@
 The uniform-resolution SE grids have the largest positive biases in all basins.
-The $Arctic-GrIS$ grid is a close second, while the FV grids have systematically smaller melt-rates.
-The ``second-place" standing of $Arctic-GrIS$ is somewhat unexpected, as this grid has the warmest lower-troposphere summer temperatures (Figure~\ref{fig:dThyps}) and greatest incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the uniform-resolution SE grids.
+The $\tt Arctic-GrIS$ grid is a close second, while the FV grids have systematically smaller melt-rates.
+The ``second-place" standing of $\tt Arctic-GrIS$ is somewhat unexpected, as this grid has the warmest lower-troposphere summer temperatures (Figure~\ref{fig:dThyps}) and greatest incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the uniform-resolution SE grids.
 
@@
 Figure~\ref{fig:basin}c shows the basin-integrated PDD melt estimate.
-In the large southeast and southwest basins (and all the other western basins), the $ne30pg3$ grid has larger PDD-based melt than the $ARCTCGRIS$ grid.
-The FV grids also have large PDD-based melt in the southwest basin, relative to $Arctic-GrIS$.
+In the large southeast and southwest basins (and all the other western basins), the $\tt ne30pg3$ grid has larger PDD-based melt than the $ARCTCGRIS$ grid.
+The FV grids also have large PDD-based melt in the southwest basin, relative to $\tt Arctic-GrIS$.
 The PDD plots indicate that the near-surface temperatures which contribute to melt are not well approximated by the summer lower-troposphere temperatures in Figure~\ref{fig:dThyps}.
@@
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a few grid cells populating the transect.
-The $f09$ grid is a bit of an exception for the K-transect, with grid cells becoming narrow in the meridional direction at high latitudes.
+The $\tt f09$ grid is a bit of an exception for the K-transect, with grid cells becoming narrow in the meridional direction at high latitudes.
 The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of the GrIS margins.
@@
 Rather, the smoothing of the raw topography, necessary to prevent the model from instigating grid-scale modes, flattens the ice sheet, causing the lower-elevation ablation zones to extend beyond the true ice sheet margin, where they lie above the actual ice surface.
-The $f19$ grid has the smoothest topography since its dynamics are coarsest (whereas $f09$, $ne30pg2$ and $ne30pg3$ use identical smoothing), and has the flattest ice sheet.
+The $\tt f19$ grid has the smoothest topography since its dynamics are coarsest (whereas $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ use identical smoothing), and has the flattest ice sheet.
 This suggests that if anything, coarser models will tend to elevate the ablation zones and depress melt rates.
@@
 
+{\color{green}{Rene - is the melt map consistent with the bin figure?
+Smallest errors are interior points which correspond to glacier_fraction bin of 1, which the largest errors.}}{\color{blue}{ARH - I can check.
+The bin figure is fractional change so its a different metric, which could explain the apparent inconsistency.}}
+
 \begin{figure}[t]
@@
 Synoptic storms are tracked using TempestExtremes atmospheric feature detection software \cite{UETAL2021}.
-As the $Arctic$ grid contains $\frac{1}{4}^{\circ}$ refinement north of about $45^{\circ}$ latitude, the storm tracker is applied to this region for the $Arctic$ and $ne30pg3$ runs to identify differences in storm characteristics due to horizontal resolution.
+As the $\tt Arctic$ grid contains $1/4^{\circ}$ refinement north of about $45^{\circ}$ latitude, the storm tracker is applied to this region for the $\tt Arctic$ and $\tt ne30pg3$ runs to identify differences in storm characteristics due to horizontal resolution.
 The composite mean precipitation maps are similar between the two grids, and exhibit the iconic comma structure of synoptic cyclones (not shown).
@@
 The PDFs are evaluated on an identical composite grid for all runs, and so storm statistics are not impacted by differences in output resolution.
-The $Arctic$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
-This is primarily due to increased resolution and not the reduced physics times-step; the $ne30pg3^{*}$ run only marginally increases the extreme precipitation rates compared with $ne30pg3$ (Figure~\ref{fig:comp-pdf}).
-
-The extreme precipitation rates in the $Arctic$ run are closer than $ne30pg3$ to the ERA5 reanalysis (Figure~\ref{fig:comp-pdf}).
-It is difficult to know how much the extreme precipitation rates in ERA5 are constrained by data assimilation, or whether these precipitation rates are due to using a similar $\frac{1}{4}^{\circ}$ model as the $Arctic$ grid.
-However, it is well documented that $\frac{1}{4}^{\circ}$ models are more skillful at simulating extreme events \cite{BetAl2013JC,OETAL2016JAMES}.
+The $\tt Arctic$ run has larger extreme precipitation rates compared to $\tt ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
+This is primarily due to increased resolution and not the reduced physics times-step; the $\tt ne30pg3^{*}$ run only marginally increases the extreme precipitation rates compared with $\tt ne30pg3$ (Figure~\ref{fig:comp-pdf}).
+
+The extreme precipitation rates in the $\tt Arctic$ run are closer than $\tt ne30pg3$ to the ERA5 reanalysis (Figure~\ref{fig:comp-pdf}).
+It is difficult to know how much the extreme precipitation rates in ERA5 are constrained by data assimilation, or whether these precipitation rates are due to using a similar $1/4^{\circ}$ model as the $\tt Arctic$ grid.
+However, it is well documented that $1/4^{\circ}$ models are more skillful at simulating extreme events \cite{BetAl2013JC,OETAL2016JAMES}.
 This is an additional benefit of the VR grids.
@@
 \end{center}
-\caption{PDFs of the total precipitation rate associated with tracked storms, by month, in the $ne30pg3$, $ne30pg3^{*}$ and $Arctic$ runs, and compared with the ERA5 dataset.}
+\caption{PDFs of the total precipitation rate associated with tracked storms, by month, in the $\tt ne30pg3$, $\tt ne30pg3^{*}$ and $\tt Arctic$ runs, and compared with the ERA5 dataset.}
 \label{fig:comp-pdf}
@@
 We developed two VR grids and introduced them into CESM2.2 as part of this work.
-Both use the spectral-element dycore; the $Arctic$ grid has $\frac{1}{4}^{\circ}$ refinement over the broader Arctic, whereas the $Arctic-GrIS$ grid is identical except for a $\frac{1}{8}^{\circ}$ patch of refinement over Greenland.
+Both use the spectral-element dycore; the $\tt Arctic$ grid has $1/4^{\circ}$ refinement over the broader Arctic, whereas the $\tt Arctic-GrIS$ grid is identical except for a $1/8^{\circ}$ patch of refinement over Greenland.
 
@@
 The FV grids have colder summers, consistent with their lower melt bias.
-However, the $Arctic-GrIS$ grid has the warmest summer temperatures of all grids, yet it has less melting than the uniform-resolution SE grids.
+However, the $\tt Arctic-GrIS$ grid has the warmest summer temperatures of all grids, yet it has less melting than the uniform-resolution SE grids.
 This suggests that grid resolution is responsible for a large fraction of the melt biases.
@@
 
-The $Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the Greenland SMB, compared to the uniform $1^{\circ}-2^{\circ}$ grids.
-The $Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but its summer temperatures are too warm.
+The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the Greenland SMB, compared to the uniform $1^{\circ}-2^{\circ}$ grids.
+The $\tt Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but its summer temperatures are too warm.
 The $1^{\circ}$ FV model gives a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids.
@@
 Yes, we can recover the fidelity of Arctic simulations using VR grids, but (so far) only at a considerable cost in cpu-hours.
-This points to the need for an intermediate resolution that is more affordable, and/or model development or tuning that reduces the biases on coarse grids.)}
-
-We are  working to develop a configuration of the $Arctic$ grid that is fully-coupled with the CESM ocean and sea ice components and the Community Ice Sheet Model (CISM), to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise.
-A fully-coupled pre-industrial control configuration has already been developed and vetted.
+This points to the need for an intermediate resolution that is more affordable, and/or model development or tuning that reduces the biases on coarse grids.)}{\color{purple}{Andrew - Maybe better to just state that higher resolution is better: 1deg better than 2deg FV, and ne30 is coarser than 1deg at Greenland latitudes so it's worse.
+But VR has higher resolution so SR-VR is better.
+It's all about resolution.}}
+
+We are  working to develop a configuration of the $\tt Arctic$ grid that is fully-coupled with the CESM ocean and sea ice components and the Community Ice Sheet Model (CISM), to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise.
+%A fully-coupled pre-industrial control configuration has already been developed and vetted.
 %, and should be incorporated into CESM in the near future.
@@
 %We are optimistic about the polar science opportunities provided by these new Arctic mesh configurations, and our open to collaborating with the CESM community to assist in anyway we can.
-These new grids and configurations will provide new opportunities for CESM polar science.
+These new grids and configurations will provide new opportunities for CESM polar science and aims to contribute to an improved understanding of the polar environment.
 {\color{blue} (WHL: I replaced the previous last sentence because it seemed too much like an advertisement.
@@
 \end{center}
-\caption{Snapshot of the lowest model level streamlines from the $Arctic-GrIS$ visualization, with color shading denoting the wind magnitude.}
+\caption{Snapshot of the lowest model level streamlines from the $\tt Arctic-GrIS$ visualization, with color shading denoting the wind magnitude.}
 \label{fig:viz}
@@
 \section{Details on spectra-element dynamical core improvements since the CESM2.0 release}
-{\color{red}{PHL working on this!}}\\
-
 Since the CESM2.0 release of the spectral-element dynamical core documented in \citeA{LetAl2018JAMES} some important algorithmic improvements have been implemented and released with CESM2.2.
@@
   T^{(ref)}&=&T_0+T_1 \Pi^{(ref)},\\
-  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}T_0}\right)}
+  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}(T_0+T_1)}\right)}, {\color{blue}{Mark - Note: new definition of T0 means this equation should also be changed}}
 \end{eqnarray}
-\cite{SJ1991QJRMS} where $g$ gravity, $T_0\equiv 288K$, $T_1=\Gamma_0 T_0 c_p^{(d)}/g$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential.
+\cite{SJ1991QJRMS} where $g$ gravity, $T_1=\Gamma_0 T_0 c_p^{(d)}/g\approx 192K$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv 288K-T_0\approx 97K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential.
 The reference Exner function is
@@
 \end{equation}
-where $\kappa=\frac{R_d}{c_p^{(d)}}$ (the reference surface pressure $p_0=1000hPa$ and at each model level the reference pressure $p^{(ref)}$ is computed from $p_s^{(ref)}$ and the standard hybrid coefficients
+where $\kappa=\frac{R^{(d)}}{c_p^{(d)}}$.
+The reference surface pressure $p_0=1000hPa$ and at each model level the reference pressure $p^{(ref)}$ is computed from $p_s^{(ref)}$ and the standard hybrid coefficients
 \begin{equation}
-    p^{(ref)}(\eta) = A(\eta)+B(\eta)p_s^{(ref)},
+    p^{(ref)}(\eta) = A(\eta)p_0+B(\eta)p_s^{(ref)},
 \end{equation}
@@
 These reference profiles are subtracted from the prognostic temperature and pressure-level-thickness states before applying hyperviscosity:
-\begin{eqnarray*}
+\begin{eqnarray}
    \hbox{CESM2.0}&\rightarrow&\hbox{CESM2.2}\\
    && \\
-  \nabla_\eta T &\rightarrow& \nabla_\eta \left(T-T^{(ref)}\right),\\ 
-    \nabla_\eta \delta p^{(d)} &\rightarrow& \nabla_\eta \left(\delta p^{(d)}-\delta p^{(ref)}\right).
-\end{eqnarray*}
-This reduces spurious transport of temperature and mass up/down-slope due to hyperviscosity.
-
-\subsection{Rewritting the pressure gradient force (PGF)}
-In the CESM2.0 we used to used the following (standard) form of the pressure gradient term:
+  \nabla^4_\eta T &\rightarrow& \nabla^4_\eta \left(T-T^{(ref)}\right),\label{eq:hyperT}\\ 
+    \nabla^4_\eta \delta p^{(d)} &\rightarrow& \nabla^4_\eta \left(\delta p^{(d)}-\delta p^{(ref)}\right).\label{eq:hyperDP}
+\end{eqnarray}
+This reduces spurious transport of temperature and mass up/down-slope due to the hyperviscosity operator.
+
+\subsection{Rewriting the pressure gradient force (PGF)}
+In the CESM2.0 the following (standard) form of the pressure gradient term was used:
 \begin{equation}
@@
 \end{equation}
-where $\Phi$ is geopotential and $\rho=\frac{R_dT_v}{p}$ is density and $\eta $ is the dry-mass vertical coordinate \cite<for details see >{LetAl2018JAMES}.
-To alleviate noise for flow over orography we switched to an Exner pressure formulation, that is, we use that (\ref{eq:pgf}) can be written in terms of the Exner pressure
+where $\Phi$ is geopotential and $\rho=\frac{R^{(d)}T_v}{p}$ is density \cite<for details see >{LetAl2018JAMES}. To alleviate noise for flow over orography, 
+ we switched to an Exner pressure formulation following \citeA{T2020JAMES}, which uses that (\ref{eq:pgf}) can be written in terms of the Exner pressure
 \begin{equation}\label{eq:pgf2}
@@
 &=& \frac{c_p^{(d)}\theta_v\kappa\Pi}{p} \nabla_{\eta }p,\\
-&=& \frac{R_d\theta_v\Pi}{p} \nabla_{\eta }p,\\
-&=& \frac{R_dT_v}{p} \nabla_{\eta }p,\\
-&=& \frac{1}{\rho} \nabla_{\eta }p,
+&=& \frac{R^{(d)}\theta_v\Pi}{p} \nabla_{\eta }p,\\
+&=& \frac{R^{(d)}T_v}{p} \nabla_{\eta }p,\\
+&=& \frac{1}{\rho} \nabla_{\eta }p.
 \end{eqnarray*}
+Using the reference states from \cite{SJ1991QJRMS},
+\begin{eqnarray}
+  \overline{T}&=&T_0+T_1 \Pi,\\
+  \overline{\theta}&=&{T_0}/\Pi+T_1,
+\end{eqnarray}
+we can define a geopotential as a function of Exner pressure
+\begin{equation}
+    \overline{\Phi} = -c_p^{(d)}\left( T_0\log \Pi+T_1\Pi-T_1\right). 
+\end{equation}
+This "balanced" geopotential obeys
+\begin{equation}
+    c_p^{(d)}\overline{\theta}\nabla \Pi+\nabla \overline{\Phi}=0 
+\end{equation}
+for any Exner pressure. Subtracting this "reference" profile from the PGF yields
+\begin{eqnarray}
+    \nabla_{\eta }\Phi+c_p^{(d)}\theta_v\nabla_{\eta }\Pi&=& \nabla_{\eta }\left(\Phi-\overline{\Phi}\right)+c_p^{(d)}\left(\theta_v-\overline{\theta}\right)\nabla_{\eta }\Pi,\nonumber \\
+    &=& \nabla_{\eta }\Phi+c_p^{(d)}\theta_v\nabla_{\eta }\Pi+c_p^{(d)}T_0\left[ \nabla_\eta \log \Pi -\frac{1}{\Pi}\nabla_\eta \Pi\right].\label{eq:pgf_ref}
+\end{eqnarray}
+In the continuum, the two formulations (left and right-hand side of (\ref{eq:pgf_ref})) are identical. But under discretization, the second formulation can have much less truncation error.
+
 \subsection{Results}
-One year averages of $OMEGA500$ (vertical pressure velocity at 500hPa) has been found to be a useful quantity to detect spurious up or down-drafts induced by steep orography.
+{\color{red}{[Adam: have you defined ne30np4 in the main text?]}}\\
+One year averages of vertical pressure velocity at 500hPa ( {\tt OMEGA500}) have been found to be a useful quantity to detect spurious up or down-drafts induced by steep orography (Figure \ref{fig:OMEGA500}).
+While the true solution is not known, strong vertical velocities aligned with element edges that are not found in the CAM-FV reference solution (Figure \ref{fig:OMEGA500}(a)) are likely not physical (spurious).
+The older CESM2.0 version of SE (Figure \ref{fig:OMEGA500}(d)) using the "traditional" discretization of the PGF, (\ref{eq:pgf_ref}), exhibits significant spurious noise patters around steep orography compared to CAM-FV (e.g., around Himalayas and Andes).
+This is strongly alleviated by switching to the Exner formulation of the PGF (\ref{eq:pgf2}; Figure \ref{fig:OMEGA500}(c)).
+By also subtracting reference profiles from pressure-level thickness and temperature, equations (\ref{eq:hyperT}) and (\ref{eq:hyperDP}) respectively, reduces strong up-down drafts further (Figure \ref{fig:OMEGA500}(d)).
+Switching to the CAM-SE-CSLAM version where physics tendencies are computed on an quasi-equal area physics grid and using the CSLAM transport scheme, marginal improvements are observed in terms of a smoother vertical velocity field (Figure \ref{fig:OMEGA500}(e,f)).
+The configuration shown in Figure \ref{fig:OMEGA500}(d) is used for the simulations shown in the main text of this paper.
+
+It is interesting to note that the noise issues and algorithmic remedies found in the real-world simulations discussed above, can be investigated by replacing all of physics with a modified version of the Held-Suarez forcing \cite{HS1994}.
+The original formulation of the Held-Suarez idealized test case used a flat Earth ($\Phi_s=0$) and a dry atmosphere.
+By simply adding the surface topography used in `real-world' simulations and removing the temperature relaxation in the lower part of domain ($\sigma>0.7$; see \citeA{HS1994} for details), surprisingly realistic vertical velocity fields (in terms of structure) result (see Figure \ref{fig:OMEGA500FHS94}).
+Since this was a very useful development tool it is shared in this manuscript.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=145mm]{OMEGA500F2000climo.pdf}
+\end{center}
+\caption{One year averages of vertical pressure velocity at 500hPa ({\tt OMEGA500}) using (a) CAM-FV (Finite-Volume dynamical core) and (b-f) various versions of the spectral-element (SE) dynamical core at approximately $1^\degree$ horizontal resolution and using 32 levels.
+(b) is equivalent to the CESM2.0 version of the SE dynamical core using the "traditional"/"old" discretization of the pressure-gradient force (PGF).
+Plot (c) is equivalent to configuration (b) but using the Exner form of the PGF.
+Plot (d) is the same as configuration (c) but also subtracting reference profiles from pressure and temperature before applying hyperviscosity operators (which is equivalent to the CESM2.2 version of SE in terms of the dynamical core).
+Plots (e) and (f) are equivalent to (c) and (d), respectively, by using the SE-CSLAM ($\tt ne30pg3$) version of the SE dynamical core (i.e.
+separate quasi-uniform physics grid and CSLAM transport scheme).}\label{fig:OMEGA500}
+\end{figure}
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=145mm]{OMEGA500FHS94.pdf}
+\end{center}
+\caption{Same as Figure \ref{fig:OMEGA500} but using modified Held-Suarez forcing and the average is over 18 months (excl. spin-up).}\label{fig:OMEGA500FHS94}
+\end{figure}
 
@@
 A.
-Herrington thanks Matt Rehme (NCAR/CISL) for his role in generating the $Arctic-GrIS$ visualization available on youtube ({\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}).
-
-The data presented in this manuscript is available at {\url{https://github.com/adamrher/2020-arcticgrids}}.
+Herrington thanks Matt Rehme (NCAR/CISL) for his role in generating the $\tt Arctic-GrIS$ visualization available on youtube ({\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}).
+
+The data presented in main part of this manuscript is available at {\url{https://github.com/adamrher/2020-arcticgrids}}.
+The source code and data for the Appendix is available at {\url{https://github.com/PeterHjortLauritzen/CAM/tree/topo-mods}}.
 

### 2022-04-18 arcticgrids.tex
coauthor mods

@@
 The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated Gauss-Lobatto-Legendre points \cite{HL2018MWR}.
-{\color{purplse}{Rene/Andrew - maybe elaborate further on the advantages of CSLAM?}}
+{\color{purple}{Rene/Andrew - maybe elaborate further on the advantages of CSLAM?}}
 
@@
 In CESM terminology, AMIP simuations use the $FHIST$ computational set and run out of the box in CESM2.2.
-{\color{grey}{Marcus - this is the only part that doesn't talk about CLM/CISM coupling.
+{\color{orange}{Marcus - this is the only part that doesn't talk about CLM/CISM coupling.
 Move somewhere else and change section name?}}
@@
 For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{Muntjewerf_EA_2021_coupling,CISM1,SETAL2019TC,KETAL2020JAMES}.
-{\color{grey}{Marcus - mention you're not running CISM.}}
+{\color{orange}{Marcus - mention you're not running CISM.}}
 
@@
 
-{\color{green}{Rene - is the melt map consistent with the bin figure?
-Smallest errors are interior points which correspond to glacier_fraction bin of 1, which the largest errors.}}{\color{blue}{ARH - I can check.
-The bin figure is fractional change so its a different metric, which could explain the apparent inconsistency.}}
+{\color{green}{Rene - is the melt map consistent with the bin figure? Smallest errors are interior points which correspond to glacier fraction bin of 1, which the largest errors.}}
+{\color{blue}{ARH - I can check. The bin figure is fractional change so its a different metric, which could explain the apparent inconsistency.}}
 
@@
   T^{(ref)}&=&T_0+T_1 \Pi^{(ref)},\\
-  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}(T_0+T_1)}\right)}, {\color{blue}{Mark - Note: new definition of T0 means this equation should also be changed}}
+  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}(T_0+T_1)}\right)},
 \end{eqnarray}
-\cite{SJ1991QJRMS} where $g$ gravity, $T_1=\Gamma_0 T_0 c_p^{(d)}/g\approx 192K$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv 288K-T_0\approx 97K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential.
+{\color{blue}{Mark - Note: new definition of T0 means this equation should also be changed}} \cite{SJ1991QJRMS} where $g$ gravity, $T_1=\Gamma_0 T_0 c_p^{(d)}/g\approx 192K$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv 288K-T_0\approx 97K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential.
 The reference Exner function is

### 2022-04-18 arcticgrids.tex
coauthor mods

@@
 \end{center}
-\caption{One year averages of vertical pressure velocity at 500hPa ({\tt OMEGA500}) using (a) CAM-FV (Finite-Volume dynamical core) and (b-f) various versions of the spectral-element (SE) dynamical core at approximately $1^\degree$ horizontal resolution and using 32 levels.
+\caption{One year averages of vertical pressure velocity at 500hPa ($\tt OMEGA500$) using (a) CAM-FV (Finite-Volume dynamical core) and (b-f) various versions of the spectral-element (SE) dynamical core at approximately $1^{\circ}$ horizontal resolution and using 32 levels.
 (b) is equivalent to the CESM2.0 version of the SE dynamical core using the "traditional"/"old" discretization of the pressure-gradient force (PGF).

### 2022-04-19 arcticgrids.tex
cleaned up citations

@@
 
-In the 1970s, the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs. 
-{\color{blue}(WHL: Add refs)}
+In the 1970s, the pole problem was largely defeated through the adoption of efficient spectral transform methods in GCMs \cite<see>[and references therein]{W2007JMSJ}.
 These methods transform grid point fields into a global, isotropic representation in wave space, where linear operators (e.g., horizontal derivatives) in the (truncated) equation set can be solved exactly.
 While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
-The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids
-{\color{blue}(define 'reduced grids')}
-and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
+The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells elongated in the longitudinal direction over the polar regions) and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
 Polar filters subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
@@
 Unstructured grids support quasi-uniform grid spacing globally, and there is no pole-problem (e.g., Figure~\ref{fig:uni-grids}c).
-Unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure.
-{\color{blue} (WHL: Add 1 or 2 refs)} This grid flexibility allows for variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
+Unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{PL2007JCP,DetAl2012IJHPCA,WETAL2013GMD}.
+This grid flexibility allows for variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
 In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).
@@
 GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation.
-The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias (Pollard 2000) \cite{VETAL2018TC}.
+The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias \cite{P2000GPC,VETAL2018TC}.
 GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
@@
 
-Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates (citations for MM5, RACMO and MAR).
-The limited area domain used by RCMs permits the use of high resolution grids able to resolve the SMB processes, and can skillfully simulate the GrIS SMB.
+Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates.
+The limited area domain used by RCMs permits the use of high resolution grids able to resolve the SMB processes, and can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}.
 However, unlike GCMs, RCMs are not a freely evolving system and boundary conditions must be prescribed at the lateral boundaries.
 The inability of the RCM solution to influence larger-scale dynamics outside the RCM domain (due to the prescribed boundary conditions) severely limits this approach from properly representing the role of the GrIS in the climate system.
+In addition, the boundary conditions are derived from a separate host model, which introduces inconsistencies due to differences in model design between the host model and the RCM.
 
@@
 This paper documents several additional enhancements to the SE dycore as part of the release of CESM2.2.
-These include three new VR configurations (Figure~\ref{fig:vr-grids}), two Arctic meshes and a Contiguous United-States mesh ($\tt CONUS$; featured in \citeA{GetAl2017JAMES,MUSCIA}).
-While there are many published studies using VR in CESM \cite<e.g., >{ZetAl2014JC,RHUZ2016JAMC,bambach}, these studies either used development code or collaborated closely with model developers.
+These include three new VR configurations (Figure~\ref{fig:vr-grids}), two Arctic meshes and a Contiguous United-States mesh ($\tt CONUS$; featured in \citeA{PETAL2020BAMS}).
+While there are dozens of published studies using VR in CESM \cite<e.g., >{ZetAl2014JC,RHUZ2016JAMC,GetAl2017JAMES,BETAL2019TAC,BETAL2021IJC}, these studies either used development code or collaborated closely with model developers.
 CESM2.2 is the first code release that contains out of the box VR functionality in CESM.
 
-This study compares the representation of high-latitude regions using the SE and FV dycores in CESM2.2 (see description below), as these two dycores treat high latitudes (i.e., the pole problem) in different ways.
+This study compares the representation of Arctic regions using the SE and FV dycores in CESM2.2 (see description below), as these two dycores treat high latitudes (i.e., the pole problem) in different ways.
 Section~\ref{sec:methods} documents the grids, dycores, and physical parameterizations used in this study, and also describes the experiments, datasets, and evaluation methods.
@@
 
-The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM;>{Gettelman2019GRL}, supports several different atmospheric dynamical cores.
+The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM;>{GETAL2019GRL,CAM63}, supports several different atmospheric dynamical cores.
 These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
-This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores {\color{green}{Andrew - maybe not mention EUL or FV3?}}.
-
-[phl: maybe mention that FV was used for IPCC simulations; SE in iHESP]
+This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores.
+CESM2 runs submitted to the Coupled Model Intercomparison Project Phase 6 \cite{CMIP6} are based on the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BETAL2018CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
 
@@
 Hyperviscous damping is applied to the divergent modes, while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC} [phl: as far as I know the Laplacian-like damping is not invoked in lower resolution setups - was implemented for 1/4 applications where the polar night jet became excessively strong].
-A polar filter damps computational instability due to the convergence of meridians, allowing a longer time step.
+A polar filter damps computational instability due to the convergence of meridians, permitting a more practical time step.
 It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the meridional direction \cite{ST1995GEOS}.
@@
 
-The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements to uniform-resolution grids: (1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
+The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over uniform-resolution setups.
+(1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
 A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements.
-(2) The topography boundary conditions need to be smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size.
+(2) The topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.
 
@@
 The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
-We also run the $1^{\circ}$ equivalent of the CAM-SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
-We run an additional $1^{\circ}$ CAM-SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d; note CSLAM is still run on the $3\times3$ control volume grid;>{HETAL2019JAMES}.
-
-Two VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic (Figure~\ref{fig:vr-grids}).
+We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
+We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMES}.
+
+Three VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic and $\tt CONUS$ (Figure~\ref{fig:vr-grids}).
 This paper serves as the official documentation of these grids.
@@
 The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution.
+The $\tt CONUS$ grid contains $1/8^{\circ}$ refinement over the United States, and $1^{\circ}$ everywhere else.
+The $\tt CONUS$ grid is not discussed any further in this paper; see \citeA{PETAL2020BAMS} for simulations using the $\tt CONUS$ grid.
 
@@
 Increased horizontal resolution permits faster vertical velocities that reduce characteristic time scales, so the physics time step is reduced to avoid large time truncation errors \cite{HR2018JAMES}.
-The $\tt Arctic$ and $\tt Arctic-GrIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time step relative to the default 1800 s time step for the standard $1^{\circ}$ and $2^{\circ}$ grids (Table~\ref{tbl:table1}).
+The $\tt Arctic$ and $\tt Arctic-GrIS$ grids are run with a $4\times$ and $8\times$ reduction in physics time step relative to the default 1800 s time step used in the $1^{\circ}$ and $2^{\circ}$ grids (Table~\ref{tbl:table1}).
 
@@
 
-All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{Gettelman2019GRL}.
+All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{GETAL2019GRL}.
 The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer, shallow convection, and cloud macrophysics scheme.
-CLUBB is coupled with the MG2 microphysics scheme \cite{MG2,Morrison2015}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
+CLUBB is coupled with the MG2 microphysics scheme \cite{GM2015JC,MG2}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
 Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
@@
 The SE dycore in CESM2.2 therefore has two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance.
-We reduced the width of the sub-grid distribution of vertical velocity ($\tt clubb_gamma$ is reduced from 0.308 to 0.270) and also reduced the strength of the damping for horizonatal component of turbulent energy ($\tt clubb_c14$ is reduced from 2.2 to 1.6) to increase cloudiness.
+We reduced the width of the sub-grid distribution of vertical velocity ($\tt clubb\_gamma=0.308\rightarrow0.270$) and also reduced the strength of the damping for horizontal component of turbulent energy ($\tt clubb\_c14=2.2\rightarrow1.6$) to increase cloudiness.
 For a thorough explanation of how CLUBB parameters impact the simulated climate, see \cite{GETAL2015JAMES}.
 
-\subsection{Experimental design}
-
-All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
-In CESM terminology, AMIP simuations use the $FHIST$ computational set and run out of the box in CESM2.2.
-{\color{orange}{Marcus - this is the only part that doesn't talk about CLM/CISM coupling.
-Move somewhere else and change section name?}}
+\subsection{Modeled SMB}
 
@@
 Ablation is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation runs off the ice sheet; liquid runoff (which is composed of both rain or ice/snow melt) can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass.
+Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass.
 These processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
@@
 Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
-For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{Muntjewerf_EA_2021_coupling,CISM1,SETAL2019TC,KETAL2020JAMES}.
+For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES,Muntjewerf_EA_2021_coupling}.
 {\color{orange}{Marcus - mention you're not running CISM.}}
@@
 
+\subsection{SMB Analysis}\label{sec:SMB}
+
+We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
+However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}).
+The area errors in Figure~\ref{fig:grisdx} may not seem large, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
+CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
+The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
+\end{center}
+\caption{The spatial properties of the GrIS as represented by different grids in this study.
+(Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
+\label{fig:grisdx}
+\end{figure}
+
+We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $\tt f19$ and $\tt ne30pg2$ grids, and integrating over these low-resolution ice masks.
+The use of low-resolution common ice masks is a conservative decision and is largely because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not reliable in the downscaling direction, i.e., mapping to a higher-resolution grid than the source grid.
+We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
+Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $\tt f19$ and $\tt ne30pg2$ masks.
+Thus, we evaluate an integrated quantity on a given grid up to four times to estimate the uncertainty due to differences in grid type and remapping algorithms.
+
+The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets.
+The SMB for total water can be expressed as:
+\begin{equation}
+SMB = precipitation + runoff + evaporation/sublimation, \label{eq:SMB}
+\end{equation}
+where all terms have consistent sign conventions (positive values contribute mass, negative values reduce mass).
+The precipitation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
+Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined snow plus ice melt.
+Note that this SMB expression is different from the internally computed SMB in CLM \cite{KETAL2020JAMES}.
+
+We consider two approaches for mapping and integrating the SMB components over the common ice masks:
+\begin{enumerate}
+\item Map the grid-cell mean quantities to the common grid, and integrate the mapped fields over the common ice masks. \label{lbl:label1}
+\item Map the patch-level quantities (i.e., the state over the ice fractional component of the grid cell) to the common grid, and integrate the mapped fields over the common ice masks.
+\label{lbl:label2}
+\end{enumerate}
+
+Note that we are mapping to low-resolution grids that have larger GrIS areas (Figure~\ref{fig:grisdx}).
+Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
+While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
+This underestimation is systematic in method~\ref{lbl:label2}; all variables are exclusive to the ice mask.
+Method~\ref{lbl:label2} is more dissipative, since mapping to the low-resolution grid will average a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
+However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.
+%The dilution of the these patch level fields caused by mapping to the common ice mask are then inflated due to integrating over the larger ice mask.
+
+The different error characteristics of the two methods are used to diversify the ensemble.
+Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
+Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
+This will degrade the SMB estimates using method~\ref{lbl:label2}, but note that we are more interested in the behavior of the processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
+
+\subsection{Experimental design}
+
+All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
+In CESM terminology, AMIP simulations use the $FHIST$ computational set and run out of the box in CESM2.2.
+
 \subsection{Observational Datasets}
@@
 We gathered SMB datasets from several sources.
-RACMO2.3 11km and RACMO2.3p2 5.5km are regional model simulations targeting Greenland, forced by ERA40, ERA interim, and ERA5 renalysis products at the domain boundaries.
-The Regional Atmospheric Climate Model (\textit{RACMO}) simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite{NETAL2015TC,NETAL2019SCIENCE}.
+Regional Atmospheric Climate Model, version 2.3 11km \cite<RACMO23;>{NETAL2015TC} and version 2.3p2 5.5km \cite<RACMO2.3p2;>{NETAL2018TC,NETAL2019SCIENCE} are RCM simulations targeting Greenland, forced by ERA40, ERA interim, and ERA5 renalysis products at the domain's lateral boundaries.
+The RACMO simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite<e.g.,>{LIVVkit,KETAL2020JAMES}.
 
@@
  \end{table*}
-
-\subsection{SMB Analysis}\label{sec:SMB}
-
-We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
-However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}).
-The area errors in Figure~\ref{fig:grisdx} may not seem large, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
-CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite<ESMF;>{ESMF}.
-The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
-\end{center}
-\caption{The spatial properties of the GrIS as represented by different grids in this study.
-(Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
-\label{fig:grisdx}
-\end{figure}
-
-We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $\tt f19$ and $\tt ne30pg2$ grids, and integrating over these low-resolution ice masks.
-The use of low-resolution common ice masks is a conservative decision and is largely because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not reliable in the downscaling direction, i.e., mapping to a higher-resolution grid than the source grid.
-We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
-Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $\tt f19$ and $\tt ne30pg2$ masks.
-Thus, we evaluate an integrated quantity on a given grid up to four times to estimate the uncertainty due to differences in grid type and remapping algorithms.
-
-The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets.
-The SMB for total water can be expressed as:
-\begin{equation}
-SMB = precipitation + runoff + evaporation/sublimation, \label{eq:SMB}
-\end{equation}
-where all terms have consistent sign conventions (positive values contribute mass, negative values reduce mass).
-The precipitation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
-Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined snow plus ice melt.
-Note that this SMB expression is different from the internally computed SMB in CLM \cite{KETAL2020JAMES}.
-
-We consider two approaches for mapping and integrating the SMB components over the common ice masks:
-\begin{enumerate}
-\item Map the grid-cell mean quantities to the common grid, and integrate the mapped fields over the common ice masks. \label{lbl:label1}
-\item Map the patch-level quantities (i.e., the state over the ice fractional component of the grid cell) to the common grid, and integrate the mapped fields over the common ice masks.
-\label{lbl:label2}
-\end{enumerate}
-
-Note that we are mapping to low-resolution grids that have larger GrIS areas (Figure~\ref{fig:grisdx}).
-Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
-While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
-This underestimation is systematic in method~\ref{lbl:label2}; all variables are exclusive to the ice mask.
-Method~\ref{lbl:label2} is more dissipative, since mapping to the low-resolution grid will average a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
-However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.
-%The dilution of the these patch level fields caused by mapping to the common ice mask are then inflated due to integrating over the larger ice mask.
-
-The different error characteristics of the two methods are used to diversify the ensemble.
-Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
-Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
-This will degrade the SMB estimates using method~\ref{lbl:label2}, but note that we are more interested in the behavior of the processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
 

### 2022-04-20 arcticgrids.tex
removed redundant files
fixed bib file!
couple bib fixes

@@
 
-\correspondingauthor{Adam Herrington}{aherring@ucar.edu}
+\correspondingauthor{Adam R. Herrington}{aherring@ucar.edu}
 
@@
 We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
-We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMES}.
+We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMESb}.
 
@@
 All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
-In CESM terminology, AMIP simulations use the $FHIST$ computational set and run out of the box in CESM2.2.
+In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2.
 
@@
 Figure~\ref{fig:dT-lores} also shows small temperature differences between $\tt ne30pg3$ and $\tt ne30pg2$, with  $\tt ne30pg3$ slightly warmer near the tropopause at high latitudes.
-This is consistent with the similar climates found for these two grids by \citeA{HETAL2019JAMES}.
+This is consistent with the similar climates found for these two grids by \citeA{HETAL2019JAMESb}.
 

### 2022-04-28 arcticgrids.tex
new grid figures, marcus edits

@@
 Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS).
-The more conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both accumulation and ablation over the GrIS.
-Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their greater degrees of freedom in representing the GrIS.
+The conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both accumulation and ablation over the GrIS.
+Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their higher degrees of freedom in representing the GrIS.
 Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, are documented as newly supported configurations in CESM2.2.
@@
 With the recent trend towards globally uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost.
-In this study we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate.
+In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate.
 We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.
@@
 While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
-The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells elongated in the longitudinal direction over the polar regions) and polar filters are necessary to ameliorate this instability \cite{JW2010LNCSE}.
-Polar filters subdue the growth of unstable modes by applying additional damping to the solution over polar regions.
+The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells elongated in the longitudinal direction over the polar regions) and polar filters are necessary to ameliorate this numerical instability \cite{JW2010LNCSE}.
+Polar filters subdue the growth of unstable computational modes by applying additional damping to the numerical solution over polar regions.
 This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid.
-We emphasize \textit{approximately} since it's at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.
+We emphasize \textit{approximately}, since it is at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.
 
@@
 \citeA{LJTN2010JAMES} has experimented with rotating lat-lon models such that their coordinate lines no longer align with an idealized, zonally balanced circulation.
-For the finite-volume lat-lon dycore considered in this paper (hereafter \textit{FV}), numerical errors were shown to be largest when the polar singularity is rotated into the baroclinic zone ($45^{\circ}$N latitude), generating spurious wave growth much earlier on in the simulation than for other rotation angles.
+For the finite-volume lat-lon dycore considered in this paper (hereafter \textit{FV}), numerical errors were shown to be largest when the polar singularity is rotated into the baroclinic zone ($45^{\circ}$N latitude), generating spurious wave growth much earlier in the simulation than for other rotation angles.
 This illustrates the advantages of coordinate surfaces aligned with latitude bands, albeit an extreme example where the polar singularity and the polar filter are also contributing to the spurious wave growth.
 The unstructured grids all generate spurious baroclinic waves earlier on in the simulations than the (unrotated) lat-lon models, although the unstructured model considered in this paper, the spectral-element dycore (hereafter \textit{SE}), holds a balanced zonal flow without wave growth appreciably longer than the rotated FV experiments, regardless of how the SE grid is rotated \cite{LJTN2010JAMES}.
+{\color{purple}{Marcus - this sentence is very complicating.}}
 
@@
 Unstructured grids support quasi-uniform grid spacing globally, and there is no pole-problem (e.g., Figure~\ref{fig:uni-grids}c).
-Unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{PL2007JCP,DetAl2012IJHPCA,WETAL2013GMD}.
-This grid flexibility allows for variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
+Conversely, unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{PL2007JCP,DetAl2012IJHPCA,WETAL2013GMD}.
+This grid flexibility allows for the adoption of nested variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
 In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).
-However, refinement comes at the cost of a smaller CFL-limited time step in the refined region.
-(The CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.)
-
-It is important to emphasize that the pole-problem is a feature of the dycore.
+However, local grid refinement comes at the cost of a smaller CFL-limited time step in the refined region; the CFL-condition --- short for Courant–Friedrichs–Lewy condition --- is a necessary condition for numerical stability when using discrete data in time and space.
+
+It is important to emphasize that the pole-problem is a distinctive feature of the dycore in atmospheric models.
 Polar filters do not directly interfere with the physical parameterizations, nor do they have any bearing on the surface models; e.g., the land model can take full advantage of the greater number of grid cells in polar regions on lat-lon grids.
-This is particularly relevant for the surface mass balance of the Greenland Ice Sheet (\textit{SMB}; the integrated sum of precipitation and runoff), which relies on hydrological processes handled by the land model.
+This is particularly relevant for the surface mass balance of the Greenland Ice Sheet (\textit{SMB}; the integrated sum of precipitation and runoff), which relies on hydrological processes represented in the land model.
 
@@
 Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates.
-The limited area domain used by RCMs permits the use of high resolution grids able to resolve the SMB processes, and can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}.
-However, unlike GCMs, RCMs are not a freely evolving system and boundary conditions must be prescribed at the lateral boundaries.
+The limited area domain used by RCMs permits the use of high resolution grids, capable of resolving SMB processes, that can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}.
+However, unlike GCMs, RCMs are not a freely evolving system and the atmospheric state must be prescribed at the lateral boundaries of the model domain.
 The inability of the RCM solution to influence larger-scale dynamics outside the RCM domain (due to the prescribed boundary conditions) severely limits this approach from properly representing the role of the GrIS in the climate system.
@@
 
-The SE dycore has been an available option since CESM, version 1, but has been under active development ever since.
+The SE dycore has been included in the model since CESM, version 1, but has been under active development ever since.
 This includes the switch to a dry-mass vertical coordinate \cite{LetAl2018JAMES} and incorporation of an accelerated multi-tracer transport scheme \cite{LTOUNGK2017MWR}, made available in CESM2.
@@
 \begin{tabular}{cc}
-         \includegraphics[width=60mm]{figs/grid-f19.pdf}&
-         \includegraphics[width=60mm]{figs/grid-f09.pdf}\\
-         \includegraphics[width=60mm]{figs/grid-ne30pg2.pdf}&
-         \includegraphics[width=60mm]{figs/grid-ne30pg3.pdf} \\
+         \includegraphics[width=50mm]{figs/grid-f19.pdf}&
+         \includegraphics[width=50mm]{figs/grid-f09.pdf}\\
+         \includegraphics[width=50mm]{figs/grid-ne30pg2.pdf}&
+         \includegraphics[width=50mm]{figs/grid-ne30pg3.pdf} \\
 \end{tabular}
@@
 \begin{center}
-\begin{tabular}{cccc}
-         \includegraphics[width=60mm]{figs/grid-ARCTIC.pdf}&
-         \includegraphics[width=60mm]{figs/grid-ARCTICGRIS.pdf} \\
+\begin{tabular}{ccc}
+         \includegraphics[width=40mm]{figs/grid-ARCTIC.pdf}&
+         \includegraphics[width=40mm]{figs/grid-ARCTICGRIS.pdf}&
+         \includegraphics[width=40mm]{figs/grid-CONUS.pdf} \\
 \end{tabular}
 \end{center}
-\caption{Spectral-element grid for the VR grids in this study, (a)$\tt Arctic$, (b) $\tt Arctic-GrIS$ and (a) $\tt CONUS$.
-Note that this is not the computational grid; each element has $3\times3$ independent grid points.
-{\color{red}{add CONUS grid}}}
+\caption{Spectral-element grid for the VR grids in this study, (a) $\tt Arctic$, (b) $\tt Arctic-GrIS$ and (c) $\tt CONUS$.
+Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
 \label{fig:vr-grids}
@@
 Hyperviscous damping is applied to the divergent modes, while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC} [phl: as far as I know the Laplacian-like damping is not invoked in lower resolution setups - was implemented for 1/4 applications where the polar night jet became excessively strong].
-A polar filter damps computational instability due to the convergence of meridians, permitting a more practical time step.
+A polar filter damps computational instability due to the convergence of meridians, permitting a longer time step.
 It takes the form of a Fourier filter in the zonal direction, with the damping coefficients increasing monotonically in the meridional direction \cite{ST1995GEOS}.
@@
 The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over uniform-resolution setups.
-(1) As the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
+Firstly, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
 A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements.
-(2) The topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.
-
-For SE grids with quasi-uniform grid spacing, The SE tracer transport scheme is replaced with the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}.
+Secondly, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.
+
+For SE grids with quasi-uniform grid spacing, the SE tracer transport scheme is replaced with the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}.
 CSLAM has improved tracer property preservation and accelerated multi-tracer transport.
@@
 
-All grids and dycores in this study use 32 levels in the vertical, with a model top of about 1~hPa or 40~km.
+All grids and dycores in this study use 32 hybrid pressure-sigma levels in the vertical, with a model top of about 1~hPa or 40~km.
 However, note that any grid or dycore can in principle be run with a higher model top or finer vertical resolution.
@@
 
-Initial simulations with the $\tt ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore.
+Initial simulations with the $\tt ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore in the standard CESM2 configuration.
 The SE dycore in CESM2.2 therefore has two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance.
 We reduced the width of the sub-grid distribution of vertical velocity ($\tt clubb\_gamma=0.308\rightarrow0.270$) and also reduced the strength of the damping for horizontal component of turbulent energy ($\tt clubb\_c14=2.2\rightarrow1.6$) to increase cloudiness.
-For a thorough explanation of how CLUBB parameters impact the simulated climate, see \cite{GETAL2015JAMES}.
-
-\subsection{Modeled SMB}
-
-All grids and dycores simulate the GrIS Surface Mass Balance (SMB).
-The SMB is the sum of mass accumulation of precipitation and mass loss from ablation.
-Ablation is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
+For a description of how CLUBB parameters impact the simulated climate, see \citeA{GETAL2015JAMES}.
+
+\subsection{Simulated surface mass balance (SMB)}
+
+All grids and dycores simulate the GrIS Surface Mass Balance (SMB), which is the sum of mass accumulation of precipitation and mass loss from ablation.
+The latter is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
 Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass.
@@
 For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES,Muntjewerf_EA_2021_coupling}.
-{\color{orange}{Marcus - mention you're not running CISM.}}
 
@@
 That is, snow accumulation above the 10-m cap contributes a positive SMB, and surface ice melting (after melting of the overlying snow) yields a negative SMB.
-Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over a 20-year $\tt Arctic$ $FHIST$ run for about 500 years.
+Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation (a model simulation with prescribed, observed sea-surface conditions) for about 500 years.
 The uniform-resolution grids are initialized with the SMB from an existing $\tt f09$ spun-up initial condition.
+In the simulations described here, the GrIS is prescribed at its observed, modern extent and thickness.
 
@@
 However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}).
-The area errors in Figure~\ref{fig:grisdx} may not seem large, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
 CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
 The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
+The area errors in Figure~\ref{fig:grisdx} may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
 
@@
 \caption{The spatial properties of the GrIS as represented by different grids in this study.
-(Left) GrIS area error, computed as the relative differences from a 4km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
+(Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
 \label{fig:grisdx}
@@
 We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $\tt f19$ and $\tt ne30pg2$ grids, and integrating over these low-resolution ice masks.
-The use of low-resolution common ice masks is a conservative decision and is largely because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not reliable in the downscaling direction, i.e., mapping to a higher-resolution grid than the source grid.
+The use of low-resolution common ice masks is a conservative decision, and is justified because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not generally reliable when mapping to a higher-resolution grid than the source grid.
 We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
@@
 \begin{equation}
-SMB = precipitation + runoff + evaporation/sublimation, \label{eq:SMB}
+SMB = accumulation + runoff + evaporation/sublimation, \label{eq:SMB}
 \end{equation}
-where all terms have consistent sign conventions (positive values contribute mass, negative values reduce mass).
-The precipitation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
-Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined snow plus ice melt.
+where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass).
+The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
+Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.
 Note that this SMB expression is different from the internally computed SMB in CLM \cite{KETAL2020JAMES}.
@@
 
-Note that we are mapping to low-resolution grids that have larger GrIS areas (Figure~\ref{fig:grisdx}).
+Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx}).
 Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
 While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
-This underestimation is systematic in method~\ref{lbl:label2}; all variables are exclusive to the ice mask.
+This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask.
 Method~\ref{lbl:label2} is more dissipative, since mapping to the low-resolution grid will average a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
@@
 
-Before delving into the simulated Arctic, we assess the global mean differences between the various grids and dycores.
+Before delving into the simulated Arctic climate conditions, we assess the global mean differences between the various grids and dycores.
 Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between uniform-resolution grids and dycores.
@@
 This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}.
-\citeA{HK2020QJRMS} have shown that this occurs in CAM due to greater resolved vertical velocities which, in turn, drive more condensational heating in the CLUBB macrophyiscs.
-The right columns in Figure~\ref{fig:dT-lores} support this interpretation, showing an increase in the climatological CLUBB heating in low ({\color{blue}{looks more like subtropics}}) and mid-latitudes on the $\tt f09$ grid.
+\citeA{HK2020QJRMS} have shown that this occurs in CAM due to higher resolved vertical velocities which, in turn, generate more condensational heating in the CLUBB macrophyiscs.
+The right column in Figure~\ref{fig:dT-lores} supports this interpretation, showing an increase in the climatological CLUBB heating in low ({\color{blue}{looks more like subtropics}}) and mid-latitudes on the $\tt f09$ grid.
 
@@
 
-Comparing the VR grids to the uniform-resolution grids is complicated because we simultaneously increase the resolution and reduce the physics time-step, both of which impact the solution \cite{W2008TELLUS}.
-We therefore run an additional $\tt ne30pg3$ simulation with the shorter physics time step used in the $\tt Arctic$ grid (450~s), referred to as $\tt ne30pg3^{*}$.
+Comparing the VR grids to the uniform-resolution grids is complicated because we simultaneously increase the resolution and reduce the physics time-step, both of which influence the solution \cite{W2008TELLUS}.
+We therefore run an additional $\tt ne30pg3$ simulation with the shorter physics time step used in the $\tt Arctic$ grid (450~s), referred to as $\tt ne30pg3^{*}$ (Table~\ref{tbl:table1}).
 Figure~\ref{fig:dT-hires}a shows the difference between $\tt ne30pg3^{*}$ and $\tt ne30pg3$ for climatological summer temperatures in zonal-mean height space.
@@
 
-It is useful to understand summer temperature biases due to their control on ice and snow melt over the Greenland Ice Sheet \cite<GrIS;>{O2001JAM}.
+It is useful to understand summer temperature biases due to their control on ice and snow melt over the GrIS \cite<>{O2001JAM}.
 Figure~\ref{fig:dThyps} shows the 1979--1998 lower troposphere summer temperature bias relative to ERA5.
@@
 All grids show a north-south gradient in bias over Greenland, with the summer temperature bias more positive for the northern part of the ice sheet.
-This pattern is also evident in the 2m summer temperature bias over Greenland (not shown).
+This pattern is also evident in the near surface temperature bias over Greenland (not shown).
+{\color{purple}{Marcus - why mention it if it's not shown?}}
 
@@
 In addition to summer temperatures, shortwave radiation is an important determinant of snow and ice melt.
-Figure~\ref{fig:FSDS} shows the summer incident shortwave radiation bias at the surface, zoomed in over Greenland.
+Figure~\ref{fig:FSDS} shows the summer incident shortwave radiation bias at the surface over Greenland and surrounding seas.
 The top panel shows the bias relative to the CERES-EBAF dataset, and the bottom panel relative to the RACMO2.3p2 dataset.
@@
 This suggests that the coarse-grid clouds are too thick in the Greenland interior and too thin around the perimeter, and that increasing horizontal resolution reduces these biases.
-This is consistent with the total summer cloud fraction bias, computed from the CALIPSO-GOCCP cloud dataset  and shown in Figure~\ref{fig:prect}.
-Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates the changes arising from any single layer due to the maximum overlap assumption used to compute this quantity.
+{\color{purple}{Marcus - This result is also consistent with the SMB bias in the GrSI interior that plagues low-res simulations.
+Is it worth commenting on that here?}}This is consistent with the total summer cloud fraction bias, computed from the CALIPSO-GOCCP cloud dataset  and shown in Figure~\ref{fig:prect}.
+Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at all other levels.
 Despite the attenuated signal, the total cloud fraction for the VR grids does indicate reduced cloud coverage in the interior and increased cloudiness around the ocean perimeter.
@@
 The accuracy of the simulated SMB is expected to be sensitive to grid resolution.
-Figure~\ref{fig:gridsdx}b shows the average grid spacing over the Greenland Ice Sheet (GrIS) in all six grids in this study.
-The $\tt ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO2.3 grid.
+Figure~\ref{fig:grisdx}b shows the average grid spacing over GrIS in all six grids in this study.
+The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160~km$, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6~km$, broadly similar to the $11~km$ grid spacing of the RACMO2.3 grid.
 The $\tt ne30pg3$ grid has an average $\Delta x=111.2~km$, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60~km$.
@@
 ({\color{blue}{Wondering if this paragraph would go better in an earlier section.
-Rene agrees}})
+Rene agrees.
+Marcus agrees (move to 2.2)}})
 
@@
 
-Large GrIS accumulation rates result from synoptic systems arriving from the south.
-These systems are orographically lifted at the ice sheet margin, especially over the steep slopes in southeast Greenland, concentrating heavy precipitation near the ice margin.
-At lower resolutions, the topography is too smooth and moisture penetrates inland, erroneously dumping precipitation onto the southern dome.
-The ability of the VR grids to more accurately simulate orographic precipitation is consistent with the cloud results above.
-As the precipitation centers move from the interior toward the margins, and even out over the ocean with increasing resolution, the cloud decks move accordingly.
-Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, and \ref{fig:prect} clearly illustrate a shift in clouds from the interior to the ocean perimeter with increasing resolution.
+Large GrIS accumulation rates result from synoptic systems arriving primarily from the south.
+These systems are orographically lifted at the ice sheet margin, especially over the steep slopes in southeast Greenland (dominated by winter precipitation), concentrating heavy precipitation near the ice margin.
+At lower resolutions, the topography is too smooth to generate a realistic orographic uplift, leading to anomalously high SMB in the GrIS interior.
+On the higher resolution grids, precipitation centers migrate from the interior toward the margins, and even out over the ocean in the case of the $\tt Arctic-GrIS$ grid.
+The more accurate representation of orographic precipitation on the higher resolution grids is consistent with the discussion on simulated cloud and radiation biases, cf.
+Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, and \ref{fig:prect}.
 
@@
 Of note is the large reduction in melt rates using the common-ice-mask approach compared to the native grid, illustrating the dissipation discussed in section~\ref{sec:SMB}.
-The errors are greatest in partially ice-covered grid cells straddling the ice sheet margins, in the ablation zone where melt rate are large.
+The errors are greatest in partially ice-covered grid cells straddling the ice sheet margins, in the ablation zone where melt rates are large.
 For integrated precipitation, the differences between the native and common-ice-mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
@@
 Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the most accumulation.
-In all basins, accumulations drops monotonically with increasing grid resolution, with some exceptions.
+In all basins, accumulation decreases monotonically with increasing grid resolution, though with some exceptions.
 The $\tt Arctic-GrIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to the RACMO precipitation in the large southwest basin.
@@
 Figure~\ref{fig:basin}c shows the basin-integrated PDD melt estimate.
-In the large southeast and southwest basins (and all the other western basins), the $\tt ne30pg3$ grid has larger PDD-based melt than the $ARCTCGRIS$ grid.
+In the large southeast and southwest basins (and all the other western basins), the $\tt ne30pg3$ grid has larger PDD-based melt than the $\tt Arctic-GrIS$ grid.
 The FV grids also have large PDD-based melt in the southwest basin, relative to $\tt Arctic-GrIS$.
-The PDD plots indicate that the near-surface temperatures which contribute to melt are not well approximated by the summer lower-troposphere temperatures in Figure~\ref{fig:dThyps}.
+The PDD plots indicate that the near-surface temperatures (which contribute to melt) are not well approximated by the summer lower-troposphere temperatures in Figure~\ref{fig:dThyps}.
 
@@
 The transects in Figure~\ref{fig:ztrans} show that the ice sheet surface on the coarse grids is not systematically lower than the true surface in ablation zones.
-Rather, the smoothing of the raw topography, necessary to prevent the model from instigating grid-scale modes, flattens the ice sheet, causing the lower-elevation ablation zones to extend beyond the true ice sheet margin, where they lie above the actual ice surface.
-The $\tt f19$ grid has the smoothest topography since its dynamics are coarsest (whereas $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ use identical smoothing), and has the flattest ice sheet.
-This suggests that if anything, coarser models will tend to elevate the ablation zones and depress melt rates.
+Rather, the smoothing and flattening of the raw topography, necessary to prevent the model from exciting grid-scale numerical modes, causes the lower-elevation ablation zones to extend beyond the true ice sheet margin, where they lie above the actual ice surface.
+The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest ($\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ use identical smoothing).
+This suggests that coarser models will tend to elevate the ablation zones and thereby depress melt rates.
 
@@
 The composite mean precipitation maps are similar between the two grids, and exhibit the iconic comma structure of synoptic cyclones (not shown).
+{\color{purple}{Marcus - no need to mention if not shown.}}
 
@@
 However, it is well documented that $1/4^{\circ}$ models are more skillful at simulating extreme events \cite{BetAl2013JC,OETAL2016JAMES}.
-This is an additional benefit of the VR grids.
+A more realistic representation of extreme precipitation events is an additional benefit of the VR grids.
 
@@
 
-Running CESM2.2 in an AMIP-style configuration, we have evaluated six grids from two dynamical cores for their performance over the Arctic and in simulating the Greenland Ice Sheet (GrIS) surface mass balance (SMB).
+Running CESM2.2 in an AMIP-style configuration, we have evaluated six grids from two dynamical cores for their performance over the Arctic and in simulating the GrIS SMB.
 The $1-2^{\circ}$ finite-volume grids have enhanced resolution over polar regions due to the convergence of meridian lines, although a polar filter is used to prevent spurious atmospheric features from forming at this higher resolution.
-Spectral-element grids comparable to the resolution of the finite-volume grids have an isotropic grid structure, meaning the grid resolution is similar over the entire domain.
+Spectral-element grids comparable to the resolution of the finite-volume grids have an isotropic grid structure where the grid resolution is similar over the entire model domain.
 We developed two VR grids and introduced them into CESM2.2 as part of this work.
 Both use the spectral-element dycore; the $\tt Arctic$ grid has $1/4^{\circ}$ refinement over the broader Arctic, whereas the $\tt Arctic-GrIS$ grid is identical except for a $1/8^{\circ}$ patch of refinement over Greenland.
+A third VR grid, $\tt CONUS$, is also available in CESM2.2.
 
 In general, the FV grids have colder summer temperatures over the Arctic compared with the SE grids (including the VR grids).
-The cloud biases in all the uniform-resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.
-The VR grids reduce the cloud biases.
-It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates; improved clouds in the Arctic do not imply improved clouds at lower latitudes.
+The cloud biases in all the uniform-resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.{\color{purple}{Marcus - hard to parse this sentence.}} The VR grids reduce the cloud biases.
+It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates; an improved representation of clouds in the Arctic does not imply improved clouds at lower latitudes.
 
@@
 This halo bias coincides with a positive cloud bias over the ice sheet interior.
-This pattern has been traced to the inadequacy of the coarser grids for resolving orographic precipitation.
+This anomaly pattern has been attributed to deficient orographic precipitation on the coarser model grids.
 With overly smooth topography on the $1-2^{\circ}$ grids, synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins.
-As a result, moisture penetrates and dumps excess precipitation into the GrIS interior, instead of being concentrated near the coastal margins as shown by observations.
+As a result, excess precipitation falls in the GrIS interior, instead of being concentrated on the steep coastal margins as shown by observations.
 This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter.
 The agreement of different observational data products on this bias lends confidence in the attribution of causes.
-The VR grids compare better to the observations and show that orographic precipitation in Greenland is largely resolved.
+The VR grids compare better to the observations and show that orographic precipitation in Greenland is largely resolved when the horizontal resolution is increased sufficiently.
 
@@
 The $1^{\circ}$ FV model gives a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids.
-In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other uniform-resolution grids.
+{\color{purple}{It is also the most heavily tuned model configuration.
+May be worth mentioning here as well.}} In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other uniform-resolution grids.
 
@@
 %, and should be incorporated into CESM in the near future.
-We have also developed a visualization of the $ARCTCGRIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}.
+We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}.
 %in order to advertise some of the stunning capabilities of the refined mesh approach.
@@
   T^{(ref)}&=&T_0+T_1 \Pi^{(ref)},\\
-  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}(T_0+T_1)}\right)},
+  p_s^{(ref)}&=&p_0\exp{\left(-\frac{\Phi_s}{R^{(d)}T_{ref}}\right)},
 \end{eqnarray}
-{\color{blue}{Mark - Note: new definition of T0 means this equation should also be changed}} \cite{SJ1991QJRMS} where $g$ gravity, $T_1=\Gamma_0 T_0 c_p^{(d)}/g\approx 192K$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv 288K-T_0\approx 97K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential.
+ \cite{SJ1991QJRMS} where $g$ gravity, $T_1=\Gamma_0 T_{ref} c_p^{(d)}/g\approx 192K$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv T_{ref}-T_1\approx 97K$; $T_{ref}=288K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential.
 The reference Exner function is
@@
 \begin{eqnarray}
-   \hbox{CESM2.0}&\rightarrow&\hbox{CESM2.2}\\
-   && \\
+   \hbox{CESM2.0}&\rightarrow&\hbox{CESM2.2}\nonumber \\
+   && \nonumber \\
   \nabla^4_\eta T &\rightarrow& \nabla^4_\eta \left(T-T^{(ref)}\right),\label{eq:hyperT}\\ 

### 2022-05-04 arcticgrids.tex
update text

@@
 This illustrates the advantages of coordinate surfaces aligned with latitude bands, albeit an extreme example where the polar singularity and the polar filter are also contributing to the spurious wave growth.
-The unstructured grids all generate spurious baroclinic waves earlier on in the simulations than the (unrotated) lat-lon models, although the unstructured model considered in this paper, the spectral-element dycore (hereafter \textit{SE}), holds a balanced zonal flow without wave growth appreciably longer than the rotated FV experiments, regardless of how the SE grid is rotated \cite{LJTN2010JAMES}.
-{\color{purple}{Marcus - this sentence is very complicating.}}
+The unstructured grids all generate spurious baroclinic waves earlier on in the simulations than the (unrotated) lat-lon models, although the unstructured model considered in this paper, the spectral-element dycore (hereafter \textit{SE}), holds a balanced zonal flow without spurious wave growth appreciably longer than the rotated FV experiments \cite{LJTN2010JAMES}.
+And unlike FV, the SE dycore has the same error characteristics regardless of how the grid is rotated.
 
@@
 Unstructured grids support quasi-uniform grid spacing globally, and there is no pole-problem (e.g., Figure~\ref{fig:uni-grids}c).
-Conversely, unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{PL2007JCP,DetAl2012IJHPCA,WETAL2013GMD}.
-This grid flexibility allows for the adoption of nested variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
+Conversely, unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}.
+This grid flexibility allows for the adoption of variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
 In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).
@@
 
-The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM;>{GETAL2019GRL,CAM63}, supports several different atmospheric dynamical cores.
+The atmospheric component of CESM2.2 \cite{CESM2}, the Community Atmosphere Model, version 6.3 \cite<CAM6;>{GETAL2019GRL,CAM63}, supports several different atmospheric dynamical cores.
 These include dycores on lat-lon grids, such as finite-volume \cite<FV;>{L2004MWR} and Eulerian spectral transform \cite<EUL;>{CETAL006JC} models, and dycores built on unstructured grids, including spectral-element \cite<SE;>{LetAl2018JAMES} and finite-volume 3 \cite<FV3;>{PL2007JCP} models.
 This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores.
-CESM2 runs submitted to the Coupled Model Intercomparison Project Phase 6 \cite{CMIP6} are based on the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BETAL2018CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
+CESM2 runs submitted to the Coupled Model Intercomparison Project Phase 6 \cite{CMIP6} used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BETAL2018CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
 
@@
 The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}.
-The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}, using a hybrid-pressure vertical coordinate.
-Hyperviscous damping is applied to the divergent modes, while Laplacian damping is applied to momentum in the top few layers, referred to as a \textit{sponge layer} \cite{L2011IJHPC} [phl: as far as I know the Laplacian-like damping is not invoked in lower resolution setups - was implemented for 1/4 applications where the polar night jet became excessively strong].
+The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}.
+Hyperviscous damping is applied to the divergent modes, and is increased in the top few layers (referred to as a \textit{sponge layer}) to prevent undesirable interactions with the model top, such as wave reflection \cite{L2011IJHPC}.
 A polar filter damps computational instability due to the convergence of meridians, permitting a longer time step.
@@
 For SE grids with quasi-uniform grid spacing, the SE tracer transport scheme is replaced with the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}.
+Atmospheric tracers have large, nearly discontinuous horizontal gradients that are difficult to represent with spectral methods, which are prone to oscillatory ``Gibbs-ringing" errors \cite{RW1990QJR}.
 CSLAM has improved tracer property preservation and accelerated multi-tracer transport.
 It uses a separate grid from the spectral-element dynamics, dividing each element into $3\times3$ control volumes with quasi-equal area.
-The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the default SE dycore in which the physics are evaluated Gauss-Lobatto-Legendre points \cite{HL2018MWR}.
-{\color{purple}{Rene/Andrew - maybe elaborate further on the advantages of CSLAM?}}
+The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the original SE dycore in which the physics are evaluated Gauss-Lobatto-Legendre points \cite{HL2018MWR}.
 
@@
 The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
-We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}c), where $ne$ refers to a grid with of $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
-We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}d;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMESb}.
-
-Three VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic and $\tt CONUS$ (Figure~\ref{fig:vr-grids}).
+We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}d), where $ne$ refers to a grid with $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
+We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}c;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMESb}.
+
+Three VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic and the United States (Figure~\ref{fig:vr-grids}).
 This paper serves as the official documentation of these grids.
-The Arctic meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
+The VR meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
 The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region.
@@
 The $\tt CONUS$ grid contains $1/8^{\circ}$ refinement over the United States, and $1^{\circ}$ everywhere else.
-The $\tt CONUS$ grid is not discussed any further in this paper; see \citeA{PETAL2020BAMS} for simulations using the $\tt CONUS$ grid.
+The $\tt CONUS$ grid is not discussed any further in this paper; see \citeA{PETAL2020BAMS} for simulations with the $\tt CONUS$ grid.
 
@@
 FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.
-We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.{\color{purple}{Rene - maybe include model cost?}}}
+We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.}
  \label{tbl:table1}
@@
 
-All grids and dycores in this study use 32 hybrid pressure-sigma levels in the vertical, with a model top of about 1~hPa or 40~km.
+All grids and dycores in this study use 32 hybrid pressure-sigma levels in the vertical, with a model top of 2~hPa or about 40~km.
 However, note that any grid or dycore can in principle be run with a higher model top or finer vertical resolution.
@@
 
-All simulations in this study use the CAM6.3 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{GETAL2019GRL}.
+All simulations in this study use the CAM6 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{GETAL2019GRL}.
 The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer, shallow convection, and cloud macrophysics scheme.
-CLUBB is coupled with the MG2 microphysics scheme \cite{GM2015JC,MG2}, with prognostic precipitation and classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
-Deep convection is parameterized using a convective quasi-equilibrium, bulk plume mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
+CLUBB is coupled with the MG2 microphysics scheme \cite{GM2015JC,MG2}, which computes prognostic precipitation and uses classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
+Deep convection is parameterized using a convective quasi-equilibrium, mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
 Boundary layer form drag is modeled after \citeA{BBW2004QJRMS}, and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale (the ridge orientation is derived from a high-resolution, global topography dataset \cite{Danielson_Gesch_2011_GMTED2010}).
@@
 
-All grids and dycores simulate the GrIS Surface Mass Balance (SMB), which is the sum of mass accumulation of precipitation and mass loss from ablation.
-The latter is the sum of evaporation/sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
+All grids and dycores simulate the GrIS SMB, which is the sum of mass accumulation of precipitation and mass loss from ablation.
+The latter is the sum of evaporation, sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
 Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass.
-These processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
-
-CLM runs on the same grid as the atmosphere, using a downscaling technique to account for sub-grid variability in SMB.
+These relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
+
+CLM runs on the same grid as the atmosphere, and uses a downscaling technique to account for sub-grid variability in SMB.
 In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (ECs), each with a distinct surface energy balance and SMB.
@@
 Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent.
-Any snow above the 10-m cap is routed to the ocean as solid runoff.
+Any snow above the 10 m cap contributes towards ice accumulation in the SMB.
 Refreezing of liquid water within the snowpack is an additional source of ice.
@@
 Changes in ice depth, but not snow depth, count toward the SMB.
-That is, snow accumulation above the 10-m cap contributes a positive SMB, and surface ice melting (after melting of the overlying snow) yields a negative SMB.
+That is, snow accumulation above the 10 m cap contributes a positive SMB, and surface ice melting (after melting of the overlying snow) yields a negative SMB.
 Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation (a model simulation with prescribed, observed sea-surface conditions) for about 500 years.
@@
 Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.
-Note that this SMB expression is different from the internally computed SMB in CLM \cite{KETAL2020JAMES}.
+Note that this SMB expression is different from the internally computed SMB described in the previous section.
 
@@
 While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
-This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask.
-Method~\ref{lbl:label2} is more dissipative, since mapping to the low-resolution grid will average a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
+This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
 However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.
-%The dilution of the these patch level fields caused by mapping to the common ice mask are then inflated due to integrating over the larger ice mask.
 
@@
 Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
-This will degrade the SMB estimates using method~\ref{lbl:label2}, but note that we are more interested in the behavior of the processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
+This will degrade the SMB estimates using method~\ref{lbl:label2}, however we are more interested in characterizing the behavior of individual processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
+
+The accuracy of the simulated SMB is expected to be sensitive to grid resolution.
+Figure~\ref{fig:grisdx}b shows the average grid spacing over GrIS in all six grids in this study.
+The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160~km$, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6~km$, broadly similar to the $11~km$ grid spacing of the RACMO2.3 grid.
+The $\tt ne30pg3$ grid has an average $\Delta x=111.2~km$, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60~km$.
+Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
+The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
+The $\tt Arctic-GrIS$ grid is about twice as expensive as the $\tt Arctic$ grid.
 
@@
 We use several observational datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations.
-We gathered SMB datasets from several sources.
-Regional Atmospheric Climate Model, version 2.3 11km \cite<RACMO23;>{NETAL2015TC} and version 2.3p2 5.5km \cite<RACMO2.3p2;>{NETAL2018TC,NETAL2019SCIENCE} are RCM simulations targeting Greenland, forced by ERA40, ERA interim, and ERA5 renalysis products at the domain's lateral boundaries.
+SMB datasets are gathered from multiple sources.
+Regional Atmospheric Climate Model, version 2.3 11km \cite<RACMO23;>{NETAL2015TC} and version 2.3p2 5.5km \cite<RACMO2.3p2;>{NETAL2018TC,NETAL2019SCIENCE} are RCM simulations targeting Greenland, forced by ERA renalyses products at the domain's lateral boundaries.
 The RACMO simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite<e.g.,>{LIVVkit,KETAL2020JAMES}.
@@
 Because of the uncertainty of comparing gridded fields to point-wise measurements, and the lack of information added with regard to model biases, we omitted these datasets from our analysis.
-{\color{purple}{Andrew - maybe do show them, briefly?}}
 
@@
 \citeA{HK2020QJRMS} have shown that this occurs in CAM due to higher resolved vertical velocities which, in turn, generate more condensational heating in the CLUBB macrophyiscs.
-The right column in Figure~\ref{fig:dT-lores} supports this interpretation, showing an increase in the climatological CLUBB heating in low ({\color{blue}{looks more like subtropics}}) and mid-latitudes on the $\tt f09$ grid.
+The right column in Figure~\ref{fig:dT-lores}a supports this interpretation, showing an increase in the climatological CLUBB heating at all latitudes in the $\tt f09$ grid, but with the largest increase in the mid-latitudes.
 
@@
 
-As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so the $\tt ne30pg3$ troposphere is modestly warmer than $\tt f09$ (Figure~\ref{fig:dT-lores}).
+As the SE dycore is less diffusive than the FV dycore, the resolved vertical velocities are larger in the SE dycore, and so the $\tt ne30pg3$ troposphere is modestly warmer than $\tt f09$ (Figure~\ref{fig:dT-lores}b).
 The stratosphere responds differently, with $\tt ne30pg3$ much cooler than $\tt f09$ in the mid-to-high latitudes.
-Figure~\ref{fig:dT-lores} also shows small temperature differences between $\tt ne30pg3$ and $\tt ne30pg2$, with  $\tt ne30pg3$ slightly warmer near the tropopause at high latitudes.
+Figure~\ref{fig:dT-lores}c also shows small temperature differences between $\tt ne30pg3$ and $\tt ne30pg2$, with  $\tt ne30pg3$ slightly warmer near the tropopause at high latitudes.
 This is consistent with the similar climates found for these two grids by \citeA{HETAL2019JAMESb}.
@@
 Figure~\ref{fig:dT-hires}a shows the difference between $\tt ne30pg3^{*}$ and $\tt ne30pg3$ for climatological summer temperatures in zonal-mean height space.
-Much of the troposphere is warmer with the reduced time step, and the mechanism is similar in that the shorter time step increases condensational heating by CLUBB.
+The troposphere is warmer with the reduced time step, and the mechanism is similar in that the shorter time step increases resolved vertical velocities (not shown) and CLUBB heating (right panel in Figure~\ref{fig:dT-hires}a).
 Figure~\ref{fig:dT-hires}b shows the difference in climatological summer temperature between the $\tt Arctic$ grid and the $\tt ne30pg3^{*}$ grid.
-With the same physics time step on each grid, the greater condensational heating and warmer temperatures are confined to the refined Arctic region.
-
-Figure~\ref{fig:dT-hires}c shows that the $\tt Arctic-GrIS$ grid is much warmer than the $\tt Arctic$ grid in the Arctic summer. This may be due, in part, to the shorter physics time step,
-%; a similar physics-time-step sensitivity in the $\tt ne30pg3^{*}$ comparison is evident in the CLUBB tendencies in the $ARCTCGRIS$ grid.
-but the temperature response is too large to be explained by the CLUBB changes alone.
+With the same physics time step, the greater condensational heating and warmer temperatures are confined to the refined Arctic region.
+
+Figure~\ref{fig:dT-hires}c shows that the $\tt Arctic-GrIS$ grid is much warmer than the $\tt Arctic$ grid in the Arctic summer.
+This may be due, in part, to the shorter physics time step, but the temperature response is too large to be explained by the CLUBB changes alone.
 This summer warming appears to be a result of variations in the stationary wave pattern, with anomalous southerly winds to the west of Greenland (not shown).
@@
 It is useful to understand summer temperature biases due to their control on ice and snow melt over the GrIS \cite<>{O2001JAM}.
-Figure~\ref{fig:dThyps} shows the 1979--1998 lower troposphere summer temperature bias relative to ERA5.
-It is computed from the 500--1000 hPa geopotential thickness, solving for the layer-mean virtual temperature using the hypsometric equation ({\color{blue}{Might be good to write this down, for those of us who studied atmospheric science a really long time ago}}).
+Figure~\ref{fig:dThyps} shows the 1979--1998 lower troposphere summer temperature bias relative to ERA5, computed by equating a layer mean virtual temperature with the 500--1000 hPa geopotential thickness.
 The results are consistent with the zonal mean height plots; increasing resolution from $\tt f19$ to $\tt f09$ warms the climate, and the $1^{\circ}$ SE grids are warmer than the FV grids.
@@
 This pattern is also evident in the near surface temperature bias over Greenland (not shown).
-{\color{purple}{Marcus - why mention it if it's not shown?}}
-
-The $\tt Arctic$ grid has summer temperatures similar to the $1^{\circ}$ SE grids, but is slightly warmer over northern Eurasia and the North Pole.
+
+The $\tt Arctic$ grid has summer temperatures similar to the $1^{\circ}$ SE grids, but is slightly warmer over northern Eurasia and the North Pole (Figure~\ref{fig:dThyps}).
 An anomalous cooling patch forms to the west of Greenland, centered over Baffin Island.
@@
 A negative bias corresponds to excessive reflection and cooling.
-The uniform grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses ({\color{blue}{please explain what SWCF in caption.}}).
+The uniform grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses.
 There is also a halo of positive bias (clouds not reflective enough) around the ocean perimeter of Greenland.
@@
 Compared to the $\tt Arctic$ grid, the $\tt Arctic-GrIS$ grid vastly reduces the cloud forcing bias over Eurasia, and also improves the bias over North America.
-In both VR grids, the halo of positive forcing bias around the perimeter of Greenland is absent.
+In both VR grids, the halo of positive shortwave cloud forcing bias around the perimeter of Greenland is absent.
 
@@
 \caption{1979-1998 Northern Hemisphere summer shortwave cloud forcing bias, relative to the CERES-EBAF gridded dataset.
+Shortwave cloud forcing is defined as the difference between all-sky and clear-sky net shortwave fluxes at the top of the atmosphere.
 Differences are computed after mapping all model output to the $1^{\circ}$ CERES-EBAF grid.}
@@
 This suggests that the coarse-grid clouds are too thick in the Greenland interior and too thin around the perimeter, and that increasing horizontal resolution reduces these biases.
-{\color{purple}{Marcus - This result is also consistent with the SMB bias in the GrSI interior that plagues low-res simulations.
-Is it worth commenting on that here?}}This is consistent with the total summer cloud fraction bias, computed from the CALIPSO-GOCCP cloud dataset  and shown in Figure~\ref{fig:prect}.
-Note that total cloud fraction characterizes the cloud field at all vertical levels, but attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at all other levels.
+This is consistent with the total summer cloud fraction bias, computed from the CALIPSO-GOCCP cloud dataset  and shown in Figure~\ref{fig:prect}.
+Note that total cloud fraction characterizes the cloud field at all vertical levels, and attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at all other levels.
 Despite the attenuated signal, the total cloud fraction for the VR grids does indicate reduced cloud coverage in the interior and increased cloudiness around the ocean perimeter.
@@
 
-The accuracy of the simulated SMB is expected to be sensitive to grid resolution.
-Figure~\ref{fig:grisdx}b shows the average grid spacing over GrIS in all six grids in this study.
-The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160~km$, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6~km$, broadly similar to the $11~km$ grid spacing of the RACMO2.3 grid.
-The $\tt ne30pg3$ grid has an average $\Delta x=111.2~km$, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60~km$.
-Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
-The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
-The $\tt Arctic-GrIS$ grid is about twice as expensive as the $\tt Arctic$ grid.
-({\color{blue}{Wondering if this paragraph would go better in an earlier section.
-Rene agrees.
-Marcus agrees (move to 2.2)}})
-
 The lower panels of Figure~\ref{fig:prect} show the summer climatological mean precipitation bias over the GrIS, expressed as the fractional difference from the RACMO2.3p2 solution.
@@
 The $\tt Arctic$ grid reduces this bias substantially, and the $\tt Arctic-GrIS$ grid reduces it further.
-This suggests that the southern dome bias arises from inadequate horizontal resolution, consistent with the original GrIS VR experiments in \citeA{VETAL2018TC}.
+This suggests that the southern dome bias arises from inadequate horizontal resolution, and consistent with the positive GrIS accumulation biases that are common to low resolution GCMs \citeA{P2000GPC,VETAL2018TC}.
 
@@
    \hline
-   $RACMO$ & 768.5 (733.5) & -347.2 (-436.4) & -221.7 (-258.5) & -36.5 (-38.8) & 510.3 (436.2) \\
+   $RACMO$ & 681.7 (733.5) & -318.6 (-436.4) & -189.1 (-258.5) & -34.5 (-38.8) & 458.1 (436.2) \\
    \hline
-   $\tt f19$ & 882.5 (913.5) & -440.3 (-546.5) & -283.7 (-284.3) & -36.6 (-37.5) & 562.2 (591.7) \\
-   $\tt f09$ & 874.8 (882.1) & -418.4 (-482.3) & -255.0 (-212.3) & -38.1 (-37.4) & 581.7 (632.4) \\
-   $\tt ne30pg2$ & 1000. (973.4) & -549.4 (-647.3) & -383.9 (-347.0) & -33.4 (-32.1) & 582.7 (594.3) \\
-   $\tt ne30pg3$ & 934.9 (909.3) & -568.8 (-686.7) & -356.2 (-330.1) & -34.4 (-32.6) & 544.3 (546.6) \\
-   $\tt Arctic$ & 795.9 (818.6) & -367.3 (-436.8) & -208.9 (-194.2) & -44.1 (-43.9) & 542.9 (580.5) \\
-   $\tt Arctic-GrIS$ & 708.7 (747.3) & -471.6 (-610.4) & -261.1 (-307.8) & -50.7 (-51.8) & 396.9 (387.7) \\
+   $\tt ne30pg2$ & 1007. (973.4) & -519.9 (-647.3) & -381.9 (-347.0) & -33.9 (-32.1) & 591.2 (594.3) \\
+   $\tt ne30pg3$ & 931.0 (909.3) & -540.8 (-686.7) & -375.8 (-330.1) & -34.1 (-32.6) & 521.2 (546.6) \\
+   $\tt f19$ & 884.9 (913.5) & -414.0 (-546.5) & -284.0 (-284.3) & -36.5 (-37.5) & 564.4 (591.7) \\
+   $\tt f09$ & 873.9 (882.1) & -389.1 (-482.3) & -256.1 (-212.3) & -37.3 (-37.4) & 580.5 (632.4) \\
+   $\tt Arctic$ & 784.1 (818.6) & -335.5 (-436.8) & -215.8 (-194.2) & -42.4 (-43.9) & 526.0 (580.5) \\
+   $\tt Arctic-GrIS$ & 693.8 (747.3) & -437.3 (-610.4) & -276.8 (-307.8) & -48.1 (-51.8) & 369.0 (387.7) \\
  \hline
@@
   \caption{1979-1998 surface mass balance of the Greenland Ice Sheet in Gt/yr.
-Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.
-{\color{red}{ARH - these numbers need updating ...
-they are only for method 1.}}}
+Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.}
  \label{tbl:table3}

### 2022-05-05 arcticgrids.tex
paper updates

@@
 
+The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution.
+Figure~\ref{fig:grisdx}b shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids in this study.
+The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid.
+The $\tt ne30pg3$ grid has an average $\Delta x=111.2$ km, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60$ km.
+Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
+The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km, and is about 10 times more expensive than the $1^{\circ}$ models.
+The $\tt Arctic-GrIS$ grid is about twice as expensive as the $\tt Arctic$ grid.
+
 The physics time step depends on the grid resolution.
@@
 This will degrade the SMB estimates using method~\ref{lbl:label2}, however we are more interested in characterizing the behavior of individual processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
-
-The accuracy of the simulated SMB is expected to be sensitive to grid resolution.
-Figure~\ref{fig:grisdx}b shows the average grid spacing over GrIS in all six grids in this study.
-The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160~km$, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6~km$, broadly similar to the $11~km$ grid spacing of the RACMO2.3 grid.
-The $\tt ne30pg3$ grid has an average $\Delta x=111.2~km$, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60~km$.
-Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
-The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models.
-The $\tt Arctic-GrIS$ grid is about twice as expensive as the $\tt Arctic$ grid.
 
@@
 Some of these temperature differences may be related to different summer shortwave cloud forcing for the various grids and dycores.
-Figure~\ref{fig:SWCF} shows the summer shortwave cloud forcing bias in six runs, using the CERES-EBAF product.
+Figure~\ref{fig:SWCF} shows the summer shortwave cloud forcing bias in six runs, using the CERES product.
 A negative bias corresponds to excessive reflection and cooling.
@@
 
-\subsection{Shortwave radiation over Greenland}
+\subsection{Clouds and precipitation over Greenland}
 
@@
 Figure~\ref{fig:FSDS} shows the summer incident shortwave radiation bias at the surface over Greenland and surrounding seas.
-The top panel shows the bias relative to the CERES-EBAF dataset, and the bottom panel relative to the RACMO2.3p2 dataset.
+The middle panel shows the bias relative to the CERES dataset, and the bottom panel relative to the RACMO2.3p2 dataset.
 The halo of excessive incident shortwave radiation around the coasts of Greenland is apparent for both datasets in relation to the coarser grids, consistent with the shortwave cloud forcing biases in Figure~\ref{fig:SWCF}.
 
-The ice sheet interior receives too little shortwave radiation on the coarser grids.
+The ice sheet interior receives too little shortwave radiation in the coarser grids.
 On the VR grids, both the interior shortwave deficit and the excessive shortwave around the ocean perimeter are improved.
-This suggests that the coarse-grid clouds are too thick in the Greenland interior and too thin around the perimeter, and that increasing horizontal resolution reduces these biases.
-This is consistent with the total summer cloud fraction bias, computed from the CALIPSO-GOCCP cloud dataset  and shown in Figure~\ref{fig:prect}.
+This suggests that the coarse-grid clouds are too thick in the Greenland interior and too thin around the perimeter.
+This is consistent with the total summer cloud fraction bias, computed from the CALIPSO cloud dataset and shown in the top panel of Figure~\ref{fig:FSDS}.
 Note that total cloud fraction characterizes the cloud field at all vertical levels, and attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at all other levels.
-Despite the attenuated signal, the total cloud fraction for the VR grids does indicate reduced cloud coverage in the interior and increased cloudiness around the ocean perimeter.
-
-The agreement of the cloud biases in and around Greenland from multiple independent datasets shows that the biases are a robust feature of the coarser grids.
-The reduced biases on the VR grids suggest that the coarse-grid biases are a result of insufficient horizontal resolution.
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_FSDS.pdf}
-\end{center}
-\caption{1979-1998 northern hemisphere summer, incident shortwave radiation bias, computed as the difference (top) from CERES-EBAF, and (bottom) RACMO2.3p2 dataset.
-The differences in the top panel are found by mapping the model output to the $1^{\circ}$ CERES-EBAF grid, and differences on the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
-Note that the averaging period for the CERES-EBAF panels, 2003-2020, is different from the averaging period for the model results.}
+Despite the attenuated signal, the cloud fractions are too high in the interior and and too low around the oceanic perimeter, in the coarser grids, while the VR grids have an overall reduction in total cloud fraction bias.
+
+The top panel of Figure~\ref{fig:prect} show the annual climatological mean precipitation bias over the GrIS, expressed as the fractional difference from the RACMO2.3p2 solution.
+The coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
+The $\tt Arctic$ grid reduces this bias substantially, and the $\tt Arctic-GrIS$ grid reduces it further, with precipitation centers migrating from the interior toward the margins.
+
+The more accurate representation of orographic precipitation on the higher resolution grids is consistent with the cloud and radiation biases, cf.
+Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, and \ref{fig:prect}.
+The agreement of the cloud and precipitation biases in and around Greenland from multiple independent datasets indicates that the biases are a robust feature of the coarser grids.
+The reduced biases on the VR grids suggest that the biases in the coarse models are a result of insufficient horizontal resolution.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=130mm]{figs/temp_contours_diffCALIPSOdiffCERESdiffRACMO.pdf}
+\end{center}
+\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias relative to the CALIPSO dataset and incident shortwave radiation bias (W/m2), computed as the difference (middle) from CERES, and (bottom) RACMO2.3p2 dataset.
+The CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
+Note that the averaging period for the CALIPSO-GOCCP and CERES-EBAF panels, 2006-2017 and 2003-2020, respectively, are different from the averaging period for the model results.}
 \label{fig:FSDS}
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_contours_diffCERESdiffRACMO_CLOUD_PRECIP.pdf}
-\end{center}
-\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias, relative to the CALIPSO-GOCCP dataset, and (bottom) precipitation rate bias, relative to the RACMO2.3p2 dataset.
-The CALIPSO-GOCCP differences are computed after mapping all model output to the $1^{\circ}$ grid, whereas the RACMO differences are computed after mapping the RACMO dataset to the individual model grids.
-Note that the averaging period for the CALIPSO-GOCCP panels (2006-2017) is different than the model averaging period.
-\color{red}{ARH - still trying to fix the layout of this figure so the label bars aren't on top of the panels}.}
+         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO}
+\end{center}
+\caption{1979-1998 (top) annual precipitation and (bottom) ice/snow melt biases relative to RACMO2.3p2, evaluated on the native model grids.
+The precipitation biases are expressed as fractional changes, whereas the melt biases are absolute changes (mm/yr).
+In the bottom panel, the \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid.
+Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $\tt f19$ and $\tt ne30pg2$ panels, in computing the basin-scale integrals.
+Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}..}
 \label{fig:prect}
@@
 \subsection{Greenland surface mass balance}
-
-The lower panels of Figure~\ref{fig:prect} show the summer climatological mean precipitation bias over the GrIS, expressed as the fractional difference from the RACMO2.3p2 solution.
-The coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
-The $\tt Arctic$ grid reduces this bias substantially, and the $\tt Arctic-GrIS$ grid reduces it further.
-This suggests that the southern dome bias arises from inadequate horizontal resolution, and consistent with the positive GrIS accumulation biases that are common to low resolution GCMs \citeA{P2000GPC,VETAL2018TC}.
-
-Large GrIS accumulation rates result from synoptic systems arriving primarily from the south.
-These systems are orographically lifted at the ice sheet margin, especially over the steep slopes in southeast Greenland (dominated by winter precipitation), concentrating heavy precipitation near the ice margin.
-At lower resolutions, the topography is too smooth to generate a realistic orographic uplift, leading to anomalously high SMB in the GrIS interior.
-On the higher resolution grids, precipitation centers migrate from the interior toward the margins, and even out over the ocean in the case of the $\tt Arctic-GrIS$ grid.
-The more accurate representation of orographic precipitation on the higher resolution grids is consistent with the discussion on simulated cloud and radiation biases, cf.
-Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, and \ref{fig:prect}.
 
@@
 To illustrate the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \citeA{RM2012GRL}.
-The uncertainty due to differences in basin area is larger than for GrIS-wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which are shown in the $\tt f19$ and $\tt ne30pg2$ panels of Figure~\ref{fig:melt}.
+The uncertainty due to differences in basin area is larger than for GrIS-wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which are shown in the $\tt f19$ and $\tt ne30pg2$ panels of Figure~\ref{fig:prect}.
 Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the most accumulation.
@@
 Whiskers span the max/min of the four ensemble members generated from the common-ice-mask approach.
-Basin definitions are after \citeA{RM2012GRL}, and are found on the common ice masks using a nearest neighbor approach, and shown in Figure~\ref{fig:melt}.}
+Basin definitions are after \citeA{RM2012GRL}, and are found on the common ice masks using a nearest neighbor approach, and shown in Figure~\ref{fig:prect}.}
 \label{fig:basin}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO_melt.pdf}
-\end{center}
-\caption{1979--1998 ice/snow melt biases (in mm/yr) relative to RACMO2.3p2, evaluated on the native model grids.
-The \citeA{RM2012GRL} basin boundaries are shown in grey for each model grid.
-Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $\tt f19$ and $\tt ne30pg2$ panels, in computing the basin-scale integrals.
-Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
-\label{fig:melt}
 \end{figure}
@@
 
-Figure~\ref{fig:melt} presents the biases in the combined ice/snow melt as map plots.
+Figure~\ref{fig:prect} presents the biases in the combined ice/snow melt as map plots.
 These plots show that the largest melt biases are on the southeast and northwest coasts, where large coarse-grid cells overlap with the ocean.
@@
 Figure~\ref{fig:ztrans} shows the representation of the ice sheet surface along two transects on the different grids, compared to the high-resolution dataset used to generate CAM topographic boundary conditions \cite{GMTED2010,gmdd-8-4623-2015}.
-The two transects are shown in Figure~\ref{fig:melt}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast.
+The two transects are shown in Figure~\ref{fig:prect}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast.
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a few grid cells populating the transect.

### 2022-05-05 arcticgrids.tex
mods

@@
 
-While the summer cloud forcing biases are consistent with the summer temperature biases in Figure~\ref{fig:dThyps} -- regions where clouds are too reflective coincide with regions that are too cold -- it is not clear whether the cold biases are caused by the cloud biases, or whether the cold biases amplify the cloud forcing bias.
-({\color{blue}{This ends abruptly.
-Anything else to add?}} {\color{purple}{Causation I agree is hard to prove due to energy transport, but in summer cloud fraction is probably playing a role.
-The likely other thing that could be playing a role is surface albedo (absorbed solar).
-Could note that.}})
+The summer cloud forcing biases are consistent with the summer temperature biases in Figure~\ref{fig:dThyps} -- regions where clouds are too reflective coincide with regions that are too cold.
+While we have not quantified the role of the clouds in contributing to the cooler Arctic temperatures, shortwave radiation is a crucial component of the Arctic energy budget during summer.
+The shortwave cloud forcing biases are on the order of 10 W/m2, which is a significant fraction of the total absorbed shortwave during Arctic summer  \cite{SETAL2007JGR} and is therefore likely a factor contributing to the cooler temperatures.
 
@@
 This is consistent with the total summer cloud fraction bias, computed from the CALIPSO cloud dataset and shown in the top panel of Figure~\ref{fig:FSDS}.
-Note that total cloud fraction characterizes the cloud field at all vertical levels, and attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at all other levels.
-Despite the attenuated signal, the cloud fractions are too high in the interior and and too low around the oceanic perimeter, in the coarser grids, while the VR grids have an overall reduction in total cloud fraction bias.
+Note that total cloud fraction characterizes the cloud field at all vertical levels, and attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at other levels.
+Despite the attenuated signal, the cloud fractions are too high in the interior and and too low around the oceanic perimeter in the coarser grids, whereas the VR grids have an overall improvement in total cloud fraction bias.
 
@@
 
-The more accurate representation of orographic precipitation on the higher resolution grids is consistent with the cloud and radiation biases, cf.
+The more accurate representation of orographic precipitation in the VR grids is consistent with the cloud and radiation biases, cf.
 Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, and \ref{fig:prect}.

### 2022-05-05 arcticgrids.tex
mods

@@
 Figure~\ref{fig:FSDS} shows the summer incident shortwave radiation bias at the surface over Greenland and surrounding seas.
-The middle panel shows the bias relative to the CERES dataset, and the bottom panel relative to the RACMO2.3p2 dataset.
+The top panel shows the bias relative to the RACMO2.3p2 dataset, and the middle panel relative to the CERES dataset.
 The halo of excessive incident shortwave radiation around the coasts of Greenland is apparent for both datasets in relation to the coarser grids, consistent with the shortwave cloud forcing biases in Figure~\ref{fig:SWCF}.
@@
 This suggests that the coarse-grid clouds are too thick in the Greenland interior and too thin around the perimeter.
-This is consistent with the total summer cloud fraction bias, computed from the CALIPSO cloud dataset and shown in the top panel of Figure~\ref{fig:FSDS}.
+This is consistent with the total summer cloud fraction bias, computed from the CALIPSO cloud dataset and shown in the bottom panel of Figure~\ref{fig:FSDS}.
 Note that total cloud fraction characterizes the cloud field at all vertical levels, and attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at other levels.

### 2022-05-06 arcticgrids.tex
mods

@@
 \citeA{HK2020QJRMS} have shown that this occurs in CAM due to higher resolved vertical velocities which, in turn, generate more condensational heating in the CLUBB macrophyiscs.
-The right column in Figure~\ref{fig:dT-lores}a supports this interpretation, showing an increase in the climatological CLUBB heating at all latitudes in the $\tt f09$ grid, but with the largest increase in the mid-latitudes.
+The right panel in Figure~\ref{fig:dT-lores}a supports this interpretation, showing an increase in the climatological CLUBB heating at all latitudes in the $\tt f09$ grid, but with the largest increase in the mid-latitudes.
 
@@
 Figure~\ref{fig:dT-hires}c shows that the $\tt Arctic-GrIS$ grid is much warmer than the $\tt Arctic$ grid in the Arctic summer.
-This may be due, in part, to the shorter physics time step, but the temperature response is too large to be explained by the CLUBB changes alone.
-This summer warming appears to be a result of variations in the stationary wave pattern, with anomalous southerly winds to the west of Greenland (not shown).
+This may be due, in part, to the shorter physics time step, but the temperature response is too large to be explained by enhanced condensational heating from CLUBB alone.
+This summer warming appears to be a result of variations in the stationary wave pattern, with a swath of anomalous southerly winds to the west of Greenland (not shown).
 This dynamic response is interesting, because other than the physics time step, the only difference between the $\tt Arctic-GrIS$ and $\tt Arctic$ runs is the doubling of resolution over Greenland.
@@
 
-Some of these temperature differences may be related to different summer shortwave cloud forcing for the various grids and dycores.
-Figure~\ref{fig:SWCF} shows the summer shortwave cloud forcing bias in six runs, using the CERES product.
+Some of these temperature differences may be related to differences in summer cloudiness.
+Figure~\ref{fig:SWCF} shows the summer shortwave cloud forcing bias in the six runs, using the CERES product.
+Shortwave cloud forcing quantifies the impact of clouds on shortwave radiation, taken as the difference between all-sky and clear-sky shortwave radiative fluxes at the top of the atmosphere.
 A negative bias corresponds to excessive reflection and cooling.
@@
 The summer cloud forcing biases are consistent with the summer temperature biases in Figure~\ref{fig:dThyps} -- regions where clouds are too reflective coincide with regions that are too cold.
-While we have not quantified the role of the clouds in contributing to the cooler Arctic temperatures, shortwave radiation is a crucial component of the Arctic energy budget during summer.
+While we have not quantified the contribution of cloud biases to the cooler Arctic temperatures, shortwave radiation is a crucial component of the Arctic energy budget during summer.
 The shortwave cloud forcing biases are on the order of 10 W/m2, which is a significant fraction of the total absorbed shortwave during Arctic summer  \cite{SETAL2007JGR} and is therefore likely a factor contributing to the cooler temperatures.
@@
 On the VR grids, both the interior shortwave deficit and the excessive shortwave around the ocean perimeter are improved.
-This suggests that the coarse-grid clouds are too thick in the Greenland interior and too thin around the perimeter.
-This is consistent with the total summer cloud fraction bias, computed from the CALIPSO cloud dataset and shown in the bottom panel of Figure~\ref{fig:FSDS}.
+This suggests that the coarse grid clouds are too thick in the interior of Greenland and too thin around the perimeter, which is consistent with the total summer cloud fraction bias, computed from the CALIPSO cloud dataset and shown in the bottom panel of Figure~\ref{fig:FSDS}.
 Note that total cloud fraction characterizes the cloud field at all vertical levels, and attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at other levels.
-Despite the attenuated signal, the cloud fractions are too high in the interior and and too low around the oceanic perimeter in the coarser grids, whereas the VR grids have an overall improvement in total cloud fraction bias.
+The VR grids exhibit an overall improvement in total cloud fraction bias, relative to the coarse grids.
 
@@
 Figures~\ref{fig:SWCF}, \ref{fig:FSDS}, and \ref{fig:prect}.
-The agreement of the cloud and precipitation biases in and around Greenland from multiple independent datasets indicates that the biases are a robust feature of the coarser grids.
-The reduced biases on the VR grids suggest that the biases in the coarse models are a result of insufficient horizontal resolution.
+The agreement of the cloud, radiation and precipitation biases in and around Greenland from multiple independent datasets indicates that the biases are a robust feature of the coarser grids.
+The reduced biases in the VR grids suggest that the deficiencies of the coarse grids are due to insufficient horizontal resolution, consistent with previous findings that coarse GCMs have large, positive precipitation biases over Greenland \cite{P2000GPC,VETAL2018TC}.
 
@@
 
-Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO {\color{purple}{Andrew - do you want to talk about the values at all?}}.
-The CESM values are averages over the ensemble of common ice masks and regridding methods described in section~\ref{sec:SMB}, and the RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}) using the same common-ice-mask approach.
-Table~\ref{tbl:table3} also contains (in parentheses) the SMB components derived from evaluating the integrals on each model's native grid and ice mask.
-Of note is the large reduction in melt rates using the common-ice-mask approach compared to the native grid, illustrating the dissipation discussed in section~\ref{sec:SMB}.
-The errors are greatest in partially ice-covered grid cells straddling the ice sheet margins, in the ablation zone where melt rates are large.
+Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO.
+The values in the table are averages over the ensemble of mapping methods to the common ice masks described in section~\ref{sec:SMB}, and the RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}).
+Table~\ref{tbl:table3} also contains (in parenthesis) the SMB components derived from evaluating the integrals on each model's native grid and ice mask.
+Of note is the large reduction in melt rates compared to the values computed on the native grid, illustrating the dissipation of this quantity discussed in section~\ref{sec:SMB}.
 For integrated precipitation, the differences between the native and common-ice-mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
 
-Figure~\ref{fig:tseries} shows time series of annually integrated precipitation and snow/ice melt over the GrIS for the various different grids and dycores, with both versions of RACMO shown in black.
+One of the most interesting aspects of the SMB calculations is that the coarse grids are characterized by too much precipitation and too much melting and runoff, compared with RACMO.
+The SMB in the coarse grids therefore have smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
+Such compensating errors highlights the importance of understanding the extent to which a model is getting the SMB right for the right reasons.
+
+Figure~\ref{fig:tseries} shows time series of annually integrated precipitation and snow/ice melt over the GrIS for the various different grids and dycores, and RACMO in black.
 The 1979-1998 climatological mean values, listed in Table~\ref{tbl:table3}, are shown as circles on the right side of the panels.
-The uniform $1^{\circ}-2^{\circ}$ grids have positive precipitation biases in the interior, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
+The uniform $1^{\circ}-2^{\circ}$ grids have positive precipitation biases, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
 The $\tt f19$ and $\tt f09$ grids perform similarly, with +110 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +165 Gt/yr and $\tt ne30pg2$ by +230 Gt/yr.
@@
 The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $\tt Arctic$ grid simulates the most realistic melt rates, with the other grids having more melt than RACMO.
-The $\tt Arctic-GrIS$ grid overpredicts melting by about 125 Gt/yr.
+The $\tt Arctic-GrIS$ grid over-predicts melting by about 125 Gt/yr.
 This is likely due to an anomalously warm lower troposphere during the summer, relative to the $\tt Arctic$ run (Figure~\ref{fig:dThyps}).
@@
 The SE grids have the largest positive melt bias, between 200--220 Gt/yr.
-It is more difficult to attribute these differences to resolution alone, since the FV grids have colder summer temperatures than the uniform-resolution SE grids.
+%It is more difficult to attribute these differences to resolution alone, since the FV grids have colder summer temperatures than the uniform-resolution SE grids.
 ({\color{blue}{Not sure I understand this sentence.
@@
 In all basins, accumulation decreases monotonically with increasing grid resolution, though with some exceptions.
-The $\tt Arctic-GrIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to the RACMO precipitation in the large southwest basin.
+The $\tt Arctic-GrIS$ grid simulates less precipitation than RACMO in the central-east and southeast basins, and is closest of all grids to RACMO in the large southwest basin.
 
@@
 
-Figure~\ref{fig:prect} presents the biases in the combined ice/snow melt as map plots.
+The bottom panel of Figure~\ref{fig:prect} presents the biases in the combined ice/snow melt as map plots.
 These plots show that the largest melt biases are on the southeast and northwest coasts, where large coarse-grid cells overlap with the ocean.
@@
 The two transects are shown in Figure~\ref{fig:prect}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast.
-The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a few grid cells populating the transect.
-The $\tt f09$ grid is a bit of an exception for the K-transect, with grid cells becoming narrow in the meridional direction at high latitudes.
+The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
+The $\tt f09$ grid is a bit of an exception, since the grid cells become narrow in the meridional direction at high latitudes, and a larger number of grid cells populate an east-west transect.
 The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of the GrIS margins.
@@
 Due to this abrupt transition, coarse grids will commonly represent the ablation zone with grid cells containing mixtures of ice-covered and ice-free regions.
-We hypothesize that coarser models have larger melt biases
-%despite the $ARCTCIGRIS$ having the warmest troposphere,
-because summer melting is confined to these mixed ice/land/ocean grid cells.
-CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees a homogenized state due to volume averaging over the sub-grid mixture.
+We hypothesize that coarser models have larger melt biases because summer melting is confined to these mixed ice/land/ocean grid cells.
+CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees the homogenized state after volume averaging over the sub-grid mixture.
 Thus, warm ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice-covered patch.
-({\color{blue}{This is an interesting conclusion pointing to the need for better treatment of surface inhomogeneity in CAM.
-This might be a way to compensate for coarse resolution in future CESM versions?}})
 
 Figure~\ref{fig:bias} shows mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region.
-Errors are computed using the common-ice-mask approach, meaning that all fields are mapped to the common masks, which define the grid cell ice fraction.
-The figure shows ({\color{blue}{Any idea why the errors are smaller in the cells with intermediate ice fraction?
-I wouldn't have expected this.}}) that coarser grids generally have two peaks in ice fraction space; a bump in positive melting errors in the 0-20$\%$ range, and another in the fully ice-covered cells.
+Errors are computed using the common ice mask approach described in section~\ref{sec:SMB}, and so the grid cell ice fraction are taken from the common masks.
+The figure shows that coarser grids generally have two peaks in ice fraction space; a bump in positive melting errors in the 0-20$\%$ range, and another in the fully ice-covered cells.
 Also shown are the $\pm 1$~standard deviation of the biases for each bin.
@@
 
+({\color{blue}{Any idea why the errors are smaller in the cells with intermediate ice fraction?  I wouldn't have expected this.}})
 {\color{green}{Rene - is the melt map consistent with the bin figure? Smallest errors are interior points which correspond to glacier fraction bin of 1, which the largest errors.}}

### 2022-05-06 arcticgrids.tex
new figs

@@
 \end{center}
-\caption{Spectral-element grid for the VR grids in this study, (a) $\tt Arctic$, (b) $\tt Arctic-GrIS$ and (c) $\tt CONUS$.
-Note that this is not the computational grid; each element has $3\times3$ independent grid points.}
+\caption{Variable-resolution grids available in CESM2.2, (a) $\tt Arctic$, (b) $\tt Arctic-GrIS$ and (c) $\tt CONUS$.
+Note what is shown is the element grid; the computational grid has $3\times3$ independent grid points per element.}
 \label{fig:vr-grids}
@@
 Figure~\ref{fig:tseries} shows time series of annually integrated precipitation and snow/ice melt over the GrIS for the various different grids and dycores, and RACMO in black.
-The 1979-1998 climatological mean values, listed in Table~\ref{tbl:table3}, are shown as circles on the right side of the panels.
+The 1979-1998 climatological mean values from Table~\ref{tbl:table3} are shown as circles on the right side of the panels.
 The uniform $1^{\circ}-2^{\circ}$ grids have positive precipitation biases, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.

### 2022-05-07 arcticgrids.tex
mods

@@
 The $\tt f19$ and $\tt f09$ grids perform similarly, with +110 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +165 Gt/yr and $\tt ne30pg2$ by +230 Gt/yr.
-The larger biases on the uniform-resolution SE grids relative to the FV grids are consistent with the coarser GrIS resolution on the SE grids (Figure~\ref{fig:grisdx}).
+The larger biases on the uniform-resolution SE grids relative to the FV grids are consistent with the coarser GrIS resolution in the SE grids (Figure~\ref{fig:grisdx}).
 
@@
 Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones.
-Positive degree-days \cite<PDD;>{B1984Z}, which accumulate the near-surface temperature in $^{\circ}C$ for days with temperature above freezing, are a more accurate proxy.
+The Positive degree-day temperature based melt index \cite<PDD;>{B1984Z}, which accumulate the near-surface temperature in $^{\circ}C$ for days with temperature above freezing, is a more accurate proxy for melting.
 PDD is nonlinear in mean monthly temperature \cite{R1991P}.
-We compute it from monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}$C and a degree-day factor of 5~mm~d$^{-1}$ $^{\circ}$C$^{-1}$.
+We compute PDD from monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}$C and a degree-day factor of 5~mm~d$^{-1}$ $^{\circ}$C$^{-1}$.
 
@@
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $\tt f09$ grid is a bit of an exception, since the grid cells become narrow in the meridional direction at high latitudes, and a larger number of grid cells populate an east-west transect.
-The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of the GrIS margins.
+The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and a larger number of grid cells populate an east-west transect.
+The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of the GrIS margin.
 
@@
 Rather, the smoothing and flattening of the raw topography, necessary to prevent the model from exciting grid-scale numerical modes, causes the lower-elevation ablation zones to extend beyond the true ice sheet margin, where they lie above the actual ice surface.
-The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest ($\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ use identical smoothing).
-This suggests that coarser models will tend to elevate the ablation zones and thereby depress melt rates.
+The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids use identical smoothing.
+This suggests that coarser models will tend to elevate the ablation zones and thereby depress melt rates, which is the opposite of what occurs in the coarse grid simulations.
 
@@
 
-Figure~\ref{fig:bias} shows mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region.
-Errors are computed using the common ice mask approach described in section~\ref{sec:SMB}, and so the grid cell ice fraction are taken from the common masks.
-The figure shows that coarser grids generally have two peaks in ice fraction space; a bump in positive melting errors in the 0-20$\%$ range, and another in the fully ice-covered cells.
+Figure~\ref{fig:bias} shows the mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region.
+Errors are computed after mapping to the common ice masks using different methods, described in section~\ref{sec:SMB}.
+The grid cell ice fractions therefore pertain to ice fractions on the low-resolution common ice masks.
 Also shown are the $\pm 1$~standard deviation of the biases for each bin.
-They indicate that the biases in 0-20$\%$ bins are mostly contained in the positive bias region (a fractional bias greater than 0), whereas the fully-covered ice cells have a wider distribution, with many grid cells also containing negative melting biases.
-The excessive melting in the 0-20$\%$ ice fraction bins supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone on coarse grids is responsible for their large melt bias.
-
-({\color{blue}{Any idea why the errors are smaller in the cells with intermediate ice fraction?  I wouldn't have expected this.}})
-{\color{green}{Rene - is the melt map consistent with the bin figure? Smallest errors are interior points which correspond to glacier fraction bin of 1, which the largest errors.}}
-{\color{blue}{ARH - I can check. The bin figure is fractional change so its a different metric, which could explain the apparent inconsistency.}}
+The figure shows that coarser grids can be characterized by a monotonic increase in melt bias as the grid cell ice fraction decreases.
+The VR grids have the smallest melt biases for small grid cell ice fractions (smaller than 50\%), the uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases.
+Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone in coarse grids is responsible for their large melt bias.
 
@@
 As the $\tt Arctic$ grid contains $1/4^{\circ}$ refinement north of about $45^{\circ}$ latitude, the storm tracker is applied to this region for the $\tt Arctic$ and $\tt ne30pg3$ runs to identify differences in storm characteristics due to horizontal resolution.
-The composite mean precipitation maps are similar between the two grids, and exhibit the iconic comma structure of synoptic cyclones (not shown).
-{\color{purple}{Marcus - no need to mention if not shown.}}
 

### 2022-05-10 arcticgrids.tex
revised conclusions

@@
 
-One of the most interesting aspects of the SMB calculations is that the coarse grids are characterized by too much precipitation and too much melting and runoff, compared with RACMO.
-The SMB in the coarse grids therefore have smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
+The coarse grids are characterized by too much precipitation and too much melting and runoff, compared with RACMO.
+The SMB in coarse grids therefore have smaller errors than their individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
 Such compensating errors highlights the importance of understanding the extent to which a model is getting the SMB right for the right reasons.
@@
 The $\tt f19$ and $\tt f09$ grids perform similarly, with +110 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +165 Gt/yr and $\tt ne30pg2$ by +230 Gt/yr.
-The larger biases on the uniform-resolution SE grids relative to the FV grids are consistent with the coarser GrIS resolution in the SE grids (Figure~\ref{fig:grisdx}).
 
@@
 The SE grids have the largest positive melt bias, between 200--220 Gt/yr.
-%It is more difficult to attribute these differences to resolution alone, since the FV grids have colder summer temperatures than the uniform-resolution SE grids.
-({\color{blue}{Not sure I understand this sentence.
-FV is both cooler and higher-resolution than SE, so one might suspect that cooler T is, in fact, connected to higher resolution.}})
 
 To illustrate the regional behavior of the SMB components, Figure~\ref{fig:basin} shows the precipitation and combined snow/ice melt integrated over the basins defined by \citeA{RM2012GRL}.
-The uncertainty due to differences in basin area is larger than for GrIS-wide integrals, owing to the differences in basin boundaries as represented by the common ice masks, which are shown in the $\tt f19$ and $\tt ne30pg2$ panels of Figure~\ref{fig:prect}.
+The uncertainty due to differences in basin area is larger than for GrIS-wide integrals, owing to the differences in basin boundaries on the common ice masks, which are shown in the $\tt f19$ and $\tt ne30pg2$ panels of Figure~\ref{fig:prect}.
 Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the most accumulation.
@@
 Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones.
-The Positive degree-day temperature based melt index \cite<PDD;>{B1984Z}, which accumulate the near-surface temperature in $^{\circ}C$ for days with temperature above freezing, is a more accurate proxy for melting.
+The Positive Degree-Day temperature based melt index \cite<PDD;>{B1984Z}, which accumulate the near-surface temperature in $^{\circ}C$ for days with temperature above freezing, is a more accurate proxy for melting.
 PDD is nonlinear in mean monthly temperature \cite{R1991P}.
@@
 The FV grids also have large PDD-based melt in the southwest basin, relative to $\tt Arctic-GrIS$.
-The PDD plots indicate that the near-surface temperatures (which contribute to melt) are not well approximated by the summer lower-troposphere temperatures in Figure~\ref{fig:dThyps}.
+The PDD plots indicate that the relationship between temperature and melt is not well approximated by the summer lower-troposphere temperatures in Figure~\ref{fig:dThyps}.
 
@@
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and a larger number of grid cells populate an east-west transect.
-The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the parabolic shape of the GrIS margin.
+The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and a larger number of grid cells can line the east-west transects.
+The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.
 
@@
 The transects in Figure~\ref{fig:ztrans} show that the ice sheet surface on the coarse grids is not systematically lower than the true surface in ablation zones.
-Rather, the smoothing and flattening of the raw topography, necessary to prevent the model from exciting grid-scale numerical modes, causes the lower-elevation ablation zones to extend beyond the true ice sheet margin, where they lie above the actual ice surface.
-The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids use identical smoothing.
-This suggests that coarser models will tend to elevate the ablation zones and thereby depress melt rates, which is the opposite of what occurs in the coarse grid simulations.
+Rather, the smoothing and flattening of the raw topography, necessary to prevent the model from exciting grid-scale numerical modes, causes the lower-elevation ablation zones to extend beyond the true ice sheet margin, causing the modeled ablation zones (which must reside within the ice sheet mask) to be elevated relative to the actual ice surface.
+The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids have similar dynamical resolution and use identical smoothing.
+This suggests that coarser models will tend to elevate the ablation zones and thereby depress melt rates, which is the opposite of the melt bias that occurs in the coarse grid simulations.
 
@@
 Figure~\ref{fig:bias} shows the mean melt bias, relative to both RACMO datasets, conditionally sampled based on grid cell ice fraction in the GrIS region.
-Errors are computed after mapping to the common ice masks using different methods, described in section~\ref{sec:SMB}.
+Errors are computed after mapping the melt rates to the common ice masks using different methods, described in section~\ref{sec:SMB}.
 The grid cell ice fractions therefore pertain to ice fractions on the low-resolution common ice masks.
@@
 The VR grids have the smallest melt biases for small grid cell ice fractions (smaller than 50\%), the uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases.
-Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone in coarse grids is responsible for their large melt bias.
+Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone of coarse grids is responsible for their large melt bias.
 
@@
 Running CESM2.2 in an AMIP-style configuration, we have evaluated six grids from two dynamical cores for their performance over the Arctic and in simulating the GrIS SMB.
-The $1-2^{\circ}$ finite-volume grids have enhanced resolution over polar regions due to the convergence of meridian lines, although a polar filter is used to prevent spurious atmospheric features from forming at this higher resolution.
-Spectral-element grids comparable to the resolution of the finite-volume grids have an isotropic grid structure where the grid resolution is similar over the entire model domain.
+The $1-2^{\circ}$ finite-volume grids have enhanced resolution over polar regions due to their convergence of meridian lines, although a polar filter is used to prevent spurious atmospheric features from forming in these regions.
+SE grids comparable to the resolution of the FV grids have an isotropic grid structure where the grid resolution is similar over the entire model domain.
 We developed two VR grids and introduced them into CESM2.2 as part of this work.
-Both use the spectral-element dycore; the $\tt Arctic$ grid has $1/4^{\circ}$ refinement over the broader Arctic, whereas the $\tt Arctic-GrIS$ grid is identical except for a $1/8^{\circ}$ patch of refinement over Greenland.
-A third VR grid, $\tt CONUS$, is also available in CESM2.2.
+Both use the SE dycore; the $\tt Arctic$ grid has $1/4^{\circ}$ refinement over the broader Arctic, whereas the $\tt Arctic-GrIS$ grid is identical except for a $1/8^{\circ}$ patch of refinement over Greenland.
+A third VR grid, $\tt CONUS$, has also been made available in CESM2.2.
 
 In general, the FV grids have colder summer temperatures over the Arctic compared with the SE grids (including the VR grids).
-The cloud biases in all the uniform-resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.{\color{purple}{Marcus - hard to parse this sentence.}} The VR grids reduce the cloud biases.
+The cloud biases in all the uniform-resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.
 It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates; an improved representation of clouds in the Arctic does not imply improved clouds at lower latitudes.
 
-At the regional level, there is a halo of negative cloud bias (/color{blue}{I got confused about signs here, because section 3 talks about a halo of positive cloud SW forcing bias, which corresponds to a negative bias in cloud amounts.
-Maybe replace 'cloud bias' with 'cloudiness bias' or something similar?}) around the ocean perimeter of Greenland on all $1-2^{\circ}$ grids, but not the VR grids.
-This halo bias coincides with a positive cloud bias over the ice sheet interior.
-This anomaly pattern has been attributed to deficient orographic precipitation on the coarser model grids.
+At the regional level, there is a halo of negative cloudiness bias around the ocean perimeter of Greenland on all $1-2^{\circ}$ grids, but not the VR grids.
+This negative cloud bias occurs while there is a positive cloud bias over the ice sheet interior.
+This anomaly pattern has been attributed to deficient orographic precipitation in the coarser model grids.
 With overly smooth topography on the $1-2^{\circ}$ grids, synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins.
-As a result, excess precipitation falls in the GrIS interior, instead of being concentrated on the steep coastal margins as shown by observations.
+As a result, excess precipitation falls in the GrIS interior, instead of being concentrated on the steep coastal margins as shown by observations \cite{P2000GPC,VETAL2018TC}.
 This results in a positive precipitation and cloud bias in the ice sheet interior, and a halo of low cloud bias about the perimeter.
@@
 We integrated the primary source and sink terms of the SMB equation over the GrIS for each of the six grids.
-The uniform $1-2^{\circ}$ grids have large positive accumulation biases because they fail to resolve orographic precipitation.
-The uniform SE grids have larger accumulation biases, suggesting that the FV grids are more skillful for precipitation due to finer resolution over Greenland, despite a polar filter.
+The uniform $1^{\circ}-2^{\circ}$ grids have large positive accumulation biases because they fail to resolve orographic precipitation.
+The uniform SE grids have larger accumulation biases, suggesting that the FV grids are more skillful for precipitation due to finer resolution over Greenland, and despite a polar filter.
 The VR grids have the most accurate accumulation rates of all the grids.
-
-The primary mass sink term of the GrIS, ice/snow melt, has similar biases.
-The uniform resolution SE grids have too much melt, while the FV grids have smaller biases.
-It is difficult to attribute these biases to grid resolution alone.
-The FV grids have colder summers, consistent with their lower melt bias.
-However, the $\tt Arctic-GrIS$ grid has the warmest summer temperatures of all grids, yet it has less melting than the uniform-resolution SE grids.
-This suggests that grid resolution is responsible for a large fraction of the melt biases.
-We propose a mechanism: Coarse grids represent ablation zones using grid cells with mixed surface types, ice-covered and ice-free.
+The primary mass sink term of the GrIS, ice/snow melt, have similar biases; the coarse grids melt too much, although the bias is worse in uniform SE grids.
+In general, on coarse grids, errors in the individual SMB terms are larger than the errors in the SMB itself, due to compensating errors.
+This observation serves as a precaution; projecting mass-loss from a glacier or ice sheet cannot be reliable if the processes representing the components of the SMB are incorrect from the start, even as the total SMB has the right magnitude.
+
+The $\tt Arctic-GrIS$ grid has the warmest summer lower troposphere of all grids, yet it has less melting than the uniform-resolution SE grids.
+This suggests that grid resolution is somehow contributing to the melt biases in coarse grids, in a way that is not obvious from the large-scale dynamics.
+We propose a mechanism: coarse grids represent ablation zones using grid cells with mixed surface types, ice-covered and ice-free.
 The warmer ice-free patches may largely determine the mean state, leading to a warm bias over the ice-covered patches of the grid cell.
@@
 
-The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the Greenland SMB, compared to the uniform $1^{\circ}-2^{\circ}$ grids.
+The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the GrIS SMB, compared to the uniform $1^{\circ}-2^{\circ}$ grids.
 The $\tt Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but its summer temperatures are too warm.
-The $1^{\circ}$ FV model gives a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids.
-{\color{purple}{It is also the most heavily tuned model configuration.
-May be worth mentioning here as well.}} In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other uniform-resolution grids.
+The $1^{\circ}$ FV model gives a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids (but perhaps also because it is the most heavily tuned model configuration in CESM).
+In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other uniform-resolution grids.
 
 As modeling systems move away from lat-lon grids towards quasi-uniform unstructured grids, it is worth taking stock of whether this will degrade the simulated polar climate.
-We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model in simulating the surface mass balance of the GrIS.
-This finding will not interrupt the ongoing transition towards unstructured grids in CESM, largely driven by gains in computational efficiency, but it has inspired us to develop alternative configurations that recover or improve on the fidelity of polar climate.
-We have shown here (and in a prior companion study \cite{VETAL2018TC}) that for CESM, Arctic-refined meshes can substantially improve the simulated mass balances of the GrIS, even compared to the $1^{\circ}$ grid.
-This should reassure the CESM modeling community that the ongoing transition away from lat-lon grids will not adversely impact CESM's usefuless as a state-of-the-art tool for simulating and understanding polar processes.
-{\color{blue} (WHL: This last sentence may be too sanguine.
-Yes, we can recover the fidelity of Arctic simulations using VR grids, but (so far) only at a considerable cost in cpu-hours.
-This points to the need for an intermediate resolution that is more affordable, and/or model development or tuning that reduces the biases on coarse grids.)}{\color{purple}{Andrew - Maybe better to just state that higher resolution is better: 1deg better than 2deg FV, and ne30 is coarser than 1deg at Greenland latitudes so it's worse.
-But VR has higher resolution so SR-VR is better.
-It's all about resolution.}}
+We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model for simulating the GrIS SMB.
+That is, the simulated GrIS SMB will be adversely impacted in future CESM versions, after the FV dycore is phased out.
+This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities.
+We therefore provide the the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.
 
 We are  working to develop a configuration of the $\tt Arctic$ grid that is fully-coupled with the CESM ocean and sea ice components and the Community Ice Sheet Model (CISM), to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise.
-%A fully-coupled pre-industrial control configuration has already been developed and vetted.
-%, and should be incorporated into CESM in the near future.
-We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}.
-%in order to advertise some of the stunning capabilities of the refined mesh approach.
+We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}, to increase awareness.
 Figure~\ref{fig:viz} shows a snapshot of this visualization, illustrating mesoscale katabatic winds descending the southeastern slopes of GrIS.
-%We are optimistic about the polar science opportunities provided by these new Arctic mesh configurations, and our open to collaborating with the CESM community to assist in anyway we can.
 These new grids and configurations will provide new opportunities for CESM polar science and aims to contribute to an improved understanding of the polar environment.
-{\color{blue} (WHL: I replaced the previous last sentence because it seemed too much like an advertisement.
-However, this new ending seems weak.
-I wonder if we should say something about future work motivated by this study, for instance investigating grids and parameterizations that provide some of the same benefits as these VR grids but at lower cost.)}
+However, we recognize the potentially prohibitive costs for some users, and so will continue to explore different grids, parameterizations and workflows that can provide some of the same benefits of the VR grids, but at a lower cost.
 
@@
 \subsection{Results}
-{\color{red}{[Adam: have you defined ne30np4 in the main text?]}}\\
 One year averages of vertical pressure velocity at 500hPa ( {\tt OMEGA500}) have been found to be a useful quantity to detect spurious up or down-drafts induced by steep orography (Figure \ref{fig:OMEGA500}).

### 2022-05-10 arcticgrids.tex
mods

@@
 We are  working to develop a configuration of the $\tt Arctic$ grid that is fully-coupled with the CESM ocean and sea ice components and the Community Ice Sheet Model (CISM), to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise.
-We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}, to increase awareness.
+We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}, to increase awareness of the capabilities of CESM2.2.
 Figure~\ref{fig:viz} shows a snapshot of this visualization, illustrating mesoscale katabatic winds descending the southeastern slopes of GrIS.

### 2022-05-10 arcticgrids.tex
fix figs

@@
 The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution.
-Figure~\ref{fig:grisdx}b shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids in this study.
+Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids in this study.
 The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid.
@@
 We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
-However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}).
+However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx})b.
 CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
 The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
-The area errors in Figure~\ref{fig:grisdx} may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
+The area errors in Figure~\ref{fig:grisdx}b may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
 
@@
 
-Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx}).
+Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx})b.
 Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
@@
 \begin{center}
-\includegraphics[width=145mm]{OMEGA500F2000climo.pdf}
+\includegraphics[width=145mm]{figs/OMEGA500F2000climo.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=145mm]{OMEGA500FHS94.pdf}
+\includegraphics[width=145mm]{figs/OMEGA500FHS94.pdf}
 \end{center}

### 2022-05-11 arcticgrids.tex
bills edits

@@
 The conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both accumulation and ablation over the GrIS.
-Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids owing to their higher degrees of freedom in representing the GrIS.
+Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS.
 Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, are documented as newly supported configurations in CESM2.2.
@@
 Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance.
-As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids.
+As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic-refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids.
 
@@
 General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change.
-GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole-problem} \cite{W2007JMSJ}.
+GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole problem} \cite{W2007JMSJ}.
 The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids).
@@
 This damping reduces the effective resolution in polar regions such that the resolved scales are \textit{approximately} the same everywhere on the grid.
-We emphasize \textit{approximately}, since it is at least conceivable that marginal increases in effective resolution occur over polar region in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.
+We emphasize \textit{approximately}, since it is conceivable that marginal increases in effective resolution occur over polar regions in lat-lon grids, despite polar filtering, since resolved waves can be represented with more grid points than at lower latitudes.
 
@@
 This illustrates the advantages of coordinate surfaces aligned with latitude bands, albeit an extreme example where the polar singularity and the polar filter are also contributing to the spurious wave growth.
-The unstructured grids all generate spurious baroclinic waves earlier on in the simulations than the (unrotated) lat-lon models, although the unstructured model considered in this paper, the spectral-element dycore (hereafter \textit{SE}), holds a balanced zonal flow without spurious wave growth appreciably longer than the rotated FV experiments \cite{LJTN2010JAMES}.
+The unstructured grids all generate spurious baroclinic waves earlier in the simulations than the (unrotated) lat-lon models, although the unstructured model considered in this paper, the spectral-element dycore (hereafter \textit{SE}), holds a balanced zonal flow without spurious wave growth appreciably longer than the rotated FV experiments \cite{LJTN2010JAMES}.
 And unlike FV, the SE dycore has the same error characteristics regardless of how the grid is rotated.
@@
 The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which have large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}.
-Unstructured grids support quasi-uniform grid spacing globally, and there is no pole-problem (e.g., Figure~\ref{fig:uni-grids}c).
-Conversely, unstructured grids are becoming increasingly common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}.
+Unstructured grids support quasi-uniform grid spacing globally, and there is no pole problem (e.g., Figure~\ref{fig:uni-grids}c).
+Conversely, unstructured grids are becoming more common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}.
 This grid flexibility allows for the adoption of variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
@@
 
-It is important to emphasize that the pole-problem is a distinctive feature of the dycore in atmospheric models.
+We emphasize that the pole problem is a distinctive feature of the dycore in atmospheric models.
 Polar filters do not directly interfere with the physical parameterizations, nor do they have any bearing on the surface models; e.g., the land model can take full advantage of the greater number of grid cells in polar regions on lat-lon grids.
@@
 GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
-These narrow ablation zones are not fully resolved in low resolution GCMs, and may further degrade the simulated SMB.
+These narrow ablation zones are not fully resolved in low-resolution GCMs, and may further degrade the simulated SMB.
 For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.
@@
 Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates.
-The limited area domain used by RCMs permits the use of high resolution grids, capable of resolving SMB processes, that can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}.
-However, unlike GCMs, RCMs are not a freely evolving system and the atmospheric state must be prescribed at the lateral boundaries of the model domain.
+The limited area domain used by RCMs permits the use of high-resolution grids capable of resolving SMB processes, and can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}.
+However, unlike GCMs, RCMs are not a freely evolving system, and the atmospheric state must be prescribed at the lateral boundaries of the model domain.
 The inability of the RCM solution to influence larger-scale dynamics outside the RCM domain (due to the prescribed boundary conditions) severely limits this approach from properly representing the role of the GrIS in the climate system.
@@
 
-In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} utilized the VR capabilities of the SE dycore in CESM, generating a grid where Greenland is represented with $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution.
+In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating a grid where Greenland is represented with $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution.
 The simulated SMB compared favorably to the SMB from RCMs and observations.
-The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios, in the freely evolving GCM framework.
-
-The SE dycore has been included in the model since CESM, version 1, but has been under active development ever since.
+The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios.
+
+The SE dycore has been included in the model since CESM version 1, but has been under active development ever since.
 This includes the switch to a dry-mass vertical coordinate \cite{LetAl2018JAMES} and incorporation of an accelerated multi-tracer transport scheme \cite{LTOUNGK2017MWR}, made available in CESM2.
@@
 While there are dozens of published studies using VR in CESM \cite<e.g., >{ZetAl2014JC,RHUZ2016JAMC,GetAl2017JAMES,BETAL2019TAC,BETAL2021IJC}, these studies either used development code or collaborated closely with model developers.
-CESM2.2 is the first code release that contains out of the box VR functionality in CESM.
+CESM2.2 is the first code release that contains out-of-the-box VR functionality.
 
@@
 
+In CESM2.2, the SE numerics have been enhanced to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}.
+
 The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over uniform-resolution setups.
-Firstly, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
+First, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
 A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements.
-Secondly, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.
+Second, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.
 
@@
 The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
-We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}d), where $ne$ refers to a grid with $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
+We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}d), where $ne$ refers to a grid with $ne \times ne$ quadrilaterial elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
 We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}c;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMESb}.
@@
 All grids and dycores in this study use 32 hybrid pressure-sigma levels in the vertical, with a model top of 2~hPa or about 40~km.
-However, note that any grid or dycore can in principle be run with a higher model top or finer vertical resolution.
+However, any grid or dycore can in principle be run with a higher model top or finer vertical resolution.
 
@@
 CLUBB is coupled with the MG2 microphysics scheme \cite{GM2015JC,MG2}, which computes prognostic precipitation and uses classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
-Deep convection is parameterized using a convective quasi-equilibrium, mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
+Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
 Boundary layer form drag is modeled after \citeA{BBW2004QJRMS}, and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale (the ridge orientation is derived from a high-resolution, global topography dataset \cite{Danielson_Gesch_2011_GMTED2010}).
@@
 
-\subsection{Simulated surface mass balance (SMB)}
-
-All grids and dycores simulate the GrIS SMB, which is the sum of mass accumulation of precipitation and mass loss from ablation.
+\subsection{Simulated surface mass balance (SMB)}\label{sec:simSMB}
+
+All grids and dycores simulate the GrIS SMB, which is the sum of mass accumulation from precipitation and mass loss from ablation.
 The latter is the sum of evaporation, sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass.
+Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, forming ice lenses.
 These relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
@@
 Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent.
-Any snow above the 10 m cap contributes towards ice accumulation in the SMB.
-Refreezing of liquid water within the snowpack is an additional source of ice.
+Any snow above the 10 m cap contributes towards ice accumulation in the SMB, and refreezing of liquid water within the snowpack is an additional source of ice.
 Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
@@
 In the simulations described here, the GrIS is prescribed at its observed, modern extent and thickness.
-
-\subsection{SMB Analysis}\label{sec:SMB}
-
-We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
-However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx})b.
-CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
-The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
-The area errors in Figure~\ref{fig:grisdx}b may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
-\end{center}
-\caption{The spatial properties of the GrIS as represented by different grids in this study.
-(Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
-\label{fig:grisdx}
-\end{figure}
-
-We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $\tt f19$ and $\tt ne30pg2$ grids, and integrating over these low-resolution ice masks.
-The use of low-resolution common ice masks is a conservative decision, and is justified because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not generally reliable when mapping to a higher-resolution grid than the source grid.
-We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
-Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $\tt f19$ and $\tt ne30pg2$ masks.
-Thus, we evaluate an integrated quantity on a given grid up to four times to estimate the uncertainty due to differences in grid type and remapping algorithms.
-
-The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets.
-The SMB for total water can be expressed as:
-\begin{equation}
-SMB = accumulation + runoff + evaporation/sublimation, \label{eq:SMB}
-\end{equation}
-where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass).
-The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
-Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.
-Note that this SMB expression is different from the internally computed SMB described in the previous section.
-
-We consider two approaches for mapping and integrating the SMB components over the common ice masks:
-\begin{enumerate}
-\item Map the grid-cell mean quantities to the common grid, and integrate the mapped fields over the common ice masks. \label{lbl:label1}
-\item Map the patch-level quantities (i.e., the state over the ice fractional component of the grid cell) to the common grid, and integrate the mapped fields over the common ice masks.
-\label{lbl:label2}
-\end{enumerate}
-
-Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx})b.
-Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
-While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
-This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
-However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.
-
-The different error characteristics of the two methods are used to diversify the ensemble.
-Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
-Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
-This will degrade the SMB estimates using method~\ref{lbl:label2}, however we are more interested in characterizing the behavior of individual processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
-
-\subsection{Experimental design}
-
-All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
-In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2.
 
@@
 
+\subsection{SMB Analysis}\label{sec:SMB}
+
+We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
+However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}b).
+CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
+The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
+The area errors in Figure~\ref{fig:grisdx}b may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
+\end{center}
+\caption{The spatial properties of the GrIS as represented by different grids in this study.
+(Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
+\label{fig:grisdx}
+\end{figure}
+
+We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $\tt f19$ and $\tt ne30pg2$ grids, and integrating over these low-resolution ice masks.
+The use of low-resolution common ice masks is a conservative decision, and is justified because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not generally reliable when mapping to a higher-resolution grid than the source grid.
+We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
+Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $\tt f19$ and $\tt ne30pg2$ masks.
+Thus, we evaluate an integrated quantity on a given grid up to four times to estimate the uncertainty due to differences in grid type and remapping algorithms.
+
+The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets.
+The SMB for total water can be expressed as:
+\begin{equation}
+SMB = accumulation + runoff + evaporation/sublimation, \label{eq:SMB}
+\end{equation}
+where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass).
+The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
+Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.
+Note that this SMB expression is different from the internally computed SMB described in section~\ref{sec:simSMB}.
+
+We consider two approaches for mapping and integrating the SMB components over the common ice masks:
+\begin{enumerate}
+\item Map the grid-cell mean quantities to the common grid, and integrate the mapped fields over the common ice masks. \label{lbl:label1}
+\item Map the patch-level quantities (i.e., the state over the ice fractional component of the grid cell) to the common grid, and integrate the mapped fields over the common ice masks.
+\label{lbl:label2}
+\end{enumerate}
+
+Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx}b).
+Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
+While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
+This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
+However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the grid-mean quantities used in method~\ref{lbl:label1}.
+
+The different error characteristics of the two methods are used to diversify the ensemble.
+Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
+Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
+This will degrade the SMB estimates using method~\ref{lbl:label2}, but we are more interested in characterizing the behavior of individual processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
+
+\subsection{Experimental design}
+
+All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
+In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2.
+
 \section{Results}\label{sec:results}
@@
 
-The top panel of Figure~\ref{fig:prect} show the annual climatological mean precipitation bias over the GrIS, expressed as the fractional difference from the RACMO2.3p2 solution.
+The top panel of Figure~\ref{fig:prect} shows the annual climatological mean precipitation bias over the GrIS, expressed as the fractional difference from the RACMO2.3p2 solution.
 The coarse $1^{\circ}-2^{\circ}$ grids have large, positive biases centered over the southern dome.
@@
 Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO.
-The values in the table are averages over the ensemble of mapping methods to the common ice masks described in section~\ref{sec:SMB}, and the RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}).
+The values in the table are averages over the ensemble of mapping methods to the common ice masks described in section~\ref{sec:SMB}, and the RACMO values refer to the average of both RACMO ensembles.
 Table~\ref{tbl:table3} also contains (in parenthesis) the SMB components derived from evaluating the integrals on each model's native grid and ice mask.
 Of note is the large reduction in melt rates compared to the values computed on the native grid, illustrating the dissipation of this quantity discussed in section~\ref{sec:SMB}.
-For integrated precipitation, the differences between the native and common-ice-mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
+For integrated precipitation, the differences between the native and common ice mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
 
 The coarse grids are characterized by too much precipitation and too much melting and runoff, compared with RACMO.
-The SMB in coarse grids therefore have smaller errors than their individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
-Such compensating errors highlights the importance of understanding the extent to which a model is getting the SMB right for the right reasons.
+The total SMB on coarse grids therefore has smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
+Such compensating errors underlie the importance of understanding the extent to which a model is getting the right SMB for the right reasons.
 
@@
 Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones.
-The Positive Degree-Day temperature based melt index \cite<PDD;>{B1984Z}, which accumulate the near-surface temperature in $^{\circ}C$ for days with temperature above freezing, is a more accurate proxy for melting.
+The Positive Degree-Day temperature-based melt index \cite<PDD;>{B1984Z}, which accumulates the near-surface temperature in $^{\circ}C$ for days with temperature above freezing, is a more accurate proxy for melting.
 PDD is nonlinear in mean monthly temperature \cite{R1991P}.
@@
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and a larger number of grid cells can line the east-west transects.
-The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.
+The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and so a larger number of grid cells fit into the east-west transects.
+The VR grids more accurately reproduce the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.
 
@@
 The $\tt Arctic$ run has larger extreme precipitation rates compared to $\tt ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year.
-This is primarily due to increased resolution and not the reduced physics times-step; the $\tt ne30pg3^{*}$ run only marginally increases the extreme precipitation rates compared with $\tt ne30pg3$ (Figure~\ref{fig:comp-pdf}).
+This is primarily due to increased resolution and not the reduced physics time step; the $\tt ne30pg3^{*}$ run only marginally increases the extreme precipitation rates compared with $\tt ne30pg3$ (Figure~\ref{fig:comp-pdf}).
 
@@
 The cloud biases in all the uniform-resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.
-It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates; an improved representation of clouds in the Arctic does not imply improved clouds at lower latitudes.
+It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates.
+An improved representation of clouds in the Arctic does not imply improved clouds at lower latitudes.
 
 At the regional level, there is a halo of negative cloudiness bias around the ocean perimeter of Greenland on all $1-2^{\circ}$ grids, but not the VR grids.
-This negative cloud bias occurs while there is a positive cloud bias over the ice sheet interior.
-This anomaly pattern has been attributed to deficient orographic precipitation in the coarser model grids.
+This negative cloud bias contrasts with a positive cloud bias over the ice sheet interior.
+This anomaly pattern is attributed to deficient orographic precipitation in the coarser model grids.
 With overly smooth topography on the $1-2^{\circ}$ grids, synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins.
@@
 The VR grids have the most accurate accumulation rates of all the grids.
-The primary mass sink term of the GrIS, ice/snow melt, have similar biases; the coarse grids melt too much, although the bias is worse in uniform SE grids.
+The primary mass sink term of the GrIS, ice/snow melt, has similar biases; the coarse grids melt too much, with a greater bias for uniform SE grids.
 In general, on coarse grids, errors in the individual SMB terms are larger than the errors in the SMB itself, due to compensating errors.
-This observation serves as a precaution; projecting mass-loss from a glacier or ice sheet cannot be reliable if the processes representing the components of the SMB are incorrect from the start, even as the total SMB has the right magnitude.
+This observation serves as a precaution; projecting mass-loss from a glacier or ice sheet cannot be reliable if the processes representing the components of the SMB are incorrect from the start, even if the total SMB has the right magnitude.
 
@@
 This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities.
-We therefore provide the the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.
+We therefore provide the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a more realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.
 
@@
 We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}, to increase awareness of the capabilities of CESM2.2.
-Figure~\ref{fig:viz} shows a snapshot of this visualization, illustrating mesoscale katabatic winds descending the southeastern slopes of GrIS.
-These new grids and configurations will provide new opportunities for CESM polar science and aims to contribute to an improved understanding of the polar environment.
+Figure~\ref{fig:viz} shows a snapshot of this visualization, illustrating mesoscale katabatic winds descending the southeastern slopes of the GrIS.
+These new grids and configurations will provide new opportunities for CESM polar science, and they aim to contribute to an improved understanding of the polar environment.
 However, we recognize the potentially prohibitive costs for some users, and so will continue to explore different grids, parameterizations and workflows that can provide some of the same benefits of the VR grids, but at a lower cost.
@@
 \appendix
-\section{Details on spectra-element dynamical core improvements since the CESM2.0 release}
+\section{Details on spectra-element dynamical core improvements since the CESM2.0 release}\label{sec:a1}
 Since the CESM2.0 release of the spectral-element dynamical core documented in \citeA{LetAl2018JAMES} some important algorithmic improvements have been implemented and released with CESM2.2.

### 2022-09-06 first_review/response-to-reviewers.tex
first commit after review

\documentclass[12pt,oneside,a4paper]{article}%{amsart}
\usepackage[margin=1in]{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{caption}
\setlength\parindent{0pt}
\usepackage[round]{natbib}
\bibliographystyle{abbrvnat}
\renewcommand{\bibsection}{}

\title{\large Response to Reviewers}
\author{\normalsize Adam R. Herrington and CoAuthors}
\date{} % delete this line to display the current date

\begin{document}

\maketitle

\section*{General Remarks}

{\color{red}{We would like to thank both reviewers for taking the time to provide thoughtful feedback on our manuscript.}} \newline

\section*{Reviewer 1}

\subsection*{General comments}

Given that the different grid and dycore choices come with considerable different computational costs, I think it would be useful to include a section addressing that specific question. Some information touching on the costs appears at different places in the manuscript. It would be good to bring that all together somewhere and make a clear comparison, e.g. including a comparative table. \newline

The way the term surface mass balance (SMB) is defined and used in the manuscript is confusing. The different components to the SMB are not consistently applied and there are two different definitions of the SMB in use. In part this confusion is related to the fact that the model itself does not calculate all the SMB components and assumptions have to be made about the separation between snow and ice, especially in the currently discussed uncoupled case. I would suggest to start with a physical definition of SMB with all its components in an equation and then explain based on that what the model is actually doing and what constraints have to be applied. In this context it may help to describe what the model would do in a coupled setup with active CISM and then apply further constraints for the diagnostic case currently described. \newline

In the discussion about the different dycores, I was wondering if there could be a notion that the latitude-longitude approaches are more mature because they have a longer history. Maybe further improvements on technical aspects like filtering and optimal solutions may be expected for the FE approaches. \newline

The relationship between atmospheric grid and land grid seems to be important. Would it be possible to say that higher resolution of the atmospheric grid improves the precipitation, while increased resolution of the land model improves the melt calculation/ablation components? \newline

\subsection*{Specific comments}

l12 I am confused about the use of the terms "uniform resolution" and "equivalent resolution". Since the lat-lon grids are also uniform (they have 1 or 2 degree resolution everywhere) I find this is not a good description to distinguish them here. Maybe this requires a few more words to make it clear to the reader what is meant here. The confusion is not fully removed in the main text either. See later comments.\newline

{\color{blue}{We take your point and agree that this is confusing. To clarify this, we've replaced ``uniform resolution" with ``quasi-uniform unstructured grids" and clarify that ``uniform" refers to isotropic grids where every grid cell has the same length in meters, as opposed to an angular distance in spherical coordinates. Whereas we had previously used ``uniform grids" to refer to both lat-lon and quasi-uniform unstructured grids, we replace ``uniform" with ``lat-lon and quasi-uniform grids" throughout the manuscript, to be more explicit about the differences between the SE and lat-lon grids.}} \newline

l29 It would be good to add one sentence about the cost here. E.g. "... recovering resolution lost in the transition to quasi-uniform grids, albeit at increased computational cost." \newline

{\color{blue}{Done.}} \newline

l37 "latitude-longitude grids" \newline

{\color{blue}{Done.}} \newline

l75 "Dycores built on lat-lon grids have some advantages over dycores built on unstructured grids" \newline

{\color{blue}{Done.}} \newline

l92 "have large" --> "has a large" \newline

{\color{blue}{Done.}} \newline

l94 The logic of "Conversely" is not clear, since I think it is referring to the first sentence in the paragraph not the preceding one. Reformulate? \newline

{\color{blue}{Sentence beginning with conversely has been reformulated to improve clarity.}} \newline

l107 Move "(SMB; the integrated sum of precipitation and runoff)" before "of the Greenland ice sheet". \newline

{\color{blue}{Done.}} \newline

l104-110 With the benefit of a high resolution land model in mind, is it possible to run the land model at a different/higher resolution than the atmosphere model? If so, this could be mentioned here. \newline

{\color{blue}{As discussed above, I'm not optimistic about the utility of a higher-res land model, and would rather not go on a tangent about custom configurations.}} \newline

l117 "(marginal regions where seasonal melting exceeds the annual mass input from precipitation)" This sentence is not correct. What defines the ablation area is a negative SMB. Problems with the statement are that melt can refreeze and precipitation can occur as snow and rain. It would maybe help to write out the whole SMB in an equation if there is confusion about these terms. \newline

{\color{blue}{Replaced with ``marginal regions where the annual SMB is negative."}} \newline

l138 "is represented with 1/4 resolution" I believe they have used two different levels of refinement (0.5 and 0.25 degree): Maybe state "is represented with up to 1/4 resolution" instead. Or give resolution in km as they do in most of the paper: 55 km and 28 km. \newline

{\color{blue}{Used reviewers ``up to" suggestion.}} \newline

l159 What I miss in this section is a short introduction to CESM. Can be brief, but should bring the reader up to speed what kind of model that is \newline

{\color{blue}{Included a short introduction to CESM. Also moved language from the experimental design section in here, as well as include a statement that the ice sheet model is not dynamically active. Our hope is that by including this at the very top of the methods section, this partly addresses the reviewers concern that it is not clear we are not running with an active ice sheet.}} \newline

l174 Is it correct to say the dycore is a model? \newline

{\color{blue}{Reworded.}} \newline

l203 "the SE numerics have been enhanced". Enhanced compared to what. An earlier model version, I suspect. Mention that. \newline

{\color{blue}{added: relative to the CESM2.0 release.}} \newline

l214 "... has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone." This seems in contrast to the very idea of refinement. Maybe this should be mentioned here? \newline

{\color{blue}{I don't follow. How is rougher topography in the refinement zone inconsistent with the very idea of refinement? From our perspective it is a requirement of refinement -- finer dynamics permits rougher topography.}} \newline

l215 "For SE grids". The grids are not SE! It is clear what you mean, but good to use proper terminology. Maybe "When using the SE dycore with quasi-uniform element sizes ..." \newline

{\color{blue}{Used the reviewers preferred wording.}} \newline

l215-219 I seem to understand from this paragraph that CSLAM is an improvement, but that it cannot be applied for VR grids. If so, please make that clear here. \newline

{\color{blue}{Added. Yes, this is unfortunately true. We do have plans to address this in VR within the next 2 years.}} \newline

l220-224 I am wondering if this separate grid could in principle be extended to match an underlying SE dynamics VR grid by further subdivision into more control volumes. \newline

{\color{blue}{The underlying issue is that the SE dynamics don't have control volumes. They have GLL nodes, which are point samples of a spectrally truncated state. In order to couple SE to other model components, we had to invent control volumes because the CESM infrastructure uses finite-volume mapping algorithms to couple to other model components. The resulting control volume grid is an anisotropic mess (see Figure 4 in my MWR 2019 paper), which was constructed by constructing control volumes consistent with each GLL node's quadrature weight. It is much preferable to split the elements into quadrilateral control volumes and use the continuous basis representation to project the SE solution onto the quadrilateral control volumes (like we do in pg3 and pg2.}} \newline

l224 add "at" after "evaluated" \newline

{\color{blue}{Done.}} \newline

l249 I don't think "RACMO" has been defined yet. Please do. \newline

{\color{blue}{Done.}} \newline

l251 The discussion about computational costs is an important one in this comparison and should get at least an own paragraph, if not an own sub-section. \newline

{\color{blue}{See introductory remarks.}} \newline

l288 It is not the grids that calculate the SMB. Maybe "CESM simulates the GrIS SMB as the sum of mass accumulation and ablation" \newline

{\color{blue}{We have replaced with the suggested language. In order to reconcile your following comments regarding this paragraph, what follows this top sentence is a description of how the SMB is computed in CESM. This requires some changes to the language, such as excluding rain from the runoff terms, and making clear only ice accumulation is contained in the accumulation terms. Having addressed some unclear statements about the definition of the SMB in the introduction, we don't feel that we need to add another equation for CESM's SMB, or a common glaciological definition plus a simplified version used by CESM. These equations can be found in the references in the sentence "For a more detailed description of the SMB is computed in CESM ...", in particular equations 4 and 5 in van Kampenhout et al. 2020. }} \newline

l289 Strictly speaking only snow accumulation is a definite positive mass source, while rain can either runoff or refreeze. Again, an equation that gives all the terms in the SMB may be useful here. \newline

{\color{blue}{See above comment.}} \newline

l290 Suggest to replace "with runoff being a combination of" by "with runoff having sources of a combination of". Again, I think runoff is not a straight sum of these, because of refreezing. \newline

{\color{blue}{This has been modified in the reworked section.}} \newline

l300 Mention here that this can change if the Greenland ice sheet is fully coupled. I generally miss already earlier a statement that you are using CESM without interactive Greenland ice sheet. \newline

{\color{blue}{This is a good catch and the language has been inserted about how this changes with an active ice sheet.}} \newline

l305 These descriptions apply to the uncoupled case, which should be clarified. It may also be interesting to explain what happens in the coupled case, since CESM has that as an option. I am a bit surprised that this is not the primary option you are describing here. \newline

{\color{blue}{After addressing your last comment, we believe the paragraph is valid for both coupled and un-coupled cases.}} \newline

l308 I think it would be interesting to show that downscaled SMB (using the elevation classes) and compare that between the different cases. That is the SMB product one would actually use either in the coupled case or when extracting the SMB from CESM for offline forcing of an ice sheet model. In addition, you could compare that result more easily between the different cases as they would all be well-defined on the output grid. \newline

{\color{blue}{We agree that would be interesting. The reason we didn't employ this approach -- which to be clear, we interpret as evaluating the SMB field downscaled to the 4km CISM grid -- is that only the SMB is downscaled. It is the resolution sensitivity of the individual components of the SMB that is the main focus of this study. It's representation on the CLM grid is still quite advanced, owing to the EC scheme. I suppose we could have tried to hack the code to downscale the SMB components to CISM, but we didn't go down that road (and no telling the number of issues that would've cropped up going down that road). That said, our approach of mapping to the common ice mask is quite complicated as well.}} \newline

l313 "Changes in ice depth, but not snow depth, count toward the SMB" seems to be in contrast with "Any snow above the 10 m cap contributes towards ice accumulation in the SMB" in the paragraph before. Please clarify. \newline

{\color{blue}{I'm not sure I see the contradiction the reviewer is referring to, but this a clearly a redundancy that we've eliminated by consolidating the paragraphs.}} \newline

l322 The datasets you use are not observational but model results. I think this should be made clear already in the heading. Maybe "Target" or "Validation" datasets. \newline

{\color{blue}{Done.}} \newline

l323 Again, the datasets in use are not observational. \newline

{\color{blue}{Done.}} \newline

l323-338 You do not seem to describe anywhere the other datasets in Table 2: ERA, CERES, Calipso. These should all be described here as well. \newline

{\color{blue}{Done.}} \newline

l331 add "e.g." after before "maintained". This is a very specific collection of the data. \newline

{\color{blue}{I'm not sure I follow, but I did clarify the second category to be less specific (remote sensing datasets instead of radar accumulation datasets).}} \newline

l343 "reference dataset" Remind us what this is and where it is sourced from. \newline

{\color{blue}{Clarified the reference dataset and added citation.}} \newline

l348 "can lead to large differences in integrated SMB" Could explain briefly why that is, because the mass loss happens in a narrow band along the margins. \newline

{\color{blue}{The Hansen paper indicates that differences in the ice masks occur in its representation of the ice margins, where lots of melting conceivable occurs. So an erroneously small ice mask may overestimate the SMB due to omitting areas where melting occurs. }} \newline

l363 Equ 1. Could be read as "evaporation divided by sublimation". Why not "SMB = accumulation + runoff + evaporation + sublimation"? \newline

{\color{blue}{Done.}} \newline

l369 This is confusing, because we are in section 2.4. If you are operating with two different SMB terms, you should have two equations to clearly distinguish them. \newline

{\color{blue}{We have reworded and moved this last sentence, indicating that the total water SMB is different than the SMB internally computed by CLM, as a topic sentence to its own paragraph immediately following. We've added clarification on the role of the internally generated CLM SMB, that we don't use this definition in this study.}} \newline

l372 "grid-cell mean quantities" Clarify that these are atmospheric/land-model grid cells. \newline

{\color{blue}{Apologies for not understanding the suggestion here. What is it that needs clarification? A topic sentence in a paragraph in the methods section begins with ``CLM runs on the same grid as the atmosphere..." Is that what you are asking about?}} \newline

l388 insert "further" before "diversify". \newline

{\color{blue}{Done.}} \newline

l439 Since you start the section in 404 with "Before delving into the simulated Arctic climate conditions ...", could mark that we are now moving from global to Arctic discussion. \newline

{\color{blue}{Good suggestion. Done}} \newline

l461 "CERES product" needs to be described in 2.5. \newline

{\color{blue}{Done.}} \newline

l478 Add "radiation" after "shortwave". \newline

{\color{blue}{Done.}} \newline

l525 "underline"? \newline

{\color{blue}{Yes. Fixed.}} \newline

l528 Did you test how much biased the results could be due to the relatively short period of 20 years in view of the large inter-annual variability? If you don't have more years, you could remove any of the 20 years and see how much the numbers change. \newline

l553 "smaller melt-rates" or smaller melt-rate biases? \newline

{\color{blue}{since both are true, replaced with ``smaller melt-rates and melt-rate biases."}} \newline

l564 Often PDD schemes have different PDD factors for snow and ice melt. Is this not not the case here? \newline

{\color{blue}{To simplify the calculation I just took an intermediate value between snow and ice (Braithwaite 1984 uses 3 mm/d/C and 7 mm/d/C for ice), so I wouldn't need to approximate how to partition at particular monthly average grid cell. Added these details in the manuscript in parenthesis.}} \newline

l557-564 Description and motivation for using the PDD model should probably be moved/added to the Methods section. \newline

{\color{blue}{Our personal opinion is that defining it on the fly, naturally arising from the analysis to explain a discrepancy between two contrasting results, is preferable instead of premeditating it earlier on in the manuscript.}} \newline

l571 Consider reordering figures, since we are jumping back from 11 to 9 here. \newline

l594 This reads like a contradiction to me "coarser models ... depress melt rates, which is the opposite of the melt bias that occurs in the coarse grid simulations". What happens in course models is the opposite to what happens in coarse models? \newline

{\color{blue}{Reworded this slightly to improve the clarity. ``This suggests that coarser models will tend to elevate the ablation zones relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and depressed melt rates, but this is opposite the melt bias that occurs in the coarse grid simulations."}} \newline

l603 "volume averaging" I would have guessed this was done by area averaging. \newline

{\color{blue}{Fair point. Changed to area averaging.}} \newline

l605 Maybe add "and more melting" after "warm bias over the ice-covered patch". \newline

{\color{blue}{Done.}} \newline

l622 Define PDFs. \newline

{\color{blue}{Done.}} \newline

l648 Add "refined over the US" after CONUS. \newline

{\color{blue}{Done.}} \newline

l735 hydrid --> hybrid \newline

{\color{blue}{Done.}} \newline

l748 Replace "shown" by "given" to avoid repeat of "showing/show". \newline

{\color{blue}{Done.}}

\subsection*{Figures}

\subsubsection*{Fig 1}
I think the grid lines could be thinner. On a printout this doesn't come out well. \newline
The color choice puts a lot of emphasis on the lakes in blue. Suggest to display the land-sea mask instead, i.e. make the lakes part of the land mask. \newline

{\color{blue}{Thinned the lines by half and changed the lake colors to tan/land.}} \newline

I would also consider using another projection that includes all of Greenland and it's surroundings. \newline

{\color{blue}{We have decided to keep the original polar projection, in order to show the polar points and Greenland.}} \newline

Caption. It is confusing to refer to all 4 grids as uniform, because they are uniform in different ways. At least distinguish them as uniform 1, 2 latitude-longitude grids (top) and quasi-uniform element size grids (bottom), or similar. \newline

{\color{blue}{Changed to new labeling discussed above.}}

\subsubsection*{Fig 10}
Better use y-axis label ICE+SNOW MELT consistent with Fig 11 and to avoid confusion ICE divided by SNOW. \newline

\subsubsection*{Fig 11}
The sign convention seems to be changed from Fig 10 with positive melt. Make consistent. \newline

\subsection*{References}

Many references are not formatted correctly in terms of upper case / lower case. There are also other problems like incomplete author lists and formatting problems. Please revise. Some examples below. \newline

l793 This is a regular paper citation that doesn't need a "Retrieved from" entry. Author list incomplete. \newline

l797 Incomplete author list. \newline

l924 Refer to final paper. Note title change: NCAR Topo v1.0 NCAR global model topography generation software for unstructured grids. https://doi.org/10.5194/gmd-8-3975-2015 \newline

l988 This is a regular paper citation that doesn't need a "Retrieved from" entry. Author list with "...". \newline

l1063 Refer to final paper. https://doi.org/10.5194/gmd-14-5023-2021 \newline

l1071 Refer to final paper. https://doi.org/10.5194/tc-13-1547-2019 \newline

\section*{Reviewer 2}

%\section*{\normalsize References}
%\setlength{\bibsep}{0pt}
%{\footnotesize
%\bibliography{bib}}

\end{document}

### 2022-09-06 first_submit/arcticgrids.tex
first commit after review

--- 2022-05-11 arcticgrids.tex
@@
 \begin{keypoints}
-\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two new Arctic refined mesh configurations.
-\item Uniform resolution grids degrade the surface mass balance of the Greenland Ice Sheet compared with equivalent low resolution latitude-longitude grids.
+\item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two Arctic refined mesh configurations.
+\item Uniform resolution grids degrade the Greenland Ice Sheet mass balance compared to equivalent resolution latitude-longitude grids.
 \item The refined Arctic meshes substantially improve the surface mass balance over all low resolution grids.
@@
 
-%\section*{Plain Language Summary}
+\section*{Plain Language Summary}
+
+The mass balance of Earth's big ice sheets is crucially important for understanding controls on global sea-level rise.
+However, the scale of the processes needed to represent ice sheet mass balance is challenging to resolve in conventional Earth System Models.
+This study evaluates the ability of different grids and atmospheric solvers (i.e, the dynamical core) in CESM2.2 to resolve the surface mass balance (SMB) of the Greenland Ice Sheet.
+We show that the ongoing transition away from latitude longitude grids, towards quasi-uniform unstructured grids in CESM2.2 leads to a degradation of the simulated SMB.
+Two variable-resolution grids with enhanced resolution over Greenland are developed and incorporated into the release of CESM2.2, which substantially improves the SMB over the latitude longitude grids.
 %[ enter your Plain Language Summary here or delete this section]
@@
 
-%{\color{blue} (WHL: Would it be worth mentioning that quasi-uniform grids also have benefits in polar regions, e.g.
-by removing elongated grid cells?)}{\color{red} (ARH: I don't think it's worth getting into that here.
-You'll see later in the manuscript that b/c the cells are narrower in the E-W direction, more cells can fit into the big ablation zone on the western margin.)}{\color{purple}(PHL: another advantage of lat-lon grids is that the coordinate lines are orthogonal!)}
-
-%{\color{blue} (WHL: This raises the question: How do lat-lon grids better resolve polar regions at a given nominal global resolution, if the filtering makes them effectively coarser?
-Does the damping degrade some processes more than others?)}{\color{red} (ARH - The phrase "approximately the same" doing a lot of heavy lifting here b/c I frankly don't know.
-The results from this paper suggest that the effective resolution is finer over polar regions, but this is a bit hand-wavy, despite my attempts to quantify it more exactly, and based on the precipitation rates having smaller biases.
-Also, to be clear, effective resolution is specific to atmospheric features; the land model is free to take advantage of these additional grids cells representing the ablation zone, and they certainly do.
-Perhaps I can add something to the effect of ``approximately the same everywhere on the grid, although marginal increases in effective resolution over the polar regions cannot be ruled out...")\\
-
-%{\color{purple}PHL: These are good points!
-I think it is useful to separate dynamics from physics in this discussion.
-Physics is called at more columns on the lat-lon grid as one approaches the poles.
-There is no "polar filtering" in physics so effectively the physics resolution is much higher than on a the cubed-sphere grid as one approaches the mid-latitudes and poles.
-Dynamics, on the other hand, filters out smaller scales so effectively it should only retain the 1 degree scale.
-That said, you will get a more accurate solution since you have more grid-points to represent waves (even though smaller scales are filtered) compared to the cubed-sphere grid.
-I think that is pretty clear in flow over orography.
-One can compare the lat-lon grid setup to a spectral model where one keeps viscosity coefficients fixed as one increased horizontal resolution - in such a setup one will get convergence of the dynamics!}}
-
-\begin{figure}[t]
-\begin{center}
-\begin{tabular}{cc}
-         \includegraphics[width=50mm]{figs/grid-f19.pdf}&
-         \includegraphics[width=50mm]{figs/grid-f09.pdf}\\
-         \includegraphics[width=50mm]{figs/grid-ne30pg2.pdf}&
-         \includegraphics[width=50mm]{figs/grid-ne30pg3.pdf} \\
-\end{tabular}
+\begin{figure}[t]
+\begin{center}
+%\begin{tabular}{cc}
+         \includegraphics[width=100mm]{grid-uniform}
+         %\includegraphics[width=50mm]{grid-f19.pdf}&
+         %\includegraphics[width=50mm]{grid-f09.pdf}\\
+         %\includegraphics[width=50mm]{grid-ne30pg2.pdf}&
+         %\includegraphics[width=50mm]{grid-ne30pg3.pdf} \\
+%\end{tabular}
 \end{center}
@@
 \begin{center}
-\begin{tabular}{ccc}
-         \includegraphics[width=40mm]{figs/grid-ARCTIC.pdf}&
-         \includegraphics[width=40mm]{figs/grid-ARCTICGRIS.pdf}&
-         \includegraphics[width=40mm]{figs/grid-CONUS.pdf} \\
-\end{tabular}
+%\begin{tabular}{ccc}
+         \includegraphics[width=140mm]{grid-varres.pdf}
+         %\includegraphics[width=40mm]{grid-ARCTIC.pdf}&
+         %\includegraphics[width=40mm]{grid-ARCTICGRIS.pdf}&
+         %\includegraphics[width=40mm]{grid-CONUS.pdf} \\
+%\end{tabular}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
+         \includegraphics[width=100mm]{temp_grisres.pdf}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-lores.pdf}
+         \includegraphics[width=130mm]{temp_dhgt_panel_STEND_CLUBB-lores.pdf}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_dhgt_panel_STEND_CLUBB-hires.pdf}
+         \includegraphics[width=130mm]{temp_dhgt_panel_STEND_CLUBB-hires.pdf}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_contours_diffERA5_Thyps.pdf}
+         \includegraphics[width=100mm]{temp_contours_diffERA5_Thyps.pdf}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_contours_diffCERES_SWCF.pdf}
+         \includegraphics[width=100mm]{temp_contours_diffCERES_SWCF.pdf}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_contours_diffCALIPSOdiffCERESdiffRACMO.pdf}
-\end{center}
-\caption{1979-1998 northern hemisphere summer (top) total cloud fraction bias relative to the CALIPSO dataset and incident shortwave radiation bias (W/m2), computed as the difference (middle) from CERES, and (bottom) RACMO2.3p2 dataset.
-The CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
+         \includegraphics[width=130mm]{temp_contours_diffCALIPSOdiffCERESdiffRACMO.pdf}
+\end{center}
+\caption{1979-1998 northern hemisphere summer surface incident shortwave radiation bias (W/m2), computed as the difference (top) from CERES, and (middle) RACMO2.3p2 dataset, and the (bottom) total cloud fraction bias relative to the CALIPSO dataset.
+CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
 Note that the averaging period for the CALIPSO-GOCCP and CERES-EBAF panels, 2006-2017 and 2003-2020, respectively, are different from the averaging period for the model results.}
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_contours_diffRACMO}
+         \includegraphics[width=130mm]{temp_contours_diffRACMO}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_tseries_GRIS.pdf}
+         \includegraphics[width=100mm]{temp_tseries_GRIS.pdf}
 \end{center}
@@
 \begin{center}
-         \includegraphics[width=100mm]{figs/temp_tseries_BASIN.pdf}
+         \includegraphics[width=100mm]{temp_tseries_BASIN.pdf}
 \end{center}
@@
 \begin{tabular}{cccc}
-         \includegraphics[width=130mm]{figs/temp_zsects.pdf}
+         \includegraphics[width=130mm]{temp_zsects.pdf}
 \end{tabular}
@@
 \caption{Model surface elevation along the (top) K-transect, and (bottom) a transect spanning the central dome down to the Kangerlussuaq glacier in southeast Greenland, for all model grids.
-The reference surface (GMTED) is a 1 km surface elevation dataset used for generating the CAM topographic boundary conditions.
+The GMTED reference surface is a 1 km surface elevation dataset \cite{GMTED2010} used for generating the CAM topographic boundary conditions.
 }
@@
 \begin{tabular}{cccc}
-         \includegraphics[width=80mm]{figs/temp_xy_diffRACMO_melt.pdf}
+         \includegraphics[width=80mm]{temp_xy_diffRACMO_melt.pdf}
 \end{tabular}
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/temp_composite_ge45N_pdf.pdf}
+         \includegraphics[width=130mm]{temp_composite_ge45N_pdf.pdf}
 \end{center}
@@
 We are  working to develop a configuration of the $\tt Arctic$ grid that is fully-coupled with the CESM ocean and sea ice components and the Community Ice Sheet Model (CISM), to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise.
-We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube\footnote{\url{https://www.youtube.com/watch?v=YwHgqDu75s8&t=4s&ab_channel=NCARVisLab}}, to increase awareness of the capabilities of CESM2.2.
+We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube (see link in Acknowledgements) to increase awareness of the capabilities of CESM2.2.
 Figure~\ref{fig:viz} shows a snapshot of this visualization, illustrating mesoscale katabatic winds descending the southeastern slopes of the GrIS.
@@
 \begin{center}
-         \includegraphics[width=130mm]{figs/Vis1472.jpg}
+         \includegraphics[width=130mm]{Vis1472.jpg}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=145mm]{figs/OMEGA500F2000climo.pdf}
+\includegraphics[width=145mm]{OMEGA500F2000climo.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=145mm]{figs/OMEGA500FHS94.pdf}
+\includegraphics[width=145mm]{OMEGA500FHS94.pdf}
 \end{center}

### 2022-09-06 second_submit/arcticgrids.tex
first commit after review

--- 2022-09-06 first_submit/arcticgrids.tex
@@
 \item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two Arctic refined mesh configurations.
-\item Uniform resolution grids degrade the Greenland Ice Sheet mass balance compared to equivalent resolution latitude-longitude grids.
-\item The refined Arctic meshes substantially improve the surface mass balance over all low resolution grids.
+\item {\color{blue}{Quasi-uniform unstructured}} grids degrade the Greenland Ice Sheet mass balance compared to latitude-longitude grids, {\color{blue}{at the conventional $1^{\circ}$ resolution}}.
+\item The refined Arctic meshes substantially improve the surface mass balance over {\color{blue}{conventional grid resolutions}}.
 \end{keypoints}
@@
 Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS).
-The conventional $1^{\circ}-2^{\circ}$ uniform resolution grids systematically overestimate both accumulation and ablation over the GrIS.
+The conventional $1^{\circ}-2^{\circ}$ resolution grids systematically overestimate both accumulation and ablation over the GrIS.
 Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS.
@@
 Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance.
-As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic-refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids.
+As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic-refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids, albeit at increased computational cost.
 
@@
 This study evaluates the ability of different grids and atmospheric solvers (i.e, the dynamical core) in CESM2.2 to resolve the surface mass balance (SMB) of the Greenland Ice Sheet.
-We show that the ongoing transition away from latitude longitude grids, towards quasi-uniform unstructured grids in CESM2.2 leads to a degradation of the simulated SMB.
+We show that the ongoing transition away from latitude{\color{blue}{-}}longitude grids, towards quasi-uniform unstructured grids in CESM2.2 leads to a degradation of the simulated SMB.
 Two variable-resolution grids with enhanced resolution over Greenland are developed and incorporated into the release of CESM2.2, which substantially improves the SMB over the latitude longitude grids.
@@
 Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain.
-With the recent trend towards globally uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost.
+With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, which is not a characteristic of lat-lon grids, which are highly anisotropic)}}.
 In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate.
@@
 
-Dycores built on lat-lon grids have some advantages over unstructured grids.
+Dycores built on lat-lon grids have some advantages over {\color{blue}{dycores built on}} unstructured grids.
 Lat-lon coordinate lines are orthogonal, and aligned with zonally symmetric circulations that characterize many large-scale features of Earth's atmosphere.
@@
 
-The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which have large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}.
+The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which {\color{blue}{has a}} large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}.
 Unstructured grids support quasi-uniform grid spacing globally, and there is no pole problem (e.g., Figure~\ref{fig:uni-grids}c).
-Conversely, unstructured grids are becoming more common due to their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}.
-This grid flexibility allows for the adoption of variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
+{\color{blue}{This is in part why}} unstructured grids are becoming more common; their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}.
+This increased grid flexibility allows for the adoption of variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
 In principle, grid refinement over polar regions can make up for any loss of resolution in transitioning away from lat-lon grids (e.g., Figure~\ref{fig:vr-grids}).
@@
 Polar filters do not directly interfere with the physical parameterizations, nor do they have any bearing on the surface models; e.g., the land model can take full advantage of the greater number of grid cells in polar regions on lat-lon grids.
-This is particularly relevant for the surface mass balance of the Greenland Ice Sheet (\textit{SMB}; the integrated sum of precipitation and runoff), which relies on hydrological processes represented in the land model.
+This is particularly relevant for the surface mass balance {\color{blue}{(\textit{SMB}; the integrated sum of precipitation and runoff)}} of the Greenland Ice Sheet, which relies on hydrological processes represented in the land model.
 
@@
 The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias \cite{P2000GPC,VETAL2018TC}.
-GrIS ablation areas (marginal regions where seasonal melting exceeds the annual mass input from precipitation) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
+GrIS ablation areas {\color{blue}{(marginal regions where the annual SMB is negative)}} are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
 These narrow ablation zones are not fully resolved in low-resolution GCMs, and may further degrade the simulated SMB.
@@
 
-In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating a grid where Greenland is represented with $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution.
+In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating {\color{blue}{grids}} where Greenland is represented with {\color{blue}{up to}} $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution.
 The simulated SMB compared favorably to the SMB from RCMs and observations.
@@
 \end{center}
-\caption{Computational grids for the uniform $1^{\circ}-2^{\circ}$ grids in this study.
+\caption{Computational grids for the $1^{\circ}-2^{\circ}$ {\color{blue}{lat-lon and quasi-uniform unstructured}} grids in this study.
 Grids names after Table~\ref{tbl:table1}, (a) $\tt f19$, (b) $\tt f09$, (c) $\tt ne30pg2$ and (d) $\tt ne30pg3$.}
@@
 \section{Methods}\label{sec:methods}
+
+{\color{blue}{CESM2.2 is a CMIP6-class \cite<Coupled Model Intercomparison Project Phase 6;>{CMIP6} Earth System Model maintained by the National Center for Atmosphere Research.
+CESM2.2 contains sub-component models for the atmosphere, land, ocean, sea-ice, and land-ice, and can be configured to run with varying degrees of complexity.
+All simulations described in this study use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
+In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2.
+The land-ice component is not dynamically active in $\tt FHIST$.
+However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the functionality to simulate the surface mass balance of the Greenland Ice Sheet.}}
+
 \subsection{Dynamical cores}
@@
 This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores.
-CESM2 runs submitted to the Coupled Model Intercomparison Project Phase 6 \cite{CMIP6} used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BETAL2018CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
+CESM2 runs submitted to CMIP6 used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BETAL2018CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
 
@@
 
-The FV dycore is a hydrostatic model that integrates the equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}.
+The FV dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}.
 The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}.
@@
 
-The SE dycore is a hydrostatic model that integrates the equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}.
+The SE dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}.
 The computational domain is a cubed-sphere grid tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
@@
 
-In CESM2.2, the SE numerics have been enhanced to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}.
-
-The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over uniform-resolution setups.
+In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to the CESM2.0 release}} to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}.
+
+The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over {\color{blue}{quasi-uniform resolution grids}}.
 First, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
@@
 
-For SE grids with quasi-uniform grid spacing, the SE tracer transport scheme is replaced with the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}.
+When using the SE dycore with quasi-uniform grid spacing, the SE tracer transport scheme is replaced with the Conservative Semi-Lagrangian Multi-tracer transport scheme (CSLAM) \cite<>{LTOUNGK2017MWR}.
 Atmospheric tracers have large, nearly discontinuous horizontal gradients that are difficult to represent with spectral methods, which are prone to oscillatory ``Gibbs-ringing" errors \cite{RW1990QJR}.
@@
 It uses a separate grid from the spectral-element dynamics, dividing each element into $3\times3$ control volumes with quasi-equal area.
-The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the original SE dycore in which the physics are evaluated Gauss-Lobatto-Legendre points \cite{HL2018MWR}.
+The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the original SE dycore in which the physics are evaluated at Gauss-Lobatto-Legendre points \cite{HL2018MWR}.
+{\color{blue}{CSLAM advection is not an available option in the VR configuration, which instead uses the standard SE tracer transport scheme with the physics evaluated at Gauss-Lobatto-Legendre points.}}
+
+\subsection{Physical parameterizations}
+
+All simulations in this study use the CAM6 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{GETAL2019GRL}.
+The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer, shallow convection, and cloud macrophysics scheme.
+CLUBB is coupled with the MG2 microphysics scheme \cite{GM2015JC,MG2}, which computes prognostic precipitation and uses classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
+Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
+Boundary layer form drag is modeled after \citeA{BBW2004QJRMS}, and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale (the ridge orientation is derived from a high-resolution, global topography dataset \cite{Danielson_Gesch_2011_GMTED2010}).
+
+Initial simulations with the SE dycore produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore in the standard CESM2 configuration.
+The SE dycore in CESM2.2 therefore has two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance.
+We reduced the width of the sub-grid distribution of vertical velocity ($\tt clubb\_gamma=0.308\rightarrow0.270$) and also reduced the strength of the damping for horizontal component of turbulent energy ($\tt clubb\_c14=2.2\rightarrow1.6$) to increase cloudiness.
+For a description of how CLUBB parameters impact the simulated climate, see \citeA{GETAL2015JAMES}.
 
@@
  \scriptsize
- \begin{tabular}{lccccc}
+ \begin{tabular}{lcccccc}
    \hline
-   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{refine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) \\ 
+   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{25 nodes, cost}} & {\color{blue}{50 nodes, cost}} \\ 
    \hline
-   $\tt f19$ & FV & 278 & - &1800 \\
-   $\tt f09$ & FV & 139 & - &1800 \\
-   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 \\
-   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 \\
-   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 \\
-   $\tt Arctic$ & SE & 111 & 28 & 450 \\
-   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 \\
+   $\tt f19$ & FV & 278 & - &1800 & {\color{blue}{436.66}} & {\color{blue}{-}} \\
+   $\tt f09$ & FV & 139 & - &1800 & {\color{blue}{1534.57}} & {\color{blue}{2024.24}} \\
+   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & {\color{blue}{1497.26}} & {\color{blue}{1683.97}} \\
+   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & {\color{blue}{1890.48}} & {\color{blue}{2090.43}} \\
+   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & {\color{blue}{-}} & {\color{blue}{-}} \\
+   $\tt Arctic$ & SE & 111 & 28 & 450 & {\color{blue}{15947.41}} & {\color{blue}{16675.45}} \\
+   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & {\color{blue}{40305.03}} & {\color{blue}{41036.67}} \\
  \hline
@@
   \caption{Grids and dycores used in this study.
-$\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{refine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step.
+$\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{fine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step.
 FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.
-We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.}
+We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.
+{\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY).
+The costs are from single month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
  \label{tbl:table1}
@@
 The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution.
-Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids in this study.
+Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids, {\color{blue}{as well as two grids pertaining to the Regional Atmospheric Climate Model \cite<RACMO;>{NETAL2018TC,NETAL2019SCIENCE}, which are used for validation purposes in this study (Table~\ref{tbl:table2})}}.
 The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid.
@@
 Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
-The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km, and is about 10 times more expensive than the $1^{\circ}$ models.
-The $\tt Arctic-GrIS$ grid is about twice as expensive as the $\tt Arctic$ grid.
+The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km.
 
@@
 
-\subsection{Physical parameterizations}
-
-All simulations in this study use the CAM6 physical parameterization package \cite<hereafter referred to as the \textit{physics};>{GETAL2019GRL}.
-The physics in CAM6 differs from its predecessors through the incorporation of high-order turbulence closure, Cloud Layers Unified by Binormals \cite<CLUBB;>{GETAL2002JAS,BOG2013JCLIM}, which jointly acts as a planetary boundary layer, shallow convection, and cloud macrophysics scheme.
-CLUBB is coupled with the MG2 microphysics scheme \cite{GM2015JC,MG2}, which computes prognostic precipitation and uses classical nucleation theory to represent cloud ice for improved cloud-aerosol interactions.
-Deep convection is parameterized using a convective quasi-equilibrium mass flux scheme \cite{ZM1995AO,NRJ2008JC} and includes convective momentum transport \cite{RSG2010JAS}.
-Boundary layer form drag is modeled after \citeA{BBW2004QJRMS}, and orographic gravity wave drag is represented with an anisotropic method informed by the orientation of topographic ridges at the sub-grid scale (the ridge orientation is derived from a high-resolution, global topography dataset \cite{Danielson_Gesch_2011_GMTED2010}).
-
-Initial simulations with the $\tt ne30pg3$ SE grid produced weaker shortwave cloud forcing relative to the tuned finite-volume dycore in the standard CESM2 configuration.
-The SE dycore in CESM2.2 therefore has two CLUBB parameter changes to provide more realistic cloud forcing and top-of-atmosphere radiation balance.
-We reduced the width of the sub-grid distribution of vertical velocity ($\tt clubb\_gamma=0.308\rightarrow0.270$) and also reduced the strength of the damping for horizontal component of turbulent energy ($\tt clubb\_c14=2.2\rightarrow1.6$) to increase cloudiness.
-For a description of how CLUBB parameters impact the simulated climate, see \citeA{GETAL2015JAMES}.
+\subsection{{\color{blue}{Computational costs}}}
+
+{\color{blue}{The last columns of Table~\ref{tbl:table1} provides cost estimates for the different grids and dycores.
+The costs, expressed as core hours per simulated year (CHPSY), are taken from single month runs of $\tt FHIST$, with no i/o, and using 25 nodes (900 tasks) and 50 nodes (1800 tasks) on the Cheyenne supercomputer \cite{cheyenne}.
+25 nodes is on the low side for a typical multi-decadal climate simulation at $1^{\circ}$ resolution, but it's the largest number of tasks that can be supported by the $\tt f19$ grid, and we chose to fix the number of tasks across all grids for the purposes of comparing their costs.
+We also provide costs using 50 nodes, excluding $\tt f19$, to provide a more practical cost estimate for longer climate integrations.
+There are probably better approaches for comparing costs across different grids and dycores, e.g., holding fixed the number of grid columns per task, but it is beyond the scope of this study.
+
+%Running with the more common $1^{\circ}$ grid for multi-decadal climate simulations, 25 nodes is on the low side, but it's the largest number of tasks that can be supported by the $\tt f19$ grid, and we chose to fix the number of tasks across all grids for the purposes of comparing their costs.
+
+The cheapest grid is the $\tt f19$ grid at 436.66 CHPSY, as this is the only grid running with $2^{\circ}$ dynamics.
+The $\tt f09$ grid costs 1534.57 CHPSY using 25 nodes, which is noticeably cheaper than $\tt ne30pg3$ at 1890.48 CHPSY.
+The $\tt ne30pg2$ grid is 20$\%$ cheaper than the $\tt ne30pg3$ grid, in both the 25 node and 50 node runs, consistent with previous estimates \cite{HR2018JAMES}.
+The FV model is known to be cheaper than SE at small core counts, whereas SE is more efficient than FV at large core counts due to its improved scalability \cite{DFSSTT2005IJHPC,DetAl2012IJHPCA}.
+In the more conventional 50 node runs, $\tt f09$ costs are much more similar to $\tt ne30pg3$, due to a 30$\%$ cost increase in $\tt f09$ relative to the 25 node run (Table~\ref{tbl:table1}).
+The $\tt Arctic$ grid is an order of magnitude more expensive than the lat-lon and quasi-uniform grids, at about 16k CHPSY, whereas the $\tt Arctic-GrIS$ grid is a more than twice that (40k CHPSY).
+Note that all timing numbers are from runs without threading.
+The $\tt f09$ grid is the only grid that runs out-of-the-box with threading; holding the number of tasks fixed leads to a $4\%-6.5\%$ reduction in CHPSY compared to runs without threading.}}
 
@@
 
-All grids and dycores simulate the GrIS SMB, which is the sum of mass accumulation from precipitation and mass loss from ablation.
-The latter is the sum of evaporation, sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, forming ice lenses.
+{\color{blue}{CESM simulates the GrIS SMB as the sum of ice accumulation and ice ablation.
+The latter contains contributions from sublimation and liquid runoff from ice melt.
+Liquid precipitation and liquid runoff may also contribute to ice accumulation by penetrating pore spaces in the snowpack/firn layer and freeze, forming ice lenses}}.
 These relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
@@
 In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (ECs), each with a distinct surface energy balance and SMB.
-The area fraction of each EC is derived from a high-resolution GrIS elevation dataset.
+The area fraction of each EC is computed from {\color{blue}{the CISM initial conditions, which are based on a high-resolution dataset of the observed, modern extent and thickness of the GrIS \cite{METAL2014NG}.
+Note that for configurations with a dynamically active ice sheet, the area fractions are continuously updated throughout the run to reflect the evolving ice sheet geometry in CISM}}.
 The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean.
 Precipitation from CAM is repartitioned into solid or liquid based on the surface temperature of the EC; precipitation falls as snow for temperatures between $T < -2^\circ$~C, as rain for $T > 0^\circ$~C,  and as a linear combination of rain and snow for temperatures between $-2^\circ$~C and $0^\circ$~C.
+
+{\color{blue}{Changes in ice depth, not snow depth, count toward the SMB.
 Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent.
-Any snow above the 10 m cap contributes towards ice accumulation in the SMB, and refreezing of liquid water within the snowpack is an additional source of ice.
-Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
+Any snow above the 10 m cap contributes towards ice accumulation, and refreezing of liquid water within the snowpack is an additional source of ice.
+Surface ice melting (after melting of the overlying snow) yields a negative SMB.}} Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
 For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES,Muntjewerf_EA_2021_coupling}.
 
-Changes in ice depth, but not snow depth, count toward the SMB.
-That is, snow accumulation above the 10 m cap contributes a positive SMB, and surface ice melting (after melting of the overlying snow) yields a negative SMB.
-Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation (a model simulation with prescribed, observed sea-surface conditions) for about 500 years.
-The uniform-resolution grids are initialized with the SMB from an existing $\tt f09$ spun-up initial condition.
-In the simulations described here, the GrIS is prescribed at its observed, modern extent and thickness.
-
-\subsection{Observational Datasets}
-
-We use several observational datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations.
+Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation for about 500 years.
+The $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids are initialized with the SMB from an existing $\tt f09$ spun-up initial condition.
+
+\subsection{{\color{blue}{Validation}} Datasets}
+
+We use several {\color{blue}{validation}} datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations.
+{\color{blue}{The ERA5 reanalysis product \cite{ERA5} is used for validating the large-scale circulation and extreme precipitation events.
+Clouds and radiation fields are validated using remote sensing products, the CERES-EBAF ED4.1 \cite{CERES-EBAF} and the CALIPSO-GOCCP \cite{CALIPSO-GOCCP} datasets, respectively.}}
+
 SMB datasets are gathered from multiple sources.
-Regional Atmospheric Climate Model, version 2.3 11km \cite<RACMO23;>{NETAL2015TC} and version 2.3p2 5.5km \cite<RACMO2.3p2;>{NETAL2018TC,NETAL2019SCIENCE} are RCM simulations targeting Greenland, forced by ERA renalyses products at the domain's lateral boundaries.
+RACMO, version 2.3 11km \cite<RACMO23;>{NETAL2015TC} and version 2.3p2 5.5km \cite<RACMO2.3p2;>{NETAL2018TC,NETAL2019SCIENCE} are RCM simulations targeting Greenland, forced by ERA renalyses products at the domain's lateral boundaries.
 The RACMO simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite<e.g.,>{LIVVkit,KETAL2020JAMES}.
-
-In-situ SMB (snow pit and ice cores) and radar accumulation datasets (e.g., IceBridge) are maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}.
+{\color{blue}{The RACMO datasets are used along with the CERES-EBAF product to validate the radiative fluxes around Greenland.}}
+
+In-situ SMB (snow pit and ice cores) and {\color{blue}{remote sensing datasets (e.g., IceBridge radar accumulation dataset)}} are maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}.
 However, these point-wise measurements are difficult to compare to model output spanning several different grids, especially since the SMB from each elevation class is not available from the model output.
@@
  \end{tabular}
-  \caption{Description of observational datasets used in this study.}
+  \caption{Description of {\color{blue}{validation}} datasets used in this study.}
  \label{tbl:table2}
@@
 We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
-However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}b).
-CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
-The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
+However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset ({\color{blue}{i.e., the GrIS initial conditions in CISM;}} Figure~\ref{fig:grisdx}b).
+CLM's dataset creation tool generates the model ice mask by mapping the reference dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
+The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2.2 grids.
 The area errors in Figure~\ref{fig:grisdx}b may seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
@@
 \begin{equation}
-SMB = accumulation + runoff + evaporation/sublimation, \label{eq:SMB}
+SMB = accumulation + runoff + evaporation + sublimation, \label{eq:SMB}
 \end{equation}
@@
 Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.
-Note that this SMB expression is different from the internally computed SMB described in section~\ref{sec:simSMB}.
+
+{\color{blue}{The total water SMB (equation~\ref{eq:SMB}) is different from the SMB internally computed by CLM and described in section~\ref{sec:simSMB}, which only tracks ice mass.
+We do not use CLM's internally computed SMB in this study.
+Rather, we utilize the components of the internally computed SMB to construct the total water SMB}}.
 
@@
 
-The different error characteristics of the two methods are used to diversify the ensemble.
+The different error characteristics of the two methods are used to {\color{blue}{further}} diversify the ensemble.
 Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
@@
 
-\subsection{Experimental design}
-
-All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
-In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2.
-
 \section{Results}\label{sec:results}
@@
 Before delving into the simulated Arctic climate conditions, we assess the global mean differences between the various grids and dycores.
-Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between uniform-resolution grids and dycores.
+Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured}} grids and dycores.
 The $\tt f09$ grid is warmer than the $\tt f19$ grid, primarily in the mid-to-high latitudes throughout the depth of the troposphere.
@@
 
-Comparing the VR grids to the uniform-resolution grids is complicated because we simultaneously increase the resolution and reduce the physics time-step, both of which influence the solution \cite{W2008TELLUS}.
+Comparing the VR grids to the {\color{blue}{lat-lon and quasi-uniform}} grids is complicated because we simultaneously increase the resolution and reduce the physics time-step, both of which influence the solution \cite{W2008TELLUS}.
 We therefore run an additional $\tt ne30pg3$ simulation with the shorter physics time step used in the $\tt Arctic$ grid (450~s), referred to as $\tt ne30pg3^{*}$ (Table~\ref{tbl:table1}).
@@
 
-It is useful to understand summer temperature biases due to their control on ice and snow melt over the GrIS \cite<>{O2001JAM}.
+{\color{blue}{Keeping our focus on the Arctic region, and in particular Greenland,}} it is useful to understand summer temperature biases due to their control on ice and snow melt over the GrIS \cite<>{O2001JAM}.
 Figure~\ref{fig:dThyps} shows the 1979--1998 lower troposphere summer temperature bias relative to ERA5, computed by equating a layer mean virtual temperature with the 500--1000 hPa geopotential thickness.
@@
 A negative bias corresponds to excessive reflection and cooling.
-The uniform grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses.
+The {\color{blue}{lat-lon and quasi-uniform}} grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses.
 There is also a halo of positive bias (clouds not reflective enough) around the ocean perimeter of Greenland.
@@
 While we have not quantified the contribution of cloud biases to the cooler Arctic temperatures, shortwave radiation is a crucial component of the Arctic energy budget during summer.
-The shortwave cloud forcing biases are on the order of 10 W/m2, which is a significant fraction of the total absorbed shortwave during Arctic summer  \cite{SETAL2007JGR} and is therefore likely a factor contributing to the cooler temperatures.
+The shortwave cloud forcing biases are on the order of 10 W/m2, which is a significant fraction of the total absorbed shortwave {\color{blue}{radiation}} during Arctic summer  \cite{SETAL2007JGR} and is therefore likely a factor contributing to the cooler temperatures.
 
@@
 The total SMB on coarse grids therefore has smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
-Such compensating errors underlie the importance of understanding the extent to which a model is getting the right SMB for the right reasons.
+Such compensating errors {\color{blue}{underline}} the importance of understanding the extent to which a model is getting the right SMB for the right reasons.
 
@@
 The 1979-1998 climatological mean values from Table~\ref{tbl:table3} are shown as circles on the right side of the panels.
-The uniform $1^{\circ}-2^{\circ}$ grids have positive precipitation biases, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
+The {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform}} grids have positive precipitation biases, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
 The $\tt f19$ and $\tt f09$ grids perform similarly, with +110 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +165 Gt/yr and $\tt ne30pg2$ by +230 Gt/yr.
@@
 The basin-integrated melt rates in Figure~\ref{fig:basin} depend on the dycore.
-The uniform-resolution SE grids have the largest positive biases in all basins.
-The $\tt Arctic-GrIS$ grid is a close second, while the FV grids have systematically smaller melt-rates.
-The ``second-place" standing of $\tt Arctic-GrIS$ is somewhat unexpected, as this grid has the warmest lower-troposphere summer temperatures (Figure~\ref{fig:dThyps}) and greatest incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the uniform-resolution SE grids.
+The {\color{blue}{quasi-}}uniform SE grids have the largest positive biases in all basins.
+The $\tt Arctic-GrIS$ grid is a close second, while the FV grids have systematically {\color{blue}{smaller melt-rates and melt-rate biases}}.
+The ``second-place" standing of $\tt Arctic-GrIS$ is somewhat unexpected, as this grid has the warmest lower-troposphere summer temperatures (Figure~\ref{fig:dThyps}) and greatest incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the quasi-uniform SE grids.
 
@@
 We compute PDD from monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}$C and a degree-day factor of 5~mm~d$^{-1}$ $^{\circ}$C$^{-1}$.
+{\color{blue}{This specific degree-day factor lies between typical values reported for snow and ice, in order to easily apply the PDD method to estimate the combined snow and ice melt.}}
 
@@
 The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids have similar dynamical resolution and use identical smoothing.
-This suggests that coarser models will tend to elevate the ablation zones and thereby depress melt rates, which is the opposite of the melt bias that occurs in the coarse grid simulations.
+This suggests that coarser models will tend to elevate the ablation zones {\color{blue}{relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and}} depressed melt rates, but this is opposite the melt bias that occurs in the coarse grid simulations.
 
@@
 We hypothesize that coarser models have larger melt biases because summer melting is confined to these mixed ice/land/ocean grid cells.
-CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees the homogenized state after volume averaging over the sub-grid mixture.
-Thus, warm ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice-covered patch.
+CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees the homogenized state after {\color{blue}{area}} averaging over the sub-grid mixture.
+Thus, warm ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice-covered patch {\color{blue}{and more melting}}.
 
@@
 The figure shows that coarser grids can be characterized by a monotonic increase in melt bias as the grid cell ice fraction decreases.
-The VR grids have the smallest melt biases for small grid cell ice fractions (smaller than 50\%), the uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases.
+The VR grids have the smallest melt biases for small grid cell ice fractions (smaller than 50\%), the {\color{blue}{quasi-}}uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases.
 Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone of coarse grids is responsible for their large melt bias.
@@
 
-Figure~\ref{fig:comp-pdf} shows monthly PDFs of the precipitation rates associated with storms.
+Figure~\ref{fig:comp-pdf} shows monthly PDF {\color{blue}{(probability density function)}} of the precipitation rates associated with storms.
 The PDFs are constructed by sampling all the precipitation rates within $30^{\circ}$ of the storm center, for each point on the storm track and for all storms.
@@
 Both use the SE dycore; the $\tt Arctic$ grid has $1/4^{\circ}$ refinement over the broader Arctic, whereas the $\tt Arctic-GrIS$ grid is identical except for a $1/8^{\circ}$ patch of refinement over Greenland.
-A third VR grid, $\tt CONUS$, has also been made available in CESM2.2.
+A third VR grid, $\tt CONUS$, {\color{blue}{with $1/8^{\circ}$ refinement over the US}}, has also been made available in CESM2.2.
 
 In general, the FV grids have colder summer temperatures over the Arctic compared with the SE grids (including the VR grids).
-The cloud biases in all the uniform-resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.
+The cloud biases in all the {\color{blue}{lat-lon and quasi-}}uniform resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.
 It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates.
@@
 We integrated the primary source and sink terms of the SMB equation over the GrIS for each of the six grids.
-The uniform $1^{\circ}-2^{\circ}$ grids have large positive accumulation biases because they fail to resolve orographic precipitation.
-The uniform SE grids have larger accumulation biases, suggesting that the FV grids are more skillful for precipitation due to finer resolution over Greenland, and despite a polar filter.
+The {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform}} grids have large positive accumulation biases because they fail to resolve orographic precipitation.
+The quasi-uniform SE grids have larger accumulation biases, suggesting that the FV grids are more skillful for precipitation due to finer resolution over Greenland, and despite a polar filter.
 The VR grids have the most accurate accumulation rates of all the grids.
-The primary mass sink term of the GrIS, ice/snow melt, has similar biases; the coarse grids melt too much, with a greater bias for uniform SE grids.
+The primary mass sink term of the GrIS, ice/snow melt, has similar biases; the coarse grids melt too much, with a greater bias for quasi-uniform SE grids.
 In general, on coarse grids, errors in the individual SMB terms are larger than the errors in the SMB itself, due to compensating errors.
@@
 
-The $\tt Arctic-GrIS$ grid has the warmest summer lower troposphere of all grids, yet it has less melting than the uniform-resolution SE grids.
+The $\tt Arctic-GrIS$ grid has the warmest summer lower troposphere of all grids, yet it has less melting than the quasi-uniform resolution SE grids.
 This suggests that grid resolution is somehow contributing to the melt biases in coarse grids, in a way that is not obvious from the large-scale dynamics.
@@
 
-The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the GrIS SMB, compared to the uniform $1^{\circ}-2^{\circ}$ grids.
+The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the GrIS SMB, compared to the {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform}} grids.
 The $\tt Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but its summer temperatures are too warm.
 The $1^{\circ}$ FV model gives a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids (but perhaps also because it is the most heavily tuned model configuration in CESM).
-In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other uniform-resolution grids.
+In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other {\color{blue}{lat-lon and quasi-}}uniform grids.
 
@@
 \end{equation}
-where $A$ and $B$ are the standard hydrid coefficients (using a dry-mass generalized vertical mass coordinate $\eta$).
+where $A$ and $B$ are the standard {\color{blue}{hybrid}} coefficients (using a dry-mass generalized vertical mass coordinate $\eta$).
 These reference profiles are subtracted from the prognostic temperature and pressure-level-thickness states before applying hyperviscosity:
@@
 \end{equation}
-The derivation showing that (\ref{eq:pgf}) and (\ref{eq:pgf2}) are equivalent is shown here:
+The derivation showing that (\ref{eq:pgf}) and (\ref{eq:pgf2}) are equivalent is {\color{blue}{given}} here:
 \begin{eqnarray*}
@@
     This material is based upon work supported by the National Center for Atmospheric Research (NCAR), which is a major facility sponsored by the NSF under Cooperative Agreement 1852977.
-Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX), were provided by the Computational and Information Systems Laboratory (CISL) at NCAR.
+Computing and data storage resources, including the Cheyenne supercomputer \cite{cheyenne}, were provided by the Computational and Information Systems Laboratory (CISL) at NCAR.
 A.

### 2022-09-08 first_review/response-to-reviewers.tex
working on reviewer 2

@@
 
+{\color{blue}{We have added a subsection on computational costs to address the reviewers concern, in the methods section.}} \newline
+
 The way the term surface mass balance (SMB) is defined and used in the manuscript is confusing.
@@
 
+{\color{blue}{We thanks the reviewer for bringing this to our attention, and agree that our constant referencing of what the SMB refers to has led to some inconsistent definitions and confusion in the text.
+We have (1) removed attempts (other than very generally) to define the SMB (and the ablation zone) in parenthesis after introducing these terms in the introduction, (2) clarified the description of the SMB that is computed in CESM (the section ``Simulated surface mass balance"), and (3) made clear in the ``SMB analysis" section that the SMB equation we are computing in this paper is different from that computed by CESM.}} \newline
+
 In the discussion about the different dycores, I was wondering if there could be a notion that the latitude-longitude approaches are more mature because they have a longer history.
@@
 
+{\color{blue}{This is a good point.
+A case in point would be the improvements to the flow over topography going from CESM2.0 to CESM2.2.
+We are still developing CAM-SE.
+We have elaborated on this point in the methods section (where we point to the CESM2.2 enhancements in the Appendix) and the conclusions section, the second to last paragraph.}} \newline
+
 The relationship between atmospheric grid and land grid seems to be important.
@@
 \newline
+
+{\color{blue}{Based on my understanding of how the atmosphere and land models are coupled, I don't think running the land model at a higher resolution (and holding the atmosphere resolution fixed) will alleviate the melt biases that tends to occur with partially ice covered grid cells.
+This is because it is the averaging of the land state over the atmospheric grid cell that is responsible for homogenizing the state.
+Even if the land cells are a lot smaller, and even binary in ice cover, If the fractional ice cover within an \textit{atmospheric grid cell} is still small, then we think that the atmospheric state that is then passed back to the land model will still be unduly influenced by the non-ice portions of the atmosphere grid cell.}}
 
@@
 
-{\color{blue}{added: relative to the CESM2.0 release.}} \newline
+{\color{blue}{Added: relative to the CESM2.0 release.}} \newline
 
@@
 
+{\color{blue}{We've done performed a Monte Carlo analysis similar to what is suggested and have found that the errors due to inter-annual variability (in Figure 10) are much smaller than our current error bars, and have therefore omitted these estimates.}} \newline
+
 l553 "smaller melt-rates" or smaller melt-rate biases? \newline
@@
 
+{\color{blue}{Done.}} \newline
+
 \subsubsection*{Fig 11}
 The sign convention seems to be changed from Fig 10 with positive melt. Make consistent. \newline
+
+{\color{blue}{Done.}} \newline
 
@@
 
+{\color{blue}{Done.}} \newline
+
 l797 Incomplete author list. \newline
@@
 l924 Refer to final paper. Note title change: NCAR Topo v1.0 NCAR global model topography generation software for unstructured grids. https://doi.org/10.5194/gmd-8-3975-2015 \newline
+
+{\color{blue}{Done.}} \newline
 
@@
 
+{\color{blue}{Done.}} \newline
+
 l1071 Refer to final paper. https://doi.org/10.5194/tc-13-1547-2019 \newline
 
+{\color{blue}{Done.}}
+
 \section*{Reviewer 2}
+
+\subsection*{General comments}
+
+The manuscript is well-written and provides an important contribution to understanding simulation of ice sheet SMB in global climate model simulations.
+The paper provides important documentation of the impact of changes to the grid and spatial resolution in CESM on ice sheet SMB.
+Some points below should be addressed before the manuscript is published but I feel these are minor overall.
+\newline
+
+One general comment: surface albedo has been shown to be an important factor in climate model simulations of GrIS SMB (e.g.
+van Angelen et al., 2012; Cullather et al., 2014; Helsen et al., 2017; Alexander et al., 2019), but this has not been discussed by the authors.
+Some of the melt differences could be influenced by the interaction between precipitation, surface albedo, and melt.
+In areas of underestimated snowfall, which tend to occur along the ice sheet margins, erroneous bare ice exposure can lead to a darker surface and enhanced melt.
+IThe authors should examine the possibility of differences in albedo influencing simulated melt.
+\newline
+
+%Cullather, R.
+I., Nowicki, S.
+M.
+J., Zhao, B., & Suarez, M.
+J.
+(2014) Evaluation of the surface representation of the Greenland Ice Sheet in a General Circulation Model.
+J.
+Climate, 27 (13), 4835-4856.
+\newline
+
+%Helsen, M.
+M., van de Wal, R.
+S.
+W., Reerink, T.
+S., Bitanja, R., Madsen, M.
+S., Yang, S., Li, Q., & Zhang, Q.
+(2017) On the importance of the albedo parameterization for the mass balance of the Greenland ice sheet in EC-Earth.
+The Cryosphere, 11, 1949-2017.
+\newline
+
+%Alexander, P.
+M., LeGrande, A.
+N., Fischer, E., Tedesco, M., Fettweis, X., Kelley, M., Nowicki, S.
+M.
+J., & Schmidt, G.
+A.
+(2019) Simulated Greenland surface mass balance in the GISS ModelE2 GCM: Role of the ice sheet surface.
+J.
+Geophys.
+Res., 124 (3), 750- 765.
+
+\subsection*{Specific comments}
+
+1.
+Lines 19-24: The language is confusing here.
+The quasi-uniform grids should be introduced with a brief description of how they are constructed, and what “quasi- uniform” means.
+Then their performance in terms of surface mass balance can be discussed, followed by the impact of grid refinement on the simulation.
+Also perhaps it can be mentioned here how one dycore is linked to two lon-lat configurations and another is linked to the quasi-uniform unstructured grid.
+\newline
+
+2. Line 38: Again, please explain the meaning of “quasi-uniform unstructured grid”  \newline
+
+3.
+Lines 52-55: Again, it would be helpful if “globally uniform unstructured grid”, “dynamical cores”, and “variable-resolution grids” can be briefly explained.
+Additionally, perhaps it should be explained here that two dycores are examined, one linked to the longitude latitude grid, and the other linked to unstructured grids.
+\newline
+
+4. Line 94: Why is “conversely” used here? It seems like a word such as “Additionally” would be more appropriate.  \newline
+
+5. Lines 115-116: Clarify that this is a positive precipitation bias in the interior.  \newline
+
+6.
+Lines 111-123: Here there is no mention of elevation classes.
+I suggest mentioning them briefly here as this is a way to increase SMB resolution (from the standpoint of runoff but not precipitation) that is used in CESM  \newline
+.
+7. Figure 1, caption: Please provide a brief description of each of the grids here for the benefit of the reader. E.g. (a) 2 latitude-longitude grid (f19), etc.  \newline
+
+8. Line 189: Briefly explain “cubed-sphere”.  \newline
+
+9. Table 1: It would be helpful to specify which grids are latitude-longitude and which are unstructured either as a column in the table or text in the caption.  \newline
+
+10. Lines 229-230: Does ne30 indicate 30 quadrilateral elements per face and pg3 indicate 3 control volumes per element? Please clarify. \newline
+
+11. Lines 239-243: I suppose degree coordinates here correspond the latitude dimension, or degrees at the equator. Please clarify.  \newline
+
+12. Line 249: This is the first time RACMO is mentioned. Please explain briefly what RACMO is here.  \newline
+
+13. Line 254: Clarify that “expensive” means “computationally expensive.” \newline
+
+14. Line 301: Though not essential it would be interesting to know the lapse rate that is assumed. \newline
+
+15. Line 314: I’m confused as to why accumulation below 10 m does not contribute to SMB. How is this addition of mass accounted for? \newline
+
+16. Lines 318-319: What is the period for the prescribed SSTs? \newline
+
+17.
+Lines 322-338: This section only includes discussion of RACMO and in-situ measurements, which are not used, while ERA5, CERES and CALIPSO are not at all discussed.
+These validation products must be documented here as well.
+\newline
+
+18.
+Line 322: RACMO2 is still a model, and does not fall under an “observational” dataset.
+I suggest changing this to read “Validation Datasets”, “Evaluation Datasets” or “Comparison Datasets”.
+\newline
+
+19. Figure 3 caption: I believe the left and right figures are switched here. Also, the fact that RACMO RCM simulations are included should be mentioned. \newline
+
+20. Formula 1: The f’s in runoff seem to have been turned into function symbols. Also I suggest using “evaporation + sublimation” to avoid confusion with division. \newline
+
+21.
+Line 406: I believe this should read “expressed as temperature differences”.
+22.
+Lines 408-416, 429-433: How can the effect be primarily ascribed to condensational heating if the temperature differences do not align in magnitude or spatial extent?
+Are there feedbacks involved?
+Please clarify in the text, or qualify the statements.
+\newline
+
+23. Lines 437-438: This also seems to be true in other cases. \newline
+
+24. Lines 440-441: This seems consistent with other results that increasing resolution led to a temperature increase. \newline
+
+25. Line 459: Perhaps the authors could mention the potential influence of clouds earlier as prior to this point, they are not mentioned. \newline
+
+26. Line 479: Clarify that cooler temperatures in the coarse resolution simulations are being referred to here. \newline
+
+27. Figure 8: The CALIPSO grid appears to be coarser resolution than the CERES grid. Is this a 2 grid? \newline
+
+28. Lines 573-575: Wouldn’t the elevation class scheme fix this problem by raising the ice-covered sub-grid portion to a higher elevation? \newline
+
+29.
+Lines 606-616: If the biases are larger at low elevations, it could also be coincidental that these areas also have lower ice fractions.
+I think the authors should also consider the possibility of other factors.
+\newline
+
+30. Lines 675-680: Perhaps the fact that clouds and precipitation move more easily inland could reduce snowfall and increase shortwave radiation along the coast, enhancing melt. \newline
+
+31.
+Lines 681-688: Again, I’m curious about the impact of elevation classes and whether precipitation-albedo feedbacks could influence the melt biases.
+I suppose the elevation classes might not entirely balance out other coarse- resolution effects.
+\newline
+
+32. Line 719: Change “spectra-element” to “spectral-element”. \newline
+
+33. Line 729: g can be defined after mentioning T1 for clarity. \newline
+
+34. Line 731: Add “is the” before “surface geopotential”. \newline
+
+35. Equation A5: How is T(ref) computed? \newline
+
+36. Lines 740-748: Here some terms should be defined, including Tv, qv and p0.
 

### 2022-09-08 second_submit/arcticgrids.tex
working on reviewer 2

@@
 Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain.
-With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, which is not a characteristic of lat-lon grids, which are highly anisotropic)}}.
+With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic)}}.
 In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate.
@@
 While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
-The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells elongated in the longitudinal direction over the polar regions) and polar filters are necessary to ameliorate this numerical instability \cite{JW2010LNCSE}.
+The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells near the polar singularity elongated in the zonal direction) and polar filters are necessary to ameliorate this numerical instability \cite{JW2010LNCSE}.
 Polar filters subdue the growth of unstable computational modes by applying additional damping to the numerical solution over polar regions.
@@
 This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores.
-CESM2 runs submitted to CMIP6 used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BETAL2018CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
+CESM2 runs submitted to CMIP6 used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BetAl2016CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
 
@@
 
-In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to the CESM2.0 release}} to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}.
+{\color{blue}{SE is a next generation of dycore, and is less mature than the FV dycore due to its shorter history.}} In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to CESM2.0}} to mitigate spurious noise over topography.
+These algorithmic changes are described in \ref{sec:a1}.
+{\color{blue}{Future versions of CESM will likely continue to include further optimizations and enhancements to the (nascent) SE dycore.}}
 
@@
 A tensor-hyperviscosity formulation is used \cite{GetAl2014GMD}, which adjusts the coefficients in two orthogonal directions to more accurately target highly distorted quadrilateral elements.
-Second, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{gmdd-8-4623-2015} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.
+Second, the topography boundary conditions are smoothed in a way that does not excite grid scale modes, and so the NCAR topography software \cite{LetAl2015GMD} has been modified to scale the smoothing radius by the local element size, resulting in rougher topography in the refinement zone.
 
@@
    \hline
-   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{25 nodes, cost}} & {\color{blue}{50 nodes, cost}} \\ 
+   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{cost(25 nodes)}} & {\color{blue}{cost(50 nodes)}} \\ 
    \hline
@@
 \end{center}
-\caption{1979-1998 northern hemisphere summer surface incident shortwave radiation bias (W/m2), computed as the difference (top) from CERES, and (middle) RACMO2.3p2 dataset, and the (bottom) total cloud fraction bias relative to the CALIPSO dataset.
+\caption{1979-1998 northern hemisphere summer surface incident shortwave radiation bias (W/m2), computed as the difference (top) from RACMO2.3p2, and (middle) the CERES-EBAF dataset, and the (bottom) total cloud fraction bias relative to the CALIPSO dataset.
 CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
@@
 Note that Figure~\ref{fig:basin} uses the basin boundaries for the two common ice masks, shown in the $\tt f19$ and $\tt ne30pg2$ panels, in computing the basin-scale integrals.
-Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}..}
+Blue lines in the $\tt f19$ panel show the location of the two transects plotted in Figure~\ref{fig:ztrans}.}
 \label{fig:prect}
@@
 One possibility is that these problematic grid cells are situated at lower elevations than the true ice sheet surface, leading to a warm bias and too much melt.
-Figure~\ref{fig:ztrans} shows the representation of the ice sheet surface along two transects on the different grids, compared to the high-resolution dataset used to generate CAM topographic boundary conditions \cite{GMTED2010,gmdd-8-4623-2015}.
+Figure~\ref{fig:ztrans} shows the representation of the ice sheet surface along two transects on the different grids, compared to the high-resolution dataset used to generate CAM topographic boundary conditions \cite{GMTED2010,LetAl2015GMD}.
 The two transects are shown in Figure~\ref{fig:prect}: the east-west ``K-transect'' in southwest Greenland and a transect extending from the central dome down to the  Kangerlussuaq glacier on the southeast coast.
@@
 The figure shows that coarser grids can be characterized by a monotonic increase in melt bias as the grid cell ice fraction decreases.
-The VR grids have the smallest melt biases for small grid cell ice fractions (smaller than 50\%), the {\color{blue}{quasi-}}uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases.
+The VR grids have the smallest melt biases for small grid cell ice fractions (less than 50\%), the {\color{blue}{quasi-}}uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases.
 Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone of coarse grids is responsible for their large melt bias.
@@
 \label{fig:bias}
+\end{figure}
+
+{\color{blue}{Another potential factor in explaining the differences across the various grids and dycores is the surface albedo.
+Bare ice has a lower albedo than snow, so it's reasonable to speculate whether the large melt biases in the coarse grid simulations are in part due to more frequent bare ice exposure.
+One logical pathway to more bare ice exposure is related to the precipitation and cloud biases.
+As the coarser models accumulate too much precipitation in the GrIS interior, the coastal ice margins receive too little snow, which can be more easily melted away to expose the underlying glacial ice.
+Figure~\ref{fig:alb} shows maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and CERES-EBAF datasets.
+The surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface.
+The lat-lon and quasi-uniform grids all have anomalously low albedo around the coasts in the southeast and northwest regions of the GrIS, which coincides with regions of anomalously low cloud cover and precipitation.
+In the VR simulations, the low albedo biases are alleviated to a large extent.
+
+In light of these albedo results, one may interpret Figure~\ref{fig:bias} differently.
+That there are larger melt biases for grid cells with only only partial ice cover, may have less to do with the partial ice cover and more to do with the enhanced bare ice exposure in the coarser grids.
+Further experimentation is required to tease out the relative contributions of these two processes.}}
+
+\begin{figure}[t]
+\begin{center}
+\begin{tabular}{cccc}
+         \includegraphics[width=130mm]{temp_contours_diffALB.pdf}
+\end{tabular}
+\end{center}
+\caption{1979-1998 northern hemisphere summer albedo bias (fraction), computed as the difference (top) from RACMO2.3p2, and (bottom) CERES-EBAF dataset.
+RACMO2.3p2 differences are found by mapping the model output to the individual model grids, and differences in the bottom panel are computed after mapping the individual model grids to the CERES-EBAF $1^{\circ}$ grid.
+Note that the averaging period for CERES-EBAF panels are 2003-2020, which is different from the averaging period for the model results.}
+\label{fig:alb}
 \end{figure}
@@
 We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model for simulating the GrIS SMB.
-That is, the simulated GrIS SMB will be adversely impacted in future CESM versions, after the FV dycore is phased out.
+{\color{blue}{The SE dycore is still under active development (e.g., \ref{sec:a1}) compared to the more mature FV dycore, and future algorithmic improvements may reduce the FV-SE GrIS SMB skill gap.
+However, such developments are unlikely to eliminate this skill gap entirely due to the fewer number of grid cells representing high latitude structures on quasi-uniform unstructured grids.}} The simulated GrIS SMB will be adversely impacted in future CESM versions, after the FV dycore is phased out.
 This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities.

### 2022-09-08 first_review/response-to-reviewers.tex
start reviewer 2

@@
 
+{\color{blue}{We agree that this is confusing and have added a sentence describing the different grids/dycores in more detail, before talking about their SMB performance.}}.
+
 2. Line 38: Again, please explain the meaning of “quasi-uniform unstructured grid”  \newline
+
+{\color{blue}{See above.}}.
 
@@
 \newline
+
+{\color{blue}{Added clarification for the meaning of quasi-uniform and variable-resolution.
+Decided against getting into the specifics of which dycore uses which grid, as this becomes apparent later on in the introduction}}.
 

### 2022-09-08 arcticgrids.tex
Updates from Overleaf

@@
 \item Uniform resolution grids degrade the surface mass balance of the Greenland Ice Sheet compared with equivalent low resolution latitude-longitude grids.
-\item The refined Arctic meshes substantially improve the surface mass balance over all low resolution grids.
+\item The refined Arctic meshes substantially improve the surface mass balance over all low-resolution grids.
 \end{keypoints}
@@
 Regional climate models (RCMs) are commonly relied upon to provide more accurate SMB estimates.
-The limited area domain used by RCMs permits the use of high-resolution grids capable of resolving SMB processes, and can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}.
+The limited area domain used by RCMs permits the use of high-resolution grids, capable of resolving SMB processes, that can skillfully simulate the GrIS SMB \cite{BETAL2004JGR,RETAL2012TC,VETAL2012TC,FETAL2013TC,METAL2017LTS,NETAL2018TC}.
 However, unlike GCMs, RCMs are not a freely evolving system, and the atmospheric state must be prescribed at the lateral boundaries of the model domain.
@@
 The simulated SMB compared favorably to the SMB from RCMs and observations.
-The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios.
+The VR approach is therefore emerging as a powerful tool for simulating and understanding the GrIS and its response to different forcing scenarios, in the freely evolving GCM framework.
 
@@
 
-In CESM2.2, the SE numerics have been enhanced to mitigate spurious noise over topography. These algorithmic changes are described in \ref{sec:a1}.
-
 The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over uniform-resolution setups.
@@
 The FV dycore is run with nominal $1^{\circ}$ and $2^{\circ}$ grid spacing, referred to as $\tt f09$ and $\tt f19$, respectively (Figure~\ref{fig:uni-grids}a,b).
-We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}d), where $ne$ refers to a grid with $ne \times ne$ quadrilaterial elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
+We also run the $1^{\circ}$ equivalent of the SE-CSLAM grid, referred to as $\tt ne30pg3$ (Figure~\ref{fig:uni-grids}d), where $ne$ refers to a grid with $ne \times ne$ elements per cubed-sphere face, and $pg$ denotes that there are $pg \times pg$  control volumes per element for computing the physics.
 We run an additional $1^{\circ}$ SE-CSLAM simulation with the physical parameterizations computed on a grid with $2\times2$ control volumes per element, $\tt ne30pg2$ \cite<Figure~\ref{fig:uni-grids}c;>[note CSLAM is still run on the $3\times3$ control volume grid]{HETAL2019JAMESb}.
@@
 
-\subsection{Simulated surface mass balance (SMB)}\label{sec:simSMB}
+\subsection{Simulated surface mass balance (SMB)}
 
@@
 The latter is the sum of evaporation, sublimation and liquid runoff, with runoff being a combination of liquid precipitation and snow and ice melt.
-Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, forming ice lenses.
+Not all liquid precipitation or snow/ice melt runs off the ice sheet; this water can penetrate pore spaces in the snowpack/firn layer and freeze, increasing the ice mass.
 These relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
@@
 Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent.
-Any snow above the 10 m cap contributes towards ice accumulation in the SMB, and refreezing of liquid water within the snowpack is an additional source of ice.
+Any snow above the 10 m cap contributes towards ice accumulation in the SMB.
+Refreezing of liquid water within the snowpack is an additional source of ice.
 Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
@@
 In the simulations described here, the GrIS is prescribed at its observed, modern extent and thickness.
+
+\subsection{SMB Analysis}\label{sec:SMB}
+
+We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
+However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx})b.
+CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
+The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
+The area errors in Figure~\ref{fig:grisdx}b may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
+
+\begin{figure}[t]
+\begin{center}
+         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
+\end{center}
+\caption{The spatial properties of the GrIS as represented by different grids in this study.
+(Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
+\label{fig:grisdx}
+\end{figure}
+
+We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $\tt f19$ and $\tt ne30pg2$ grids, and integrating over these low-resolution ice masks.
+The use of low-resolution common ice masks is a conservative decision, and is justified because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not generally reliable when mapping to a higher-resolution grid than the source grid.
+We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
+Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $\tt f19$ and $\tt ne30pg2$ masks.
+Thus, we evaluate an integrated quantity on a given grid up to four times to estimate the uncertainty due to differences in grid type and remapping algorithms.
+
+The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets.
+The SMB for total water can be expressed as:
+\begin{equation}
+SMB = accumulation + runoff + evaporation/sublimation, \label{eq:SMB}
+\end{equation}
+where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass).
+The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
+Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.
+Note that this SMB expression is different from the internally computed SMB described in the previous section.
+
+We consider two approaches for mapping and integrating the SMB components over the common ice masks:
+\begin{enumerate}
+\item Map the grid-cell mean quantities to the common grid, and integrate the mapped fields over the common ice masks. \label{lbl:label1}
+\item Map the patch-level quantities (i.e., the state over the ice fractional component of the grid cell) to the common grid, and integrate the mapped fields over the common ice masks.
+\label{lbl:label2}
+\end{enumerate}
+
+Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx})b.
+Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
+While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
+This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
+However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the grid-mean quantities used in method~\ref{lbl:label1}.
+
+The different error characteristics of the two methods are used to diversify the ensemble.
+Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
+Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
+This will degrade the SMB estimates using method~\ref{lbl:label2}, but we are more interested in characterizing the behavior of individual processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
+
+\subsection{Experimental design}
+
+All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
+In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2.
 
@@
  \end{table*}
-
-\subsection{SMB Analysis}\label{sec:SMB}
-
-We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
-However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}b).
-CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
-The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids.
-The area errors in Figure~\ref{fig:grisdx}b may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.
-
-\begin{figure}[t]
-\begin{center}
-         \includegraphics[width=100mm]{figs/temp_grisres.pdf}
-\end{center}
-\caption{The spatial properties of the GrIS as represented by different grids in this study.
-(Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
-\label{fig:grisdx}
-\end{figure}
-
-We have taken a common-ice-mask approach by mapping all model fields to the lowest-resolution grids, i.e., the $\tt f19$ and $\tt ne30pg2$ grids, and integrating over these low-resolution ice masks.
-The use of low-resolution common ice masks is a conservative decision, and is justified because we seek to use first-order remapping algorithms to map fields to the common ice mask, which is not generally reliable when mapping to a higher-resolution grid than the source grid.
-We use two remapping algorithms: ESMF first-order conservative and the TempestRemap \cite{TempestRemap} high-order monotone algorithm.
-Since mapping errors are sensitive to grid type, we evaluate all quantities on both common ice masks, the $\tt f19$ and $\tt ne30pg2$ masks.
-Thus, we evaluate an integrated quantity on a given grid up to four times to estimate the uncertainty due to differences in grid type and remapping algorithms.
-
-The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets.
-The SMB for total water can be expressed as:
-\begin{equation}
-SMB = accumulation + runoff + evaporation/sublimation, \label{eq:SMB}
-\end{equation}
-where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass).
-The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink.
-Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.
-Note that this SMB expression is different from the internally computed SMB described in section~\ref{sec:simSMB}.
-
-We consider two approaches for mapping and integrating the SMB components over the common ice masks:
-\begin{enumerate}
-\item Map the grid-cell mean quantities to the common grid, and integrate the mapped fields over the common ice masks. \label{lbl:label1}
-\item Map the patch-level quantities (i.e., the state over the ice fractional component of the grid cell) to the common grid, and integrate the mapped fields over the common ice masks.
-\label{lbl:label2}
-\end{enumerate}
-
-Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx}b).
-Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid.
-While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term.
-This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask.
-However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the grid-mean quantities used in method~\ref{lbl:label1}.
-
-The different error characteristics of the two methods are used to diversify the ensemble.
-Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
-Unfortunately, the patch-level values of evaporation/sublimation are not available from the model output, and we estimate their contribution by zeroing out the field for grid cells that have no ice, prior to mapping to the common ice mask.
-This will degrade the SMB estimates using method~\ref{lbl:label2}, but we are more interested in characterizing the behavior of individual processes across grids and dycores, expressed as the components of the SMB, rather than the SMB itself.
-
-\subsection{Experimental design}
-
-All simulations described here use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
-In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2.
 
@@
 Table~\ref{tbl:table3} shows the 1979-1998 climatological SMB components for each grid, compared with RACMO.
-The values in the table are averages over the ensemble of mapping methods to the common ice masks described in section~\ref{sec:SMB}, and the RACMO values refer to the average of both RACMO ensembles.
+The values in the table are averages over the ensemble of mapping methods to the common ice masks described in section~\ref{sec:SMB}, and the RACMO values are averages over both RACMO datasets (Table~\ref{tbl:table2}).
 Table~\ref{tbl:table3} also contains (in parenthesis) the SMB components derived from evaluating the integrals on each model's native grid and ice mask.
 Of note is the large reduction in melt rates compared to the values computed on the native grid, illustrating the dissipation of this quantity discussed in section~\ref{sec:SMB}.
-For integrated precipitation, the differences between the native and common ice mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
+For integrated precipitation, the differences between the native and common-ice-mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
 
@@
 The total SMB on coarse grids therefore has smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
-Such compensating errors underlie the importance of understanding the extent to which a model is getting the right SMB for the right reasons.
+Such compensating errors highlight the importance of understanding the extent to which a model is getting the right SMB for the right reasons.
 
@@
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and so a larger number of grid cells fit into the east-west transects.
-The VR grids more accurately reproduce the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.
+The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and a larger number of grid cells can line the east-west transects.
+The VR grids are more skillful at reproducing the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.
 
@@
 This negative cloud bias contrasts with a positive cloud bias over the ice sheet interior.
-This anomaly pattern is attributed to deficient orographic precipitation in the coarser model grids.
+This anomaly pattern has been attributed to deficient orographic precipitation in the coarser model grids.
 With overly smooth topography on the $1-2^{\circ}$ grids, synoptic systems moving into Greenland are not sufficiently lifted when encountering the steep ice margins.
@@
 This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities.
-We therefore provide the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a more realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.
+We therefore provide the the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a more realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.
 
@@
 \appendix
-\section{Details on spectra-element dynamical core improvements since the CESM2.0 release}\label{sec:a1}
+\section{Details on spectra-element dynamical core improvements since the CESM2.0 release}
 Since the CESM2.0 release of the spectral-element dynamical core documented in \citeA{LetAl2018JAMES} some important algorithmic improvements have been implemented and released with CESM2.2.

### 2022-09-10 first_review/response-to-reviewers.tex
updates

@@
 
+{\color{blue}{We thank the reviewer for this point.
+We had not thought too much about albedo and so it's for the best that this was brought up.
+This is an interesting hypothesis, that in the coarser grids, the negative precip biases along the ice sheet margins may lead to more frequent bare ice exposure, and through it's lesser albedo, more melt.
+We have explored whether there is any evidence for this in our simulations, and do find that there is a negative albedo anomaly around the southeast, southwest and northwest coasts in the coarse grids.
+However, this is only with respect to the RACMO simulations.}}
+
 \subsection*{Specific comments}
@@
 
-{\color{blue}{We agree that this is confusing and have added a sentence describing the different grids/dycores in more detail, before talking about their SMB performance.}}.
+{\color{blue}{We agree that this is confusing and have added a sentence describing the different grids/dycores in more detail, before talking about their SMB performance.}}. \newline
 
@@
 
-{\color{blue}{See above.}}.
+{\color{blue}{See above.}} \newline
 
@@
 Decided against getting into the specifics of which dycore uses which grid, as this becomes apparent later on in the introduction}}.
+\newline
 
@@
 
+{\color{blue}{Sentence has been reworded.}} \newline
+
 5. Lines 115-116: Clarify that this is a positive precipitation bias in the interior.  \newline
+
+{\color{blue}{Done.}} \newline
 
@@
 I suggest mentioning them briefly here as this is a way to increase SMB resolution (from the standpoint of runoff but not precipitation) that is used in CESM  \newline
-.
+
+{\color{blue}{We agree that the idea of ECs should be introduced here, and have added a sentence.}} \newline
+
 7. Figure 1, caption: Please provide a brief description of each of the grids here for the benefit of the reader. E.g. (a) 2 latitude-longitude grid (f19), etc.  \newline
+
+{\color{blue}{Done.}} \newline
 
@@
 
+{\color{blue}{Done.}} \newline
+
 9. Table 1: It would be helpful to specify which grids are latitude-longitude and which are unstructured either as a column in the table or text in the caption.  \newline
+
+{\color{blue}{Added to the text which grids are lat-lon and which are unstructured.}} \newline
 
@@
 
+{\color{blue}{The text indicates that ne30 indicates are 30x30 elements per cube-face, and pg3 refers to 3x3 control volumes per element.
+The global number of grid columns is then 6 faces * 30x30 elements per face * 3x3 control volumes per element.}} \newline
+
 11. Lines 239-243: I suppose degree coordinates here correspond the latitude dimension, or degrees at the equator. Please clarify.  \newline
+
+{\color{blue}{Done.}} \newline
 
@@
 
+{\color{blue}{Fixed.}} \newline
+
 13. Line 254: Clarify that “expensive” means “computationally expensive.” \newline
+
+{\color{blue}{This information has been absorbed into a new sub-section on computational costs.
+References to ``cheaper" or ``more expensive" are more obviously referring computational costs as that is the title of the section.}} \newline
 
@@
 
+{\color{blue}{6.5K/km. I'd refer to reviewer to the Sellevold et al paper referenced, for more details on the choice of lapse rate.}} \newline
+
 15. Line 314: I’m confused as to why accumulation below 10 m does not contribute to SMB. How is this addition of mass accounted for? \newline
 
+{\color{blue}{The internally computed SMB is a bit strange, in that only tracks ice.
+It does not consider the overlying 10m of snow to contribute to the SMB.
+In response to the other reviewer, we've clarified this in sections 2.5.}} \newline
+
 16. Lines 318-319: What is the period for the prescribed SSTs? \newline
+
+{\color{blue}{Perpetual 1979. Added this in parantheses.}} \newline
 
@@
 
+{\color{blue}{Fixed.}} \newline
+
 18.
@@
 
+{\color{blue}{Changed to validation dataset throughout the manuscript.}} \newline
+
 19. Figure 3 caption: I believe the left and right figures are switched here. Also, the fact that RACMO RCM simulations are included should be mentioned. \newline
+
+{\color{blue}{Fixed, and included references to the RACMO grids in the text and figure caption.}} \newline
 
@@
 
-21.
-Line 406: I believe this should read “expressed as temperature differences”.
+{\color{blue}{Switched to type writer font. Used reviewers suggestion to replace slash with a plus sign.}} \newline
+
+21. Line 406: I believe this should read “expressed as temperature differences”.
+
+{\color{blue}{Did something similar, as temperature was missing from the sentence.}} \newline
+
 22.
@@
 
+{\color{blue}{The imperfect alignment of the clubb heating anomalies and the temperature anomalies has a lot to do with averaging the clubb heating over all time.
+If you look at Figure 10 in Herrington and Reed QJRMS paper, the alignment is strongest when you sample the clubb heating when/where the dycore has upward vertical velocities.
+Whereas there is no correspondence with the temperature anomalies for descending grid columns.
+So since our clubb heating is just a climatology, we lose that alignment with the temperature anomalies.
+But it's clear that a strong heating signal is piercing through in the climatology anyway, and so I'm strongly implying that this is consistent with QJRMS mechanism.
+I've added an entire paragraph discussing this misalignment.}} \newline
+
 23. Lines 437-438: This also seems to be true in other cases. \newline
+
+{\color{blue}{See above comment.}} \newline
 
@@
 
+{\color{blue}{This response appears to be different from the other cases; with clear changes to the planetary waves that are not seen in the other simulations.}} \newline
+
 25. Line 459: Perhaps the authors could mention the potential influence of clouds earlier as prior to this point, they are not mentioned. \newline
+
+{\color{blue}{See above comment.}} \newline
 
@@
 
+{\color{blue}{Decided to leave it as.}} \newline
+
 27. Figure 8: The CALIPSO grid appears to be coarser resolution than the CERES grid. Is this a 2 grid? \newline
 
+{\color{blue}{Yes, CALIPSO is on a 2 degree grid. Corrected Table 2. Great catch!}} \newline
+
 28. Lines 573-575: Wouldn’t the elevation class scheme fix this problem by raising the ice-covered sub-grid portion to a higher elevation? \newline
+
+{\color{blue}{That's a good point that we overlooked.
+We have added a sentence to the end of the next paragraph clarifying that differences between the grid cell mean and actual ice surface are not necessarily problematic due to the EC scheme.}} \newline
 
@@
 
+{\color{blue}{We've added references to this coincidence in response to your related questions in the general remarks.}} \newline
+
 30. Lines 675-680: Perhaps the fact that clouds and precipitation move more easily inland could reduce snowfall and increase shortwave radiation along the coast, enhancing melt. \newline
+
+{\color{blue}{See response to general remarks.}} \newline
 
@@
 
+{\color{blue}{See response to general remarks.}} \newline
+
 32. Line 719: Change “spectra-element” to “spectral-element”. \newline
+
+{\color{blue}{Done.}} \newline
 
@@
 
+{\color{blue}{Done.}} \newline
+
 34. Line 731: Add “is the” before “surface geopotential”. \newline
+
+{\color{blue}{Done.}} \newline
 
@@
 
-36. Lines 740-748: Here some terms should be defined, including Tv, qv and p0.
+{\color{blue}{Defined in equation A1.}} \newline
+
+36. Lines 740-748: Here some terms should be defined, including Tv, qv and p0. \newline
+
+{\color{blue}{Added definitions for Tv and thetav. p0 is defined just below equation A3. I don't see a ``qv".}}
 

### 2022-09-10 second_submit/arcticgrids.tex
updates

@@
 Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS).
-The conventional $1^{\circ}-2^{\circ}$ resolution grids systematically overestimate both accumulation and ablation over the GrIS.
-Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS.
-Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, are documented as newly supported configurations in CESM2.2.
+{\color{blue}{The finite-volume dynamical core uses structured, latitude-longitude grids, whereas the spectral-element dynamical core is built on unstructured meshes, permitting grid flexibility such as quasi-uniform grid spacing globally.}} The $1^{\circ}-2^{\circ}$ {\color{blue}{latitude-longitude and quasi-uniform unstructured}} grids systematically overestimate both accumulation and ablation over the GrIS.
+Of these $1^{\circ}-2^{\circ}$ grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS.
+Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, were developed for the spectral-element dynamical core and are documented here as newly supported configurations in CESM2.2.
 The Arctic meshes substantially improve the simulated clouds and precipitation rates in the Arctic.
@@
 Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain.
-With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic)}}.
-In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate.
+With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic due to the polar singularity)}}.
+In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids {\color{blue}{(i.e., grids with enhanced resolution over a particular region)}}, to understand their impacts on the simulated Arctic climate.
 We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.
@@
 GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation.
-The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias \cite{P2000GPC,VETAL2018TC}.
+The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias {\color{blue}{in the interior}} \cite{P2000GPC,VETAL2018TC}.
 GrIS ablation areas {\color{blue}{(marginal regions where the annual SMB is negative)}} are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
 These narrow ablation zones are not fully resolved in low-resolution GCMs, and may further degrade the simulated SMB.
-For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.
+{\color{blue}{More recently, GCMs such as CESM and the NASA Goddard Institute for Space Studies GCM \cite{AETAL2019JGR}, have implemented an elevation class downscaling scheme for computing the SMB.
+The downscaling helps to resolve to these narrow ablation zones in GCMs \cite{SETAL2019TC}, but large SMB biases remain.}} For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.
 
@@
 \begin{center}
-%\begin{tabular}{cc}
          \includegraphics[width=100mm]{grid-uniform}
-         %\includegraphics[width=50mm]{grid-f19.pdf}&
-         %\includegraphics[width=50mm]{grid-f09.pdf}\\
-         %\includegraphics[width=50mm]{grid-ne30pg2.pdf}&
-         %\includegraphics[width=50mm]{grid-ne30pg3.pdf} \\
-%\end{tabular}
-\end{center}
-\caption{Computational grids for the $1^{\circ}-2^{\circ}$ {\color{blue}{lat-lon and quasi-uniform unstructured}} grids in this study.
-Grids names after Table~\ref{tbl:table1}, (a) $\tt f19$, (b) $\tt f09$, (c) $\tt ne30pg2$ and (d) $\tt ne30pg3$.}
+\end{center}
+\caption{Computational grids for the $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids in this study.
+Grids names after Table~\ref{tbl:table1}, (a) $2^{\circ}$ lat-lon grid ($\tt f19$), (b) $1^{\circ}$ lat-lon grid ($\tt f09$), (c) $1^{\circ}$ quasi-uniform unstructured grid with reduced physics resolution ($\tt ne30pg2$) and (d) $1^{\circ}$ quasi-uniform unstructured grid ($\tt ne30pg3$).}
 \label{fig:uni-grids}
@@
 \begin{center}
-%\begin{tabular}{ccc}
          \includegraphics[width=140mm]{grid-varres.pdf}
-         %\includegraphics[width=40mm]{grid-ARCTIC.pdf}&
-         %\includegraphics[width=40mm]{grid-ARCTICGRIS.pdf}&
-         %\includegraphics[width=40mm]{grid-CONUS.pdf} \\
-%\end{tabular}
 \end{center}
@@
 The SE dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}.
-The computational domain is a cubed-sphere grid tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
+The computational domain is a cubed-sphere grid ({\color{blue}{A cubed-sphere grid is obtained by projecting each face of a cube onto a sphere}}) tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
 Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre quadrature points.
@@
 The VR meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
-The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region.
+The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region ({\color{blue}{when using degrees to describe the resolution of unstructured grids, degrees refers to degrees at the equator}}).
 The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution.
@@
 FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.
-We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.
+{\color{blue}{The FV dycore uses lat-lon grids, whereas the SE and SE-CSLAM dycores run on unstructured grids.}} We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.
 {\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY).
@@
 
-Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation for about 500 years.
+Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation ({\color{blue}{with perpetual 1979 boundary conditions}}) for about 500 years.
 The $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids are initialized with the SMB from an existing $\tt f09$ spun-up initial condition.
@@
    CERES-EBAF ED4.1 & 2003-2020 & $1^{\circ}$ & \citeA{CERES-EBAF} \\
-   CALIPSO-GOCCP & 2006-2017 & $1^{\circ}$ & \citeA{CALIPSO-GOCCP} \\
+   CALIPSO-GOCCP & 2006-2017 & {\color{blue}{$2^{\circ}$}} & \citeA{CALIPSO-GOCCP} \\
    RACMO2.3 & 1979-1998 & 11 km & \citeA{NETAL2015TC} \\
@@
 \caption{The spatial properties of the GrIS as represented by different grids in this study.
-(Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
+(Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from the reference dataset described in the text.
+{\color{blue}{Also included are the two RACMO grids used for validation (see Table~\ref{tbl:table2})}}.}
 \label{fig:grisdx}
@@
 \begin{equation}
-SMB = accumulation + runoff + evaporation + sublimation, \label{eq:SMB}
+\tt SMB = accumulation + runoff + evaporation + sublimation, \label{eq:SMB}
 \end{equation}
@@
 Before delving into the simulated Arctic climate conditions, we assess the global mean differences between the various grids and dycores.
-Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured}} grids and dycores.
+Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots {\color{blue}{of temperature}}, expressed as differences between {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured}} grids and dycores.
 The $\tt f09$ grid is warmer than the $\tt f19$ grid, primarily in the mid-to-high latitudes throughout the depth of the troposphere.
@@
 With the same physics time step, the greater condensational heating and warmer temperatures are confined to the refined Arctic region.
+
+{\color{blue}{We note that there is not perfect alignment of the CLUBB heating anomalies and the temperature anomalies in Figures~\ref{fig:dT-lores} $\&$ \ref{fig:dT-hires}.
+This leaves some uncertainty as to whether CLUBB is solely responsible for the temperature anomalies caused by changing grid resolution or dycore, as asserted above.
+In the deep Tropics, gravity waves are expected to propagate temperature anomalies far from their heating source, and so the misalignment is less troublesome for this region.
+In the mid-latitudes, we speculate that this misalignment is due to averaging the CLUBB heating over all time.
+\citeA{HK2020QJRMS} have shown that only the CLUBB heating coinciding with upward resolved vertical velocities tends to align spatially with the temperature response, whereas heating anomalies associated with descending grid columns had no alignment with the temperature response.
+Therefore the average over all time will dilute the heating signal, and potentially lead to misalignments between the heating rates and temperature anomalies.}}
 
@@
 This suggests that coarser models will tend to elevate the ablation zones {\color{blue}{relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and}} depressed melt rates, but this is opposite the melt bias that occurs in the coarse grid simulations.
+{\color{blue}{Also, the EC downscaling should be able to correct for situations where model ice surface is elevated relative to the true ice surface, although for very large elevation differences/errors the EC scheme may not be adequate}}.
 
@@
 \appendix
-\section{Details on spectra-element dynamical core improvements since the CESM2.0 release}\label{sec:a1}
+\section{Details on {\color{blue}{spectral}}-element dynamical core improvements since the CESM2.0 release}\label{sec:a1}
 Since the CESM2.0 release of the spectral-element dynamical core documented in \citeA{LetAl2018JAMES} some important algorithmic improvements have been implemented and released with CESM2.2.
@@
 \end{eqnarray}
- \cite{SJ1991QJRMS} where $g$ gravity, $T_1=\Gamma_0 T_{ref} c_p^{(d)}/g\approx 192K$ with standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv T_{ref}-T_1\approx 97K$; $T_{ref}=288K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ surface geopotential.
+ \cite{SJ1991QJRMS} where $T_1=\Gamma_0 T_{ref} c_p^{(d)}/g\approx 192K$, {\color{blue}{with gravity $g$,}} and standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv T_{ref}-T_1\approx 97K$; $T_{ref}=288K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ {\color{blue}{is the}} surface geopotential.
 The reference Exner function is
@@
 \end{equation}
+{\color{blue}{and virtual temperature is
+\begin{equation}
+T_v =T\left( \frac{1+\frac{1}{\epsilon}m^{(wv)}}{\sum_{\ell \in \mathcal{L}_{all}}m^{(\ell)}}\right),\label{eq:tv}
+\end{equation}
+where $m^{(\ell)}$ is dry mixing ratio of component of moist air $\ell$; $\mathcal{L}_{all}$ is the set of all components of moist air and, in particular, $'wv'$ is water vapor.}}
+
 The derivation showing that (\ref{eq:pgf}) and (\ref{eq:pgf2}) are equivalent is {\color{blue}{given}} here:
@@
 \end{eqnarray*}
-Using the reference states from \cite{SJ1991QJRMS},
+{\color{blue}{where virtual potential temperature is $T_v/\Pi$.}} Using the reference states from \cite{SJ1991QJRMS},
 \begin{eqnarray}

### 2022-09-10 second_submit/arcticgrids.tex
updates

@@
 
-{\color{blue}{Another potential factor in explaining the differences across the various grids and dycores is the surface albedo.
+{\color{blue}{Another potential factor in explaining the melt differences across the grids and dycores is the surface albedo.
 Bare ice has a lower albedo than snow, so it's reasonable to speculate whether the large melt biases in the coarse grid simulations are in part due to more frequent bare ice exposure.
-One logical pathway to more bare ice exposure is related to the precipitation and cloud biases.
-As the coarser models accumulate too much precipitation in the GrIS interior, the coastal ice margins receive too little snow, which can be more easily melted away to expose the underlying glacial ice.
-Figure~\ref{fig:alb} shows maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and CERES-EBAF datasets.
-The surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface.
-The lat-lon and quasi-uniform grids all have anomalously low albedo around the coasts in the southeast and northwest regions of the GrIS, which coincides with regions of anomalously low cloud cover and precipitation.
-In the VR simulations, the low albedo biases are alleviated to a large extent.
-
-In light of these albedo results, one may interpret Figure~\ref{fig:bias} differently.
-That there are larger melt biases for grid cells with only only partial ice cover, may have less to do with the partial ice cover and more to do with the enhanced bare ice exposure in the coarser grids.
-Further experimentation is required to tease out the relative contributions of these two processes.}}
+One reasonable pathway to more bare ice exposure is related to the precipitation and cloud biases.
+As the lat-lon and quasi-uniform grids accumulate too much precipitation in the GrIS interior, it's possible the coastal ice margins receive too little snow, which may be more easily melted away to expose the underlying glacial ice.
+The top row of Figure~\ref{fig:alb} shows the summer snow cover bias relative to ERA5.
+There is no indication that the coastal regions have lower snow fractions, and more bare ice exposure, compared with the VR runs.
+In southeast Greenland, the coarse runs have anomalously high snow cover relative to the VR runs.
+In southwest and central-west Greenland, the snow cover is anomalously low, but arguably no more so than occurs in the VR runs.
+And in combining these results wih the precipitation biases (Figure~\ref{fig:bias}), there just isn't much support for a reduction in snowfall and snow cover around the coastal margins in the coarser runs.
+
+The middle and bottom rows of Figure~\ref{fig:alb} show maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and the processed MODIS albedo dataset of \citeA{MODIS}.
+The model surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface.
+While RACMO2.3p2 indicates that the albedo near the southwest/southeast coastal margins is lower in the lat-lon and quasi-uniform grids relative to the VR grids, the MODIS dataset does not show that the VR grids have a substantially different albedo from the coarse grids.
+We do observe a persistent positive albedo bias in the northwest in all grids, in both validation dataset, which is consistent with a low melt bias in that region \cite{LETAL2020JAMES}.}}
 
@@
 \end{center}
-\caption{1979-1998 northern hemisphere summer albedo bias (fraction), computed as the difference (top) from RACMO2.3p2, and (bottom) CERES-EBAF dataset.
-RACMO2.3p2 differences are found by mapping the model output to the individual model grids, and differences in the bottom panel are computed after mapping the individual model grids to the CERES-EBAF $1^{\circ}$ grid.
-Note that the averaging period for CERES-EBAF panels are 2003-2020, which is different from the averaging period for the model results.}
+\caption{1979-1998 northern hemisphere summer albedo bias (fraction), computed as the difference (top) from RACMO2.3p2, and (middle) the MODIS dataset of \citeA{MODIS}.
+The bottom row is 1979-1998 bias in summer snow cover (fraction), relative to ERA5.
+As all observational datasets in this plot are on high-resolution grids, they've all been mapped to the individual model grids, before taking differences.}
 \label{fig:alb}
@@
 The warmer ice-free patches may largely determine the mean state, leading to a warm bias over the ice-covered patches of the grid cell.
-This mechanism is supported by analysis of melt biases binned by grid-cell ice fraction.
+This mechanism is supported by analysis of melt biases binned by grid-cell ice fraction, {\color{blue}{although further testing of this hypothesis is needed.}}
 

### 2022-09-11 first_review/response-to-reviewers.tex
updates

@@
 We had not thought too much about albedo and so it's for the best that this was brought up.
-This is an interesting hypothesis, that in the coarser grids, the negative precip biases along the ice sheet margins may lead to more frequent bare ice exposure, and through it's lesser albedo, more melt.
-We have explored whether there is any evidence for this in our simulations, and do find that there is a negative albedo anomaly around the southeast, southwest and northwest coasts in the coarse grids.
-However, this is only with respect to the RACMO simulations.}}
+We have made a good faith effort to explore this mechanism in the simulations, which has resulted in an additional figure (Figure 14) and two additional paragraphs at the end of section 2.3.
+In general we don't find a great deal of evidence for the reviewers proposed mechanism.
+We believe this is largely because the positive interior precipitation biases are not offset by a negative precipitation bias at the coastal margins.
+We then took a closer look at albedo, where one validation dataset indicated the coarse grids had anomalously low albedo around the southern coasts, the other validation dataset indicated the low albedo anomaly occurs in the VR grids as well.}}
 
@@
 
-{\color{blue}{We've added references to this coincidence in response to your related questions in the general remarks.}} \newline
+{\color{blue}{We have added a caveat to the conclusions, that our hypothesis needs more testing.}} \newline
 

### 2022-09-13 second_submit/arcticgrids.tex
update

--- 2022-09-10 second_submit/arcticgrids.tex
@@
 
-The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales that are difficult to represent in GCMs \cite{P2010CC}.
+The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales {\color{blue}\cite{Fyke_EA_2018_Rev_Geophys}} that are difficult to represent in GCMs \cite{P2010CC}.
 GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation.
@@
 CESM2.2 contains sub-component models for the atmosphere, land, ocean, sea-ice, and land-ice, and can be configured to run with varying degrees of complexity.
-All simulations described in this study use an identical transient 1979-1998 Atmospheric Model Inter-comparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
+All simulations described in this study use an identical transient 1979-1998 Atmospheric Model Intercomparison Project (AMIP) configuration, with prescribed monthly sea-surface temperature and sea ice following \citeA{CESMSST}.
 In CESM terminology, AMIP simulations use the $\tt FHIST$ computational set and run out of the box in CESM2.2.
 The land-ice component is not dynamically active in $\tt FHIST$.
-However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the functionality to simulate the surface mass balance of the Greenland Ice Sheet.}}
+However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the functionality to simulate the SMB of the Greenland Ice Sheet.}}
 
@@
 The SE dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}.
-The computational domain is a cubed-sphere grid ({\color{blue}{A cubed-sphere grid is obtained by projecting each face of a cube onto a sphere}}) tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
+The computational domain is a cubed-sphere grid ({\color{blue}{obtained by projecting each face of a cube onto a sphere}}) tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
 Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre quadrature points.
@@
 
-{\color{blue}{SE is a next generation of dycore, and is less mature than the FV dycore due to its shorter history.}} In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to CESM2.0}} to mitigate spurious noise over topography.
+{\color{blue}{SE is a next-generation dycore, and is less mature than the FV dycore due to its shorter history.}} In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to CESM2.0}} to mitigate spurious noise over topography.
 These algorithmic changes are described in \ref{sec:a1}.
-{\color{blue}{Future versions of CESM will likely continue to include further optimizations and enhancements to the (nascent) SE dycore.}}
+{\color{blue}{Future versions of CESM will likely include further optimizations and enhancements to the (nascent) SE dycore.}}
 
@@
 The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the original SE dycore in which the physics are evaluated at Gauss-Lobatto-Legendre points \cite{HL2018MWR}.
-{\color{blue}{CSLAM advection is not an available option in the VR configuration, which instead uses the standard SE tracer transport scheme with the physics evaluated at Gauss-Lobatto-Legendre points.}}
+{\color{blue}{CSLAM advection is not available in the VR configuration, which instead uses the standard SE tracer transport scheme with the physics evaluated at Gauss-Lobatto-Legendre points.}}
 
@@
 The VR meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
-The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region ({\color{blue}{when using degrees to describe the resolution of unstructured grids, degrees refers to degrees at the equator}}).
+The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region ({\color{blue}{when using degrees to describe the resolution of unstructured grids, we refer to degrees at the equator}}).
 The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution.
@@
 {\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY).
-The costs are from single month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
+The costs are from single-month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
  \label{tbl:table1}
@@
 The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution.
-Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids, {\color{blue}{as well as two grids pertaining to the Regional Atmospheric Climate Model \cite<RACMO;>{NETAL2018TC,NETAL2019SCIENCE}, which are used for validation purposes in this study (Table~\ref{tbl:table2})}}.
+Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids, {\color{blue}{as well as two grids pertaining to the Regional Atmospheric Climate Model \cite<RACMO;>{NETAL2018TC,NETAL2019SCIENCE}, which are used for validation (Table~\ref{tbl:table2})}}.
 The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid.
@@
 
-{\color{blue}{The last columns of Table~\ref{tbl:table1} provides cost estimates for the different grids and dycores.
-The costs, expressed as core hours per simulated year (CHPSY), are taken from single month runs of $\tt FHIST$, with no i/o, and using 25 nodes (900 tasks) and 50 nodes (1800 tasks) on the Cheyenne supercomputer \cite{cheyenne}.
-25 nodes is on the low side for a typical multi-decadal climate simulation at $1^{\circ}$ resolution, but it's the largest number of tasks that can be supported by the $\tt f19$ grid, and we chose to fix the number of tasks across all grids for the purposes of comparing their costs.
-We also provide costs using 50 nodes, excluding $\tt f19$, to provide a more practical cost estimate for longer climate integrations.
-There are probably better approaches for comparing costs across different grids and dycores, e.g., holding fixed the number of grid columns per task, but it is beyond the scope of this study.
+{\color{blue}{The last columns of Table~\ref{tbl:table1} provide cost estimates for the different grids and dycores.
+The costs, expressed as core hours per simulated year (CHPSY), are taken from single-month runs of $\tt FHIST$, with no i/o, and using 25 nodes (900 tasks) and 50 nodes (1800 tasks) on the Cheyenne supercomputer \cite{cheyenne}.
+For a typical multi-decadal climate simulation at $1^{\circ}$ resolution, 25 nodes are on the low side.
+However, this is the largest number of tasks supported by the $\tt f19$ grid, and we fixed the number of tasks across all grids for the purpose of comparing costs.
+We also show costs for 50 nodes, excluding $\tt f19$, to provide a more practical estimate for longer climate integrations.
+Other approaches for comparing costs across different grids and dycores, e.g., holding fixed the number of grid columns per task, are beyond the scope of this study.
 
@@
 
-The cheapest grid is the $\tt f19$ grid at 436.66 CHPSY, as this is the only grid running with $2^{\circ}$ dynamics.
+The cheapest grid is the $\tt f19$ grid at 436.66 CHPSY, as this is the only grid with $2^{\circ}$ dynamics.
 The $\tt f09$ grid costs 1534.57 CHPSY using 25 nodes, which is noticeably cheaper than $\tt ne30pg3$ at 1890.48 CHPSY.
@@
 The FV model is known to be cheaper than SE at small core counts, whereas SE is more efficient than FV at large core counts due to its improved scalability \cite{DFSSTT2005IJHPC,DetAl2012IJHPCA}.
-In the more conventional 50 node runs, $\tt f09$ costs are much more similar to $\tt ne30pg3$, due to a 30$\%$ cost increase in $\tt f09$ relative to the 25 node run (Table~\ref{tbl:table1}).
-The $\tt Arctic$ grid is an order of magnitude more expensive than the lat-lon and quasi-uniform grids, at about 16k CHPSY, whereas the $\tt Arctic-GrIS$ grid is a more than twice that (40k CHPSY).
-Note that all timing numbers are from runs without threading.
-The $\tt f09$ grid is the only grid that runs out-of-the-box with threading; holding the number of tasks fixed leads to a $4\%-6.5\%$ reduction in CHPSY compared to runs without threading.}}
+In the more conventional 50-node runs, $\tt f09$ costs are more similar to $\tt ne30pg3$, due to a 30$\%$ cost increase in $\tt f09$ relative to the 25-node run (Table~\ref{tbl:table1}).
+The $\tt Arctic$ grid is an order of magnitude more expensive than the lat-lon and quasi-uniform grids, at about 16k CHPSY, whereas the $\tt Arctic-GrIS$ grid adds another factor of $\sim2$ (40k CHPSY).
+All timing numbers are from runs without threading.
+The $\tt f09$ grid is the only grid that runs out-of-the-box with threading; holding the number of tasks fixed reduces CHPSY by $4\%-6.5\%$ reduction compared to runs without threading.}}
 
@@
 {\color{blue}{CESM simulates the GrIS SMB as the sum of ice accumulation and ice ablation.
-The latter contains contributions from sublimation and liquid runoff from ice melt.
-Liquid precipitation and liquid runoff may also contribute to ice accumulation by penetrating pore spaces in the snowpack/firn layer and freeze, forming ice lenses}}.
-These relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
+The latter includes sublimation, as well as liquid runoff from ice melt.
+Liquid precipitation and runoff may also contribute to ice accumulation by penetrating pore spaces in the snowpack/firn layer and freezing, forming ice lenses}}.
+The relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
 
@@
 The area fraction of each EC is computed from {\color{blue}{the CISM initial conditions, which are based on a high-resolution dataset of the observed, modern extent and thickness of the GrIS \cite{METAL2014NG}.
-Note that for configurations with a dynamically active ice sheet, the area fractions are continuously updated throughout the run to reflect the evolving ice sheet geometry in CISM}}.
+For configurations with a dynamically active ice sheet, the area fractions are continuously updated throughout the run to reflect the evolving ice sheet geometry in CISM}}.
 The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean.
@@
 Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent.
-Any snow above the 10 m cap contributes towards ice accumulation, and refreezing of liquid water within the snowpack is an additional source of ice.
+Any snow above the 10 m cap contributes to ice accumulation, and refreezing of liquid water within the snowpack is an additional source of ice.
 Surface ice melting (after melting of the overlying snow) yields a negative SMB.}} Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
@@
 We use several {\color{blue}{validation}} datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations.
-{\color{blue}{The ERA5 reanalysis product \cite{ERA5} is used for validating the large-scale circulation and extreme precipitation events.
-Clouds and radiation fields are validated using remote sensing products, the CERES-EBAF ED4.1 \cite{CERES-EBAF} and the CALIPSO-GOCCP \cite{CALIPSO-GOCCP} datasets, respectively.}}
+{\color{blue}{The ERA5 reanalysis product \cite{ERA5} is used to validate the large-scale circulation and extreme precipitation events.
+Clouds and radiation fields are validated using remote sensing products, the CERES-EBAF ED4.1 \cite{CERES-EBAF} and the CALIPSO-GOCCP \cite{CALIPSO-GOCCP} datasets, respectively.
+We also use a MODIS GrIS surface albedo product, mapped to a 25 km regional climate model grid \cite{MODIS}.}}
 
@@
    CALIPSO-GOCCP & 2006-2017 & {\color{blue}{$2^{\circ}$}} & \citeA{CALIPSO-GOCCP} \\
+   {\color{blue}{MODIS-MAR}} & {\color{blue}{2000-2013}} & {\color{blue}{25 km}} & {\color{blue}{\citeA{MODIS}}}\\
    RACMO2.3 & 1979-1998 & 11 km & \citeA{NETAL2015TC} \\
@@
 We do not use CLM's internally computed SMB in this study.
-Rather, we utilize the components of the internally computed SMB to construct the total water SMB}}.
+Rather, we use the components of the internally computed SMB to construct the total water SMB}}.
 
@@
 
-{\color{blue}{We note that there is not perfect alignment of the CLUBB heating anomalies and the temperature anomalies in Figures~\ref{fig:dT-lores} $\&$ \ref{fig:dT-hires}.
+{\color{blue}{There is not perfect alignment of the CLUBB heating anomalies and the temperature anomalies in Figures~\ref{fig:dT-lores} and \ref{fig:dT-hires}.
 This leaves some uncertainty as to whether CLUBB is solely responsible for the temperature anomalies caused by changing grid resolution or dycore, as asserted above.
-In the deep Tropics, gravity waves are expected to propagate temperature anomalies far from their heating source, and so the misalignment is less troublesome for this region.
-In the mid-latitudes, we speculate that this misalignment is due to averaging the CLUBB heating over all time.
-\citeA{HK2020QJRMS} have shown that only the CLUBB heating coinciding with upward resolved vertical velocities tends to align spatially with the temperature response, whereas heating anomalies associated with descending grid columns had no alignment with the temperature response.
-Therefore the average over all time will dilute the heating signal, and potentially lead to misalignments between the heating rates and temperature anomalies.}}
+In the deep tropics, gravity waves are expected to propagate temperature anomalies far from their heating source, and so some misalignment is expected for this region.
+In the mid-latitudes, we speculate that the misalignment is due to averaging the CLUBB heating over all time.
+\citeA{HK2020QJRMS} have shown that only the CLUBB heating coinciding with upward resolved vertical velocities tends to align spatially with the temperature response, whereas heating anomalies associated with descending grid columns have no alignment with the temperature response.
+Therefore the average over all time will dilute the heating signal, potentially leading to misalignments.}}
 
@@
 {\color{blue}{Keeping our focus on the Arctic region, and in particular Greenland,}} it is useful to understand summer temperature biases due to their control on ice and snow melt over the GrIS \cite<>{O2001JAM}.
-Figure~\ref{fig:dThyps} shows the 1979--1998 lower troposphere summer temperature bias relative to ERA5, computed by equating a layer mean virtual temperature with the 500--1000 hPa geopotential thickness.
+Figure~\ref{fig:dThyps} shows the 1979--1998 lower-troposphere summer temperature bias relative to ERA5, computed by equating a layer mean virtual temperature with the 500--1000 hPa geopotential thickness.
 The results are consistent with the zonal mean height plots; increasing resolution from $\tt f19$ to $\tt f09$ warms the climate, and the $1^{\circ}$ SE grids are warmer than the FV grids.
@@
 All grids show a north-south gradient in bias over Greenland, with the summer temperature bias more positive for the northern part of the ice sheet.
-This pattern is also evident in the near surface temperature bias over Greenland (not shown).
+This pattern is also evident in the near-surface temperature bias over Greenland (not shown).
 
@@
 On the VR grids, both the interior shortwave deficit and the excessive shortwave around the ocean perimeter are improved.
-This suggests that the coarse grid clouds are too thick in the interior of Greenland and too thin around the perimeter, which is consistent with the total summer cloud fraction bias, computed from the CALIPSO cloud dataset and shown in the bottom panel of Figure~\ref{fig:FSDS}.
+This suggests that the coarse-grid clouds are too thick in the interior of Greenland and too thin around the perimeter, which is consistent with the total summer cloud fraction bias, computed from the CALIPSO cloud dataset and shown in the bottom panel of Figure~\ref{fig:FSDS}.
 Note that total cloud fraction characterizes the cloud field at all vertical levels, and attenuates the changes arising from any single layer due to the occurrence of overlapping clouds at other levels.
@@
 The total SMB on coarse grids therefore has smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
-Such compensating errors {\color{blue}{underline}} the importance of understanding the extent to which a model is getting the right SMB for the right reasons.
+Such compensating errors {\color{blue}{underscore}} the importance of understanding whether a model is getting the right SMB for the right reasons.
 
@@
 The uncertainty due to differences in basin area is larger than for GrIS-wide integrals, owing to the differences in basin boundaries on the common ice masks, which are shown in the $\tt f19$ and $\tt ne30pg2$ panels of Figure~\ref{fig:prect}.
-Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show the southeast and southwest basins have the most accumulation.
+Nonetheless, the regional totals in Figure~\ref{fig:basin} correctly show that the southeast and southwest basins have the most accumulation.
 In all basins, accumulation decreases monotonically with increasing grid resolution, though with some exceptions.
@@
 
-Lower troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones.
+Lower-troposphere temperature is not a strict proxy for melting; e.g., it may not capture microclimate effects as a result of a better representation of the low-elevation ablation zones.
 The Positive Degree-Day temperature-based melt index \cite<PDD;>{B1984Z}, which accumulates the near-surface temperature in $^{\circ}C$ for days with temperature above freezing, is a more accurate proxy for melting.
@@
 The $1^{\circ}-2^{\circ}$ grids are noticeably coarse, with only a handful of grid cells populating the transect.
-The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and so a larger number of grid cells fit into the east-west transects.
+The $\tt f09$ grid is a bit of an exception since the grid cells narrow in the zonal direction at high latitudes, and so more grid cells fit into the east-west transects.
 The VR grids more accurately reproduce the steep margins of the ice sheet, capturing the characteristic parabolic shape of the GrIS margin.
@@
 The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids have similar dynamical resolution and use identical smoothing.
-This suggests that coarser models will tend to elevate the ablation zones {\color{blue}{relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and}} depressed melt rates, but this is opposite the melt bias that occurs in the coarse grid simulations.
-{\color{blue}{Also, the EC downscaling should be able to correct for situations where model ice surface is elevated relative to the true ice surface, although for very large elevation differences/errors the EC scheme may not be adequate}}.
+This suggests that coarser models will tend to elevate the ablation zones {\color{blue}{relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and}} depressed melt rates.
+This is opposite the melt bias in the coarse grid simulations.
+{\color{blue}{Also, the EC downscaling should be able to correct for situations where model ice surface is elevated relative to the true ice surface, although for very large elevation differences or errors, the EC scheme may not be adequate}}.
 
@@
 We hypothesize that coarser models have larger melt biases because summer melting is confined to these mixed ice/land/ocean grid cells.
-CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees the homogenized state after {\color{blue}{area}} averaging over the sub-grid mixture.
+CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees the homogenized state after {\color{blue}{area-}}averaging over the sub-grid mixture.
 Thus, warm ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice-covered patch {\color{blue}{and more melting}}.
@@
 The figure shows that coarser grids can be characterized by a monotonic increase in melt bias as the grid cell ice fraction decreases.
-The VR grids have the smallest melt biases for small grid cell ice fractions (less than 50\%), the {\color{blue}{quasi-}}uniform SE grids and $\tt f19$ have the largest melt biases and the $\tt f09$ grid melt biases lie between these two cases.
+The VR grids have the smallest melt biases for small grid cell ice fractions (less than 50\%), the {\color{blue}{quasi-}}uniform SE grids and $\tt f19$ have the largest melt biases, and the $\tt f09$ grid melt biases lie between these two cases.
 Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone of coarse grids is responsible for their large melt bias.
@@
 
-{\color{blue}{Another potential factor in explaining the differences across the various grids and dycores is the surface albedo.
-Bare ice has a lower albedo than snow, so it's reasonable to speculate whether the large melt biases in the coarse grid simulations are in part due to more frequent bare ice exposure.
-One logical pathway to more bare ice exposure is related to the precipitation and cloud biases.
-As the coarser models accumulate too much precipitation in the GrIS interior, the coastal ice margins receive too little snow, which can be more easily melted away to expose the underlying glacial ice.
-Figure~\ref{fig:alb} shows maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and CERES-EBAF datasets.
-The surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface.
-The lat-lon and quasi-uniform grids all have anomalously low albedo around the coasts in the southeast and northwest regions of the GrIS, which coincides with regions of anomalously low cloud cover and precipitation.
-In the VR simulations, the low albedo biases are alleviated to a large extent.
-
-In light of these albedo results, one may interpret Figure~\ref{fig:bias} differently.
-That there are larger melt biases for grid cells with only only partial ice cover, may have less to do with the partial ice cover and more to do with the enhanced bare ice exposure in the coarser grids.
-Further experimentation is required to tease out the relative contributions of these two processes.}}
+{\color{blue}{Changes in surface albedo may also explain the melt differences across the grids and dycores.
+The top and middle rows of Figure~\ref{fig:alb} show maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and the processed MODIS albedo dataset of \citeA{MODIS}.
+The model surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface.
+Both RACMO2.3p2 and MODIS datasets indicate that the albedo in the quasi-uniform grids is lower in the central-west and northwest region, relative to the other grids, whereas the albedo is anomalously low at the southwest and southeast coastal margins in both lat-lon and quasi-uniform grids, compared with the VR grids.
+All grids and dycores have a positive albedo anomaly in northeast Greenland.
+
+The lower albedo values in the lat-lon and quasi-uniform grids is consistent with their larger melt rates.
+Since bare ice has a lower albedo than snow, lower snow fractions over the GrIS will expose larger areas of bare ice and reduce the surface albedo.
+The bottom panel of Figure~\ref{fig:alb} shows the summer snow fraction bias, computed relative to the ERA5 reanalysis dataset.
+There is broad agreement with the albedo patterns.
+The quasi-uniform grids have low snow fractions in the central-west and north-west coastal regions, and both lat-lon and quasi-uniform grids have anomalously low snow fractions along the southeast and southwestern coasts.
+In contrast, the $\tt Arctic$ grid has larger snow fractions around the south and western coasts, in better agreement with ERA5.
+The $\tt Arctic-GrIS$ grid has low snow fractions along the western coasts, similar to the lat-lon and quasi-uniform grids.
+
+To the extent that the snow fraction changes are responsible for the more salient albedo in anomalies in the lat-lon and quasi-uniform grids, we seek to understand the snow cover changes.
+As the lat-lon and quasi-uniform grids accumulate too much precipitation in the GrIS interior, it is possible that the coastal ice margins --- where seasonal melt typically occurs --- receive too little snow, which may be more easily melted away to expose the underlying glacial ice.
+However, Figure~\ref{fig:prect} does not support a reduction in precipitation along the coastal margins, even for the solid precipitation (not shown).
+Instead it seems more likely that the warmer near surface environment, especially in the quasi-uniform grids, leads to anomalously low snow cover and consequent albedo changes.}}
 
@@
 \end{center}
-\caption{1979-1998 northern hemisphere summer albedo bias (fraction), computed as the difference (top) from RACMO2.3p2, and (bottom) CERES-EBAF dataset.
-RACMO2.3p2 differences are found by mapping the model output to the individual model grids, and differences in the bottom panel are computed after mapping the individual model grids to the CERES-EBAF $1^{\circ}$ grid.
-Note that the averaging period for CERES-EBAF panels are 2003-2020, which is different from the averaging period for the model results.}
+\caption{(Middle and bottom) 1979-1998 northern hemisphere summer albedo bias (fraction), computed as the difference (middle) from RACMO2.3p2, and (bottom) the MODIS dataset of \citeA{MODIS}.
+(Bottom) 1979-1998 bias in summer snow cover (fraction) relative to ERA5.
+All model and validation data have been mapped to the $\tt f19$ grid to facilitate the comparison across the different grids.}
 \label{fig:alb}
@@
 Both use the SE dycore; the $\tt Arctic$ grid has $1/4^{\circ}$ refinement over the broader Arctic, whereas the $\tt Arctic-GrIS$ grid is identical except for a $1/8^{\circ}$ patch of refinement over Greenland.
-A third VR grid, $\tt CONUS$, {\color{blue}{with $1/8^{\circ}$ refinement over the US}}, has also been made available in CESM2.2.
-
-In general, the FV grids have colder summer temperatures over the Arctic compared with the SE grids (including the VR grids).
-The cloud biases in all the {\color{blue}{lat-lon and quasi-}}uniform resolution grids, whether FV or SE, are similar, in general being too cloudy over Arctic land masses.
+A third VR grid, $\tt CONUS$, {\color{blue}{with $1/8^{\circ}$ refinement over the continental US}}, has also been made available in CESM2.2.
+
+In general, the FV grids simulate colder summer temperatures over the Arctic compared with the SE grids (including the VR grids).
+The cloud biases in all the {\color{blue}{lat-lon and quasi-}}uniform resolution grids {\color{blue}(FV and SE)} are similar, in general being too cloudy over Arctic land masses.
 It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates.
@@
 The $\tt Arctic-GrIS$ grid has the warmest summer lower troposphere of all grids, yet it has less melting than the quasi-uniform resolution SE grids.
-This suggests that grid resolution is somehow contributing to the melt biases in coarse grids, in a way that is not obvious from the large-scale dynamics.
-We propose a mechanism: coarse grids represent ablation zones using grid cells with mixed surface types, ice-covered and ice-free.
+This suggests that grid resolution (in coarse grids) is contributing to the melt biases in a way that is not obvious from the large-scale dynamics.
+The mechanism we propose is that coarse grids represent ablation zones using grid cells with mixed surface types, ice-covered and ice-free.
 The warmer ice-free patches may largely determine the mean state, leading to a warm bias over the ice-covered patches of the grid cell.
 This mechanism is supported by analysis of melt biases binned by grid-cell ice fraction.
+{\color{blue}{We leave further analysis of this hypothesis to future work.}}
 
 The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the GrIS SMB, compared to the {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform}} grids.
-The $\tt Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but its summer temperatures are too warm.
-The $1^{\circ}$ FV model gives a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids (but perhaps also because it is the most heavily tuned model configuration in CESM).
-In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other {\color{blue}{lat-lon and quasi-}}uniform grids.
+The $\tt Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but the summer temperatures are too warm.
+The $1^{\circ}$ FV model simulates a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids (but perhaps also because it is the most heavily tuned model configuration in CESM).
+In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other {\color{blue}{(coarser) lat-lon and quasi-}}uniform grids.
 
@@
 {\color{blue}{The SE dycore is still under active development (e.g., \ref{sec:a1}) compared to the more mature FV dycore, and future algorithmic improvements may reduce the FV-SE GrIS SMB skill gap.
-However, such developments are unlikely to eliminate this skill gap entirely due to the fewer number of grid cells representing high latitude structures on quasi-uniform unstructured grids.}} The simulated GrIS SMB will be adversely impacted in future CESM versions, after the FV dycore is phased out.
-This finding will not interrupt the ongoing transition towards unstructured grids in CESM, which is largely driven by gains in computational efficiency and grid refinement capabilities.
+However, such developments are unlikely to eliminate this skill gap entirely, because quasi-uniform unstructured grids have fewer grid cells representing high-latitude structures.}} Thus, the simulated GrIS SMB will likely be adversely impacted in future CESM versions, after the FV dycore is phased out.
+This finding will not interrupt the ongoing transition towards unstructured grids in CESM however, which is largely driven by gains in computational efficiency and grid refinement capabilities.
 We therefore provide the Arctic refined-meshes to the community by way of CESM2.2, providing users the option to simulate a more realistic GrIS SMB, although at a substantial computational premium relative to conventional $1^{\circ}-2^{\circ}$ grids.
@@
 We are  working to develop a configuration of the $\tt Arctic$ grid that is fully-coupled with the CESM ocean and sea ice components and the Community Ice Sheet Model (CISM), to provide multi-century projections of the state of the GrIS and its contribution to sea-level rise.
-We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on youtube (see link in Acknowledgements) to increase awareness of the capabilities of CESM2.2.
+We have also developed a visualization of the $\tt Arctic-GrIS$ run, now available on YouTube (see link in Acknowledgements) to increase awareness of the capabilities of CESM2.2.
 Figure~\ref{fig:viz} shows a snapshot of this visualization, illustrating mesoscale katabatic winds descending the southeastern slopes of the GrIS.

### 2022-09-16 first_review/response-to-reviewers.tex
update

@@
 
-{\color{red}{We would like to thank both reviewers for taking the time to provide thoughtful feedback on our manuscript.}} \newline
+{\color{blue}{We would like to thank both reviewers for carefully reviewing our manuscript and providing detailed feedback.
+We have taken seriously the reviewers questions/concerns and hope they are satisfied with our responses.
+Some of the larger issues we addressed are as follows.
+We have added a section on computational costs, and added cost metrics to the main Table 1.
+A few paragraphs (and a figure) were added to understand the impact of albedo changes on the melt biases.
+It was not entirely clear whether we are running the ice sheet model (we are not), and so we tried to hammer that point home at the very top of the methods section.
+There were several definitions of the surface mass balance (SMB) throughout, and not all of them the same.
+This has been cleaned up for clarity, making clear that the internally computed SMB in CESM is a bit unconventional, and that we use the components of the internally computed SMB to construct a more complete SMB, which we evaluate throughout the paper.}}
 
@@
 {\color{blue}{We thank the reviewer for this point.
-We had not thought too much about albedo and so it's for the best that this was brought up.
-We have made a good faith effort to explore this mechanism in the simulations, which has resulted in an additional figure (Figure 14) and two additional paragraphs at the end of section 2.3.
+We had not thought too much about albedo and so it's for the best that the reviewer brings this up.
+We have added a Figure and three paragraphs addressing the albedo changes in the simulations, which are significant.
+We have made a good faith effort to explore the precip-albedo mechanism that the reviewer describes.
 In general we don't find a great deal of evidence for the reviewers proposed mechanism.
 We believe this is largely because the positive interior precipitation biases are not offset by a negative precipitation bias at the coastal margins.
-We then took a closer look at albedo, where one validation dataset indicated the coarse grids had anomalously low albedo around the southern coasts, the other validation dataset indicated the low albedo anomaly occurs in the VR grids as well.}}
+We think it is more likely that the warmer near surface temperatures in the coarse grids are clawing back the snow fractions, and therefore the albedo.}}
 

### 2022-09-16 second_submit/arcticgrids.tex
update

@@
 \item The CESM2.2 release includes several enhancements to the spectral-element dynamical core, including two Arctic refined mesh configurations.
-\item {\color{blue}{Quasi-uniform unstructured}} grids degrade the Greenland Ice Sheet mass balance compared to latitude-longitude grids, {\color{blue}{at the conventional $1^{\circ}$ resolution}}.
-\item The refined Arctic meshes substantially improve the surface mass balance over {\color{blue}{conventional grid resolutions}}.
+\item Quasi-uniform unstructured grids degrade the Greenland Ice Sheet mass balance compared to latitude-longitude grids, at the conventional $1^{\circ}$ resolution.
+\item The refined Arctic meshes substantially improve the surface mass balance over conventional grid resolutions.
 \end{keypoints}
@@
 Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS).
-{\color{blue}{The finite-volume dynamical core uses structured, latitude-longitude grids, whereas the spectral-element dynamical core is built on unstructured meshes, permitting grid flexibility such as quasi-uniform grid spacing globally.}} The $1^{\circ}-2^{\circ}$ {\color{blue}{latitude-longitude and quasi-uniform unstructured}} grids systematically overestimate both accumulation and ablation over the GrIS.
+The finite-volume dynamical core uses structured, latitude-longitude grids, whereas the spectral-element dynamical core is built on unstructured meshes, permitting grid flexibility such as quasi-uniform grid spacing globally.
+The $1^{\circ}-2^{\circ}$ latitude-longitude and quasi-uniform unstructured grids systematically overestimate both accumulation and ablation over the GrIS.
 Of these $1^{\circ}-2^{\circ}$ grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS.
@@
 This study evaluates the ability of different grids and atmospheric solvers (i.e, the dynamical core) in CESM2.2 to resolve the surface mass balance (SMB) of the Greenland Ice Sheet.
-We show that the ongoing transition away from latitude{\color{blue}{-}}longitude grids, towards quasi-uniform unstructured grids in CESM2.2 leads to a degradation of the simulated SMB.
+We show that the ongoing transition away from latitude-longitude grids, towards quasi-uniform unstructured grids in CESM2.2 leads to a degradation of the simulated SMB.
 Two variable-resolution grids with enhanced resolution over Greenland are developed and incorporated into the release of CESM2.2, which substantially improves the SMB over the latitude longitude grids.
@@
 Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain.
-With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic due to the polar singularity)}}.
-In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids {\color{blue}{(i.e., grids with enhanced resolution over a particular region)}}, to understand their impacts on the simulated Arctic climate.
+With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost (hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic due to the polar singularity).
+In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids (i.e., grids with enhanced resolution over a particular region), to understand their impacts on the simulated Arctic climate.
 We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.
@@
 While spectral transform methods are still used today, local numerical methods have become desirable for their ability to run efficiently on massively parallel systems.
-The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids, with cells near the polar singularity elongated in the zonal direction) and polar filters are necessary to ameliorate this numerical instability \cite{JW2010LNCSE}.
+The pole problem has thus re-emerged in contemporary climate models that use lat-lon grids, and some combination of reduced grids (modified lat-lon grids in which cells around the polar singularity are elongated in the zonal direction) and polar filters are necessary to ameliorate this numerical instability \cite{JW2010LNCSE}.
 Polar filters subdue the growth of unstable computational modes by applying additional damping to the numerical solution over polar regions.
@@
 
-Dycores built on lat-lon grids have some advantages over {\color{blue}{dycores built on}} unstructured grids.
+Dycores built on lat-lon grids have some advantages over dycores built on unstructured grids.
 Lat-lon coordinate lines are orthogonal, and aligned with zonally symmetric circulations that characterize many large-scale features of Earth's atmosphere.
@@
 
-The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which {\color{blue}{has a}} large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}.
+The polar filter in the FV model impedes efficiency at large processor (CPU) counts because it requires a spectral transform, which has a large communication overhead \cite{ST1995GEOS,DetAl2012IJHPCA}.
 Unstructured grids support quasi-uniform grid spacing globally, and there is no pole problem (e.g., Figure~\ref{fig:uni-grids}c).
-{\color{blue}{This is in part why}} unstructured grids are becoming more common; their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}.
+This is in part why unstructured grids are becoming more common; their improved performance on massively parallel systems and lack of constraints on grid structure \cite{TTI1997JCP,PL2007JCP,WETAL2013GMD}.
 This increased grid flexibility allows for the adoption of variable-resolution grids (e.g., Figure~\ref{fig:vr-grids}; hereafter abbreviated as \textit{VR}), sometimes referred to as regional grid refinement.
@@
 Polar filters do not directly interfere with the physical parameterizations, nor do they have any bearing on the surface models; e.g., the land model can take full advantage of the greater number of grid cells in polar regions on lat-lon grids.
-This is particularly relevant for the surface mass balance {\color{blue}{(\textit{SMB}; the integrated sum of precipitation and runoff)}} of the Greenland Ice Sheet, which relies on hydrological processes represented in the land model.
-
-The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales {\color{blue}\cite{Fyke_EA_2018_Rev_Geophys}} that are difficult to represent in GCMs \cite{P2010CC}.
+This is particularly relevant for the surface mass balance (\textit{SMB}; the integrated sum of precipitation and runoff) of the Greenland Ice Sheet, which relies on hydrological processes represented in the land model.
+
+The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales \cite{Fyke_EA_2018_Rev_Geophys} that are difficult to represent in GCMs \cite{P2010CC}.
 GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation.
-The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias {\color{blue}{in the interior}} \cite{P2000GPC,VETAL2018TC}.
-GrIS ablation areas {\color{blue}{(marginal regions where the annual SMB is negative)}} are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
+The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias in the interior \cite{P2000GPC,VETAL2018TC}.
+GrIS ablation areas (marginal regions where the annual SMB is negative) are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation.
 These narrow ablation zones are not fully resolved in low-resolution GCMs, and may further degrade the simulated SMB.
-{\color{blue}{More recently, GCMs such as CESM and the NASA Goddard Institute for Space Studies GCM \cite{AETAL2019JGR}, have implemented an elevation class downscaling scheme for computing the SMB.
-The downscaling helps to resolve to these narrow ablation zones in GCMs \cite{SETAL2019TC}, but large SMB biases remain.}} For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.
+More recently, GCMs such as CESM and the NASA Goddard Institute for Space Studies GCM \cite{AETAL2019JGR}, have implemented an elevation class downscaling scheme for computing the SMB.
+The downscaling helps to resolve these narrow ablation zones in GCMs \cite{SETAL2019TC}, but large SMB biases remain.
+For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.
 
@@
 
-In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating {\color{blue}{grids}} where Greenland is represented with {\color{blue}{up to}} $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution.
+In order to retain the benefits of RCMs in a GCM, \citeA{VETAL2018TC} used the VR capabilities of the SE dycore in CESM, generating grids where Greenland is represented with up to $1/4^{\circ}$ resolution, and elsewhere with the more conventional $1^{\circ}$ resolution.
 The simulated SMB compared favorably to the SMB from RCMs and observations.
@@
 \end{center}
-\caption{Computational grids for the $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids in this study.
+\caption{Computational grids for the $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids in this study, using a northern hemisphere polar projection.
 Grids names after Table~\ref{tbl:table1}, (a) $2^{\circ}$ lat-lon grid ($\tt f19$), (b) $1^{\circ}$ lat-lon grid ($\tt f09$), (c) $1^{\circ}$ quasi-uniform unstructured grid with reduced physics resolution ($\tt ne30pg2$) and (d) $1^{\circ}$ quasi-uniform unstructured grid ($\tt ne30pg3$).}
@@
 
-{\color{blue}{CESM2.2 is a CMIP6-class \cite<Coupled Model Intercomparison Project Phase 6;>{CMIP6} Earth System Model maintained by the National Center for Atmosphere Research.
+CESM2.2 is a CMIP6-class \cite<Coupled Model Intercomparison Project Phase 6;>{CMIP6} Earth System Model maintained by the National Center for Atmosphere Research.
 CESM2.2 contains sub-component models for the atmosphere, land, ocean, sea-ice, and land-ice, and can be configured to run with varying degrees of complexity.
@@
 The land-ice component is not dynamically active in $\tt FHIST$.
-However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the functionality to simulate the SMB of the Greenland Ice Sheet.}}
+However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the necessary functionality to simulate the SMB of the Greenland Ice Sheet.
 
@@
 This study compares the performance of the SE and FV dycores, omitting the EUL and FV3 dycores.
-CESM2 runs submitted to CMIP6 used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g., >{SETAL2014JAMES,BetAl2016CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
+CESM2 runs submitted to CMIP6 used the FV dycore, whereas the SE dycore is often used for global high-resolution simulations \cite<e.g.,>{SETAL2014JAMES,BetAl2016CC,IHESP} due to its higher throughput on massively parallel systems \cite{DetAl2012IJHPCA}.
 
@@
 
-The FV dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}.
+The FV dycore integrates the hydrostatic equations of motion using a finite-volume discretization on a spherical lat-lon grid \cite{LR1997QJR}.
 The 2D dynamics evolve in floating Lagrangian layers that are periodically mapped to an Eulerian reference grid in the vertical \cite{L2004MWR}.
@@
 
-The SE dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}.
-The computational domain is a cubed-sphere grid ({\color{blue}{obtained by projecting each face of a cube onto a sphere}}) tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
+The SE dycore integrates the hydrostatic equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}.
+The computational domain is a cubed-sphere grid (obtained by projecting each face of a cube onto a sphere) tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}).
 Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre quadrature points.
@@
 
-{\color{blue}{SE is a next-generation dycore, and is less mature than the FV dycore due to its shorter history.}} In CESM2.2, the SE numerics have been enhanced {\color{blue}{relative to CESM2.0}} to mitigate spurious noise over topography.
+SE is a next-generation dycore, and is less mature than the FV dycore due to its shorter history.
+In CESM2.2, the SE numerics have been enhanced relative to CESM2.0 to mitigate spurious noise over topography.
 These algorithmic changes are described in \ref{sec:a1}.
-{\color{blue}{Future versions of CESM will likely include further optimizations and enhancements to the (nascent) SE dycore.}}
-
-The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over {\color{blue}{quasi-uniform resolution grids}}.
+Future versions of CESM will likely include further optimizations and enhancements to the (more junior) SE dycore.
+
+The SE dycore supports regional grid refinement via its VR configuration, requiring two enhancements over quasi-uniform resolution grids.
 First, as the numerical viscosity increases with resolution, explicit hyperviscosity relaxes according to the local element size, reducing in strength  by an order of magnitude per halving of the grid spacing.
@@
 The physical parameterizations are computed from the state on the CSLAM grid, which has clear advantages over the original SE dycore in which the physics are evaluated at Gauss-Lobatto-Legendre points \cite{HL2018MWR}.
-{\color{blue}{CSLAM advection is not available in the VR configuration, which instead uses the standard SE tracer transport scheme with the physics evaluated at Gauss-Lobatto-Legendre points.}}
+CSLAM advection is not available in the VR configuration, which instead uses the standard SE tracer transport scheme with the physics evaluated at Gauss-Lobatto-Legendre points.
 
@@
 The VR meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}).
-The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region ({\color{blue}{when using degrees to describe the resolution of unstructured grids, we refer to degrees at the equator}}).
+The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region (when using degrees to describe the resolution of unstructured grids, we refer to degrees at the equator).
 The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution.
@@
    \hline
-   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{cost(25 nodes)}} & {\color{blue}{cost(50 nodes)}} \\ 
+   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & cost(25 nodes) & cost(50 nodes) \\ 
    \hline
-   $\tt f19$ & FV & 278 & - &1800 & {\color{blue}{436.66}} & {\color{blue}{-}} \\
-   $\tt f09$ & FV & 139 & - &1800 & {\color{blue}{1534.57}} & {\color{blue}{2024.24}} \\
-   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & {\color{blue}{1497.26}} & {\color{blue}{1683.97}} \\
-   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & {\color{blue}{1890.48}} & {\color{blue}{2090.43}} \\
-   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & {\color{blue}{-}} & {\color{blue}{-}} \\
-   $\tt Arctic$ & SE & 111 & 28 & 450 & {\color{blue}{15947.41}} & {\color{blue}{16675.45}} \\
-   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & {\color{blue}{40305.03}} & {\color{blue}{41036.67}} \\
+   $\tt f19$ & FV & 278 & - &1800 & 436.66 & - \\
+   $\tt f09$ & FV & 139 & - &1800 & 1534.57 & 2024.24 \\
+   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & 1497.26 & 1683.97 \\
+   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & 1890.48 & 2090.43 \\
+   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & - & - \\
+   $\tt Arctic$ & SE & 111 & 28 & 450 & 15947.41 & 16675.45 \\
+   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & 40305.03 & 41036.67 \\
  \hline
@@
 FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.
-{\color{blue}{The FV dycore uses lat-lon grids, whereas the SE and SE-CSLAM dycores run on unstructured grids.}} We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.
-{\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY).
-The costs are from single-month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
+The FV dycore uses lat-lon grids, whereas the SE and SE-CSLAM dycores run on unstructured grids.
+We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$.
+The last columns provide the computational costs in core hours per simulated year (CHPSY).
+The costs are from single-month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}
  \label{tbl:table1}
@@
 The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution.
-Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids, {\color{blue}{as well as two grids pertaining to the Regional Atmospheric Climate Model \cite<RACMO;>{NETAL2018TC,NETAL2019SCIENCE}, which are used for validation (Table~\ref{tbl:table2})}}.
+Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids, as well as two grids pertaining to the Regional Atmospheric Climate Model \cite<RACMO;>{NETAL2018TC,NETAL2019SCIENCE}, which are used for validation (Table~\ref{tbl:table2}).
 The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid.
 The $\tt ne30pg3$ grid has an average $\Delta x=111.2$ km, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60$ km.
-Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
+Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs (Table~\ref{tbl:table1}), the convergence of meridians on the FV grid enhances the resolution over the GrIS.
 The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km.
@@
 
-\subsection{{\color{blue}{Computational costs}}}
-
-{\color{blue}{The last columns of Table~\ref{tbl:table1} provide cost estimates for the different grids and dycores.
+\subsection{Computational costs}
+
+The last columns of Table~\ref{tbl:table1} provide cost estimates for the different grids and dycores.
 The costs, expressed as core hours per simulated year (CHPSY), are taken from single-month runs of $\tt FHIST$, with no i/o, and using 25 nodes (900 tasks) and 50 nodes (1800 tasks) on the Cheyenne supercomputer \cite{cheyenne}.
@@
 Other approaches for comparing costs across different grids and dycores, e.g., holding fixed the number of grid columns per task, are beyond the scope of this study.
-
-%Running with the more common $1^{\circ}$ grid for multi-decadal climate simulations, 25 nodes is on the low side, but it's the largest number of tasks that can be supported by the $\tt f19$ grid, and we chose to fix the number of tasks across all grids for the purposes of comparing their costs.
 
@@
 All timing numbers are from runs without threading.
-The $\tt f09$ grid is the only grid that runs out-of-the-box with threading; holding the number of tasks fixed reduces CHPSY by $4\%-6.5\%$ reduction compared to runs without threading.}}
+The $\tt f09$ grid is the only grid that runs out-of-the-box with threading; holding the number of tasks fixed reduces CHPSY by $4\%-6.5\%$ reduction compared to runs without threading.
 
@@
 
-{\color{blue}{CESM simulates the GrIS SMB as the sum of ice accumulation and ice ablation.
+CESM simulates the GrIS SMB as the sum of ice accumulation and ice ablation.
 The latter includes sublimation, as well as liquid runoff from ice melt.
-Liquid precipitation and runoff may also contribute to ice accumulation by penetrating pore spaces in the snowpack/firn layer and freezing, forming ice lenses}}.
+Liquid precipitation and runoff may also contribute to ice accumulation by penetrating pore spaces in the snowpack/firn layer and freezing, forming ice lenses.
 The relevant SMB processes are represented by different CESM components, but it is the Community Land Model, version 5 \cite<CLM;>{CLM5}, that aggregates these processes and computes the SMB.
@@
 In short, the ice sheet patch in a CLM grid cell is subdivided into 10 elevation classes (ECs), each with a distinct surface energy balance and SMB.
-The area fraction of each EC is computed from {\color{blue}{the CISM initial conditions, which are based on a high-resolution dataset of the observed, modern extent and thickness of the GrIS \cite{METAL2014NG}.
-For configurations with a dynamically active ice sheet, the area fractions are continuously updated throughout the run to reflect the evolving ice sheet geometry in CISM}}.
+The area fraction of each EC is computed from the CISM initial conditions, which are based on a high-resolution dataset of the observed, modern extent and thickness of the GrIS \cite{METAL2014NG}.
+For configurations with a dynamically active ice sheet, the area fractions are continuously updated throughout the run to reflect the evolving ice sheet geometry in CISM.
 The near-surface air temperature, humidity, and air density are calculated for each EC using an assumed lapse rate and the elevation difference from the grid-cell mean.
@@
 
-{\color{blue}{Changes in ice depth, not snow depth, count toward the SMB.
+Changes in ice depth, not snow depth, count toward the SMB.
 Snow accumulation in each EC is limited to a depth of 10 m liquid water equivalent.
 Any snow above the 10 m cap contributes to ice accumulation, and refreezing of liquid water within the snowpack is an additional source of ice.
-Surface ice melting (after melting of the overlying snow) yields a negative SMB.}} Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
+Surface ice melting (after melting of the overlying snow) yields a negative SMB.
+Integrating over all ECs, weighting by the area fractions, provides a more accurate SMB than would be found using the grid-cell mean elevation.
 For a more detailed description of how the SMB is computed in CESM, we refer the reader to \citeA{CISM1,SETAL2019TC,KETAL2020JAMES,Muntjewerf_EA_2021_coupling}.
 
-Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation ({\color{blue}{with perpetual 1979 boundary conditions}}) for about 500 years.
+Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation (with perpetual 1979 boundary conditions) for about 500 years.
 The $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids are initialized with the SMB from an existing $\tt f09$ spun-up initial condition.
 
-\subsection{{\color{blue}{Validation}} Datasets}
-
-We use several {\color{blue}{validation}} datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations.
-{\color{blue}{The ERA5 reanalysis product \cite{ERA5} is used to validate the large-scale circulation and extreme precipitation events.
+\subsection{Validation Datasets}
+
+We use several validation datasets (Table~\ref{tbl:table2}) to assess the performance of the simulations.
+The ERA5 reanalysis product \cite{ERA5} is used to validate the large-scale circulation and extreme precipitation events.
 Clouds and radiation fields are validated using remote sensing products, the CERES-EBAF ED4.1 \cite{CERES-EBAF} and the CALIPSO-GOCCP \cite{CALIPSO-GOCCP} datasets, respectively.
-We also use a MODIS GrIS surface albedo product, mapped to a 25 km regional climate model grid \cite{MODIS}.}}
+We also use a MODIS GrIS surface albedo product, mapped to a 25 km regional climate model grid \cite{MODIS}.
 
@@
 The RACMO simulations have been shown to perform skillfully against observations and are often used as modeling targets \cite<e.g.,>{LIVVkit,KETAL2020JAMES}.
-{\color{blue}{The RACMO datasets are used along with the CERES-EBAF product to validate the radiative fluxes around Greenland.}}
-
-In-situ SMB (snow pit and ice cores) and {\color{blue}{remote sensing datasets (e.g., IceBridge radar accumulation dataset)}} are maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}.
+The RACMO datasets are used along with the CERES-EBAF product to validate the radiative fluxes around Greenland.
+
+In-situ SMB (snow pit and ice cores) and remote sensing datasets (e.g., IceBridge radar accumulation dataset) are maintained in The Land Ice Verification and Validation toolkit (LIVVkit), version 2.1 \cite{LIVVkit}.
 However, these point-wise measurements are difficult to compare to model output spanning several different grids, especially since the SMB from each elevation class is not available from the model output.
@@
    CERES-EBAF ED4.1 & 2003-2020 & $1^{\circ}$ & \citeA{CERES-EBAF} \\
-   CALIPSO-GOCCP & 2006-2017 & {\color{blue}{$2^{\circ}$}} & \citeA{CALIPSO-GOCCP} \\
-   {\color{blue}{MODIS-MAR}} & {\color{blue}{2000-2013}} & {\color{blue}{25 km}} & {\color{blue}{\citeA{MODIS}}}\\
+   CALIPSO-GOCCP & 2006-2017 & $2^{\circ}$ & \citeA{CALIPSO-GOCCP} \\
+   MODIS-MAR & 2000-2013 & 25 km & \citeA{MODIS}\\
    RACMO2.3 & 1979-1998 & 11 km & \citeA{NETAL2015TC} \\
@@
  \end{tabular}
-  \caption{Description of {\color{blue}{validation}} datasets used in this study.}
+  \caption{Description of validation datasets used in this study.}
  \label{tbl:table2}
@@
 We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
-However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset ({\color{blue}{i.e., the GrIS initial conditions in CISM;}} Figure~\ref{fig:grisdx}b).
+However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset ({i.e., the GrIS initial conditions in CISM; Figure~\ref{fig:grisdx}b).
 CLM's dataset creation tool generates the model ice mask by mapping the reference dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
@@
 (Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from the reference dataset described in the text.
-{\color{blue}{Also included are the two RACMO grids used for validation (see Table~\ref{tbl:table2})}}.}
+Also included are the two RACMO grids used for validation (see Table~\ref{tbl:table2}).}
 \label{fig:grisdx}
@@
 
-{\color{blue}{The total water SMB (equation~\ref{eq:SMB}) is different from the SMB internally computed by CLM and described in section~\ref{sec:simSMB}, which only tracks ice mass.
+The total water SMB (equation~\ref{eq:SMB}) is different from the SMB internally computed by CLM and described in section~\ref{sec:simSMB}, which only tracks ice mass.
 We do not use CLM's internally computed SMB in this study.
-Rather, we use the components of the internally computed SMB to construct the total water SMB}}.
+Rather, we use the components of the internally computed SMB to construct the total water SMB.
 
@@
 
-The different error characteristics of the two methods are used to {\color{blue}{further}} diversify the ensemble.
+The different error characteristics of the two methods are used to further diversify the ensemble.
 Each of the four regridding combinations (with conservative and high-order remapping to the $\tt f09$ and $\tt ne30pg2$ grids) are repeated with each method, resulting in (up to) eight values for each integrated quantity.
@@
 Before delving into the simulated Arctic climate conditions, we assess the global mean differences between the various grids and dycores.
-Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots {\color{blue}{of temperature}}, expressed as differences between {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured}} grids and dycores.
+Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots of temperature, expressed as differences between $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids and dycores.
 The $\tt f09$ grid is warmer than the $\tt f19$ grid, primarily in the mid-to-high latitudes throughout the depth of the troposphere.
@@
 
-Comparing the VR grids to the {\color{blue}{lat-lon and quasi-uniform}} grids is complicated because we simultaneously increase the resolution and reduce the physics time-step, both of which influence the solution \cite{W2008TELLUS}.
+Comparing the VR grids to the lat-lon and quasi-uniform grids is complicated because we simultaneously increase the resolution and reduce the physics time-step, both of which influence the solution \cite{W2008TELLUS}.
 We therefore run an additional $\tt ne30pg3$ simulation with the shorter physics time step used in the $\tt Arctic$ grid (450~s), referred to as $\tt ne30pg3^{*}$ (Table~\ref{tbl:table1}).
@@
 
-{\color{blue}{There is not perfect alignment of the CLUBB heating anomalies and the temperature anomalies in Figures~\ref{fig:dT-lores} and \ref{fig:dT-hires}.
+There is not perfect alignment of the CLUBB heating anomalies and the temperature anomalies in Figures~\ref{fig:dT-lores} and \ref{fig:dT-hires}.
 This leaves some uncertainty as to whether CLUBB is solely responsible for the temperature anomalies caused by changing grid resolution or dycore, as asserted above.
@@
 \citeA{HK2020QJRMS} have shown that only the CLUBB heating coinciding with upward resolved vertical velocities tends to align spatially with the temperature response, whereas heating anomalies associated with descending grid columns have no alignment with the temperature response.
-Therefore the average over all time will dilute the heating signal, potentially leading to misalignments.}}
+Therefore the average over all time will dilute the heating signal, potentially leading to misalignments.
 
@@
 
-{\color{blue}{Keeping our focus on the Arctic region, and in particular Greenland,}} it is useful to understand summer temperature biases due to their control on ice and snow melt over the GrIS \cite<>{O2001JAM}.
+Keeping our focus on the Arctic region, and in particular Greenland, it is useful to understand summer temperature biases due to their control on ice and snow melt over the GrIS \cite<>{O2001JAM}.
 Figure~\ref{fig:dThyps} shows the 1979--1998 lower-troposphere summer temperature bias relative to ERA5, computed by equating a layer mean virtual temperature with the 500--1000 hPa geopotential thickness.
@@
 A negative bias corresponds to excessive reflection and cooling.
-The {\color{blue}{lat-lon and quasi-uniform}} grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses.
+The lat-lon and quasi-uniform grids have similar biases, with the clouds reflecting 20--40 W/m$^2$ too much shortwave radiation over a wide swath of the Arctic, primarily the land masses.
 There is also a halo of positive bias (clouds not reflective enough) around the ocean perimeter of Greenland.
@@
 While we have not quantified the contribution of cloud biases to the cooler Arctic temperatures, shortwave radiation is a crucial component of the Arctic energy budget during summer.
-The shortwave cloud forcing biases are on the order of 10 W/m2, which is a significant fraction of the total absorbed shortwave {\color{blue}{radiation}} during Arctic summer  \cite{SETAL2007JGR} and is therefore likely a factor contributing to the cooler temperatures.
+The shortwave cloud forcing biases are on the order of 10 W/m2, which is a significant fraction of the total absorbed shortwave radiation during Arctic summer  \cite{SETAL2007JGR} and is therefore likely a factor contributing to the cooler temperatures.
 
@@
 The total SMB on coarse grids therefore has smaller errors than the individual components (Table~\ref{tbl:table3}), because large errors in the source and sink terms offset one another when added together.
-Such compensating errors {\color{blue}{underscore}} the importance of understanding whether a model is getting the right SMB for the right reasons.
+Such compensating errors underscore the importance of understanding whether a model is getting the right SMB for the right reasons.
 
@@
 The 1979-1998 climatological mean values from Table~\ref{tbl:table3} are shown as circles on the right side of the panels.
-The {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform}} grids have positive precipitation biases, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
+The $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform grids have positive precipitation biases, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
 The $\tt f19$ and $\tt f09$ grids perform similarly, with +110 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +165 Gt/yr and $\tt ne30pg2$ by +230 Gt/yr.
@@
 The basin-integrated melt rates in Figure~\ref{fig:basin} depend on the dycore.
-The {\color{blue}{quasi-}}uniform SE grids have the largest positive biases in all basins.
-The $\tt Arctic-GrIS$ grid is a close second, while the FV grids have systematically {\color{blue}{smaller melt-rates and melt-rate biases}}.
+The quasi-uniform SE grids have the largest positive biases in all basins.
+The $\tt Arctic-GrIS$ grid is a close second, while the FV grids have systematically smaller melt-rates and melt-rate biases.
 The ``second-place" standing of $\tt Arctic-GrIS$ is somewhat unexpected, as this grid has the warmest lower-troposphere summer temperatures (Figure~\ref{fig:dThyps}) and greatest incident shortwave radiation (Figure~\ref{fig:FSDS}), yet it has less melting than the quasi-uniform SE grids.
@@
 We compute PDD from monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}$C and a degree-day factor of 5~mm~d$^{-1}$ $^{\circ}$C$^{-1}$.
-{\color{blue}{This specific degree-day factor lies between typical values reported for snow and ice, in order to easily apply the PDD method to estimate the combined snow and ice melt.}}
+This specific degree-day factor lies between typical values reported for snow and ice, in order to apply the PDD method to estimate the combined snow and ice melt.
 
@@
 The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids have similar dynamical resolution and use identical smoothing.
-This suggests that coarser models will tend to elevate the ablation zones {\color{blue}{relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and}} depressed melt rates.
+This suggests that coarser models will tend to elevate the ablation zones relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and depressed melt rates.
 This is opposite the melt bias in the coarse grid simulations.
-{\color{blue}{Also, the EC downscaling should be able to correct for situations where model ice surface is elevated relative to the true ice surface, although for very large elevation differences or errors, the EC scheme may not be adequate}}.
+Also, the EC downscaling should be able to correct for situations where model ice surface is elevated relative to the true ice surface, although for very large elevation differences or errors, the EC scheme may not be adequate.
 
@@
 We hypothesize that coarser models have larger melt biases because summer melting is confined to these mixed ice/land/ocean grid cells.
-CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees the homogenized state after {\color{blue}{area-}}averaging over the sub-grid mixture.
-Thus, warm ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice-covered patch {\color{blue}{and more melting}}.
+CLM deals with land heterogeneity in a complex and sophisticated manner, but CAM only sees the homogenized state after area-averaging over the sub-grid mixture.
+Thus, warm ice-free land patches in a grid cell may unduly influence the climate over the entire grid cell, causing a warm bias over the ice-covered patch and more melting.
 
@@
 The figure shows that coarser grids can be characterized by a monotonic increase in melt bias as the grid cell ice fraction decreases.
-The VR grids have the smallest melt biases for small grid cell ice fractions (less than 50\%), the {\color{blue}{quasi-}}uniform SE grids and $\tt f19$ have the largest melt biases, and the $\tt f09$ grid melt biases lie between these two cases.
+The VR grids have the smallest melt biases for small grid cell ice fractions (less than 50\%), the quasi-uniform SE grids and $\tt f19$ have the largest melt biases, and the $\tt f09$ grid melt biases lie between these two cases.
 Figure~\ref{fig:bias} generally supports our hypothesis that the prevalence of mixed-grid cells in the ablation zone of coarse grids is responsible for their large melt bias.
@@
 
-{\color{blue}{Changes in surface albedo may also explain the melt differences across the grids and dycores.
-The top and middle rows of Figure~\ref{fig:alb} show maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and the processed MODIS albedo dataset of \citeA{MODIS}.
-The model surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface.
+Changes in surface albedo may also explain the melt differences in the grids and dycores.
+The top and middle rows of Figure~\ref{fig:alb} show maps of summer surface albedo biases in the simulations, relative to the RACMO2.3p2 and the processed MODIS albedo dataset of \citeA{MODIS}.
+The model surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface..
+All grids and dycores have been mapped to the $\tt f19$ grid in the figure, and only cells overlapping with the $\tt f19$ GrIS ice mask are shown.
 Both RACMO2.3p2 and MODIS datasets indicate that the albedo in the quasi-uniform grids is lower in the central-west and northwest region, relative to the other grids, whereas the albedo is anomalously low at the southwest and southeast coastal margins in both lat-lon and quasi-uniform grids, compared with the VR grids.
-All grids and dycores have a positive albedo anomaly in northeast Greenland.
+All grids and dycores have a positive albedo anomaly in northeast Greenland, which may explain the depressed melt rates and unrealistic ice sheet expansion in this region when coupled to the dynamic ice sheet \cite{LETAL2020JAMES}.
 
@@
 
-To the extent that the snow fraction changes are responsible for the more salient albedo in anomalies in the lat-lon and quasi-uniform grids, we seek to understand the snow cover changes.
+To the extent that the snow fraction changes are responsible for the albedo anomalies in the lat-lon and quasi-uniform grids, we seek to understand the snow cover changes.
 As the lat-lon and quasi-uniform grids accumulate too much precipitation in the GrIS interior, it is possible that the coastal ice margins --- where seasonal melt typically occurs --- receive too little snow, which may be more easily melted away to expose the underlying glacial ice.
-However, Figure~\ref{fig:prect} does not support a reduction in precipitation along the coastal margins, even for the solid precipitation (not shown).
-Instead it seems more likely that the warmer near surface environment, especially in the quasi-uniform grids, leads to anomalously low snow cover and consequent albedo changes.}}
+However, Figure~\ref{fig:prect} does not support a reduction in precipitation along the coastal margins, even when only plotting the solid precipitation (not shown).
+Instead it seems more likely that the warmer near surface environment, especially in the quasi-uniform grids, leads to anomalously low snow cover and consequent albedo changes.
 
@@
 
-Figure~\ref{fig:comp-pdf} shows monthly PDF {\color{blue}{(probability density function)}} of the precipitation rates associated with storms.
+Figure~\ref{fig:comp-pdf} shows monthly PDF (probability density function) of the precipitation rates associated with storms.
 The PDFs are constructed by sampling all the precipitation rates within $30^{\circ}$ of the storm center, for each point on the storm track and for all storms.
@@
 Both use the SE dycore; the $\tt Arctic$ grid has $1/4^{\circ}$ refinement over the broader Arctic, whereas the $\tt Arctic-GrIS$ grid is identical except for a $1/8^{\circ}$ patch of refinement over Greenland.
-A third VR grid, $\tt CONUS$, {\color{blue}{with $1/8^{\circ}$ refinement over the continental US}}, has also been made available in CESM2.2.
+A third VR grid, $\tt CONUS$, with $1/8^{\circ}$ refinement over the continental US, has also been made available in CESM2.2.
 
 In general, the FV grids simulate colder summer temperatures over the Arctic compared with the SE grids (including the VR grids).
-The cloud biases in all the {\color{blue}{lat-lon and quasi-}}uniform resolution grids {\color{blue}(FV and SE)} are similar, in general being too cloudy over Arctic land masses.
+The cloud biases in all the lat-lon and quasi-uniform resolution grids (FV and SE) are similar, in general being too cloudy over Arctic land masses.
 It should be emphasized that our analysis is specific to the Arctic summer because of its relevance to GrIS melt rates.
@@
 We integrated the primary source and sink terms of the SMB equation over the GrIS for each of the six grids.
-The {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform}} grids have large positive accumulation biases because they fail to resolve orographic precipitation.
+The $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform grids have large positive accumulation biases because they fail to resolve orographic precipitation.
 The quasi-uniform SE grids have larger accumulation biases, suggesting that the FV grids are more skillful for precipitation due to finer resolution over Greenland, and despite a polar filter.
@@
 This mechanism is supported by analysis of melt biases binned by grid-cell ice fraction.
-{\color{blue}{We leave further analysis of this hypothesis to future work.}}
-
-The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the GrIS SMB, compared to the {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform}} grids.
+We leave further analysis of this hypothesis to future work.
+
+The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the GrIS SMB, compared to the {$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform grids.
 The $\tt Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but the summer temperatures are too warm.
 The $1^{\circ}$ FV model simulates a surprisingly realistic SMB, likely due to the relatively fine resolution of Greenland on lat-lon grids (but perhaps also because it is the most heavily tuned model configuration in CESM).
-In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other {\color{blue}{(coarser) lat-lon and quasi-}}uniform grids.
+In particular, a greater number of grid cells in the ablation zone reduces the influence of mixed ice-covered/ice-free grid cells that represent ablation poorly on the other (coarser) lat-lon and quasi-uniform grids.
 
@@
 We have found that the $1^{\circ}$ FV model has clear advantages over the $1^{\circ}$ SE model for simulating the GrIS SMB.
-{\color{blue}{The SE dycore is still under active development (e.g., \ref{sec:a1}) compared to the more mature FV dycore, and future algorithmic improvements may reduce the FV-SE GrIS SMB skill gap.
-However, such developments are unlikely to eliminate this skill gap entirely, because quasi-uniform unstructured grids have fewer grid cells representing high-latitude structures.}} Thus, the simulated GrIS SMB will likely be adversely impacted in future CESM versions, after the FV dycore is phased out.
+The SE dycore is still under active development (e.g., \ref{sec:a1}) compared to the more mature FV dycore, and future algorithmic improvements may reduce the FV-SE GrIS SMB skill gap.
+However, such developments are unlikely to eliminate this skill gap entirely, because quasi-uniform unstructured grids have fewer grid cells representing high-latitude structures.
+Thus, the simulated GrIS SMB will likely be adversely impacted in future CESM versions, after the FV dycore is phased out.
 This finding will not interrupt the ongoing transition towards unstructured grids in CESM however, which is largely driven by gains in computational efficiency and grid refinement capabilities.
@@
 \appendix
-\section{Details on {\color{blue}{spectral}}-element dynamical core improvements since the CESM2.0 release}\label{sec:a1}
+\section{Details on spectral-element dynamical core improvements since the CESM2.0 release}\label{sec:a1}
 Since the CESM2.0 release of the spectral-element dynamical core documented in \citeA{LetAl2018JAMES} some important algorithmic improvements have been implemented and released with CESM2.2.
@@
 \end{eqnarray}
- \cite{SJ1991QJRMS} where $T_1=\Gamma_0 T_{ref} c_p^{(d)}/g\approx 192K$, {\color{blue}{with gravity $g$,}} and standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv T_{ref}-T_1\approx 97K$; $T_{ref}=288K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ {\color{blue}{is the}} surface geopotential.
+ \cite{SJ1991QJRMS} where $T_1=\Gamma_0 T_{ref} c_p^{(d)}/g\approx 192K$, with gravity $g$, and standard lapse rate $\Gamma_0\equiv 6.5K/km$ and $T_0\equiv T_{ref}-T_1\approx 97K$; $T_{ref}=288K$ ($c_p^{(d)}$ specific heat of dry air at constant pressure; $R^{(d)}$ gas constant for dry air), and $\Phi_s$ is the surface geopotential.
 The reference Exner function is
@@
 \end{equation}
-where $A$ and $B$ are the standard {\color{blue}{hybrid}} coefficients (using a dry-mass generalized vertical mass coordinate $\eta$).
+where $A$ and $B$ are the standard hybrid coefficients (using a dry-mass generalized vertical mass coordinate $\eta$).
 These reference profiles are subtracted from the prognostic temperature and pressure-level-thickness states before applying hyperviscosity:
@@
 \end{equation}
-{\color{blue}{and virtual temperature is
+and virtual temperature is
 \begin{equation}
@@
 \end{equation}
-where $m^{(\ell)}$ is dry mixing ratio of component of moist air $\ell$; $\mathcal{L}_{all}$ is the set of all components of moist air and, in particular, $'wv'$ is water vapor.}}
-
-The derivation showing that (\ref{eq:pgf}) and (\ref{eq:pgf2}) are equivalent is {\color{blue}{given}} here:
+where $m^{(\ell)}$ is dry mixing ratio of component of moist air $\ell$; $\mathcal{L}_{all}$ is the set of all components of moist air and, in particular, $'wv'$ is water vapor.
+
+The derivation showing that (\ref{eq:pgf}) and (\ref{eq:pgf2}) are equivalent is given here:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-{\color{blue}{where virtual potential temperature is $T_v/\Pi$.}} Using the reference states from \cite{SJ1991QJRMS},
+where virtual potential temperature is $T_v/\Pi$. Using the reference states from \cite{SJ1991QJRMS},
 \begin{eqnarray}

### 2022-09-17 first_review/response-to-reviewers.tex
second submit complete

@@
 \title{\large Response to Reviewers}
-\author{\normalsize Adam R. Herrington and CoAuthors}
+\author{\normalsize Adam R. Herrington and Coauthors}
 \date{} % delete this line to display the current date
@@
 There were several definitions of the surface mass balance (SMB) throughout, and not all of them the same.
-This has been cleaned up for clarity, making clear that the internally computed SMB in CESM is a bit unconventional, and that we use the components of the internally computed SMB to construct a more complete SMB, which we evaluate throughout the paper.}}
+This has been cleaned up for clarity, making clear that the internally computed SMB in CESM is a bit unconventional, and that we use the components of the internally computed SMB to construct a more complete SMB, which we evaluate throughout the paper.
+\newline
+
+Note that a bug was found in the script that computes the components of the SMB, and therefore Figure 10 and 11, and Table 3 have been updated with the bug fix.
+This bug does not impact our analysis or conclusions.}}
 

### 2022-09-17 second_submit/arcticgrids.tex
second submit complete

@@
 We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget.
-However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset ({i.e., the GrIS initial conditions in CISM; Figure~\ref{fig:grisdx}b).
+However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (i.e., the GrIS initial conditions in CISM; Figure~\ref{fig:grisdx}b).
 CLM's dataset creation tool generates the model ice mask by mapping the reference dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}.
@@
 \caption{1979-1998 northern hemisphere summer surface incident shortwave radiation bias (W/m2), computed as the difference (top) from RACMO2.3p2, and (middle) the CERES-EBAF dataset, and the (bottom) total cloud fraction bias relative to the CALIPSO dataset.
-CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
+CERES and CALIPSO differences are found by mapping the model output to their $1^{\circ}$ and $2^{\circ}$ grids, respectively, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
 Note that the averaging period for the CALIPSO-GOCCP and CERES-EBAF panels, 2006-2017 and 2003-2020, respectively, are different from the averaging period for the model results.}
@@
    \hline
-   grid name & accumulation & total melt & runoff & sublimation & SMB \\ 
+   grid name & accumulation & total melt & runoff & evap+subl & SMB \\ 
    \hline
@@
    \hline
-   $\tt ne30pg2$ & 1007. (973.4) & -519.9 (-647.3) & -381.9 (-347.0) & -33.9 (-32.1) & 591.2 (594.3) \\
-   $\tt ne30pg3$ & 931.0 (909.3) & -540.8 (-686.7) & -375.8 (-330.1) & -34.1 (-32.6) & 521.2 (546.6) \\
-   $\tt f19$ & 884.9 (913.5) & -414.0 (-546.5) & -284.0 (-284.3) & -36.5 (-37.5) & 564.4 (591.7) \\
-   $\tt f09$ & 873.9 (882.1) & -389.1 (-482.3) & -256.1 (-212.3) & -37.3 (-37.4) & 580.5 (632.4) \\
-   $\tt Arctic$ & 784.1 (818.6) & -335.5 (-436.8) & -215.8 (-194.2) & -42.4 (-43.9) & 526.0 (580.5) \\
-   $\tt Arctic-GrIS$ & 693.8 (747.3) & -437.3 (-610.4) & -276.8 (-307.8) & -48.1 (-51.8) & 369.0 (387.7) \\
+   $\tt ne30pg2$ & 1013.6 (986.9) & -560.6 (-561.4) & -410.0 (-380.6) & -39.0 (-37.2) & 564.6 (569.1) \\
+   $\tt ne30pg3$ & 930.6 (909.7) & -583.1 (-543.6) & -403.2 (-337.9) & -39.1 (-37.6) & 488.3 (534.2) \\
+   $\tt f19$ & 888.2 (919.7) & -446.3 (-472.1) & -304.5 (-307.6) & -41.9 (-42.9) & 541.8 (569.2) \\
+   $\tt f09$ & 878.6 (891.6) & -419.5 (-401.5) & -274.4 (-230.6) & -42.3 (-42.4) & 561.9 (618.6) \\
+   $\tt Arctic$ & 786.3 (823.5) & -361.2 (-350.5) & -231.4 (-199.1) & -47.0 (-48.7) & 507.9 (575.7) \\
+   $\tt Arctic-GrIS$ & 695.9 (751.6) & -471.6 (-513.7) & -297.7 (-301.8) & -53.2 (-57.0) & 345.0 (392.8) \\
  \hline
@@
   \caption{1979-1998 surface mass balance of the Greenland Ice Sheet in Gt/yr.
-Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.}
+Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.
+evap+subl refers to the combined evaporation and sublimation term.}
  \label{tbl:table3}
@@
 Table~\ref{tbl:table3} also contains (in parenthesis) the SMB components derived from evaluating the integrals on each model's native grid and ice mask.
-Of note is the large reduction in melt rates compared to the values computed on the native grid, illustrating the dissipation of this quantity discussed in section~\ref{sec:SMB}.
-For integrated precipitation, the differences between the native and common ice mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
+For the accumulation term in the quasi-uniform SE grids, the native grid values are actually less than common ice mask values, likely due to the inflationary impacts of method 1 discussed in section~\ref{sec:SMB}.
+For the total melt term, the common ice mask values aren't nearly as dissipative as expected, with only the $\tt Arctic-GrIS$ grid showing a large reduction in melt rates relative to the native grid values.
 
@@
 The $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform grids have positive precipitation biases, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
-The $\tt f19$ and $\tt f09$ grids perform similarly, with +110 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +165 Gt/yr and $\tt ne30pg2$ by +230 Gt/yr.
+The $\tt f19$ and $\tt f09$ grids perform similarly, with +200 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +250 Gt/yr and $\tt ne30pg2$ by +330 Gt/yr.
 
 The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $\tt Arctic$ grid simulates the most realistic melt rates, with the other grids having more melt than RACMO.
-The $\tt Arctic-GrIS$ grid over-predicts melting by about 125 Gt/yr.
+The $\tt Arctic-GrIS$ grid over-predicts melting by about 150 Gt/yr.
 This is likely due to an anomalously warm lower troposphere during the summer, relative to the $\tt Arctic$ run (Figure~\ref{fig:dThyps}).
-The $\tt f19$ and $\tt f09$ melting rates are improved over $\tt Arctic-GrIS$, overestimating melt by only 70--90 Gt/yr.
-The SE grids have the largest positive melt bias, between 200--220 Gt/yr.
+The $\tt f19$ and $\tt f09$ melting rates are improved over $\tt Arctic-GrIS$, overestimating melt by only 100--120 Gt/yr.
+The SE grids have the largest positive melt bias, between 240--265 Gt/yr.
 
@@
 The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids have similar dynamical resolution and use identical smoothing.
-This suggests that coarser models will tend to elevate the ablation zones relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and depressed melt rates.
+This suggests that coarser models will tend to elevate the ablation zones {relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and depressed melt rates.
 This is opposite the melt bias in the coarse grid simulations.
@@
 
-The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the GrIS SMB, compared to the {$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform grids.
+The $\tt Arctic$ grid substantially improves the simulated Arctic climate, including precipitation extremes and the GrIS SMB, compared to the $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform grids.
 The $\tt Arctic-GrIS$ grid has the most realistic cloud and precipitation fields, but the summer temperatures are too warm.

### 2022-09-17 second_submit/arcticgrids_trackchanges.tex
second submit complete

--- 2022-09-13 second_submit/arcticgrids.tex
@@
 {\color{blue}{More recently, GCMs such as CESM and the NASA Goddard Institute for Space Studies GCM \cite{AETAL2019JGR}, have implemented an elevation class downscaling scheme for computing the SMB.
-The downscaling helps to resolve to these narrow ablation zones in GCMs \cite{SETAL2019TC}, but large SMB biases remain.}} For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.
+The downscaling helps to resolve these narrow ablation zones in GCMs \cite{SETAL2019TC}, but large SMB biases remain.}} For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.
 
@@
 The land-ice component is not dynamically active in $\tt FHIST$.
-However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the functionality to simulate the SMB of the Greenland Ice Sheet.}}
+However, the surface mass balance is computed by the land model before being passed to the land-ice component; $\tt FHIST$ includes the necessary functionality to simulate the SMB of the Greenland Ice Sheet.}}
 
@@
 These algorithmic changes are described in \ref{sec:a1}.
-{\color{blue}{Future versions of CESM will likely include further optimizations and enhancements to the (nascent) SE dycore.}}
+{\color{blue}{Future versions of CESM will likely include further optimizations and enhancements to the (more junior) SE dycore.}}
 
@@
 The $\tt ne30pg3$ grid has an average $\Delta x=111.2$ km, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60$ km.
-Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS.
+Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs (Table~\ref{tbl:table1}), the convergence of meridians on the FV grid enhances the resolution over the GrIS.
 The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km.
@@
 \caption{1979-1998 northern hemisphere summer surface incident shortwave radiation bias (W/m2), computed as the difference (top) from RACMO2.3p2, and (middle) the CERES-EBAF dataset, and the (bottom) total cloud fraction bias relative to the CALIPSO dataset.
-CALIPSO and CERES differences are found by mapping the model output to the $1^{\circ}$ grid, and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
+{\color{blue}{CERES and CALIPSO differences are found by mapping the model output to their $1^{\circ}$ and $2^{\circ}$ grids, respectively,}} and differences in the bottom panel are computed after mapping the RACMO2.3p2 dataset to the individual model grids.
 Note that the averaging period for the CALIPSO-GOCCP and CERES-EBAF panels, 2006-2017 and 2003-2020, respectively, are different from the averaging period for the model results.}
@@
    \hline
-   grid name & accumulation & total melt & runoff & sublimation & SMB \\ 
+   grid name & accumulation & total melt & runoff & {\color{blue}{evap+subl}} & SMB \\ 
    \hline
@@
   \caption{1979-1998 surface mass balance of the Greenland Ice Sheet in Gt/yr.
-Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.}
+Values shown are using the common ice mask approach described in the methods section, whereas values in parentheses are from integrating over the native grid and ice mask.
+{\color{blue}{evap+subl refers to the combined evaporation and sublimation term.}}}
  \label{tbl:table3}
@@
 Table~\ref{tbl:table3} also contains (in parenthesis) the SMB components derived from evaluating the integrals on each model's native grid and ice mask.
-Of note is the large reduction in melt rates compared to the values computed on the native grid, illustrating the dissipation of this quantity discussed in section~\ref{sec:SMB}.
-For integrated precipitation, the differences between the native and common ice mask approaches are much smaller, since the combined solid/liquid precipitation rates are not directly tied to the ice mask.
+{\color{blue}{For the accumulation term in the quasi-uniform SE grids, the native grid values are actually less than common ice mask values, likely due to the inflationary impacts of method 1 discussed in section~\ref{sec:SMB}.
+For the total melt term, the common ice mask values aren't nearly as dissipative as expected, with only the $\tt Arctic-GrIS$ grid showing a large reduction in melt rates relative to the native grid values.}}
 
@@
 The {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform}} grids have positive precipitation biases, whereas the VR grids have the smallest biases, with precipitation comparable to RACMO.
-The $\tt f19$ and $\tt f09$ grids perform similarly, with +110 Gt/yr bias, whereas $\tt ne30pg3$ is biased by about +165 Gt/yr and $\tt ne30pg2$ by +230 Gt/yr.
+The $\tt f19$ and $\tt f09$ grids perform similarly, with {\color{blue}{+200}} Gt/yr bias, whereas $\tt ne30pg3$ is biased by about {\color{blue}{+250 Gt/yr}} and $\tt ne30pg2$ by {\color{blue}{+330}} Gt/yr.
 
 The combined annual snow/ice melt shown in the bottom panel of Figure~\ref{fig:tseries} indicates that the $\tt Arctic$ grid simulates the most realistic melt rates, with the other grids having more melt than RACMO.
-The $\tt Arctic-GrIS$ grid over-predicts melting by about 125 Gt/yr.
+The $\tt Arctic-GrIS$ grid over-predicts melting by about {\color{blue}{150}} Gt/yr.
 This is likely due to an anomalously warm lower troposphere during the summer, relative to the $\tt Arctic$ run (Figure~\ref{fig:dThyps}).
-The $\tt f19$ and $\tt f09$ melting rates are improved over $\tt Arctic-GrIS$, overestimating melt by only 70--90 Gt/yr.
-The SE grids have the largest positive melt bias, between 200--220 Gt/yr.
+The $\tt f19$ and $\tt f09$ melting rates are improved over $\tt Arctic-GrIS$, overestimating melt by only {\color{blue}{100--120}} Gt/yr.
+The SE grids have the largest positive melt bias, between {\color{blue}{240--265}} Gt/yr.
 
@@
 We compute PDD from monthly mean 2-meter temperature using the method of \citeA{CG2005JG}, assuming a fixed monthly mean standard deviation of 3$^{\circ}$C and a degree-day factor of 5~mm~d$^{-1}$ $^{\circ}$C$^{-1}$.
-{\color{blue}{This specific degree-day factor lies between typical values reported for snow and ice, in order to easily apply the PDD method to estimate the combined snow and ice melt.}}
+{\color{blue}{This specific degree-day factor lies between typical values reported for snow and ice, in order to apply the PDD method to estimate the combined snow and ice melt.}}
 
@@
 
-{\color{blue}{Changes in surface albedo may also explain the melt differences across the grids and dycores.
-The top and middle rows of Figure~\ref{fig:alb} show maps of summer albedo biases in the simulations, relative to the RACMO2.3p2 and the processed MODIS albedo dataset of \citeA{MODIS}.
-The model surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface.
+{\color{blue}{Changes in surface albedo may also explain the melt differences in the grids and dycores.
+The top and middle rows of Figure~\ref{fig:alb} show maps of summer surface albedo biases in the simulations, relative to the RACMO2.3p2 and the processed MODIS albedo dataset of \citeA{MODIS}.
+The model surface albedo is diagnosed from the incident and net shortwave radiative fluxes at the surface..
+All grids and dycores have been mapped to the $\tt f19$ grid in the figure, and only cells overlapping with the $\tt f19$ GrIS ice mask are shown.
 Both RACMO2.3p2 and MODIS datasets indicate that the albedo in the quasi-uniform grids is lower in the central-west and northwest region, relative to the other grids, whereas the albedo is anomalously low at the southwest and southeast coastal margins in both lat-lon and quasi-uniform grids, compared with the VR grids.
-All grids and dycores have a positive albedo anomaly in northeast Greenland.
+All grids and dycores have a positive albedo anomaly in northeast Greenland, which may explain the depressed melt rates and unrealistic ice sheet expansion in this region when coupled to the dynamic ice sheet \cite{LETAL2020JAMES}.
 
@@
 
-To the extent that the snow fraction changes are responsible for the more salient albedo in anomalies in the lat-lon and quasi-uniform grids, we seek to understand the snow cover changes.
+To the extent that the snow fraction changes are responsible for the albedo anomalies in the lat-lon and quasi-uniform grids, we seek to understand the snow cover changes.
 As the lat-lon and quasi-uniform grids accumulate too much precipitation in the GrIS interior, it is possible that the coastal ice margins --- where seasonal melt typically occurs --- receive too little snow, which may be more easily melted away to expose the underlying glacial ice.
-However, Figure~\ref{fig:prect} does not support a reduction in precipitation along the coastal margins, even for the solid precipitation (not shown).
+However, Figure~\ref{fig:prect} does not support a reduction in precipitation along the coastal margins, even when only plotting the solid precipitation (not shown).
 Instead it seems more likely that the warmer near surface environment, especially in the quasi-uniform grids, leads to anomalously low snow cover and consequent albedo changes.}}
